### 2019-04-30 inst/collusion_chess_draws.Rmd
start to write this up more formally

---
title: "Collusion at the 1962 chess world championships"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{collusion at the 1962 world chess championships}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

First load the package:
```{r}
library("hyper2")
options("width" = 90)
```

Now generate the dataset:

```{r}
jj <- read.table("stockholm1962.txt",header=FALSE)
results <- as.matrix(jj[,-(1:2)])
players <- as.character(jj$V1)
nationality <- as.character(jj$V2)
rownames(results) <- players
colnames(results) <- substr(players,1,1)
results
```

(this taken from Wikipedia
\url{https://en.wikipedia.org/wiki/World_Chess_Championship_1963}).
In this dataset, "0" means a loss, "1" a draw and "2" a win.  Thus
Fisher (row 1) beat Korchnoi (column 4).  As a consistency check we
verify that the results table is antisymmetric:

```{r}
all(t(results) == 2-results)
which(results != 2-t(results),arr.ind=TRUE)
```

Thus Barcza vs Portisch is problematic.  Glossing over this

```{r}
H <- hyper2(pnames=c("draw","collusion",players))

for(i in seq_len(nrow(results)-1)){  # i=row
    for(j in seq(from=i+1,to=ncol(results))){ # j = col
        o <- results[i,j]
        if((nationality[i]=="USSR") & (nationality[j]=="USSR")){
            monster <- "collusion"
        } else {
            monster <- "draw" 
        }
        if(o==2){ # row player wins
            H[players[i]] %<>% inc(1)
        } else if(o==0){  # col player wins
            H[players[j]] %<>% inc(1)
        } else if(o==1){ # game drawn
            H[monster] %<>% inc(1)
        } else {
            stop("this cannot happen")
        }
        H[c(players[i],players[j],monster)] %<>% dec(1)
    }  # j loop closes
} # i loop closes
```
and we have

```{r}
head(H)
dotchart(maxp(H),pch=16)
dotchart(log(maxp(H)),pch=16)
```

which would indicate strongly that collusion is indeed a strong
feature of the tournament.  Now some optimization.  First optimize freely:

```{r}
max_support_free <- maxp(H,give=TRUE)$value
ml_p   <- maxp(H)

s <- indep(equalp(H))
small <- 1e-3
s[1] <- s[1] + small
s[2] <- s[2] - small
```
Perform the constrained optimization:
```{r}
ml_p_constrained <- maxp(H,fcm=c(1,-1,rep(0,22)),fcv=0,startp=s)
max_support_constrained <- loglik(H,indep(ml_p_constrained))
max_support_constrained
```

Now compare the supports:

```{r}
extra_support <- max_support_free - max_support_constrained
extra_support
```

well in excess of Edwards's two-unit criterion.  Alternatively, we may
apply Wilks's theorem and observe that the test statistic is in the
tail region of its asymptotic distribution:

```{r}
pchisq(2*extra_support,df=1,lower.tail=FALSE)
```