### 2019-05-02 inst/collusion_chess_draws.Rmd
more work

---
title: "Collusion at the 1962 chess world championships"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{collusion at the 1962 world chess championships}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

First load the package:
```{r}
library("hyper2")
options("width" = 90)
```

In inference problems where the dataset is outcomes of paired
comparisons, such as chess, The Bradley-Terry model can be augmented
to accommodate draws in a number of ways.  Here I present a new
augmentation of Bradley-Terry in which a drawn chess match is regarded
as a three-way competition between the players and an entity that wins
if the match is drawn.  The resulting likelihood functions are easily
optimised numerically and I present a number of results.

Consider first the standard Bradley-Terry model for $n$ players.
First we ignore the effect of playing white.  We wish to find numbers
("strengths") $p_i,1\leqslant i\leqslant n$ with $p_i\geqslant 0, \sum
p_i=1$ such that the probability of $i$ beating $j\neq i$ is
$\frac{p_i}{p_i+p_j}$.  A suitable likelihood function, on the
assumption of independence of trials, would be

$$
\prod_{i<j}
\left(\frac{p_i}{p_i+p_j}\right)^{w_{ij}}
\left(\frac{p_j}{p_i+p_j}\right)^{l_{ij}}
$$

where $w_{ij}$ is the number of times $i$ beats $j$ and $l_{ij}$ is
the number of times $i$ loses to $j$.  Incorporation of draws is
straightforward: we consider a chess match to be a three-way trial
between the two players and an entity that wins if the match is drawn.
We thus seek a likelihood function for non-negative
$p_1,\ldots,p_n,p_D$ $p_D+\sum p_i=1$ and this would be

\[
\prod_{i<j}
\left(\frac{p_i}{p_i+p_j+p_D}\right)^{W_{ij}}
\left(\frac{p_j}{p_i+p_j+p_D}\right)^{L_{ij}}
\left(\frac{p_D}{p_i+p_j+p_D}\right)^{D_{ij}}
\]

where $W_{ij}$ is the number of games won by $i$ when playing $j$,
$L_{ij}$ is games lost and $D_{ij}$ is games drawn.  However, this
approach ignores the extra strength conferred by playing white.  We
thus consider a chess match between two players to be a three-way
entity but with a friendly ghost with strength $p_W$ whose strength is
added to the white player.

\[
\prod_{i<j}
\left(\frac{p_i+p_W}{p_i+p_j+p_D+p_W}\right)^{W_{ij}^\mathrm{white}}
\left(\frac{p_j}{p_i+p_j+p_D+p_W    }\right)^{L_{ij}^\mathrm{white}}
\left(\frac{p_i}{p_i+p_j+p_D+p_W}\right)^{W_{ij}^\mathrm{black}}
\left(\frac{p_j+p_W}{p_i+p_j+p_D+p_W    }\right)^{L_{ij}^\mathrm{black}}
\left(\frac{p_D}{p_i+p_j+p_D}\right)^{D_{ij}}
\]

where $W_{ij}^\mathrm{white}$ is the number of games won by $i$ when
playing white against $j$, and so on.  The final term is algebraically
equal to

\[
\prod_{i<j}
\left(\frac{p_D}{p_i+p_j+p_D}\right)^{D_{ij}^\mathrm{white}}
\left(\frac{p_D}{p_i+p_j+p_D}\right)^{D_{ij}^\mathrm{black}}
\]

which might be preferable on grounds of clarity.  We now apply this
likelihood model to a real dataset.

```{r}
jj <- read.table("stockholm1962.txt",header=FALSE)
results <- as.matrix(jj[,-(1:2)])
players <- as.character(jj$V1)
nationality <- as.character(jj$V2)
rownames(results) <- players
colnames(results) <- substr(players,1,1)
results
```

(this taken from Wikipedia
\url{https://en.wikipedia.org/wiki/World_Chess_Championship_1963}).
In this dataset, "0" means a loss, "1" a draw and "2" a win.  Thus
Fisher (row 1) beat Korchnoi (column 4).  As a consistency check we
verify that the results table is antisymmetric:

```{r}
all(t(results) == 2-results)
which(results != 2-t(results),arr.ind=TRUE)
```

Thus Barcza vs Portisch is problematic.  Glossing over this

```{r}
H <- hyper2(pnames=c("draw","collusion",players))

for(i in seq_len(nrow(results)-1)){  # i=row
    for(j in seq(from=i+1,to=ncol(results))){ # j = col
        o <- results[i,j]
        if((nationality[i]=="USSR") & (nationality[j]=="USSR")){
            monster <- "collusion"
        } else {
            monster <- "draw" 
        }
        if(o==2){ # row player wins
            H[players[i]] %<>% inc(1)
        } else if(o==0){  # col player wins
            H[players[j]] %<>% inc(1)
        } else if(o==1){ # game drawn
            H[monster] %<>% inc(1)
        } else {
            stop("this cannot happen")
        }
        H[c(players[i],players[j],monster)] %<>% dec(1)
    }  # j loop closes
} # i loop closes
```
and we have

```{r}
head(H)
dotchart(maxp(H),pch=16)
dotchart(log(maxp(H)),pch=16)
```

which would indicate strongly that collusion is indeed a strong
feature of the tournament.  Now some optimization.  First optimize freely:

```{r}
max_support_free <- maxp(H,give=TRUE)$value
ml_p   <- maxp(H)

s <- indep(equalp(H))
small <- 1e-3
s[1] <- s[1] + small
s[2] <- s[2] - small
```
Perform the constrained optimization:
```{r}
ml_p_constrained <- maxp(H,fcm=c(1,-1,rep(0,22)),fcv=0,startp=s)
max_support_constrained <- loglik(H,indep(ml_p_constrained))
max_support_constrained
```

Now compare the supports:

```{r}
extra_support <- max_support_free - max_support_constrained
extra_support
```

well in excess of Edwards's two-unit criterion.  Alternatively, we may
apply Wilks's theorem and observe that the test statistic is in the
tail region of its asymptotic distribution:

```{r}
pchisq(2*extra_support,df=1,lower.tail=FALSE)
```

# Now Curacao

This is the 5th candidates Curacao 1962 results table:

```{r}
jj <- read.table("curacao1962.txt",header=FALSE)
curacao <- as.matrix(jj[,-(1:2)])
players <- as.character(jj$V1)
nationality <- as.character(jj$V2)
rownames(curacao) <- players
colnames(curacao) <- players
curacao[is.na(curacao)] <- ""
curacao <- noquote(curacao)
curacao
```

Now convert to a likelihood function:
```{r}
Hcur <- hyper2(pnames=c("draw","collusion",players))
for(i in seq_len(nrow(curacao)-1)){  # i=row
  for(j in seq(from=i+1,to=ncol(curacao))){ # j = col
    r <- strsplit(curacao[i,j],'')[[1]]
    if((nationality[i]=="USSR") & (nationality[j]=="USSR")){
      monster <- "collusion"
    } else {
      monster <- "draw"
    }
    if(any(r=="W")){  # row player wins
      wins <- sum(r=="W")
      Hcur[  players[i]                    ] %<>% inc(wins)
      Hcur[c(players[i],players[j],monster)] %<>% dec(wins)
    }
    if(any(r=="L")){  # row player loses
      losses <- sum(r=="L")
      Hcur[             players[j]         ] %<>% inc(losses)
      Hcur[c(players[i],players[j],monster)] %<>% dec(losses)
    }
    if(any(r=="W")){  # row player draws
      draws <- sum(r=="D")
      Hcur[                        monster ] %<>% inc(draws)
      Hcur[c(players[i],players[j],monster)] %<>% dec(draws)
    }
  }  # j loop closes
} # i loop closes
Hcur
```

```{r}
dotchart(maxp(Hcur))
```