### 2019-05-20 inst/collusion_chess_draws.Rnw
work

% -*- mode: noweb; noweb-default-code-mode: R-mode; -*-
\documentclass[nojss]{jss}
\usepackage{dsfont}
\usepackage{bbm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{yfonts}
\usepackage{booktabs}
\usepackage{units}
\usepackage{xcolor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% just as usual
\author{Robin K. S. Hankin\\Auckland University of Technology}
\title{A generalization of the Bradley-Terry model for draws in chess}

%\VignetteIndexEntry{chess draws}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Robin K. S. Hankin}
\Plaintitle{Chess Draws}
\Shorttitle{Chess Draws}

%% an abstract and keywords
\Abstract{

In inference problems where the dataset is outcomes of paired
comparisons, such as chess, the Bradley-Terry model can be augmented
to chess-specific considerations in a number of ways.  Here I present
a new generalization of Bradley-Terry in which a chess game is
regarded as a three-way competition between the two players and an
entity that wins if the game is drawn.  Bradley-Terry can be further
modified to account for the white player advantage in a number of
ways.  Such reification affords insight into players' strengths,
response to playing black or white, and risk-aversion as manifested by
probability of drawing.  The likelihood functions arising are easily
optimised numerically.  I analyse a number of datasets of chess
results, including the infamous World Chess Championships 1963 in
which the Soviets were accused of collusion.  I conclude, on the basis
of a dataset that includes only the scorelines at the event itself,
that the scoreline of the Candidates Tournament (Cura\c{c}ao 1962)
exhibits evidence of collusion ($p<10^{-5}$), in agreement with
previous work that uses additional information such as historical
results including number of moves.}

\Keywords{Dirichlet distribution, hyperdirichlet distribution,
  combinatorics, \proglang{R}, multinomial distribution, constrained
optimization, Bradley-Terry}
\Plainkeywords{Dirichlet distribution, hyperdirichlet distribution,
  combinatorics, R, multinomial distribution, constrained
optimization, Bradley-Terry}

%% publication information
%% NOTE: This needs to filled out ONLY IF THE PAPER WAS ACCEPTED.
%% If it was not (yet) accepted, leave them commented.
%% \Volume{13}
%% \Issue{9}
%% \Month{September}
%% \Year{2004}
%% \Submitdate{2004-09-29}
%% \Acceptdate{2004-09-29}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Robin K. S. Hankin\\
  Auckland University of Technology\\
  E-mail: \email{hankin.robin@gmail.com}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand\headercell[1]{%
   \smash[b]{\begin{tabular}[t]{@{}c@{}} #1 \end{tabular}}}

\SweaveOpts{}
\begin{document}

<<echo=FALSE,print=FALSE>>=
ignore <- require(magrittr,quietly=TRUE)
ignore <- require(hyper2,quietly=TRUE)
options("width" = 90)
@

\section{Introduction: the Bradley-Terry model}

\cite{davidson1970}

\cite{joe1990}

Many situations involve pairwise comparisons that amount to a
Bernoulli trial.  The canonical application would be to two-way
competitive situations such as tennis in which one or other player is
idenfied as the winner, but the system has been applied to many
non-competitive situations such mating preference~\citep{west2008} and
consumer behaviour.  The Bradley-Terry model~\citep{bradley1952}
considers $n$ players and assigns non-negative numbers
$p_1,\ldots,p_n$ with unit sum to each player.  Then player $i$ beats
player $j$ with probability $\frac{p_i}{p_i+p_j}$.  This simple system
has been modified and extended many times, for example to account for
multi-way choices such as mating preferences~\citep{hankin2010}.

Bradley-Terry is not directly applicable to Chess for two reasons:
firstly, the white player appears to be at a considerable advantage;
and secondly, many chess games end in a draw: white wins about
$37.5\%$ of all games, black wins $27.6\%$, and about $39.4\%$ are
drawn.  We consider the first-mover advantage first, then consider the
effect of drawn games.

\paragraph{First-mover advantage} Traditionally White moves first in
any game, and this feature introduces an asymmetry into play.
Historically, we see about 1.35 white wins for every black win.
However, interpretations of this statistic vary and the issue is hotly
debated with opinions ranging from that of~\cite{adams2007}---``white
to play and win''---to~\citet{adorjan1988}---``black is OK!''.

\paragraph{Draws} Chess games may be won by either player, or may be
drawn; the implication of a draw is that neither player is better than
the other.  Before 1867, tournament games that were drawn were
replayed until one or other player won, thus reducing chess game to a
Bernoulli trial, almost surely.  This system became unworkable due to
the large number of draws, resulting in the modern practice of
awarding each player of a drawn game half a point, rendering chess a
zero-sum game.  Drawn games are generally considered to be undesirable
as they are usually poorer in spectacular or striking elements such as
sharp tactical combinations or sacrifice.  Excessive drawing dissuades
sponsors, leaves spectators feeling cheated, and gives commentators
less material to discuss.  Players themselves do not seem to value
draws in the wider context of the game: \citeauthor{fischer1969}, for
example, includes only 8 draws in his compilation of 60 memorable
games, surely not representative of tournament play in which about
39.4\% of games are drawn (binomial test, $p<10^{-5}$).

\cite{moul2009} conceptualise a player as being able to adopt a
strategy on a spectrum ranging from conservative to aggressive, thus
being able to exert some control over the probability of a draw and
suggest that risk-averse players may pursue less vigorous strategies,
trading expected payout for reduced variance.  Further,
\cite{langer2013} point out that rational players in an ongoing
competitive situation modify their strategies depending on whether
they are leading or lagging: the lagging player will adopt a higher
risk strategy (`nothing to lose'), while the leading player is
incentivised to protect the lead by lowering the risk of losing.  In a
chess context, this means the leading player will aim for a draw.

Such arguments suggest considering a drawn chess game as a
qualitatively different outcome from a statement of equality of skill:
the play itself tends to be different, the drawing players have a
different mindset, and drawn games have less value in the wider
context of the game.

\section{Generalized Bradley-Terry}

Consider first the standard Bradley-Terry model for a finite number of
chess players.  First we ignore the effect of playing white.  We wish
to find numbers (``strengths") $p_i$ for $1\leqslant i\leqslant n$
with $p_i\geqslant 0, \sum p_i=1$ such that the probability of $i$
beating $j\neq i$ is $\frac{p_i}{p_i+p_j}$.  A suitable likelihood
function, on the assumption of independence of trials, would be

\begin{equation}
\prod_{i<j}
\left(\frac{p_i}{p_i+p_j}\right)^{W_{ij}}
\left(\frac{p_j}{p_i+p_j}\right)^{L_{ij}}
\end{equation}

where $W_{ij}$ is the number of times $i$ beats $j$ and $L_{ij}$ is
the number of times $i$ loses to $j$.  Here I present a new
generalization of Bradley-Terry in which a chess game is regarded as
a three-way competition between the two players and an entity that
wins if the game is drawn.  The system is further modified to account
for first-move advantage by positing an entity whose strength is added
to that of the white player.  This methodology affords insight into
players' strengths, response to playing black or white, and
risk-aversion as manifested by probability of drawing.

If we consider a chess game to be a three-way trial, we define a
likelihood function for non-negative $p_1,\ldots,p_n,D$ with
$D+\sum p_i=1$ and this would be

\begin{equation}
\prod_{i<j}
\left(\frac{p_i}{p_i+p_j+D}\right)^{W_{ij}}
\left(\frac{p_j}{p_i+p_j+D}\right)^{L_{ij}}
\left(\frac{D  }{p_i+p_j+D}\right)^{D_{ij}}
\end{equation}

where $W_{ij}$ is the number of games won by $i$ when playing $j$,
$L_{ij}$ is games lost and $D_{ij}$ is games drawn.  However, this
approach ignores the extra strength conferred by the first-move
advantage.  We thus consider another entity whose strength is added to
the white player.  The strength of the white player is thus the sum of
the player's strength and another entity, the White Monster.
Recalling that a game might be a draw, we obtain the probabilities as
given in Table~\ref{probabilities}.  The corresponding likelihood
function is given below in equation \ref{likelihood_WD}:

\newcommand{\oo}{\vphantom{W^{W^{W^W}}}}

\begin{table}
\centering
\begin{tabular}{l|cc}\\
  &$i$ plays white & $j$ plays white\\ \hline
$\Prob{(\mbox{$i$ wins})}$  & $\frac{p_i+W\oo}{p_i+p_j+W+D}$ & $\frac{p_i}{p_i+p_j+W+D}$  \\
$\Prob{(\mbox{draw})}$      & $\frac{D\oo}{p_i+p_j+W+D}$     & $\frac{D}{p_i+p_j+W+D}$    \\
$\Prob{(\mbox{$i$ loses})}$ & $\frac{p_j\oo}{p_i+p_j+W+D}$   & $\frac{p_j+W}{p_i+p_j+W+D}$\\
\end{tabular}
\caption{Probabilities of a win, draw, loss for the two players under a generalized Bradley-Terry model\label{probabilities}}
\end{table}

\begin{align}\nonumber
\prod_{i<j}&{}
\left(\frac{p_i+W}{p_i+p_j+D+W}  \right)^{W_{ij}^\mathrm{white}}
\left(\frac{p_j}{p_i+p_j+D+W    }\right)^{L_{ij}^\mathrm{white}}\\
&{} \left(\frac{p_i}{p_i+p_j+D+W}\right)^{W_{ij}^\mathrm{black}}\label{likelihood_WD}
\left(\frac{p_j+W}{p_i+p_j+D+W  }\right)^{L_{ij}^\mathrm{black}}\\
&{} \left(\frac{D}{p_i+p_j+D    }\right)^{D_{ij}}
\nonumber
\end{align}

where $W_{ij}^\mathrm{white}$ is the number of games won by $i$ when
playing white against $j$, and so on.  The final term is algebraically
equal to

\begin{equation}
\prod_{i<j}
\left(\frac{D}{p_i+p_j+D+W}\right)^{D_{ij}^\mathrm{white}}
\left(\frac{D}{p_i+p_j+D+W}\right)^{D_{ij}^\mathrm{black}}
=
\prod_{i<j}
\left(\frac{p_D}{p_i+p_j+D}\right)^{D_{ij}^\mathrm{white}+D_{ij}^\mathrm{black}}
\end{equation}

which might be preferable on grounds of clarity.  We now apply this
and related likelihood models to real chess datasets.  The general
approach follows~\citet{edwards1972}.  To test any hypothesis $H$ on a
likelihood function $\mathcal{L}$ one first identifies the constraint
on hypothesis space corresponding to $H$.  One then finds the
unconstrained maximum likelihood $\mathcal{L}_\mathrm{free}$ and
compares it with the maximum likelihood $\mathcal{L}_\mathrm{cons}$
when the optimization is constrained by $H$.  The test statistic is
then the likelihood ratio
$\mathcal{L}_\mathrm{free}/\mathcal{L}_\mathrm{cons}$ or, equivalently
the support $\mathcal{S}=\log\mathcal{R}$.  Wilks's theorem states
that the asymptotic distribution of $2\mathcal{S}$ is $\chi^2_n$ where
$n$ is the number of degrees of freedom implemented by $H$.

The above likelihood functions are objective and reasonable
representations of chess results.  However, they differ markedly from
previous likelihood functions of draws in that they include reified
entities representing proclivity to draw $D$, and white or first-mover
advantage ($W$).

\subsection{Previous work}

\cite{davidson1970} uses a threshhold approach to interpret a draw as
performance of the players the same to within a threshhold.  Using a
logistic expression for the difference in players' performance, he
effectively uses, in current notation, $D=\nu\sqrt{p_ip_j}$.  Here we
note that such a model automatically give stronger players a higher
draw potential, and also note that it is difficult to account for
first-player advantage in this framework.  Such an approach is
numerically challenging in the current framework as the unit-sum
constraint becomes nonlinear and likelihood maximization becomes
difficult.

\subsection{Numerical considerations}

The Method of Support \citep{edwards1972} involves only numerical
optimization (not integration which is computationally expensive).
The analysis presented here uses the framework of the \pkg{hyper2}
package~\citep{hankin2017} as a convenient and unified interface.
Because gradients are available, convergence to the evaluate is
generally rapid and hypothesis-specific constraints on parameter space
are, in general, straightforward to implement.

\section{Anand, Karpov, Kasparov}

Grandmasters Anand, Karpov, Kasparov have long-standing rivalry and
have played many games amongst themselves.  A summary may be found
in Table~\ref{kka}.  Equation~\ref{likelihood_WD} becomes

\begin{table}
\centering
\begin{tabular}{@{} *{4}{c} @{}}
\headercell{\small{Plays White,}\\ \small{wins}} & \multicolumn{3}{c@{}}{\small{Plays black, loses}}\\
\cmidrule(l){2-4}
& An &  Kr & Ks        \\ 
\midrule
  An & -  &  18 &   6  \\
  Kr &  7 &  -  &  18  \\
  Ks & 15 &  30 &  -   \\
\end{tabular}\quad
\begin{tabular}{@{} *{4}{c} @{}}
\headercell{\small{Plays W,}\\ \small{draws}} & \multicolumn{3}{c@{}}{\small{Plays black, draws}}\\
 \cmidrule(l){2-4}
& An &  Kr & Ks        \\ 
\midrule
  An & -  &  20 &  20  \\
  Kr & 29 &  -  &  72  \\
  Ks & 26 &  57 &  -   \\
\end{tabular}\quad
\begin{tabular}{@{} *{4}{c} @{}}
\headercell{\small{Plays W,}\\ \small{loses}} & \multicolumn{3}{c@{}}{\small{Plays black, wins}}\\
\cmidrule(l){2-4}
& An &  Kr & Ks        \\ 
\midrule
  An & -  &   5 &   11  \\
  Kr & 13 &  -  &    9  \\
  Ks &  2 &   7 &  -   \\
\end{tabular}
\caption{Results of 365 chess games \label{kka} between Grandmaster
  Anand, Grandmaster Karpov, and Grandmaster Kasparov.  Thus, for
  example, Anand played white against Karpov a total of $18+20+13=51$
  times and won 18, drew 20, lost 13}
\end{table}

\begin{equation}\label{kka_likelihood}
\mathcal{L}\left(p_\mathrm{An},p_\mathrm{Kr},p_\mathrm{Ks},W,D\right)=
\frac{
p_\mathrm{An}^{15}
p_\mathrm{Kr}^{12} 
p_\mathrm{Ks}^{20}
D^{224}
\left(p_\mathrm{Kr} + W\right)^{25}
\left(p_\mathrm{Ks} + W\right)^{45}
\left(p_\mathrm{An} + W\right)^{24}
}{
\left(p_\mathrm{Kr} + p_\mathrm{Ks} + W + D\right)^{193}
\left(p_\mathrm{Kr} + p_\mathrm{An} + W + D\right)^{92}
\left(p_\mathrm{Ks} + p_\mathrm{An} + W + D\right)^{80}
}
\end{equation}

<<echo=FALSE,print=FALSE,results=hide>>=
maxp(karpov_kasparov_anand)
@

\begin{figure}[t]
\begin{center}
<<ddc,echo=FALSE,fig=TRUE>>=
dotchart(maxp(karpov_kasparov_anand),pch=16,xlim=c(0,0.55))
@
\caption{Maximum likelihood estimate for the strengths of the 3 Grandmasters in Table~\ref{kka} 
\label{maxp_karpov_kasparov_anand} under a likelihood defined as Equation~\ref{likelihood_WD}}
  \end{center}
\end{figure}

This likelihood function is easily maximized numerically
and this is shown visually in figure~\ref{maxp_karpov_kasparov_anand},
which suggests a number of interesting and testable statistical
hypotheses.  Firstly, we observe that $D=0$ cannot possibly be the
case: the likelihood for this hypothesis would be identically zero,
owing to the nonzero term in the numerator of the last term of
Equation~\ref{likelihood_WD}.  However, we see that the maximum
likelihood estimate for draw strength $\hat{D}\simeq 0.53$: compare
the na\"ive estimate of $\frac{224}{365}\simeq 0.61$ given by
observing 224 draws among 265 games.  The MLE is lower because the
maximization process automatically adjusts for the different players'
strengths.  Secondly, we may make inferences about $W$, the extra
strength conferred by playing white: it is conceivable that $W=0$ as
this term does not appear on its own in the numerator of any term.  We
may test $H_0\colon W=0$ using the method of support by maximizing the
likelihood subject the constraint $W=0$ and comparing with the
likelihood when the optimization is free.

<<maxlikekka,echo=FALSE,print=FALSE,results=hide>>=
small <- 1e-4

e <- function(white){
    maxp(
        karpov_kasparov_anand,
        startp=c(1/3-small,1/3-small,1/3-small,small),
        fcm=-cbind(0, 0, 0, 1),fcv= -white,give=TRUE)$value
}
ignore1 <- e(0.001)
ignore2 <- e(0.002)
ignore3 <- e(0.003)
ignore4 <- maxp(karpov_kasparov_anand,give=TRUE)$value
support <- 328.2975-335.6025
@

<<ignore34diff,echo=FALSE,print=FALSE,results=hide>>=
ignore3
ignore4
ignore3-ignore4
@

The result of numerical optimization is about 7.01 units of support,
or a likelihood ratio of $e^{7.01}\simeq 1107.7$: that is, well over
the two-units-of-support per degree of freedom suggested by
\citeauthor{edwards1972}; we can convincingly reject the hypothesis
that $W=0$ and infer that playing white confers a significant
advantage.  This is consistent with expectation, but it is now
possible to go significantly beyond mere existence of the effect and
to quantify it, accounting for the differential strengths of the
players.  Taking the MLE above, and considering a game between say
Karpov and Anand, the strengths are Karpov=0.064, Anand=0.151, white
=0.104 and draw =0.53.  Thus we would estimate that Karpov playing
white against Kasparov would win with probability
$\frac{0.064+0.104}{0.064+0.104+0.148+0.532}\simeq 0.198$, draw with
probability $\frac{532}{64+104+148+532}\simeq 0.625$, and lose with
probability $\simeq 0.174$ (compare $+0.075\ {=}0.627\ {-}0.297$ if
Karpov plays black).

<<kka_equal_strength, print=FALSE,echo=FALSE>>=
objective <- function(V){
  x <- V[1]  # strength of all three players
  w <- V[2]  # white player advantage
  D <- 1-3*x-w  # Draw monster
  return(  
      -loglik(karpov_kasparov_anand,c(x,x,x,w))  # negative as we minimize
  )
}

jj <-
constrOptim(
    theta=c(0.1,0.4),
    f=objective,
    grad=NULL,
    ui = rbind(diag(2),-diag(2),c(-3,-1)),
    ci = c(0,0,-1,-1,-1)
)

unconstrained <- maxp(karpov_kasparov_anand,give=TRUE)$value
constrained <- jj$value

extra_support <-   unconstrained + constrained  # negative used
pvalue <- pchisq(2*extra_support,df=2,lower.tail=FALSE)
@

One very natural and interesting hypothesis would be that the three
players are of equal strength
%equal_strength_is_assessed_in_chunk_kka_equal_strength_above____________
and we might hypothesise that
$p_\mathrm{Kr}=p_\mathrm{Ks}=p_\mathrm{An}$.  This too is readily
testable by the device of maximizing the support, subject to the
constraint that $p_\mathrm{Kr}=p_\mathrm{Ks}=p_\mathrm{An}$.  In this
case the extra support in going from the constrained to the free
optimization is about 4.93, or a likelihood ratio of about $138$.
However, this time the constraint corresponds to a loss of {\em two}
degrees of freedom so we would need 4 units of support, which is
attained in this case.  Alternatively we might follow Wilks, and
observe that the value of $2\mathcal{S}=2*4.92=9.84$ lies in the tail
region of its asymptotic $\chi^2_2$ distribution with a $p$-value of
about 0.0072.  We can justifiably assert that the players do indeed
have different strengths, even if we account for white advantage and
the proclivity to draw.

\subsection{Each player has a draw monster}

As argued above, players with a cautious style tend to draw and this
is an important and interesting component of the game.  To investigate
this we need a likelihood function that can account for possible
differences in draw proclivity among the players.  This may be
achieved by considering a game between two players to be a contest
between five entities: the two real players, the white advantage, and
two {\em personalised} draw entities, one for each real player; a
drawn game is counted as a win for either of the draw ``ghosts''.  Under
these assumptions, table~\ref{kka} translates into the following
likelihood function:

%\begin{figure}[htbp]
%\begin{center}
%<<ddcc,echo=TRUE,fig=TRUE>>=
%pie(maxp(kka_3draws))
%@
%\caption{Maximum likelihood estimate for the strengths of the 3 Grandmasters in Table~\ref{kka} 
%under a likelihood function that admits differential proclivities to draw}
%  \end{center}
%\end{figure}

\begin{equation}\label{KKA_personalised_drawmonster}
\mathcal{L}\left(
p_\mathrm{An},p_\mathrm{Kr},p_\mathrm{Ks}
,W,
d_\mathrm{An},d_\mathrm{Kr},d_\mathrm{Ks}
\right)=
\frac{
\parbox{3in}{$
p_\mathrm{An}^{15}
p_\mathrm{Kr}^{12}
p_\mathrm{Ks}^{20}
\left(p_\mathrm{Kr} + W\right)^{25}
\left(p_\mathrm{Ks} + W\right)^{45} 
\left(p_\mathrm{An} + W\right)^{24}\\ \rule{10mm}{0mm}
\left(d_\mathrm{Kr} + d_\mathrm{Ks}\right)^{129}
\left(d_\mathrm{Kr} + d_\mathrm{An}\right)^{49}
\left(d_\mathrm{Ks} + d_\mathrm{An}\right)^{46}
$}}{\parbox{3in}{$
\left(p_\mathrm{Kr} + p_\mathrm{Ks} + W + d_\mathrm{Kr} + d_\mathrm{Ks}\right)^{193}\\ \rule{10mm}{0mm}
\left(p_\mathrm{Kr} + p_\mathrm{An} + W + d_\mathrm{Kr} + d_\mathrm{An}\right)^{92}\\ \rule{20mm}{0mm}
\left(p_\mathrm{Ks} + p_\mathrm{An} + W + d_\mathrm{Ks} + d_\mathrm{An}\right)^{80}
$}}
\end{equation}

\begin{figure}[t]
\begin{center}
<<echo=FALSE,fig=TRUE>>=
dotchart(maxp(kka_3draws),pch=16)
@ 
\caption{Maximum likelihood estimate for \label{kka_3draws} equation \ref{KKA_personalised_drawmonster} equation} 
  \end{center}
\end{figure}

In figure~\ref{kka_3draws}, we see considerable variability in
proclivity to draw and again this may be tested with the Method of
Support, which gives a $p$-value of about 0.039, borderline
significant: an interesting result, given that in this case no draw
monster appears alone anywhere in the likelihood function.

<<echo=FALSE,print=FALSE,results=hide>>=
objective <- function(V){
  karpov <- V[1]
  kasparov <- V[2]
  anand <- V[3]
  white <- V[4]

  draw <- (1-karpov-kasparov-anand-white)/3

    return(  
        -loglik(kka_3draws,
                c(karpov,kasparov,anand,white,draw,draw)
                )
        )
}

jj <-
constrOptim(
    theta=c(0.1,0.1,0.1,0.1),
    f=objective,
    grad=NULL,
    ui <- rbind(
        diag(nrow=4),
        c(-1,-1,-1,-1)
        ),
    ci = c(
        rep(0,4),
        -1
        )
)

unconstrained <- maxp(kka_3draws,give=TRUE)$value
constrained <- jj$value
extra_support <-   unconstrained + constrained
pchisq(2*extra_support,df=2,lower.tail=FALSE)
@

\subsection{Each player has a distinct White Monster}

We now consider the possibility that the strength conferred by playing
white, is specific to each player and we denote this $w_\mathrm{An}$,
$w_\mathrm{Kr}$ and $w_\mathrm{Ks}$.  The corresponding likelihood
function would be

\begin{equation}\label{KKA_personalised_whitemonster}
%\mathcal{L}\left(
%p_\mathrm{An},p_\mathrm{Kr},p_\mathrm{Ks},
%w_\mathrm{An},w_\mathrm{Kr},w_\mathrm{Ks},D
%\right)=
\frac{
%\parbox{3in}
%       {$
p_\mathrm{An}^{15}
p_\mathrm{Kr}^{12}
p_\mathrm{Ks}^{20}
D^{224}
\left(p_\mathrm{An} + w_\mathrm{An}\right)^{24}%\\ \rule{10mm}{1mm}
\left(p_\mathrm{Kr} + w_\mathrm{Kr}\right)^{25}
\left(p_\mathrm{Ks} + w_\mathrm{Ks}\right)^{45}
%$}
}{\parbox{4in}{$
\left(p_\mathrm{An} + p_\mathrm{Kr} + w_\mathrm{An} +  D\right)^{43}
\left(p_\mathrm{An} + p_\mathrm{Kr} + w_\mathrm{Kr} +  D\right)^{49}\\ \rule{10mm}{0mm}
\left(p_\mathrm{An} + p_\mathrm{Ks} + w_\mathrm{An} +  D\right)^{37}
\left(p_\mathrm{An} + p_\mathrm{Ks} + w_\mathrm{Ks} +  D\right)^{43}\\ \rule{20mm}{0mm}
\left(p_\mathrm{Kr} + p_\mathrm{Ks} + w_\mathrm{Kr} +  D\right)^{99}
\left(p_\mathrm{Kr} + p_\mathrm{Ks} + w_\mathrm{Ks} +  D\right)^{94}
$}}
\end{equation}

In the likelihood function above we see, for example, that Anand won a
total of $18+6=24$ games playing white, $13+2=15$ games playing
Black, and that Anand played white against Karpov a total of
$18+20+5=43$ times.  We fail to reject the null of
$w_\mathrm{An}=w_\mathrm{Kr}=w_\mathrm{Ks}$, having a likelihood ratio
of $4.5$, or an asymptotic $p$-value of $0.076$.  There is no strong
evidence that the players differ in regard to their white advantage.

\section{Collusion in the World Chess Championship 1963}

The method of support allows one to account for proclivity to draw
while accounting for players' strengths and first-move advantage.  We
now consider one famous incident, following the World Chess
Championships 1963, in which Fischer famously and publicly claimed
that the Soviet players colluded.  The accusation was that Soviet
players agreed ahead of time to draw when playing one another, in
order to conserve energy for subsequent matches.

\citet{moul2009} consider this accusation from a statistical
perspective and present a detailed analysis of Cura\c{c}ao.  Unlike
the present analysis, which uses a dataset restricted to game result
(win/loss/draw), \citet{moul2009} included number of moves in each
game\footnote{The reasoning was that hard-fought games that happened
  to end in a draw are more likely to last longer than collusive or
  pre-arranged draws.} in their analysis, and used a dataset of 7355
games played in the period 1940-1978 to calibrate the ability of the
players.  In contrast, the analysis presented uses a dataset
restricted to game outcome (win-loss-draw) in the competition itself.

 \begin{table}
 \centering
 \begin{tabular}{ll|cccccccc}
 &
 &%\rotatebox{-90}{Petrosian} 
 &\rotatebox{-90}{Keres} 
 &\rotatebox{-90}{Geller}
 &\rotatebox{-90}{Fischer} 
 &\rotatebox{-90}{Korchnoi}
 &\rotatebox{-90}{Benko}
 &\rotatebox{-90}{Tal}
 &\rotatebox{-90}{Filip}\\
 Petrosian &USSR    &             &DDDD    &DDDD   &DWDD   &DDWW   &DDWD   &WWD    &DWWD\\
 Keres     &USSR    &             & -      &DDDD   &LDWD   &DDWD   &WWWL   &WDW    &DWWD\\
 Geller    &USSR    &             & -      & -     &WWDL   &DDWD   &DDDW   &DWW    &DWWD\\
 Fischer   &USA     &             & -      & -     & -     &LWLD   &LWDW   &DWD    &WDWD\\
 Korchnoi  &USSR    &             & -      & -     & -     & -     &DDDL   &WLD    &WWWW\\
 Benko     &USA     &             & -      & -     & -     & -     & -     &WLD    &LWWD\\
 Tal       &USSR    &             & -      & -     & -     & -     & -     & -     &WLD \\
 Filip     &TCH     &             & -      & -     & -     & -     & -     & -     & -  \\
 \end{tabular}
 \caption{Cura\c{c}ao 1962 \label{candidates} Candidates, World Chess Championship 1963}
 \end{table}

Here I analyse the original subject of the complaint, the Cura\c{c}ao
1962 Candidates, Table~\ref{candidates}.  Fischer alleged that three
of the Soviet players (Keres, Petrosian, and Geller) had agreed to
draw amongst themselves.  Observe that all twelve of the games between
the three accused players ended in a draw, a significant result given
51 winning games and 54 draws (Fisher exact test, one-sided,
$p=0.0079$).  The likelihood function for this dataset would have

\begin{equation}\parbox{4in}{$
P^5\,
{Ke}^5\,
G^4
\mathit{Fis}\,  % Fischer, cf Filip
\mathit{Ko}^5
B
T^2  % Tal
\mathit{Fil}\,
D^{42}C^{12}\oo\\ \rule{10mm}{0mm}
\left(P+W\right)^3
\left(\mathit{Ke}+W\right)^4
\left(G+W\right)^4\
\left(\mathit{Fis}+W\right)^7\oo\\ \rule{20mm}{0mm}
\left(\mathit{Ko}+W\right)^2
\left(B+W\right)^5
\left(T+W\right)
\left(\mathit{Fil}+W\right)\oo
$}\end{equation}

\noindent on the numerator (where $C$ is the strength of a collusive
draw), and a total of 36 terms such as
$\left(P+\mathit{Fis}+W+D\right)^{4}$ and
$\left(P+\mathit{Ke}+W+C\right)^{4}$ on the denominator.  The
likelihood ratio for the hypothesis that collusive draws have the same
strength as regular draws isabout $e^{11.8}\simeq 10^5$: we conclude
that the high number of draws among the three accused players was
unlikely to be due to chance.

One plausible and charitable explanation for this result is that
Soviet players have a higher propensity to draw than western players.
We can investigate this by considering the interzonal tournament which
took place in Stockholm in 1962.

\subsection{Chess collusion: interzonal Stockholm 1962}

\newcommand{\halfw}{$\nicefrac{1}{2}$}
\newcommand{\halfb}{\colorbox{black}{\textcolor{white}{$\nicefrac{{\bf 1}}{{\bf 2}}$}}}
\newcommand{\zerow}{$0$}
\newcommand{\zerob}{\colorbox{black}{\textcolor{white}{$0$}}}
\newcommand{\oneww}{$1$}
\newcommand{\onebb}{\colorbox{black}{\textcolor{white}{$1$}}}

\begin{table}
\centering
\begin{tabular}{l|cccccc}
&\rotatebox{-90}{Fischer} 
&\rotatebox{-90}{Geller} 
&\rotatebox{-90}{Petrosian\ }
&\rotatebox{-90}{Korchnoi} 
&\rotatebox{-90}{Filip}
& $\cdots$
\\ \hline %       Fischer   Geller    Petrosian Korchnoi    Filip
Fischer    (USA)  &   -     & \halfb  & \halfw  & \oneww    &\halfb    &\\
Geller    (USSR)  & \halfw  &    -    & \halfb  & \halfw    & \halfw   &\\
Petrosian (USSR)  & \halfb  & \halfw  &    -    & \halfb    & \halfw   &\\
Korchnoi  (USSR)  & \zerob  & \halfb  & \halfw  &   -       & \oneww   &\\
Filip      (TCH)  & \halfw  & \halfb  & \halfb  & \zerob    &  -       &\\
$\vdots$&&&&&\\
\end{tabular}
\caption{Subset of the Fifth Interzonal round-robin, Stockholm 1962.\label{interzonal}
  Results $\in\{0,\halfw,1\}$ and box colour refer to row
  player.  Thus, for example, Fischer played black against Geller and drew}
\end{table}

For the results shown in table~\ref{interzonal}, we would have

\begin{equation}
\mathcal{L}\left(F,G,P,K,L,W,D\right)=
%                D/(F+G+W+D)  (F+P+D)/(F+P+D+W)   (F+W)/(F+W+D+K)   D/(F+W+L+D)
%                             D/(G+P+D+W)         D/(G+K+W+D)       D/(G+F+W+D)
%                                                 D/(P+K+W+D)       D/(P+L+W+D)
%                                                                   (K+W)/(K+L+W+D)
\frac{
  D^8(F+W)(K+W)
}{\parbox{4in}{$
    (F+G+W+D)(F+P+W+D)\\ \rule{10mm}{0mm}(F+K+W+D)  (F+L+W+D)\\ \rule{20mm}{0mm}
         (G+P+W+D)  (G+K+W+D)\\ \rule{30mm}{0mm}(G+F+W+D)
                    (P+K+W+D)\\ \rule{40mm}{0mm}(P+L+W+D)
                               (K+L+W+D)
 $}}
\end{equation}

\noindent (the numerator includes 8 draws and 2 wins, one for Fischer
and one for Korchnoi, both playing white); but the full likelihood
function for the interzonal has 297 terms, including $D^{110}$.  To
investigate the possibility of Soviet collusion we introduce a new
draw entity: a soviet Draw monster $D_s$, who wins when two Soviet
players draw, and is inactive whenever a game has at least one
non-Soviet player.  Then $H_N\colon D_s=D$ corresponds to no collusion
and the likelihood function for table~\ref{interzonal} would be

\begin{equation}\label{collusion}
\mathcal{L}\left(F,G,P,K,L,W,D,D_s\right)=
%                D/(F+G+W+D)  (F+P+D)/(F+P+D+W)   (F+W)/(F+W+D+K)   D/(F+W+L+D)
%                             D/(G+P+D+W)         D/(G+K+W+D)       D/(G+F+W+D)
%                                                 D/(P+K+W+D)       D/(P+L+W+D)
%                                                                   (K+W)/(K+L+W+D)
\frac{
  D_s^3D^5(F+W)(K+W)
}{\parbox{4in}{$
    (F+G+W+D)(F+P+W+D)\\ \rule{10mm}{0mm}(F+K+W+D)  (F+L+W+D)\\ \rule{20mm}{0mm}
         (G+P+W+D_s)  (G+K+W+D_s)\\ \rule{30mm}{0mm}(G+F+W+D)
                    (P+K+W+D_s)\\ \rule{40mm}{0mm}(P+L+W+D)
                               (K+L+W+D)
 $}}
\end{equation}

The maximum likelihood estimate is $D\simeq 0.043, D_s=0.354$; a
constrained maximization gives a likelihood ratio of about 5096, or a
$p$-value of about $3.6\times 10^{-5}$.  Indeed we could go further
and plot a profile likelihood curve for the log-odds ratio
$\ln\left(D_s/D\right)$ which, if Dirichlet, would be accurately
parabolic over a wide range~\cite[page 343]{ohagan2004}.  This is
illustrated in Figure~\ref{proflike}, that shows a credible interval
of about 1.8-2.4.

<<echo=FALSE,print=FALSE>>=
do_from_scratch  <- FALSE  # set to TRUE to recalculate (slow, maybe 2 mins)
hotstart <- indep(maxp(interzonal_collusion))

f <- function(LO,give=FALSE){
    small <- 1e-5

    eq <- hotstart
    m <- sum(eq[2:3])
    a <- exp(LO)
    eq[2:3] <- m/(1+a)*c(1,a)  # exact

    eq[2:3] <- eq[2:3] + c(-small,small)

    out <-
        maxp(collusion,
             startp=eq,
             fcm=c(0, -1, exp(LO),rep(0,22)),
             fcv=0,
             give=TRUE)

    if(give){return(out)}else{return(out$value)}
}

x <- seq(from= 1.5,to = 2.7,len=32)
if(do_from_scratch){
    jj <- sapply(x,f)
} else {
    jj <-
        c(-246.6939, -245.7192, -244.7981, -243.9319, -243.1223, -242.3704, 
          -241.6776, -241.0452, -240.4744, -239.9661, -239.5216, -239.1417, 
          -238.8274, -238.5795, -238.3987, -238.2858, -238.2414, -238.2660, 
          -238.3601, -238.5240, -238.7582, -239.0629, -239.4382, -239.8843, 
          -240.4013, -240.9891, -241.6476, -242.3767, -243.1762, -244.0458, 
          -244.9853, -245.9941)
}
@

\begin{figure}[htbp]
\begin{center}
<<proflike,echo=FALSE,fig=TRUE>>=
plot(x,jj-max(jj),type='b',xlab="log-odds",ylab="support")
abline(h=c(0,-2))
@
\caption{Profile likelihood \label{proflike} for odds ratio
  $\log\left(D_s/D\right)$: note the accurately parabolic figure in
  the region plotted}
  \end{center}
\end{figure}

\section{Conclusions and further work}

The present work details an approach to chess analysis based on
reification of entities representing propensity to draw, and the
first-mover (or white player) advantage.  Likelihood functions for the
players' and reified entities' strengths are defined using only
specific tournament outcomes; further, the analysis uses only game
outcome, not length.

In this context we note, following \cite{langer2013}, that a player
may well adopt a riskier strategy after a sequence of defeats,
adopting a ``nothing-to-lose'' mentality.  Such a phenomenon would be
detectable, at least in principle, by generalizing the personalised
draw likelihood function \ref{KKA_personalised_drawmonster} and
seeking evidence for the draw strength changing as a tournament
progresses: it would become stronger for a player on a winning streak,
and weaker for a losing player.

\cite{moul2009} present detailed Monte Carlo simulations
to assess the likely impact of any Soviet cartel on tournament outcomes.
In principle, the present system of likelihood and reified entities
would admit such simulations and might be interesting to compare results.

\bibliography{chess}
\end{document}