### 2019-09-23 inst/order.Rmd
some reflections on order

---
title: "Order statistics and numerical optimization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The `hyper2` package

We consider the likelihood function given in section 5 (Men's sculling
in the 2016 Olympics) of the package vignette.  Here we investigate
some numerical problems arising from the use of `optim()` and
`constrOptim()`.  We will consider $n=10$ competitors and without loss
of generality, the competitors finish in order $1,2,3,\ldots,10$.  In
the vignette we see a formula

\[
{\mathcal L}\left(p_1,\ldots,p_{10}\right)=
\frac{p_1}{p_1+\cdots+p_{10}}\cdot
\frac{p_2}{p_2+\cdots+p_{10}}\cdot
\frac{p_3}{p_3+\cdots+p_{10}}\cdot
\frac{p_4}{p_4+\cdots+p_{10}}\cdot
\frac{p_5}{p_5+\cdots+p_{10}}\cdot\\
\frac{p_6}{p_6+\cdots+p_{10}}\cdot
\frac{p_7}{p_7+\cdots+p_{10}}\cdot
\frac{p_8}{p_8+p_9+p_{10}}\cdot
\frac{p_9}{p_9+p_{10}}\cdot
\frac{p_{10}}{p_{10}}
\]

We can calculate the evaluate (that is, the maximum likelihood
estimate) from the formula by inspection.  It is clear from the
penultimate term that $p_{10}=0$.  Similarly from the preceding term
it is easy to see that $p_9=0$ and so on.  The evaluate is thus
$p=(1,0,\ldots,0)$.  Let's see if we can reproduce this using
numerical methods.

The `hyper2` idiom for creating a likelihood function is as follows:

```{r defineH}
library("hyper2")
H <- order_likelihood(rbind(1:10))
H
```
And then maximizing it is carried out using the `maxp()` function:

```{r}
maxp(H)
loglik(H,indep(maxp(H)))
```

```{r}
dotchart(maxp(H))
```

So the numerics have not done a particularly good job in identifying
the evaluate, presumably as a result of the difficulty of honouring
the constraints.  However, if we give the system a hand by starting at
a point closer to the evaluate we can do much better:

```{r}
f <- function(p,n){c(p,rep((1-p)/n,n-1))}
sss <- f(0.9999,9)  # sensible start point, close to the evaluate
maxp(H,startp=sss)
loglik(H,indep(maxp(H,startp=sss)))
```

```{r}

H <- hyper2(
    pnames=c("Rahul","Kim-Joy","Ruby","Briony","Manon","Jon",
             "Dan","Karen","Terry","Antony","Luke","Imelda")
)
H
```

```{r}
	ranks <- c(Antony=12, Briony=2, Dan=4, Imelda=11, Jon=5, Karen=6, KimJoy=10, Luke=8, Manon=3, Rahul=7, Ruby=1, Terry=9)
```