### 2019-12-08 inst/loser.Rmd
slight copyediting

---
title: "Weakly informative observations: identifying the weakest player"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(hyper2)
```

# Partial rank observations: who lost?

Here we consider a situation in which we make only a rather
uninformative observation: namely, the identity of the _loser_ (which
is here taken to be me, player `a`).  We have six competitors
`a,b,c,d,e,f` with nonnegative strengths $p_1,p_2,p_3,p_4,p_5,p_6,
\sum p_i=1$.  Competitor `a` has strength $p_1$, and loses.  We are
interested in how informative this observation is with regard to
various assertions about the strength of `a`.  We have three
hypotheses:

* $H_1\colon p_1=0$
* $H_2\colon\exists i\neq 1 \mbox{ with } p_1\geq p_i>0$
* $H_3\colon p_1=p_2=\ldots=p_n=\frac{1}{n}$

(observe that $H_2$ implies that someone is worse than me).  All three
hypotheses correspond to a restriction on allowable parameter space.
The first step is to define a suitable likelihood function for our
observation:

```{r}
n <- 6 # number of competitors
H <- hyper2(pnames=letters[1:6])
L <- ggol(H,letters[2:n],letters[1]) 
length(L)
L[[1]]
L[[2]]
```

In the above, `L` is a log-likelihood function for the observation
that competitor `a` came last.  In R idiom, it is a list of hyper2
objects corresponding to all $5!=120$ possible observed order
statistics (only the first two, `L[[1]]` and `L[[2]]`, are printed,
which correspond to orders `bcdefa` and `bcdfea` respectively).  Any
order is permissible, so long as player `a` is bottom.

## Hypothesis 1

Now we maximize the likelihood
under $H_1$, that is, perform an unconstrained maximization:

```{r}
(ans_unconstrained <- maxp(L,give=TRUE))
```

The constraint embodied in $H_1$ does not need to be enforced: the
evaluate $\hat{p}$ has $p_1=0$, consistent with competitor `a` coming
last; actually we have
$\hat{p}=\left(0,\frac{1}{5},\frac{1}{5},\frac{1}{5},\frac{1}{5},\frac{1}{5}\right)$
analytically, and the numerics are not too far from this.  The
likelihood at the evaluate is 1 (technically, this is wrong because
likelihood is defined only up to a multiplicative constant.  A correct
statement is: the likelihood function evaluated by
`like_single_list()` returns 1 at $\hat{p}$ returns 1).

## Hypothesis 2

The next step is to perform a constrained maximization, the constraint
being $H_2$.  This says that I am better than at least one other
person, here chosen to be, without loss of generality, competitor `b`:

```{r}
small <- 1e-2  # numerical necessity
ans_constrained <-
maxp(L,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,-1,rep(0,n-3))),fcv=0,give=TRUE)
ans_constrained
```

Observe that the constraint is active: $p_1=p_2$ in the numerical
result.  Analytically, the evaluate is indeterminate in
$p_3,p_4,p_5,p_6$, but has $p_1=p_2=0$ (numerically this means that
$p_1=p_2=\epsilon$, where $\epsilon$ is some small number chosen for
smallness and numerical expediency).

## Hypothesis 3

Recall that $H_3\colon p_1=p_2=\ldots=p_n=\frac{1}{n}$; this is
equivalent to a random chance.  Because this is a point hypothesis, we
do not need to use any maximization techniques; it is sufficient to
evaluate the likelihood at this point:

```{r}
like_single_list(indep(equalp(L[[1]])),L)
```

The likelihood can be shown to be $\frac{1}{n}$ (which is clear on
intuitive grounds: $H_3$ says that all players are equivalent and the
order statistic is totally random.  If this is correct, then one would
identify player `a` as the worst with a probability $\frac{1}{n}$).
Note again that this is loose language as likelihood is only defined
up to a multiplicative constant.

# Discussion

We are now in a position to define a likelihood function on the three
hypotheses $H_1,H_2,H_3$:

$$
{\mathcal L}(H_1)=1\qquad
{\mathcal L}(H_2)=\frac{1}{2}\qquad
{\mathcal L}(H_3)=\frac{1}{n}
$$

(recall that $n$ is the number of competitors).  We can see that the
likelihood ratio ${\mathcal L}(H_2)/{\mathcal L}(H_1)$ is 0.5 for an
observation that competitor `a` comes last.  This is exactly the same
as the likelihood ratio for the following situation: we toss a coin
which is either fair ($H_\mbox{fair}$) or two-headed
($H_\mbox{2heads}$).  We observe a head.  Then ${\mathcal
L}(H_\mbox{2heads})/{\mathcal L}(H_\mbox{fair}) = 0.5$ as well (this
is Edwards's valet and black and white ball thought experiment).

## Repeated observations

Suppose we make $r$ independent observations, all of which are that
competitor `a` comes last.  Then the likelihood function becomes

$$
{\mathcal L}(H_1)=1\qquad
{\mathcal L}(H_2)=\frac{1}{2^r}\qquad
{\mathcal L}(H_3)=\frac{1}{n^r}
$$

The support for $H_1$ against $H_2$ would be $\log(2^r)=r\log 2$.  To
exceed Edwards's criterion of 2 units of support per degree of freedom
would require $2/\log(2)\simeq 2.9$ observations, or 3 rounded up;
alternatively, we might require the log-likelihood to lie in the tail
area of its asymptotic $\chi^2_1$ distribution, that is

```{r}
qchisq(0.95,1)
```

that is $2r\log 2\geq 3.84$ or $r\geq 2.77$.

The support for $H_1$ against $H_3$ would be $r\log n$ but this time
we need $n-1$ degrees of freedom.  The likelihood criterion would be
$r\log n\geq 2(n-1)$, or $r\geq 2(n-1)/\log n$.  Alternatively we
might adopt a frequentist view and require
$2\log(n^r)\geq\chi^2_{n-1}(0.95)$, or $r\geq Q/(2\log n)$, where Q is
the 95th percentile of the chi-square distribution with $n$ degrees of
freedom, `qchisq(0.95,df=n)`.  It might be interesting to compare
these two approaches.