### 2019-12-08 inst/counterstrike.Rmd
nice Rmd file

---
title: "A statistical analysis of a counterstrike game: Boston 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("partitions") ## needed for perms()
library("magrittr") 
```
# Abstract

In this short vignette I use the `hyper2` package to analyse an
influential Counter-Strike match between two professional teams (_FaZe
Clan_ and _Cloud9_) in the major championships at Boston 2018.  I use
likelihood methods to demonstrate that one player, Skadoodle, has a
legitimate claim to be the strongest player on the field; but the data
is indistinguishable from a null of equal player strength.
Counter-strike dataset kindly supplied by Zachary Hankin.

# Counterstrike

E-sports are a form of competition using video games.  E-sports are
becoming increasingly popular, with high-profile tournaments attracting
over 400 million viewers, and prize pools exceeding US$20m.

_Counter Strike: Global Offensive_ (CS-GO) is a multiplayer
first-person shooter game in which two teams of five compete in an
immersive virtual reality combat environment.  CS-GO is distinguished
by the ability to download detailed gamefiles in which every (sic)
aspect of an entire match is recorded, it being possible to replay the
match at will.

Statistical analysis of such gamefiles is extremely difficult,
primarily due to complex gameplay features such as cooperative
teamwork, within-team communication, and real-time strategic fluidity.

## Boston 2018

"ELEAGUE Major: Boston 2018", here denoted "Boston 2018", was a major
championship held in Boston Massachusetts from 26-28 January 2018
featuring 24 professional teams; the purse was US$1000000.

The final pitted _FaZe Clan_ against _Cloud9_, with _Cloud9_ beating
favourites _FaZe Clan_ 2-1.

# Dataset

The dataset used here is a list whose six elements correspond to six
rounds of play, which are assumed to be statistically independent.
The R idiom is:

```{r, zachslist}
zachslist <- list(
    round1 =
      c("Skadoodle","olofmeister","tarik",
        "GuardiaN", "RUSH", "rain","Stewie2k",
        "karrigan","autimatic","NiKo"
        ),
    round2 =
      c("karrigan", "NiKo", "Stewie2K",
        "RUSH", "rain", "GuardiaN",
        "autimatic", "olofmeister"
        ),
    round3 =
      c("rain","tarik","autimatic",
        "karrigan","RUSH","GuardiaN",
        "Stewie2K","NiKo","olofmeister"
        ),
    round4 =
      c("rain","GuardiaN","karrigan",
        "NiKo","olofmeister"
        ),
    round5 =
      c("olofmeister","rain","karrigan",
        "tarik","Stewie2K","autimatic"
        ),
    round6 =
      c("GuardiaN","karrigan")
)
```

(data kindly supplied by Zachary Hankin).  Thus in round 1, the death
order was Skadoodle first, followed by olofmeister, then tarik.

## Likelihood function for a simple deathorder statistic

Suppose `team1 = (a,b,c,d)` and `team2 = (x,y,z)`; suppose the
deathorder were `(a,b,y,c)`.  Then the likelihood function for `a`'s
death would be `(x+y+z)/(a+b+c+d+x+y+z)` [that is, the likelihood
function for a Bernoulli trial between team `(x,y,z)` and
`(a,b,c,d)`].  Note the appearance of (`a`) in the denominator: he
[surely!] was on the losing team.  The likelihood function for the
full deathorder is then

```
 (x+y+z)/(a+b+c+d+x+y+z) * (x+y+z)/(b+c+d+x+y+z) * (c+d)/(c+d+x+y+z) * (x+z)/(c+d+x+z)
```

where the four terms correspond to the deaths of `a,b,y,c`
respectively.  See how the denominator gets shorter as the teams die
one by one.

# Likelihood function for the observed dataset

```{r, teamsetup}
team1  <- c("autimatic","tarik", "Skadoodle","Stewie2k","RUSH")   #Cloud9
team2 <- c("NiKo","olofmeister","karrigan","GuardiaN","rain")  # FaZe Clan
```

Consider the following function:

```{r, deffun}
`counterstrike_maker` <- function(team1,team2,deathorder){

  if(identical(sort(c(team1,team2)),sort(deathorder))){
    deathorder <- deathorder[-length(deathorder)]  # last player not killed
  }

  H <- hyper2(pnames=c(team1,team2))

  for(killed_player in deathorder){
    if(killed_player %in% team1){
      H[team2] %<>% inc
      H[c(team1,team2)] %<>% dec
      team1 %<>% "["(team1 != killed_player)      
    } else {
      H[team1] %<>% inc
      H[c(team1,team2)] %<>% dec
      team2 %<>% "["(team2 != killed_player)      
    }
  }
  return(H)
}
```

Function `counterstrike_maker()` defined above returns a
log-likelihood function for the strengths of the players in `team1`
and `team2` above.  The function needs the identities of all players
on each team.  It assumes that players are always killed by the
opposing team; if a player is killed by his own team, function
`counterstrike_maker()` may return an error.  The function does not
have strong error-checking functionality.

In the function, the `deathorder` argument specifies the order in
which players were killed (the first element is the first player to be
killed and so on).  Note that the identity of the killer is not needed
as it is assumed that a death is due to the combined strength of the
team, rather than the individual shooter who actually fired the shot.
Note that it is not required for all the players to be killed;
short rounds are OK (but are not as statistically informative).

We can use it:

```{r, useit}
H <- hyper2(pnames=c(team1,team2))
for(i in zachslist){
  H <- H + counterstrike_maker(team1,team2, deathorder=i)
}
```

then plot it:

```{r, plotit}
dotchart(maxp(H),pch=16,main='observed data')
```

showing that Skadoodle has the highest estimated strength.

# Random data

it is possible to analyse synthetic data generated randomly by
function `rdeath()`:

```{r, randomdata}

`rdeath` <- function(team1,team2){
  out <- NULL
  while(length(team1)>0 & length(team2)>0){
    if(runif(1)>length(team1)/(length(c(team1,team2)))){ ## killed person is on team1
      killed_player <- sample(team1,1)
      team1 %<>% "["(team1 != killed_player)      
    } else { ## killed person is on team2
      killed_player <- sample(team2,1)
      team2 %<>% "["(team2 != killed_player) 
    }
    out <- c(out,killed_player)
  }
  return(out)
}

Hrand <- hyper2(pnames=c(team1,team2))
for(i in 1:6){
  pp <- rdeath(team1,team2)
  Hrand <- Hrand + counterstrike_maker(team1,team2, pp)
}

dotchart(maxp(Hrand),pch=16,main='synthetic data')
```

We now run $n=1000$ trials on the assumption of equal player strength
(that is, random deathorder) and for each trial calculate the
log-likelihood ratio [for the hypothesis p=equal_strengths, vs the
hypothesis p=maxp(Hrand)].  It then draws a histogram of the
loglikelihood ratio thus obtained, and superimposes a line showing the
observed loglikelihood ratio.

```{r, random2}
equalstrengths <- rep(0.1, 9)
datapoint <- loglik(indep(maxp(H)),H) - loglik(equalstrengths,H)

n <- 100  # takes about 5 minutes for n=1000 trials
distrib <- rep(0,n)
for(i in seq_len(n)){
  Hrand <- hyper2(pnames=c(team1,team2))
  for(jj in 1:6){
    pp <- rdeath(team1,team2)
    Hrand <- Hrand + counterstrike_maker(team1,team2, pp)
	}
  distrib[i] <- loglik(indep(maxp(Hrand)),Hrand) - loglik(equalstrengths,Hrand)
}

hist(distrib)
abline(v=datapoint,lwd=6)
```

Thus we can see that the observation is not significant: from this
perspective, the deathorder may as well be totally random.