### 2020-02-03 inst/notthewinner.Rmd
more work

---
title: "Weakly informative observations: not the winner"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(hyper2)
```

# Partial rank observations: someone didn't win

Here we consider a situation in which we make the observation that a
particular player *does not* win.  We are ignorant of the order
statistic except for the fact that (wlog) player 1 is not first
placed.

As in `loser.Rmd`, we have three hypotheses:

* $H_1\colon p_1=0$
* $H_2\colon\exists i\neq 1 \mbox{ with } p_1\geq p_i>0$
* $H_3\colon p_1=p_2=\ldots=p_n=\frac{1}{n}$

(observe that $H_2$ implies that someone is worse than me).  All three
hypotheses correspond to a restriction on allowable parameter space.
The first step is to define a suitable likelihood function for our
observation.  Given that $P(D|H_1)=1$, $P(D|H_2)\leq\frac{1}{2}$ and
$P(D|H_3) =\frac{n-1}{n}$, this gives us a likelihood function for our
three hypotheses.

What we will do is combine this observation with other, perhaps more
informative, observations which we will take to be (conditionally)
independent of one another.  Firstly we will consider some highly
informative order statistics:

```{r}
(H <- hyper2(pnames=paste("p",1:5,sep="")) +
    order_likelihood(c(5,2,1,4)) + 
    order_likelihood(c(2,4,1)) + 
    order_likelihood(c(4,3,1)) + 
    order_likelihood(c(3,5)))
(mH <- maxp(H))
loglik(indep(mH),H)
```

Note the low estimated strength of `p1`.

## Hypothesis 1

```{r}
small <- 1e-3
(mH1 <- maxp(H, startp=rep(small/2,4),fcm=cbind(-1,0,0,0),fcv = -small))
loglik(indep(mH1),H)
```

and for a consistency we will use slightly different values of `small`:

```{r}
small <- 5e-4
(mH2 <- maxp(H, startp=rep(small/2,4),fcm=cbind(-1,0,0,0),fcv = -small))
loglik(indep(mH2),H)
```

```{r}
small <- 3e-4
(mH2 <- maxp(H, startp=rep(small/2,4),fcm=cbind(-1,0,0,0),fcv = -small))
loglik(indep(mH2),H)
```

Observe that decreasing the value of variable `small` corresponds to a
more restrictive constraint on the likelihood.  And as the value of
`small` decreases the likelihood at the evaluate decreases quite
substantially (the constraint is active as we can see from the first
element of the evaluate).  Usually this does not matter but in this
case we can deduce using elementary arguments that the likelihood of
$H_2$ is zero: if $p_1=0$, then $p_4=0$ from the first line, and then
$p_1=0$ from line 2; then $p_3=0$ from line 3, $p_5=0$ from line 4 and
finally $p+2=0$ from line 1.

Thus we can say for sure that $H_1$ has zero likelihood and we may
discount it.

## Hypothesis 2

```{r}
n <- 5 # number of players
small <- 1e-3  # numerical necessity
(ans_constrained <-
maxp(H,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,-1,rep(0,n-3))),fcv=0,give=TRUE)
)
small <- 5e-4
(ans_constrained <-
maxp(H,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,-1,rep(0,n-3))),fcv=0,give=TRUE)
)
```

In the above, note that the value of `small` is immaterial and the
evaluate has $p+1=p_2\neq 0$.

## Hypothesis 3

```{r}
loglik(indep(equalp(H)),H)
```

The support for $H_2$ against $H_3$ is about $-6.846-(-7.455) = 1.391$.

# Additional observation

Suppose we now make another observation: person 1 did not win.  How
does this change our relative support for $H_2$ and $H_1$?

```{r}
library("partitions")
M <- perms(5)
M <- M[,M[1,] != 1]
SL <- list()
for(i in seq_len(ncol(M))){
	SL[[i]] <- H + order_likelihood(M[,i,drop=TRUE])
}
SL <- as.suplist(SL)
1e4*like_single_list(indep(equalp(SL)),SL)
1e4*like_single_list(indep(maxp(SL)),SL)
```

Something is very wrong here because the support should be very
similar to the support applied to the order statistic without the
additional observation.