### 2020-02-20 inst/eurovision.Rmd
slight rewording and reworking

---
title: "Eurovision Song Contest: semifinal 1, 2009"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
abbreviated <- TRUE  # change to FALSE for full names
```

## The Eurovision Song contest

The dataset `wiki_matrix`, defined below, is copied from "Eurovision
Song Contest 2009," Wikipedia, accessed May 13, 2018.  It refers to
semi-final 1.  More documentation is given in `euro.Rd` [type
`help(euro2009)` at the `R` prompt].

First we specify the matrix as appearing in the Wikipedia page:

```{r, specifymatrix}
wiki_matrix <- matrix(c(
#   ME  CZ  BE  BY  SW  AM  AD  CH  TR  IL  BG  IS  MK  RO  FI  PT  MT  BA  DE  UK
    NA, 00, 00, 03, 00, 05, 01, 02, 05, 01, 00, 00, 08, 00, 00, 01, 06, 10, 02, 00, # ME
    00, NA, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, # CZ
    00, 00, NA, 00, 00, 01, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, # BE
    02, 01, 00, NA, 01, 04, 00, 00, 00, 04, 01, 01, 06, 00, 04, 00, 01, 00, 00, 00, # BY
    00, 06, 04, 07, NA, 08, 07, 04, 04, 07, 00, 10, 03, 04, 10, 08, 08, 04, 04, 07, # SW
    04, 12, 10, 10, 05, NA, 00, 01, 10, 10, 08, 02, 02, 08, 01, 00, 00, 01, 10, 05, # AM
    00, 00, 00, 00, 00, 00, NA, 00, 01, 00, 00, 00, 00, 00, 00, 04, 03, 00, 00, 00, # AD
    00, 00, 00, 02, 02, 00, 02, NA, 00, 00, 00, 00, 00, 00, 05, 02, 00, 02, 00, 00, # CH
    08, 05, 12, 06, 07, 10, 05, 12, NA, 06, 12, 07, 12, 12, 07, 05, 10, 12, 12, 12, # TR
    05, 04, 03, 04, 06, 07, 08, 05, 03, NA, 04, 06, 01, 03, 06, 00, 04, 00, 05, 01, # IL
    00, 00, 00, 00, 00, 00, 00, 00, 02, 00, NA, 00, 05, 00, 00, 00, 00, 00, 00, 00, # BG
    07, 10, 07, 12, 12, 12, 10, 07, 08, 12, 06, NA, 04, 10, 12, 12, 12, 07, 06, 08, # IS
    10, 03, 00, 00, 00, 00, 00, 06, 06, 00, 10, 00, NA, 02, 00, 00, 00, 08, 00, 00, # MK
    06, 00, 02, 01, 00, 02, 04, 00, 07, 08, 05, 04, 07, NA, 00, 10, 02, 06, 01, 02, # RO
    03, 00, 01, 00, 10, 00, 03, 00, 00, 00, 00, 12, 00, 01, NA, 03, 05, 00, 00, 04, # FI
    00, 02, 06, 00, 03, 00, 12, 10, 00, 02, 02, 08, 00, 07, 02, NA, 00, 03, 07, 06, # PT
    01, 07, 08, 08, 04, 03, 06, 03, 00, 05, 03, 05, 00, 06, 03, 06, NA, 05, 03, 10, # MT
    12, 08, 05, 05, 08, 06, 00, 08, 12, 03, 07, 03, 10, 05, 08, 07, 07, NA, 08, 03) # BA
, nrow=18,byrow=TRUE)
```

Each row corresponds to a competitor and each column to a judge.  Note
that the matrix is not square: the last two columns correspond to
Germany and the UK, who were judges but not competitors.  I have
removed the first column in the Wikipedia table (which gave the total
of the points), replaced self-voting entries with `NA`, and replaced
blanks with `00`; every entry is padded to two digits.  Thus the first
row corresponds to votes cast for `ME`, Montenegro.  We see that `BE`
(Belgium) awarded them three points, `am` (Armenia) five points, and
so on.

```{r,pointsforplaces}
points <- c(12,10,8,7,6,5,4,3,2,1)
```

Variable `points` gives the number of points awarded, under Eurovision
rules, to voters' first, second, third, etc choice.  Points for any
competitor are added and the winner is the competitor with the most
points.  However, in the `hyper2` model, the numerical values
themselves do not affect the likelihood function; only the order of
the voters' preferences matters.  The following R idiom translates
`wiki_matrix` into a form suitable for analysis with `hyper2`:

```{r,prefsetter}
preference <- wiki_matrix*0  
for(i in seq_along(points)){ preference[wiki_matrix == points[i]] <- i }

countries <- data.frame(
    fullname =
        c("Montenegro", "Czech rep", "Belgium", "Belarus", "Sweden",
          "Armenia", "Andorra", "Switzerland", "Turkey", "Israel",
          "Bulgaria", "Iceland", "Macedonia", "Romania", "Finland",
          "Portugal", "Malta", "Bosnia Herz", "Germany", "UK"),
    abbreviation = c("ME","CZ","BE", "BY", "SW", "AM", "AD", "CH",
                     "TR", "IL", "BG", "IS", "MK", "RO", "FI", "PT",
                     "MT", "BA", "DE", "UK")
)

if(abbreviated){ 
    jj <- countries$abbreviation
} else { 
    jj <- countries$fullname
}

competitors <- jj[1:18]
colnames(preference) <- jj      # voters; 20 countries (18 + DE + UK)
rownames(preference) <- competitors

rownames(wiki_matrix) <- jj[1:18]
colnames(wiki_matrix) <- jj
```

In the above, matrix `preference` records voters' first, second, third,
etc choice.  A zero entry means no points (nul punkte!) and `NA` means
that voter was forbidden from voting for that player (countries cannot
vote for themselves).

The competitors were the first 18 countries (the last two countries,
Germany and the UK, voted but did not compete).  We need the voters'
choices to be the *rows* for consistency with the rest of the package
(and, for that matter, the aylmer package and indeed the emulator
package), so we take the transpose:

```{r,taketranspose}
preference <- t(preference)
preference
```

Now, take the first row of `preference`.  This represents the votes
cast *by* (sic) Montenegro (`"ME"`).  Their favourite was [last
column] Bosnia & Herzegovina (`"BA"`), who they gave rank 1 to.  Their
second favourite was Macedonia, their third was Turkey, and so on.
They were not allowed to vote for themselves, which is why the first
column is `NA`.  So the order was: `BA` (1), `MK` (2) , `TR` (3), `IS`
(4), `RO` (5), `IL` (6), `AM` (7), `FI` (8), `BY` (9), `MT` (10).  The
other countries (`CZ`, `BE`, `SW`, `CH`, `BG`, `PT`) did not receive
any points from Montenegro.  We need to convert matrix `preference`
into a likelihood function, here `euro2009`:

```{r,morestuff}
euro2009 <- hyper2(d=18)

for(i in seq_len(nrow(preference))){   # cycle through the rows; each row is a voter
    d <- preference[i,,drop=TRUE]
    d[is.na(d)] <- -1  # kludge: make the voting country ineligible to vote.
    while(any(d>0)){
        eligible <- which(d>=0)   # This is why we set NA values to -1
        euro2009[which(d==1)] %<>% "+"(1)
                                        # The first choice among
                                        # eligible players has +1
                                        # power on the numerator

        euro2009[eligible] %<>% dec()   # denominator of all eligible players

        d[d==1] <- -1  # once you've won, you are ineligible to be chosen again
                       # NB, pipe idiom is ugly:  d %<>% `[<-`(. == 1, -1)

        d[d>0] %<>% dec()  # everyone moves down the list, so who
                            # *was* second choice becomes first
                            # choice, who *was* third choice becomes
                            # second, and so on.

    } # while() loop closes
} # i loop closes

## syntatic sugar:
pnames(euro2009) <- competitors
```

Now we can check null of equal strengths:

```{r,checknull}
m <- maxp(euro2009)  # MLE
print(loglik(indep(m),euro2009))
print(loglik(indep(equalp(euro2009)),euro2009))
```

Thus the difference is about 122 units of support, surely significant.

```{r,orderplot}
points_scored <- rowSums(wiki_matrix,na.rm=TRUE)  # points
ox <- rank(-points_scored)
oy <- rank(-m)   # m = MLE
par(pty='s') # square plot
plot(ox,oy,asp=1,pty='s',xlim=c(0,18),ylim=c(0,18),pch=16,
xlab="official order",ylab="my order",main="Eurovision 2009")
par(xpd=TRUE) # allow country names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(m)[ox[i]],pos=4,col='gray') }
par(xpd=FALSE) # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

In the graph above, we see that both systems agree on the first four places.
We can also plot the numerical values:

```{r}
plot(log(rowSums(wiki_matrix,na.rm=TRUE)),log(m),
     pch=16,xlab="points scored",ylab="MLE strength")
```