### 2020-03-10 inst/formula1_extra.Rmd
initial work on pole position

---
title: "formula1: further tests"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
```

## Formula 1

(this document follows on from `formula1.Rmd`).

Here I analyse results from the 2017 series of Formula 1 motor racing
for two interesting hypotheses: that Lewis Hamilton does indeed
perform better in the wet, and also whether pole position makes a
difference.

Consider the following dataset, taken from Wikipedia:

```{r readformula1}
f2017 <- read.table("formula1_2017.txt",header=TRUE)
head(f2017)
```

Thus, reading the top row, Hamilton placed second in Australia, first
in China, second in Bahrain, and so on.  Reading the first column, we
see that in Australia, Hamilton came second, Vettel first, and so on.

```{r,readcond}
(cond <- read.table("formula1_conditions.txt",header=TRUE))
```

Thus from the fourth row, for example, we see that in Russia the
conditions were dry and Vettel drove in pole position.  We might ask whether

```{r dotable}
TT <- table(wins=as.vector(f2017[1,1:20]==1),pole=cond$pole == "Hamilton")
fisher.test(TT)
```

We are going to use a generalization of function
\code{ordertable2supp()} that incorporates pole position and condition
information:

```{r,applydata}
```

See how the rows of `a2017` correspond to the columns of `f2017`.  In
the following, we will define a `hyper2` object `H` which has a
reified player `pole` whose strength represents the advantage
conferred by driving pole position.

```{r definegeneralizedordertable2supp}
f2017 <- read.table("formula1_2017.txt",header=TRUE)
cond <- read.table("formula1_conditions.txt",header=TRUE)

x <- read.table("formula1_2017.txt",header=TRUE)
noscore <- c("Ret", "WD", "DNS", "DSQ", "DNP", "NC")
venues <- as.vector(colnames(x)[-ncol(x)])
jj <- apply(x[,-ncol(x)], 2, function(y){`[<-`(y,y %in% noscore,0)})

fmat <- matrix(as.numeric(jj), byrow = TRUE, ncol = nrow(x))
colnames(fmat) <- rownames(x)
rownames(fmat) <- venues

out <- hyper2(pnames = c(colnames(fmat),"pole"))

for(i in seq_len(nrow(fmat))){
  jj1 <- ordervec2supp(fmat[i,, drop=TRUE])
  pwa(jj1,as.vector(cond$pole[i]),"pole")
  out  <- out + pwa(jj1,as.vector(cond$pole[i]),"pole")
}
(mo <- maxp(out))
pie(mo)
```

Now we can use it to create a profile likelihood curve along the same
lines as table tennis