### 2020-03-12 inst/formula1_extra.Rmd
just the strongest 8

---
title: "formula1: further tests"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
```

## Formula 1

(this document follows on from `formula1.Rmd`).

Here I analyse results from the 2017 series of Formula 1 motor racing
for two interesting hypotheses: that Lewis Hamilton does indeed
perform better in the wet, and also whether pole position makes a
difference.

Consider the following dataset, taken from Wikipedia:

```{r readformula1}
f2017 <- read.table("formula1_2017.txt",header=TRUE)
head(f2017)
```

Thus, reading the top row, Hamilton placed second in Australia, first
in China, second in Bahrain, and so on.  Reading the first column, we
see that in Australia, Hamilton came second, Vettel first, and so on.

Now we can consider only the strongest say 8 competitors.  There are
two ways to proceed.  Firstly, we can retain just the 8 strongest
competitors as estimated from the full likelihood function:

```{r,strongestnineanotherapproach,cache=TRUE}
mo8a <- sort(mo,decreasing=TRUE)[seq_len(8)]
pie(mo8a)
```

Or secondly, we can simply erase everyone but the first 8 from the
likelihood function `out`  and use that:

```{r,strongestnine,cache=TRUE}
out8 <- keep(out,seq_len(8))
head(out8)
mo8 <- maxp(out8)
dotchart(log(mo8))
pie(mo8)
```

Note how the two approaches are not equivalent: the second approach
has redistributed the estimated strengths and, for example, Ricciardo
has an estimated strength of essentially zero (because he consistently
loses to the other 7 players).

# Pole position

Driving pole position is an important thing:

```{r,readcond}
(cond <- read.table("formula1_conditions.txt",header=TRUE))
```

Thus from the fourth row, for example, we see that in Russia the
conditions were dry and Vettel drove in pole position.  We might ask
whether Hamilton's winning is associated with his driving at pole
position:

```{r makehamiltontable}
(TT <- table(wins=as.vector(f2017[1,1:20]==1),pole=cond$pole == "Hamilton"))
```

Thus from the above we see that, of Hamilton's 9 wins, 8 of them were
from pole; conversely, of his 11 pole races, he lost only 3.

```{r,fishertestpoleham}
fisher.test(TT)
```

We are going to use a generalization of function `ordertable2supp()`
that incorporates pole position and condition information:

See how the rows of `a2017` correspond to the columns of `f2017`.  In
the following, we will define a `hyper2` object `H` which has a
reified player `pole` whose strength represents the advantage
conferred by driving pole position.

```{r definegeneralizedordertable2supp}
f2017 <- read.table("formula1_2017.txt",header=TRUE)
cond <- read.table("formula1_conditions.txt",header=TRUE)

x <- read.table("formula1_2017.txt",header=TRUE)
noscore <- c("Ret", "WD", "DNS", "DSQ", "DNP", "NC")
venues <- as.vector(colnames(x)[-ncol(x)])
jj <- apply(x[,-ncol(x)], 2, function(y){`[<-`(y,y %in% noscore,0)})

fmat <- matrix(as.numeric(jj), byrow = TRUE, ncol = nrow(x))
colnames(fmat) <- rownames(x)
rownames(fmat) <- venues

out <- hyper2(pnames = c(colnames(fmat),"pole"))

for(i in seq_len(nrow(fmat))){  # iterate through the races
  jj1 <- ordervec2supp(fmat[i,, drop=TRUE])      # jj1 is a LF for a single race...
  jj2 <- pwa(jj1,as.vector(cond$pole[i]),"pole") # jj2 includes a pole ghost racer
  out  <- out + jj2                              # now increment 'out' with ghost LF 
}
head(out)
```

In the above, we see ghost driver `pole`.

```{r,maximizeoutandpie,cache=TRUE}
(mo <- maxp(out))
pie(mo)
```

# Profile likelihood of pole strength

Now we can use it to create a profile likelihood curve along the same
lines as table tennis, but for simplicity we will keep just the
strongest 9 competitors:

```{r,justnine}
4+5
```