### 2020-03-23 inst/soling.Rmd
start work on issue #46

---
title: "Soling at the 2000 Summer Olympics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Soling is a type of boat raced in the 2000 Summer Olympics.  Here I analyse race data taken from

  \url{https://en.wikipedia.org/w/index.php?title=Sailing_at_the_2000_Summer_Olympics_\%E2\%80\%93_Soling&oldid=945362535}

using a Plackett-Luce extension of the Bradley-Terry model.

```{r loadlib}
library("hyper2")
data(soling)
soling_table_2000
ordertable_to_ranktable(soling_table_2000)
head(soling2000)
```

We can also consider observations from the round robins and quarter finals:

```{r,rrqf}
soling2000_rr1
soling2000_rr2
soling2000_qf
```

Thus, for example, taking `soling2000_rr1[1,2] == TRUE` we see that
Germany beat Sweden in round robin 1.  To incorporate this information
we need a little bespoke function:

```{r, definef}
rr <- function(H,M){
	for(i in seq_len(nrow(M)-1)){
		for(j in (i+1):ncol(M)){
			if(M[i,j]){
				winner <- i
				loser  <- j
			} else {
				winner <- j
				loser <- i
			}
			H[winner] %<>% inc
			H[c(winner,loser)] %<>% dec
		}
	}
	return(H)
}
```

now

```{r}
before <- maxp(soling2000)
soling2000 %<>% rr(soling2000_rr1)
soling2000 %<>% rr(soling2000_rr2)
soling2000 %<>% rr(soling2000_qf)
after <- maxp(soling2000)
par(pty="s")
plot(before,after,asp=1,pch=16,xlim=c(0,0.3),ylim=c(0,0.3))
abline(0,1)
```