### 2020-04-01 inst/formula1.Rmd
Hamilton in the wet

---
title: "formula1"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

## Formula 1

Formula 1 racing is an important and prestigious motor sport.  Here I
analyse a series of racing results using the `hyper` package which
implements the Plackett-Luce likelihood function.

Consider the following dataset, taken from Wikipedia:

```{r cars}
f2017 <- read.table("formula1_2017.txt",header=TRUE)
head(f2017)
```

(in the package this is object `F1_table_2017`).  Each row is a driver
and each column (after the first) a venue.  We see that Hamilton, the
first row, came second in Australia, first in China, second in
Bahrain, fourth in Russia, and so on.  In the first column, we see the
result from Australia (AUS) in which Hamilton came second, Vettel
first, Bottas third, and so on.  Notation used also includes six
different classes of no-score such as "Ret" for retired, "WD" for
withdrawn, and so on.

We can represent this data in a different form using
`wikitable_to_ranktable()`:

```{r}
wikitable_to_ranktable(f2017)
```

Looking at the first column ("c1" is read "came first") we see who
came first at each venue: Vettel in Australia (AUS), Hamilton in China
(CHN), Vettel in Bahrain (BHR), and so on.  I now use bespoke R
function \code{ordertable2supp()}, that translates data of this type
into a log-likelihood.  The function is similar to, but slightly
different from, the analysis in `eurovision.R`.

```{r,applydata}
L2017 <- ordertable2supp(f2017)
head(L2017,3)
```

Above, we see a small part of the Placket-Luce likelihood function for
the 2017 results.  But from now on we will use built-in dataset
`formula1`, which contains extra information including non-finishing
competitors.  We can find the maximum likelihood estimator by using
the `maxp()` function:

```{r,maxp2017,cache=TRUE}
rm(L2017)
mL2017 <- maxp(formula1)
mL2017
dotchart(mL2017,pch=16,main='2017 Formula 1')
dotchart(log(mL2017)[1:23],pch=16,main='2017 Formula 1, log scale')
```

(in the diagram above, we remove the bottom-ranked two players, Resta
and Button, whose estimated strength is zero).

# Likelihood scoring vs points scoring

The championship winning order is determined by a points system: the
winner is awarded 25 points, second place 18 points, and so on.  We
can now check to see how the points total matches up with the
likelihood estimation.

```{r,likelihoodandscore}
points <- f2017$points
names(points) <- f2017$driver

m <- mL2017   # maxp(formula1)
ox <- rank(-points) # negative because number 1 = largest number of points
oy <- rank(-m)      # negative because number 1 = largest strength
par(pty='s')        # square plot
plot(ox,oy,asp=1,pty='s',xlim=c(0,25),ylim=c(0,25),pch=16,
xlab="official order",ylab="my order",main='Formula 1, 2017 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray') }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

In the graph above, we see a certain amount of disagreement.  For
example, using F1 points scored places Hamilton first (363), Vettel
second (317), and Bottas third (305); while `hyper2` places Hamilton
first (0.29), Bottas second (0.17), and Vettel third (0.14).  Note
also the large number of players with zero points (who rank equal 23rd
by the F1 points system) but are nevertheless separated by
Bradley-Terry.  These players appear as a line of points at $x=23$:
Resta to Ericcson.  Note also Palmer and Alonso who managed to score a
F1 point but rank below Ericcson according to Bradley-Terry.

But overall there
seems to be some degree of correlation between the two, and
differences in detail.

# Analysis of nulls:

## Hypothesis: all players have equal strengths

Here we write $H_0\colon p_1=p_2=\cdots =p_n=\frac{1}{n}$.

```{r,equalstrength,cache=TRUE}
equalp.test(formula1)
```

highly signficant!

## Null of Hamilton being average

We can consider Hamilton and ask various questions about his playing
strength.  One natural hypothesis would be that Hamilton's strength is
equal to the mean strength, or $\frac{1}{25}$.

* $H_A\colon\sum p_i=1$
* $H_0\colon p_\mathrm{Hamilton} < \frac{1}{25}$

(we attempt to reject the null $H_0$).  Noting that the evaluate for
Hamilton's strength $\widehat{p_\mathrm{Hamilton}}$ is about 0.29, we
ask how much support is lost, relative to the support at the
unconstrained evaluate.

```{r,specifichamiltonstrength,cache=TRUE}
specificp.test(formula1,"Hamilton")  # default value is 1/25=0.04
```

(note that the constrained maximum is on the boundary of the
admissible region with $p_\mathrm{Hamilton}\approxequal\frac{1}{25}$).

## Null of Hamilton being better than other players

We may now test

* $H_4\colon p_\mathrm{Hamilton} > \max_{i\neq\mathrm{Hamilton}} p_i$

This is quite difficult to evaluate formally but we can evaluate it
with no loss in accuracy by observing that the second strongest
competitor is Bottas and considering instead

* $H_4'\colon p_\mathrm{Hamilton} > p_\mathrm{Bottas}$.

```{r,L4like,cache=TRUE}
small <- 0.01
L4 <- maxp(formula1, startp=c(small, rep((1-small)/24,23)),fcm=rbind(c(-1,1,rep(0,22))),fcv=0,give=TRUE)
L4
```

Calculating the support and asymptotic p-values:
```{r,l4val}
#L2 - L4$value 
#pchisq(2*(L2 - L4$value),df=1,lower.tail=FALSE)
2+3
```

## All years data

We will consider four years data:

```{r,fouryearsdata, cache=TRUE}
F1_2014 <- "formula1_2014.txt" %>% read.table(header=TRUE) %>% ordertable2supp
F1_2015 <- "formula1_2015.txt" %>% read.table(header=TRUE) %>% ordertable2supp
F1_2016 <- "formula1_2016.txt" %>% read.table(header=TRUE) %>% ordertable2supp
F1_2017 <- "formula1_2017.txt" %>% read.table(header=TRUE) %>% ordertable2supp
a <- list(F1_2014, F1_2015, F1_2016, F1_2017)
alldrivers <- all_pnames(a)

F1_total <- hyper2(pnames=alldrivers)

F1_total %<>% add(change_pnames(F1_2014,alldrivers)) 
F1_total %<>% add(change_pnames(F1_2015,alldrivers)) 
F1_total %<>% add(change_pnames(F1_2016,alldrivers)) 
F1_total %<>% add(change_pnames(F1_2017,alldrivers))

head(F1_total)
```

So object `F1_total` is a likelihood function for all four years data,
on the assumption of independent observations and in this case driver
strength being independent of year.  We can find the evaluate in the same way:

```{r,findmallyears,cache=TRUE}
mallyears <- maxp(F1_total)
```

and the plot:

```{r,allyearplot,fig.height=7}
dotchart(mallyears,pch=16,main='Formula 1, 2014-7')
```

We see that Hamilton has the highest estimated strength.

# Conditions

```{r,loadtable}
(details <- read.table("formula1_conditions.txt",header=TRUE))
drivers <- pnames(formula1)
```

We cannot use `ordertable2supp()` here because we somehow need to
incorporate pole position and conditions into our likelihood function.
The first step is to modify `f2017` so that non-finishers have a zero
position:

At this point, we introduce a new player: `Ham_wet`, for
"Hamilton in the wet", and we can test the hypothesis that the
`Hamilton == Ham_wet`.  We will cycle through all the venues and,
if it is wet (according to dataframe `conditions`) we will credit
Hamilton's placing to `Ham_wet`.

```{r,hamiltonwetanddry,cache=TRUE}
f2017 <- read.table("formula1_2017.txt",header=TRUE)
f2017 <- f2017[,-ncol(f2017)]
details <- read.table("formula1_conditions.txt",header=TRUE)

drivers <- rownames(f2017)
alldrivers <- c("Ham_dry","Ham_wet",drivers[-1])

L2017 <- hyper2(pnames=alldrivers) # uninformative
for(i in seq_len(ncol(f2017))){
	venue <- colnames(f2017)[i]
	conditions <- details[which(details$venue == venue),4]
	o <- f2017[,i]
        suppressWarnings(o <- as.numeric(levels(o))[o])
        o[is.na(o)] <- 0
	names(o) <- rownames(f2017)
	if(conditions == "wet"){
	   names(o)[names(o) == "Hamilton"] <- "Ham_wet"
	} else if(conditions == "dry"){
	   names(o)[names(o) == "Hamilton"] <- "Ham_dry"
	} else {
	   stop("neither wet nor dry?")
	}
	L2017 <- order_obs(L2017,o)
}  # for(i) loop closes
head(L2017)
```

We can now test the hypothesis that Hamilton's performance in the wet
is the same as his performance in the dry:

```{r,hamwetdrytest,cache=TRUE}
options(width = 80)
samep.test(L2017,c("Ham_wet","Ham_dry"))
```