### 2020-04-12 inst/jester.Rmd
incorporate the jester dataset

---
title: "The Jester joke recommender system"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goldberg et al present a dataset in which respondents rated a number
of jokes.  Here, I analyse a small portion of this dataset using the
`hyper2` package.  This document is intended to illustrate an
extremely challenging application of the `hyper2` package and (without
cache) takes a long time to process.  Goldberg's dataset has 24938
lines, one per respondent, and 101 columns, one per joke (the first
column shows the number of jokes rated by each respondent); here I use
150 lines and 99 jokes (the 100th joke was not funny).

```{r }
library("hyper2",quietly=TRUE)
a <- read.csv("jester-data-3.csv",head=FALSE) # File is 150 lines only
a <- as.matrix(a[,-c(1,100)])
a[a==99] <- NA
colnames(a) <- paste("joke",sprintf("%02d",seq_len(ncol(a))),sep="")
a[1,]
```

Row 1 of `a` is displayed and shows that the first respondent rated
joke 5 at -1.65, joke 7 at -0.78, and so on; `NA` means that that
respondent did not rank that joke.  We can perform some summary
statistics of `a`:

```{r}
plot(rowSums(!is.na(a)),xlab="respondent")
```

The above plot shows how many jokes each of the 100 respondents rated.

```{r}
plot(colSums(!is.na(a)),xlab="joke index")
```

The above shows how many respondents rated each joke.  It would make
sense to remove the jokes that were not rated:

```{r}
dim(a)
a <- a[,colSums(!is.na(a))>1]
dim(a)
```

showing that 91 jokes were rated by at least one respondent.  We need
to transform the dataset:

```{r}
f <- function(x){
    x <- x[!is.na(x)]
    x[order(x,decreasing=TRUE)] <- seq_along(x)
    return(x)
}
f(a[1,])
```

Thus we see that this respondent rated `joke08` to be the funniest, having rank 1.

```{r,calculate_jester,cache=TRUE}
jester <- hyper2(pnames=colnames(a))
system.time(for(i in seq_len(nrow(a))){jester %<>% order_obs(f(a[i,]))})
```

now

```{r,cache=TRUE}
system.time(mjest <- maxp(jester,n=1))
```

and

```{r}
mjest
plot(mjest)
```

```{r,cache=TRUE}
equalp.test(H,startp=indep(mjest))
```

## Reference

Eigentaste: A Constant Time Collaborative Filtering Algorithm. Ken
Goldberg, Theresa Roeder, Dhruv Gupta, and Chris Perkins. _Information
Retrieval_, 4(2), 133-151. July 2001.