### 2020-04-17 inst/formula1_points.Rmd
initial stab at points system

---
title: "formula1: the points system"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
options(width = 140)
```

## Formula 1: top 9

Formula 1 racing is an important and prestigious motor sport.  Here I
analyse a series of racing results using the `hyper` package which
implements the Plackett-Luce likelihood function.

Consider the following dataset, taken from Wikipedia:

```{r cars}
f2017 <- read.table("formula1_2017.txt",header=TRUE)
no_of_drivers <- 9  # top 9
f2017 <- f2017[seq_len(no_of_drivers),]
f2017
```

Now process it to give the ranks:

```{r}
f <- function(x){
	noscore <- c("Ret", "WD", "DNS", "DSQ", "DNP", "NC")
	x <- as.character(x)
	x[x %in% noscore] <-  "0"
	x <- as.numeric(x)
	wanted <- x != 0
	xw <- x[wanted]
	xw[order(xw,decreasing=FALSE)] <- seq_along(xw)
	x[wanted] <- xw
	x
}
o  <- apply(f2017[,seq_len(20)],2,f)
rownames(o) <- rownames(f2017)
(f2017 <- o)
```

Each row is a driver and each column (after the first) a venue.  We
see that Hamilton, the first row, came second in Australia, first in
China, second in Bahrain, fourth in Russia, and so on.  In the first
column, we see the result from Australia (AUS) in which Hamilton came
second, Vettel first, Bottas third, and so on.  Notation used also
includes six different classes of no-score such as "Ret" for retired,
"WD" for withdrawn, and so on.

Now calculate points:

```{r}
points <- c(25,18,15,12,10,8,6,4,2,1,0,0,0,0,0,0,0,0,0,0)
stopifnot(all(diff(points)<=0))
points <- points/points[1]
jj <- f2017
jj[jj==0] <- length(points)
jj[] <- points[jj]
jj
real_points <- rowSums(jj)
real_points
sort(real_points,decreasing=TRUE)
names(sort(real_points,decreasing=TRUE))
```

Now do some random sampling.  First create a likelihood function:

```{r,calcH,cache=TRUE}
H <- hyper2(pnames=rownames(o))
 for(i in seq_len(ncol(o))){
	H <- H + ordervec2supp(o[,i])
}
head(H)
mL2017 <- maxp(H)
mL2017
dotchart(mL2017,pch=16,main='2017 Formula 1')
```

Now use random sampling:

```{r}
random_table <- rrank(n=ncol(f2017), p=mL2017)
rownames(random_table) <- colnames(f2017)
random_table
```

then see how close it is to reality:

```{r}
maxp(rank_likelihood(random_table)) # should be close to mL2017
```

```{r}
a <- random_table
class(a) <- "matrix"
a[] <- points[a]
a
p <- colSums(a)
p
rando <- order(p,decreasing=TRUE)
rando
```

Thus we see that

```{r}
jj <- sort(p,decreasing=TRUE)
jj
names(jj)
```

```{r,dothework,cache=TRUE}
n <- 200
goodness <- rep(-Inf,n)
synthetic_winner_points     <- rep("",n)
synthetic_winner_likelihood <- rep("",n)
f <- function(x,y){sum(cumprod(x==y))}
for(i in seq_len(n)){
	a <- rrank(n=ncol(f2017), p=mL2017)
	jj <- a
	class(jj) <- "matrix"
	jj[] <- points[jj]
	p <- colSums(jj)  # synthetic points
	names(p) <- names(mL2017)
	synthetic_rank_points <- names(sort(p,decreasing=TRUE))
	synthetic_winner_points[i] <- synthetic_rank_points[1]

	mH <- maxp(rank_likelihood(a))
	synthetic_rank_likelihood <- names(sort(mH,decreasing=TRUE))
	synthetic_winner_likelihood[i] <- synthetic_rank_likelihood[1]

	goodness[i] <- f(synthetic_rank_likelihood,synthetic_rank_points)
}
plot(table(goodness))
table(synthetic_winner_likelihood,synthetic_winner_points)
```

So from the table we see about half the times the winner was correctly
predicted by the points system.