### 2020-05-04 inst/gradient.Rmd
gradientn vignette

---
title: "Numerical verification of function `gradientn()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2",quietly=TRUE)
```

We will numerically verify the `gradientn()` function for two `hyper2`
objects.

## Chess dataset

The chess dataset is as follows:

```{r}
chess
```

We will define two gradient functions: `chess_gradient_theoretical()`,
which calculates the partial derivatives using straightforward algebra
(specifically elementary calculus, $\partial(x+y)^n/\partial
x=n(x+y)^{n-1}$), and `chess_gradient_theoretical()`, which uses `C` code:

```{r}
chess_gradient_theoretical <- function(x){
    Topalov <- x[1]
    Anand   <- x[2]
    Karpov  <- x[3]

    out <- c(
        Topalov =  30/Topalov -35/(Topalov+Anand) -18/(Topalov+Karpov),
        Anand   = -35/(Topalov+Anand) +36/Anand -35/(Anand+Karpov),
        Karpov  = -18/(Topalov+Karpov) -35/(Anand+Karpov) +22/Karpov
    )

    names(out) <- pnames(chess)
    out

}
chess_gradient_numerical  <- function(x){gradientn(chess,x)}
```

So in the above we have two ways of calculating the gardient: one
theoretical and one numerical.  Do they agree?

```{r}
p_chess <- 
	c(
		Topalov = 0.5,
		Anand   = 0.3,
		Karpov  = 0.2
	)
chess_gradient_theoretical(p_chess) - chess_gradient_numerical(p_chess)
```

Yes, they agree

# Icons

The `icons` dataset is:

```{r}
icons
```

We can do the same thing but the algebra is more involved:

```{r}
icons_gradient_theoretical <- function(x){
    NB <- x[1]
    L <- x[2]
    PB <- x[3]
    THC <- x[4]
    OA <- x[5]
    WAIS <- x[6]

	out <- c(
		NB = (
			+32/NB
            -20/(NB + L + THC + OA)
            -15/(NB + L + THC + WAIS)
            -09/(NB + L + OA + WAIS)
            -18/(NB + PB + THC + OA)
            -18/(NB + PB + THC + WAIS)
            -08/(NB + PB + OA + WAIS)
        ),
        L = (
            -20/(NB + L + THC + OA)
            -15/(NB + L + THC + WAIS)
            -09/(NB + L + OA + WAIS)
            +24/L
            -11/(L + PB + THC + OA)
            -16/(L + PB + THC + WAIS)
            -18/(L + PB + OA + WAIS)
        ),
        PB = (
            -18/(NB + PB + THC + OA)
            -18/(NB + PB + THC + WAIS)
            -08/(NB + PB + OA + WAIS)
            -11/(L + PB + THC + OA)
            -16/(L + PB + THC + WAIS)
            -18/(L + PB + OA + WAIS)
            +30/PB
        ),
        THC = (
            -20/(NB + L + THC + OA)
            -15/(NB + L + THC + WAIS)
            -18/(NB + PB + THC + OA)
            -18/(NB + PB + THC + WAIS)
            -11/(L + PB + THC + OA)
            -16/(L + PB + THC + WAIS)
            +24/THC
        ),
        OA = (
            -20/(NB + L + THC + OA)
            -09/(NB + L + OA + WAIS)
            -18/(NB + PB + THC + OA)
            -08/(NB + PB + OA + WAIS)
            -11/(L + PB + THC + OA)
            -18/(L + PB + OA + WAIS)
            +14/OA
        ),
        WAIS = (
            -15/(NB + L + THC + WAIS)
            -09/(NB + L + OA + WAIS)
            -18/(NB + PB + THC + WAIS)
            -08/(NB + PB + OA + WAIS)
            -16/(L + PB + THC + WAIS)
            -18/(L + PB + OA + WAIS)
            +9/WAIS
        )
    )
	names(out) <- pnames(icons)
	return(out)
}

icons_gradient_numerical  <- function(x){gradientn(icons,x)}
```

Then

```{r}
p_icons <- c(NB=0.3, L=0.1, PB=0.2, THC=0.15, OA=0.15, WAIS=0.1)
```

```{r}
icons_gradient_theoretical(p_icons)
icons_gradient_theoretical(p_icons) - icons_gradient_numerical(p_icons)
```

also agreeing.

## Difference between `gradientn()` and `gradient()`

Taking the chess dataset we have

```{r}
rbind(indep(p_chess), gradient(chess,indep(p_chess)))
rbind(p_chess,gradientn(chess,p_chess))
```

Indeed, because the powers of `chess` sum to zero, we can test the
fact that the derivative in the direction parallel to `p_chess` is
zero:

```{r}
sum(powers(chess))
sum(p_chess*gradientn(chess,p_chess))
```

which is correct to numerical precision.  Further,

```{r}
sum(powers(icons))
sum(p_icons*gradientn(icons,p_icons))
```