### 2020-05-27 inst/ordertable_to_ranktable.Rmd
more documentation

---
title: "The print method for class ranktable"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

## Rank tables

Rank tables can be confusing, here is a tiny example that should
clarify things.

```{r}
OT <- cbind(1:5,c(1,5,2:4))
rownames(OT) <- letters[seq_len(nrow(OT))]
colnames(OT) <- c("J1","J2")
OT
```

Object `a` is a straightforward order table, just like
`skating_table`; we see, for example, that both judges put competitor
`a` first, but put `b` second and fifth respectively.  The judges
second choice was `b` and `c` respectively.  We can coerce order table
`OT` to a rank table `RT`:

```{r}
RT <- ordertable_to_ranktable(OT)
RT
```

Object `RT` gives the same information but in a different form.  Note
that above we see the print method in action, which uses
`ranktable_to_printable_object()`.  However, it is possible to view
`RT` directly by removing class `ranktable`:

```{r}
class(RT)
class(RT) <- "matrix"  # removes "ranktable" from class list
RT
```

Note carefully that the second row of `RT` (which corresponds to the
second judge) does *not* match the second column of `OT`.  The way to
read the second row of `RT` is as follows.  We see the entries `1 3 4
5 2` indicating that the first choice was competitor 1 [viz `a`, the
second choice competitor 3 [viz `c`], the third choice competitor 4
[viz `d`], fourth choice competitor 5 [`e`] and fifth choice
competitor 2 [`b`].  So the rows are *ranks* (see the discussion in
`rrank.Rd`).  But it is very tempting to read the second row of `RT`
as "judge 2 placed competitor `b` in third place", which is wrong
because this is interpreting the row as an *order*.  This is why the
print method is there.