### 2020-08-24 inst/cook_repeated.Rmd
consistentify

---
title: "Judging domestic cooking ability: repeated incomplete rank observations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Suppose there are five people in a family.  Each day for five days,
one person cooks a meal.  After the fifth day, each person considers
the four meals cooked by other family members, and puts them in order
of preference.  This process is repeated and here I consider how to
interpret such data.

```{r loadlib,echo=FALSE}
library(hyper2,quietly=TRUE)
```

```{r data}
cooking_table <- as.matrix(read.table("cooking_weekly.txt", header=TRUE))
head(cooking_table)
tail(cooking_table)
```
Thus from the first row we find that one judge considered Lesley to be
fourth best, Alice third, Zac second and Annabel best (this must have
been Robin: no-one is allowed to vote for themselves and a zero means
no vote).  This is the transpose of an order table such as
`skating.txt`.  In this context a zero contains no information (unlike
the formula 1 tables in which zeros signify did not finish).  Document
`curling.Rmd` contains a discussion and here we will use the same
method.  First we will consider the support function for the first row only:

### First row of the dataset

```{r}
(day1 <- cooking_table[1,])
H_row1 <- rank_likelihood(day1[day1>0])
```

In the above, we calculate the Plackett-Luce likelihood function;
observe that it is uninformative about competitor number 1, which I
show in two ways.  Firstly:

```{r}
p <- c(0.3,0.2,0.1,0.3)
# Now create p2 which is the same but competitor 1 is stronger:
p2 <- p
alpha <- 1.3
beta <- (1-alpha*p[1])/(1-p[1])
p2[1] <- p2[1]*alpha   # person 1 becomes stronger...
p2[-1] <- p2[-1]*beta  # ...and everyone else becomes weaker

loglik(indep(p),H_row1)
loglik(indep(p2),H_row1)
```

## Complete dataset

It is easy to iterate over the rows of the dataset:

```{r}
H <- hyper2()

for(i in seq_len(nrow(cooking_table))){  # iterate over rows
jj <- cooking_table[i,] 
    H <- H + rank_likelihood(jj[jj>0])
}
H
```

(compare file `curling.Rmd` in which we iterate over columns not rows).
Standard techniques may be used:

```{r}
equalp.test(H)
mH <- maxp(H)
mH
pie(mH)
```

# Preferential voting

It seems that the boys tend to prefer the boys' cooking and the girls
prefer the girls' cooking.  But simple things first:

```{r}
a <- cooking_table # saves typing
a <- a[a[,4]>0,]  # just lines where zac does not vote
jj <- table(robin_votes=a[,1]==0,zac_last=a[,4]==4)
jj
```

The table above shows that Robin cast a total of `r sum(jj[2,])` votes
for Zac, and everyone else combined cast a total of `r sum(jj[1,])`
votes for Zac.  Everyone else voted Zac last a total of `r jj[1,2]`
times and higher than last `r jj[1,1]` times:

```{r}
fisher.test(jj,alternative="less")
```

So it looks like Robin is less likely to vote Zac last than everyone
else.  Does he do likewise for me?

```{r}
a <- cooking_table # saves typing
a <- a[a[,1]>0,]  # just lines where robin does not vote
jj <- table(zac_votes=a[,4]==0,robin_last=a[,1]==4)
jj
fisher.test(jj,alternative="less")
```

No evidence for that.

We can do slightly better, and test a more general hypothesis,
specifically that the girls systematically vote other girls higher,
and boys systematically vote boys higher, than warranted.  To test
this hypothesis, we posit two reified entities, `girl` and `boy`.
Entity `girl` helps a girl when another girl is voting for her, and
entity `boy` helps a boy when another boy is voting for him.

```{r,cache=TRUE}
cooking_table <-
	as.matrix(read.table("cooking_weekly.txt", header=TRUE))
cooking_table <- cooking_table[rowSums(cooking_table==0)==1,]
boys <-c("robin","zac")
girls <- c("lesley","alice","annabel")
cooks <- colnames(cooking_table)
P <- hyper2()
P0 <- hyper2()
for(i in seq_len(nrow(cooking_table))){  # iterate over rows
    x <- cooking_table[i,] 
	if(sum(x==0)>1){break}
    names(x) <- cooks
	if(names(x[x==0]) %in% boys){  # if the voter is a boy
		x <- x[x>0]  # remove the voter
		Pjj <- pwa(rank_likelihood(x),names(x)[names(x) %in% boys], chameleon="boy")
	} else if (names(x[x==0]) %in% girls){ # voter a girl
		x <- x[x>0]
		Pjj <- pwa(rank_likelihood(x),names(x)[names(x) %in% girls], chameleon="girl")
	} else {
		stop()
	}
    P <- P + Pjj
}
mP <- maxp(P)
```

```{r}
mP
pie(mP)
```

Now test whether the girls preferentially vote:

```{r,cache=TRUE}
specificp.gt.test(P,"girl",0)
```