### 2020-08-26 inst/formula1_allyears.Rmd
new years and new analysis

---
title: "formula1_allyears"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("hyper2")
```

## Introduction

Here I analyse Formula 1 results 2010-2019 using Plackett-Luce
likelihoods as implemented in the `hyper2` package.  Files like
`formula1_2017.txt` are directly copied from Wikipedia (with slight
whitespace changes).

```{r,label=f1yearsseparate,cache=TRUE}
f <- function(M){M[,-ncol(M)]} # removes 'points' column
o <- function(k){ordertable2supp(f(read.table(k)))}
F1_2010 <- o("formula1_2010.txt")
F1_2011 <- o("formula1_2011.txt")
F1_2012 <- o("formula1_2012.txt")
F1_2013 <- o("formula1_2013.txt")
F1_2014 <- o("formula1_2014.txt")
F1_2015 <- o("formula1_2015.txt")
F1_2016 <- o("formula1_2016.txt")
F1_2017 <- o("formula1_2017.txt")
F1_2018 <- o("formula1_2018.txt")
F1_2019 <- o("formula1_2019.txt")
```

```{r,label=combineallyears,cache=TRUE}
F1_list<- list(F1_2010, F1_2011, F1_2012, F1_2013, F1_2014, F1_2015, F1_2016, F1_2017, F1_2018, F1_2019)
F1_total <- F1_2010 + F1_2011 + F1_2012 + F1_2013 + F1_2014 + F1_2015 + F1_2016 + F1_2017 + F1_2018 + F1_2019
```

```{r,calculatemaxlikelhood,cache=TRUE}
m <- maxp(F1_total)
pie(m)
dotchart(m,pch=16,main='Formula 1, 2012-9')
```

```{r,workwithM}
M <- matrix(NA,length(pnames(F1_total)),2019-2010+1)
rownames(M) <- pnames(F1_total)
colnames(M) <- paste("y",10:19,sep="")

for(i in seq_len(nrow(M))){
    for(j in seq_len(ncol(M))){
        if(rownames(M)[i] %in% pnames(F1_list[[j]])){M[i,j] <- 1}
    }
}
jj <- order(rowSums(M,na.rm=T),decreasing=T)
M <- M[jj,]

image(x=2010:2019,y=seq_along(pnames(F1_total)),t(M))
```