### 2020-09-14 inst/jebo_chess_rebuttal.txt
revision of JEBO paper, ancillary files at ~/Dropbox/draft_papers/

“A generalization of the Bradley-Terry model for draws in chess”
JEBO-D-20-00916
Referee Report
This paper develops extensions to the Bradley-Terry model appropriate to Bernoulli trials
so that the model can accommodate the chess-specific needs of draws and the first-mover
advantage possessed by White. The author takes this extended model to various sets of
exclusively game-level outcomes and estimates time-invariant player strengths as well as the
importance of draws and first-mover advantage. After estimating common parameters for draws
and first-mover advantage, the author allows for (time-invariant) player-specific parameters on
these margins. The paper then concludes by considering distinct draw parameters for Soviet and
non-Soviet players. Using data from the top chess events of 1962 and 1963, the author
determines that Soviet players draw more frequently against each other than in other gamescenarios.
The paper does a nice job of clearly laying out the various reified “monsters” that extend
the original Bradley-Terry model to accommodate chess. This framework is plausible and
intuitive. I also appreciated the paper’s brevity in making its point. The allowance for playerspecific tendencies to draw or to gain excess-advantage from moving first also strikes me as an
innovation to the literature. Lastly, the paper is well-written and a fun read.
My comments on the rest of the paper are detailed below by decreasing importance.
1) The author never properly explores the paper’s identification strategy. As I read it, the
identifying assumption, besides the functional form assumption, is that player strengths
are constant over the duration of the considered data. This seems highly plausible within
a specific event (e.g., the Candidates tournament of 1963, the Interzonals of 1962), but
much less so with the Anand-Karpov-Kasparov example. The author does not specify the
timespan over which these 365 games were played, but I imagine that the timespan must
exceed 10 years. To the extent that all players have matured to full strength and none
have yet begun to decline, this assumption would be benign, but it should be supported.
Precise estimation will hinge on a large number of games, but this will be in tension with
the assumption of time-invariant player-strengths. The author should explore this tension.
2) The author neglects to motivate the paper beyond its application within the chess world.
The tournament literature’s applications lie primarily, though definitely not exclusively,
within compensation and labor. How could this extension be used within that context?
Where might economists see the first-mover advantage in action? Where might the
extension to allow draws be especially lucrative? Etc.
3) As the author notes, the work of Moul and Nye (2009) is directly on point. The author’s
discussion of this paper, however, is overly sparse and at times misleading. Moul and

Nye use existing observed player strengths (variations of commonly used scores based on
the Elo rating system that goes back to 1960). Moul and Nye then use these strengths
within an ordered probit to estimate parameters that relate back to draw likelihood and
White’s first-mover advantage. The sharp difference between that paper and this is then
that the author is estimating player relative strengths (conditional upon a pool of players)
while Moul and Nye use “observed” strengths.1 The latter approach is far less
parametrically demanding, but leans much more on the assumption of normally
distributed disturbances (basically the differing assumption of functional form). The latter
approach also allows individuals’ player strength to vary across time and is far more
flexible to changes in the composition of the pool of players.
Furthermore, Moul and Nye estimate several versions of their ordered probit without data
on the number of moves (ply-count). This is contrary to the implication of the paper (p. 6,
immediately below Table 2), which suggests that the ply-count data are essential to that
paper’s findings. All of that paper’s core results (significance of draws, first-mover
advantage, a distinct draw outcome for round-robin international tournament games that
involve two Soviets) are originally derived without ply-count data. The ply-count data are
incorporated to explore a potential mechanism by which players in a draw-cartel might
benefit.
4) The paper concludes that the estimates derived from the 1962 Interzonals and the 1963
Candidates strongly suggest that the “draw monster” won games involving two Soviet
players more frequently than games that involved at least one non-Soviet. This is
consistent with a Soviet cartel, but it could also be the result of Soviet-specific styles of
play. Moul and Nye (2009) test the Soviet style-of-play hypothesis by considering
international tournaments when the format changed from all-play-all to knockout and allSoviet tournaments. Both comparisons yield results consistent with a Soviet draw-cartel
in international all-play-all tournaments and rejecting the Soviet style-of-play hypothesis.
By only considering tournaments that were international and all-play-all, the author
forgoes this critical source of identification.
In the cartel example, the author also neglects the advantage that presumably accrued to
Soviet players when they played non-Soviet players in all-play-all international
tournaments. This would suggest another “monster” to be estimated, one that benefits the
Soviet player when playing a non-Soviet opponent. Without such an addition, Soviet
strengths will be overstated.

1

Those relative strengths are highly predictive, with t-statistics of around 20.