### 2020-09-14 inst/jebo_chess_rebuttal.txt
line breaks, also now seven-bit clean ascii

"A generalization of the Bradley-Terry model for draws in chess"

JEBO-D-20-00916

Referee Report

This paper develops extensions to the Bradley-Terry model appropriate
to Bernoulli trials so that the model can accommodate the
chess-specific needs of draws and the first-mover advantage possessed
by White. The author takes this extended model to various sets of
exclusively game-level outcomes and estimates time-invariant player
strengths as well as the importance of draws and first-mover
advantage. After estimating common parameters for draws and
first-mover advantage, the author allows for (time-invariant)
player-specific parameters on these margins. The paper then concludes
by considering distinct draw parameters for Soviet and non-Soviet
players. Using data from the top chess events of 1962 and 1963, the
author determines that Soviet players draw more frequently against
each other than in other game scenarios.

The paper does a nice job of clearly laying out the various reified
"monsters" that extend the original Bradley-Terry model to accommodate
chess. This framework is plausible and intuitive. I also appreciated
the paper's brevity in making its point. The allowance for
playerspecific tendencies to draw or to gain excess-advantage from
moving first also strikes me as an innovation to the
literature. Lastly, the paper is well-written and a fun read.

My comments on the rest of the paper are detailed below by decreasing
importance.

1) The author never properly explores the paper's identification
strategy. As I read it, the identifying assumption, besides the
functional form assumption, is that player strengths are constant over
the duration of the considered data. This seems highly plausible
within a specific event (e.g., the Candidates tournament of 1963, the
Interzonals of 1962), but much less so with the Anand-Karpov-Kasparov
example. The author does not specify the timespan over which these 365
games were played, but I imagine that the timespan must exceed 10
years. To the extent that all players have matured to full strength
and none have yet begun to decline, this assumption would be benign,
but it should be supported.

Precise estimation will hinge on a large number of games, but this
will be in tension with the assumption of time-invariant
player-strengths. The author should explore this tension.

2) The author neglects to motivate the paper beyond its application
within the chess world.  The tournament literature's applications lie
primarily, though definitely not exclusively, within compensation and
labor. How could this extension be used within that context?  Where
might economists see the first-mover advantage in action? Where might
the extension to allow draws be especially lucrative? Etc.

3) As the author notes, the work of Moul and Nye (2009) is directly on
point. The author's discussion of this paper, however, is overly
sparse and at times misleading. Moul and Nye use existing observed
player strengths (variations of commonly used scores based on the Elo
rating system that goes back to 1960). Moul and Nye then use these
strengths within an ordered probit to estimate parameters that relate
back to draw likelihood and White's first-mover advantage. The sharp
difference between that paper and this is then that the author is
estimating player relative strengths (conditional upon a pool of
players) while Moul and Nye use "observed" strengths [footnote: Those
relative strengths are highly predictive, with t-statistics of around
20]. The latter approach is far less parametrically demanding, but
leans much more on the assumption of normally distributed disturbances
(basically the differing assumption of functional form). The latter
approach also allows individuals' player strength to vary across time
and is far more flexible to changes in the composition of the pool of
players.

Furthermore, Moul and Nye estimate several versions of their ordered
probit without data on the number of moves (ply-count). This is
contrary to the implication of the paper (p. 6, immediately below
Table 2), which suggests that the ply-count data are essential to that
paper's findings. All of that paper's core results (significance of
draws, first-mover advantage, a distinct draw outcome for round-robin
international tournament games that involve two Soviets) are
originally derived without ply-count data. The ply-count data are
incorporated to explore a potential mechanism by which players in a
draw-cartel might benefit.

4) The paper concludes that the estimates derived from the 1962
Interzonals and the 1963 Candidates strongly suggest that the "draw
monster" won games involving two Soviet players more frequently than
games that involved at least one non-Soviet. This is consistent with a
Soviet cartel, but it could also be the result of Soviet-specific
styles of play. Moul and Nye (2009) test the Soviet style-of-play
hypothesis by considering international tournaments when the format
changed from all-play-all to knockout and allSoviet tournaments. Both
comparisons yield results consistent with a Soviet draw-cartel in
international all-play-all tournaments and rejecting the Soviet
style-of-play hypothesis.  By only considering tournaments that were
international and all-play-all, the author forgoes this critical
source of identification.

In the cartel example, the author also neglects the advantage that
presumably accrued to Soviet players when they played non-Soviet
players in all-play-all international tournaments. This would suggest
another "monster" to be estimated, one that benefits the Soviet player
when playing a non-Soviet opponent. Without such an addition, Soviet
strengths will be overstated.