### 2020-09-14 inst/jebo_chess_rebuttal.txt
insert chevrons to mark referee's comments

"A generalization of the Bradley-Terry model for draws in chess"

JEBO-D-20-00916

Referee Report

> This paper develops extensions to the Bradley-Terry model appropriate
> to Bernoulli trials so that the model can accommodate the
> chess-specific needs of draws and the first-mover advantage possessed
> by White.  The author takes this extended model to various sets of
> exclusively game-level outcomes and estimates time-invariant player
> strengths as well as the importance of draws and first-mover
> advantage.  After estimating common parameters for draws and
> first-mover advantage, the author allows for (time-invariant)
> player-specific parameters on these margins.  The paper then concludes
> by considering distinct draw parameters for Soviet and non-Soviet
> players.  Using data from the top chess events of 1962 and 1963, the
> author determines that Soviet players draw more frequently against
> each other than in other game scenarios.
> 
> The paper does a nice job of clearly laying out the various reified
> "monsters" that extend the original Bradley-Terry model to accommodate
> chess.  This framework is plausible and intuitive.  I also appreciated
> the paper's brevity in making its point. The allowance for
> playerspecific tendencies to draw or to gain excess-advantage from
> moving first also strikes me as an innovation to the literature.
> Lastly, the paper is well-written and a fun read.
> 
> 
> My comments on the rest of the paper are detailed below by decreasing
> importance.
> 
> 1) The author never properly explores the paper's identification
> strategy.  As I read it, the identifying assumption, besides the
> functional form assumption, is that player strengths are constant over
> the duration of the considered data. This seems highly plausible
> within a specific event (e.g., the Candidates tournament of 1963, the
> Interzonals of 1962), but much less so with the Anand-Karpov-Kasparov
> example.  The author does not specify the timespan over which these
> 365 games were played, but I imagine that the timespan must exceed 10
> years.  To the extent that all players have matured to full strength
> and none have yet begun to decline, this assumption would be benign,
> but it should be supported.
> 
> Precise estimation will hinge on a large number of games, but this
> will be in tension with the assumption of time-invariant
> player-strengths.  The author should explore this tension.
> 
> 2) The author neglects to motivate the paper beyond its application
> within the chess world.  The tournament literature's applications lie
> primarily, though definitely not exclusively, within compensation and
> labor.  How could this extension be used within that context?  Where
> might economists see the first-mover advantage in action? Where might
> the extension to allow draws be especially lucrative? Etc.
> 
> 3) As the author notes, the work of Moul and Nye (2009) is directly on
> point.  The author's discussion of this paper, however, is overly
> sparse and at times misleading.  Moul and Nye use existing observed
> player strengths (variations of commonly used scores based on the Elo
> rating system that goes back to 1960).  Moul and Nye then use these
> strengths within an ordered probit to estimate parameters that relate
> back to draw likelihood and White's first-mover advantage.  The sharp
> difference between that paper and this is then that the author is
> estimating player relative strengths (conditional upon a pool of
> players) while Moul and Nye use "observed" strengths [footnote: Those
> relative strengths are highly predictive, with t-statistics of around
> 20].  The latter approach is far less parametrically demanding, but
> leans much more on the assumption of normally distributed disturbances
> (basically the differing assumption of functional form).  The latter
> approach also allows individuals' player strength to vary across time
> and is far more flexible to changes in the composition of the pool of
> players.
> 
> Furthermore, Moul and Nye estimate several versions of their ordered
> probit without data on the number of moves (ply-count).  This is
> contrary to the implication of the paper (p. 6, immediately below
> Table 2), which suggests that the ply-count data are essential to that
> paper's findings.  All of that paper's core results (significance of
> draws, first-mover advantage, a distinct draw outcome for round-robin
> international tournament games that involve two Soviets) are
> originally derived without ply-count data.  The ply-count data are
> incorporated to explore a potential mechanism by which players in a
> draw-cartel might benefit.
> 
> 4) The paper concludes that the estimates derived from the 1962
> Interzonals and the 1963 Candidates strongly suggest that the "draw
> monster" won games involving two Soviet players more frequently than
> games that involved at least one non-Soviet.  This is consistent with
> a Soviet cartel, but it could also be the result of Soviet-specific
> styles of play.  Moul and Nye (2009) test the Soviet style-of-play
> hypothesis by considering international tournaments when the format
> changed from all-play-all to knockout and all Soviet tournaments.
> Both comparisons yield results consistent with a Soviet draw-cartel in
> international all-play-all tournaments and rejecting the Soviet
> style-of-play hypothesis.  By only considering tournaments that were
> international and all-play-all, the author forgoes this critical
> source of identification.
> 
> In the cartel example, the author also neglects the advantage that
> presumably accrued to Soviet players when they played non-Soviet
> players in all-play-all international tournaments.  This would suggest
> another "monster" to be estimated, one that benefits the Soviet player
> when playing a non-Soviet opponent.  Without such an addition, Soviet
> strengths will be overstated.
>