### 2020-09-29 inst/curacao1962_threeplayers_rest_monster.Rmd
fixes #65

---
title: "curacao_threeplayers_rest_monster"
author: "Robin Hankin"
date: "28/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This work extends that of `curacao1962_threeplayers.R` by including
a rest monster.

```{r}
library("hyper2")

a <- read.table("curacao1962_candidates.txt",header=FALSE)
colnames(a) <- c("W", "B", "result", "round")
head(a)
```

specify the players:

```{r}
  players <- 
    c("Petrosian", "Keres", "Geller", "Fischer", "Korchnoi",
      "Benko", "Tal", "Filip")
  names(players) <-   # accused of collusion or not
    c("coll", "coll", "coll", "USA", "USSR", "USA", "USSR", "TCH")
collusive_players <-     c("Petrosian", "Keres", "Geller")
 ```

```{r}
d <- as.matrix(cbind(as.character(a$V1),as.character(a$V2)))
colnames(d) <- c("white","black")

stopifnot(all(c(d) %in% players))

jj <- matrix(0,28,8)
colnames(jj) <- players

cumulative_matches <- jj
cumulative_collusive_matches <- jj
cumulative_noncollusive_matches <- jj

for(i in seq_len(28)){
  for(j in seq_along(players)){
    p <- players[j]
    jj <- subset(a,a$round < i)
    cm <- 0
    ncm <- 0
    for(k in seq_len(nrow(jj))){

      w <- jj$W[k]
      b <- jj$B[k]

      this_match <- c(w,b)
      if((players[j] %in% c(w,b))){
        if(players[j] %in% collusive_players){  # this player is collusive
          if((w %in% collusive_players) & (b %in% collusive_players)){  # both players collusive
            cm <- cm+1
          } else { # non-collusive match
            ncm <- ncm+1
          }
        } else { # noncollusive player
          ncm <- ncm+1
        }
      } 
    } # k loop closes
    cumulative_collusive_matches[i,j] <- cm
    cumulative_noncollusive_matches[i,j] <- ncm
    cumulative_matches[i,j] <- cm + ncm
  } # j loop closes
} # i loop closes
```

So we have three matrices `cumulative_matches`,
`cumulative_collusive_matches` and `cumulative_noncollusive_matches`:

```{r}
cumulative_matches
cumulative_collusive_matches
cumulative_noncollusive_matches
```

In the above three matrices, each row is a round, and the entries show
the number of matches, collusive matches, and noncollusive matches
played up to and including that round.

```{r}
a <- cbind(a,cncmp_white=0,cncmp_black=0)
for(i in seq_len(nrow(a))){
  r <- a$round[i]
  jj_w <- which(colnames(cumulative_noncollusive_matches)==a$W[i])
  jj_b <- which(colnames(cumulative_noncollusive_matches)==a$B[i])
  ncmp_w <- cumulative_noncollusive_matches[r, jj_w]
  ncmp_b <- cumulative_noncollusive_matches[r, jj_b]
  a$cncmp_white[i] <- ncmp_w
  a$cncmp_black[i] <- ncmp_b
}
head(a)
tail(a)
```

Thus the final match, between Petrosian and Filip was played in round
28, ended in a draw, and by that time Petrosian had played 18
noncollusive matches and Filip had played 26.

We can see the difference between the two players:

```{r}
diff <- a$cncmp_white-a$cncmp_black  # difference in each game
table(diff)
table(abs(diff))
cumsum(table(abs(diff)))
```

Now create a hyper2 object with appropriate reified monsters:

```{r}
problem <- function(...){stop("must be a white win or a black win or a draw")}

H <- hyper2(pnames=c(players,"white","draw","rest"))
for(i in seq_len(nrow(a))){
  white_player <- a[i,1]
  black_player <- a[i,2]

  collusive <- (white_player %in% collusive_players) & (black_player %in% collusive_players)
  if(collusive){
    drawmonster <- "draw"
  } else {
    drawmonster <- "draw"
  }

  white_player_rested <- a[i,5] - a[i,6] > 0
  black_player_rested <- a[i,6] - a[i,5] > 0

  result <- a[i,3]

  if(white_player_rested){  
    if(       result == "1-0"){ 
      H[c(white_player,             "white","rest"            )] %<>% inc
    } else if(result == "0-1"){ 
      H[c(             black_player                           )] %<>% inc
    } else if(result == "1/2-1/2"){                         
      H[c(                                         drawmonster)] %<>% inc
    } else { problem() }
      H[c(white_player,black_player,"white","rest",drawmonster)] %<>% dec
  } else if(black_player_rested){
    if(       result == "1-0"){ 
      H[c(white_player,             "white"                   )] %<>% inc
    } else if(result == "0-1"){ 
      H[c(             black_player,        "rest"            )] %<>% inc
    } else if(result == "1/2-1/2"){                         
      H[c(                                         drawmonster)] %<>% inc
    } else { problem() }
      H[c(white_player,black_player,"white","rest",drawmonster)] %<>% dec
  } else { # neither player rested, rest monster inactive
    if(       result == "1-0"){ 
      H[c(white_player,             "white"                   )] %<>% inc
    } else if(result == "0-1"){ 
      H[c(             black_player                           )] %<>% inc
    } else if(result == "1/2-1/2"){                         
      H[c(                                         drawmonster)] %<>% inc
    } else { problem() }
      H[c(white_player,black_player,"white",       drawmonster)] %<>% dec
  }
}

H
summary(H)
mH <- maxp(H)
mH
pie(mH)
```