### 2020-10-26 inst/ordertable_to_ranktable.Rmd
more work and rephrasing

---
title: "The print method for class ranktable"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

## Rank tables

Rank tables can be confusing, here is a tiny example that should
clarify things.

```{r}
OT <- cbind(1:5,c(1,5,2:4))
rownames(OT) <- letters[seq_len(nrow(OT))]
colnames(OT) <- c("J1","J2")
OT
```

Object `OT` is a straightforward order table, just like
`skating_table` but smaller and simpler.  We have two columns, one for
each judge (named `J1` and`J2`), and five competitors, `a` through
`e`.  The first row refers to competitor `a`; both judges put
competitor `a` first.  The second row shows that `J1` put `b` second
and `J2` put competitor `b` fifth.  If instead we ask who was placed
third by each judge, this is not directly given; but by looking for
`3` in the table we see that `J1` put `c` third and `J2` put `d`
third.  To answer this and similar questions directly, we can coerce
order table `OT` to a rank table `RT`:

```{r}
RT <- ordertable_to_ranktable(OT)
RT
```

Object `RT` gives the same information as `OT` but in a different
form.  Note that above we see the print method in action, which uses
`ranktable_to_printable_object()`.  Here the _rows_ correspond to the
two judges.  Each row is a rank vector.  Column headings refer to
placings (`c1` is supposed to be read "came first", `c2` "came second"
etc).  Looking at the fifth column (`c5`) we see that `J1` and `J2`
placed `e` and `b` fifth respectively.

Note that it is possible to view object `RT` directly by removing
class `ranktable`:

```{r}
class(RT)
class(RT) <- "matrix"  # removes "ranktable" from class list
RT
```

Note carefully that the second row of `RT` (which corresponds to the
second judge) does *not* match the second column of `OT`.  The way to
read the second row of `RT` is as follows.  We see the entries `1 3 4
5 2` indicating that the first choice was competitor 1 [viz `a`], the
second choice was competitor 3 [viz `c`], the third choice was
competitor 4 [viz `d`], fourth choice was competitor 5 [`e`] and fifth
choice was competitor 2 [`b`].  So the rows are *ranks* (see the
discussion in `rrank.Rd`).  But it is very tempting to read the second
row of `RT` as "judge 2 placed competitor `b` in third place", which
is wrong because this is interpreting the row as an *order*.  This is
why the print method is there.