### 2021-01-19 inst/rock_paper_scissors_monster.Rmd
new vignette

---
title: "Quantifying the rock-paper-scissors effect"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
abbreviated <- TRUE  # change to FALSE for full names
```

## Abstract

Here I quantify the strength of a circular rock-paper-scissors effect
using the `hyper2` package, improving on previous results which merely
established its existence.

## Rock Paper Scissors

Consider the following dataset:

```{r,chesstable}
o <- matrix(c(22,NA,8,13,23,NA,NA,12,10),3,3)
dimnames(o) <- list(
    match=c("T_vs_A","A_vs_K","T_vs_K"),
    player=c("Topalov","Anand","Karpov"))
o
```

This is documented in `chess.Rd` and shows that Topalov played Anand
$22+13=35$ times, winning 22 and losing 13 matches.  The corresponding
likelihood function is `chess`:

```{r,showchess}
chess
maxp(chess)
```

(we refer to this as the "TAK" model).  We might ask whether there is
evidence for the players having differing strengths:

```{r}
equalp.test(chess)
```

Thus we fail to reject $H_0\colon p_T=p_A=p_K$ under T-A-K as the
p-value is about 0.39.

## RPS in chess

The existence of a rock-paper-scissors relationship between the
players was discussed in Hankin (2010) and the null of no RPS
phenomena could be rejected using the Alymer test.

Here we improve on that work by introducing a monster, `M`, that
accounts for the rock-paper-scissors nature of the players.  The RPS
monster helps players that benefit from the RPS nature of the result.
The simplest nontrivial probability model is that all three real
players have the same strength: $p_T=p_A=p_K$, which we write $T=A=K$.
Under $T=A=K$, writing $p_T=p_A=p_K=\alpha$ we can use a binomial test
to calculate the p-value for the hypothesis that $p_M=0$:

```{r,binomtest}
binom.test(sum(diag(o)),sum(o,na.rm=TRUE),alternative="greater")
```

Estimating $p_M$ is also straightforward:

$$
Prob(\mathrm{procyclic}\ \mathrm{win})=\frac{\alpha+M}{2\alpha+M}
$$

(a "procyclic win" is a win that would be expected under a strong
monster).  Here we also have $3\alpha+M=0$, so to estimate the
strength of the monster we would solve

$$
\frac{\alpha+M}{2\alpha+M}=p,\qquad{3\alpha + M}=1
$$

for $\alpha$ and $M$, where $p$ is the probability of a procyclic win.
This gives us

$$
\alpha=\frac{1-p}{2-p},\qquad M=\frac{2p-1}{2-p}.
$$

If we use the maximum likelihood estimate value for a procyclic win of
$\frac{55}{88}=0.625$ we find that $\hat{M}=0.18$ approximately.
Further, we may express $p$, the probability of a pro-cyclic win, as
$p=\frac{1+2M}{2+M}$.  If we observe $a$ procyclic wins and $b$
countercyclic wins a likelihood function for $M$ would then be

$$
\mathcal{L}(M)= p^a(1-p)^b=\left(\frac{1+2M}{2+M}\right)^a\cdot
\left(\frac{1-M}{2+M}\right)^b
$$

(a graph is given below).  We may relax the assumption of equal
strength of the real players by using a more sophisticated hyper2
construction:

```{r,setupH}
H <- hyper2()
H[c("Topalov",                 "M")] %<>% inc(22   )
H[c(          "Anand"             )] %<>% inc(   13)
H[c("Topalov","Anand",         "M")] %<>% dec(22+13)

H[c(          "Anand",         "M")] %<>% inc(   23)
H[c(                  "Karpov"    )] %<>% inc(   12)
H[c(          "Anand","Karpov","M")] %<>% dec(23+12)

H[c("Topalov"                     )] %<>% inc( 8   )
H[c(                  "Karpov","M")] %<>% inc(   10)
H[c("Topalov",        "Karpov","M")] %<>% dec( 8+10)
H
```

We can call this the TAKM model, and find the evaluate:

```{r}
maxp(H)
```

We see that, under TAKM, the monster has strength of about 0.16,
slightly less than the 0.18 obtained for $M$ under the equality
assumption $p_T=p_A=p_K$.  Further, under TAKM, the three "real"
players are again all of approximately equal strength, and we may test
this:

```{r,testequalTAKM}
samep.test(H,c("Topalov","Anand","Karpov"))
```

(note the estimated monster strengths of 0.16 and 0.18 agree with the
previous estimate).  Thus the p-value of about 0.77 indicates failure
to reject the null (compare 0.39 under TAK).  Now we can test
$H_0\colon p_M=0$ under TAKM:

```{r}
specificp.gt.test(H,"M",0)
```

(in the test above, note that the players may appear in a different
order from previously).  The test shows that we may reject the
hypothesis that $p_M=0$ under TAKM with a p-value of $0.04$ (compare
0.012 from the binomial test).

### Profile likelihood

We may plot a profile likelihood curve for $M$:

```{r}
M <- seq(from=0.001,to=0.4,len=40)
L1 <- profsupp(H,"M",M)
L1 <- L1-max(L1)

L2 <- 55*log(1+2*M)+33*log(1-M)-88*log(2+M)
L2 <- L2-max(L2)

plot(M,L1,type="l",col="black",ylab="log likelihood",xlab="Monster strength")
points(M,L2,type="l",col="red")
abline(h=c(0,-2),lty=2)
legend("topright",col=c("black","red"),lty=1,legend=c("TAKM","T=A=K"))
```

Above we see support curves for RPS monster strength using TAKM and
$T=A=K$, showing very similar results.