### 2021-01-21 inst/karate.Rmd
consistentify karate

---
title: "karate"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

Here I analyse a [karate](https://en.wikipedia.org/wiki/Karate)
dataset taken from
[https://en.wikipedia.org/wiki/2018_World_Karate_Championships].

```{r readkaratetable}
karate_table  <- read.table("karate.txt",header=TRUE)
head(karate_table)
```

Converting it to a likelihood function is straightforward:

```{r converttolikelihood}
karate <- hyper2()
for(i in seq_len(nrow(karate_table))){

  p1 <- karate_table$karateka1[i]
  p2 <- karate_table$karateka2[i]

  w1 <- karate_table$wins1[i]
  w2 <- karate_table$wins2[i]

  karate[p1] %<>% inc(w1)
  karate[p2] %<>% inc(w2)
  karate[c(p1,p2)] %<>% dec(w1+w2)
}
karate
karate_maxp <- maxp(karate)
karate_maxp
```

### Package dataset

Following lines create `karate.rda`, residing in the `data/` directory
of the package.

```{r,label=savekaratedataset}
save(karate_table,karate,karate_maxp,file="karate.rda")
```