### 2021-01-29 inst/handover.Rmd
consistentification

---
title: "handover"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("aylmer")
library("magrittr")
options("digits" = 5)
```

This short document is based on Altham and Hankin 2010, which
discussed an interesting dataset originally presented by Lin
et. al. (2009).  It uses more sophisticated idiom furnished by the
`hyper2` package, which is an improved and efficient upgrade of the
primitive `hyperdirichlet` package.  The dataset arose from 69 medical
malpractice claims and refers to the two surgeon–reviewers' answers to
the question "was there a communication breakdown in the hand-off
between physicians caring for the patient?".  The rows of Table 1
correspond to the answers that were given by reviewer 1, and the
columns to the answers given by reviewer 2.

```{r readcurlingtable}
handover_table <- as.matrix(read.table("handover.txt"))
names(dimnames(handover_table)) <- c("reviewer1","reviewer2")
handover_table[3,3] <- NA # structural zero
handover_table
```

Thus we see that in 26 cases, both reviewers answered "yes", in 18
<both said "no".  The table also records partial responses that include
observations missed by one or both reviewers.  Responses missed by
both reviewers are uninformative and automatically rejected:
structural zeros, in `aylmer` terminology.  We first apply the Aylmer
test to assess whether the two reviewers give the same proportion of
"yes" responses:

```
> aylmer.test(a,alternative=function(a){a[1,2]-a[2,1]})

	Aylmer functional test for count data

data:  a
p-value = 0.169
alternative hypothesis: test function exceeds observed

> 
```

(Compare the McNemar test which gives an exact $p$-value of
$7/64\simeq 0.1094$).  We follow the previous workers and assume that
cases are missing at random and seek a likelihood function on
$\theta_{11},\theta_{10},\theta_{01},\theta_{00}$ subject to
$\theta_{ij}\geq 0,\sum_{i,j \in\left\lbrace
0,1\right\rbrace}\theta_{ij}=1$.  Here "1" means "yes" and "0"
means "no", so $\theta_{01}$ is the probability of reviewer 1
responding "no" and reviewer 2 responding "yes".

```{r}
handover <- hyper2()
handover["t11"] %<>% inc(handover_table[1,1])
handover["t10"] %<>% inc(handover_table[1,2])
handover["t01"] %<>% inc(handover_table[2,1])
handover["t00"] %<>% inc(handover_table[2,2])

handover[c("t11","t10")] %<>% inc(handover_table[1,3])  # yes-yes plus yes-no  = 2 
handover[c("t01","t00")] %<>% inc(handover_table[2,3])  #  no-yes plus  no-no  = 9 
handover[c("t11","t01")] %<>% inc(handover_table[3,1])  # yes-yes plus  no-yes = 4 
handover[c("t10","t00")] %<>% inc(handover_table[3,2])  # yes-no  plus  no-no  = 4 
handover
handover_maxp <- maxp(handover)
handover_maxp
```

We may simply test the hypothesis that $\theta_{01} = \theta_{10}$:

```{r}
samep.test(handover,c("t01","t10"))
```

Alternatively we might use the Bayesian paradigm and calculate the
probability of $\theta_{01} > \theta_{10}$:

```{r}
probability(handover,disallowed=function(p){p[2]<p[3]},tol=0.05)
```

the result differing slightly from the previous value, probably due to
numerical issues.

We can now consider the distribution of
$\psi=\theta_{10}/(\theta_{10}+\theta_{01})$ under the posterior
distribution induced by the complete cases:

```{r,samplepsi,message=FALSE,cache=TRUE}
jj <- rp(300000,H=handover)
head(jj)
psifun <- function(x){x[3]/(x[2]+x[3])}
logoddsfun <- function(x){log(x[2]/x[3])}
psi <- apply(jj,1,psifun)
log_odds <- apply(jj,1,logoddsfun)
psi_complete <- rbeta(length(psi),2,6)
```

We can now reproduce the four graphics given by Altham and Hankin:

Sample of $\psi=\theta_{10}/(\theta_{10}+\theta_{01})$ under the
posterior induced by the whole dataset (histogram), together with the
posterior distribution of $\psi$ induced by the complete cases only
(a beta distribution with parameters 2 and 6).

```{r histpsi,echo=FALSE,fig.cap="Sample of $\\psi=\\theta_{10}/(\\theta_{10}+\\theta_{01})$ under the posterior induced  by the whole dataset (histogram), together with the posterior distribution of $\\psi$ induced by the complete cases only (a beta distribution with parameters 2 and 6)"}
hist(psi,nclass=70,probability=TRUE,main=expression(paste("PDF of ",psi)))
p <- seq(from=0,to=1,len=100)
points(p,dbeta(p,2,6),type="l")
abline(v=psifun(handover_maxp),col="red",lwd=5)
segments(x0=0.5,y0=0,y1=1.5,col="blue",lwd=5)
legend("topright",lty=1,lwd=c(1,5,5),col=c("black","red","blue"),
       legend=c(
           "complete cases",
           expression(hat(psi) * " (complete cases)"),
           expression(psi * " = 0.5")
           ))
```

```{r qqplotpsi,echo=FALSE,fig.cap="Quantile-quantile plot of $\\psi=\\theta_{10}/(\\theta_{10}+\\theta_{01})$ under the posterior induced  by the whole dataset against the posterior distribution of $\\psi$ induced by the complete cases only (a beta distribution with parameters 2 and 6)",cache=TRUE}
par(pty='s')
qqplot(psi,psi_complete,asp=1,xlim=0:1,ylim=0:1)
abline(0,1)
```

```{r ecdfpsi,echo=FALSE,fig.cap="Empirical cumulative distribution function of $\\psi=\\theta_{10}/(\\theta_{10}+\\theta_{01})$ under the posterior induced  by the whole dataset, and that of $\\psi$ induced by the complete cases only (a beta distribution with parameters 2 and 6)",cache=TRUE}
plot(psi,seq_along(psi)/length(psi),type="l",col="black",xlab=expression(psi),ylab="quantile")
points(psi_complete,seq_along(psi_complete)/length(psi_complete),type="l",col="red")
legend("topleft",lty=1,col=c("black","red"),legend=c("whole dataset","complete cases"))
```

```{r logoddsqqplot,echo=FALSE,fig.cap="QQ plot of $\\phi=\\log(\\theta_{01}/\\theta_{10})$",cache=TRUE}
qqnorm(log_odds,main="QQ plot of log-odds")
abline(mean(log_odds),sd(log_odds))
```

Figures \@ref(fig:histpsi) to  \@ref(fig:logoddsqqplot) show different aspects of the dataset.

### References {-}

* PME Altham and RKS Hankin 2010.  "Using recently developed software on
a 2x2 table of matched pairs with incompletely classified data".
_Journal of the Royal Statistical Society_, 59:2):377-379
* L West and RKS Hankin 2008. "Exact tests for two-way contingency
tables with structural zeros".  _Journal of Statistical Software_
28(11)
* RKS Hankin 2010. "A generalization of the Dirichlet distribution". 
 _Journal of Statistical Software_ 33(11)
* RKS Hankin 2017. "Partial rank data with the hyper2 package:
  likelihood functions for generalized Bradley-Terry models".  _The R
  Journal_ 9(2)

### Package dataset {-}

Following lines create `handover.rda`, residing in the `data/` directory of the package.

```{r,label=savecurlingdataset}
save(handover_table,handover_maxp,handover,file="handover.rda")
```