### 2021-01-29 inst/moto.Rmd
consistentification

---
title: "Grand Prix motorcycle racing"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
options(width = 120)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
```

## The 2019 FIM motoGP world championship

Here I analyse a series of Grand Prix motorcycling race results using
the `hyper2` package which implements the Plackett-Luce likelihood
function.  Consider the following dataset, taken from Wikipedia:

```{r cars}
moto_table <- read.table("motoGP_2019.txt")
head(moto_table)
```

Each row is a driver (rider?) and each column (after the first) a
venue.  Taking the first row we see that Marquez came second in Qatar
(QAT), first in Argentina, retired in Austin, and so on.  In the first
column, we see the result from Qatar (QAT) in which Marquez came
second, Dovizioso first, Vinales third, and so on.  Notation used also
includes six different classes of no-score such as "Ret" for retired,
"WD" for withdrawn, and so on.  I now use package function
`ordertable2supp()`, that translates data of this type into a
log-likelihood.  In the package, object `moto` is the corresponding
Plackett-Luce likelihood function; we drop the final column which is
the points scored by each rider.

```{r,showmotolikelihoodhead}
moto <- ordertable2supp(moto_table[,-ncol(moto_table)])
head(moto)
```

Above, we see a small part of the Placket-Luce likelihood function for
the 2019 results.  We can find the maximum likelihood estimator by
using the `maxp()` function:

```{r,maxpcalculator,cache=TRUE}
moto_maxp <- maxp(moto)
moto_maxp
```

```{r dotchartmoto, echo=TRUE,fig.cap="Dot chart of riders' strengths"}
dotchart(moto_maxp,pch=16)
```

```{r piechartmoto,echo=TRUE,fig.cap="Pie chart of riders' strengths"}
pie(moto_maxp)
```

In figures \@ref(fig:dotchartmoto) and \@ref(fig:piechartmoto), note
the dominance of Marquez.  It is sometimes preferable to plot the
estimated strengths on a log scale:

```{r dotchartlogmoto, echo=TRUE,fig.cap="Dot chart of riders' strengths, log scale"}
dotchart(log10(moto_maxp),pch=16)
```

(in figure \@ref(fig:dotchartlogmot) observe that the bottom-ranked
player, Lecuona, has a ML strength of zero but the estimate is
slightly positive for numerical reasons).

### Package dataset {-}

Following lines create `moto.rda`, residing in the `data/` directory
of the package.

```{r,label=savemotodataset}
save(moto,moto_table,moto_maxp,file="moto.rda")
```