### 2021-01-30 inst/ordertrans.Rmd
bugfix and typos

---
title: "The `ordertrans()` function"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
```

# Introduction

The `ordertrans()` function can be difficult to understand and this
short document provides some sensible use-cases.  The manpage provides
a very simple example, but here we are going to use an even simpler
example:

```{r,label=simpexample}
x <- c(d=2,a=3,b=1,c=4)
x
```

In the above, object `x` is a named vector with elements
`seq_along(x)` in some order.  It means that competitor `d` came
second, competitor `a` came third, `b` came first and `c` came fourth.
Technically `x` is an order vector because it answers the question
"where did a particular competitor come?"  However, it might equally
be a rank vector because it answers the question "who came first? who
came second?"  (see also the discussion at `rrank.Rd`).

But note that `x` is not helpfully structured to answer either of
these questions: you have to go searching through the names or the
ranks respectively---both of which may appear in any order---to find
the competitor or rank you are interested in.  What we want is to
rearrange the elements of `x` so that the names are in alphabetical
order.  That way we can see straighaway where competitor `a` placed
(in this case, third).  This nontrivial task is accomplished by
function `ordertrans()`:

```{r,label=useordertrans}
o <- ordertrans(x) # by default, sorts names() into alphabetical order
o
```

Observe that objects `x` and `o` are equal in the sense that they are
a rearrangement of one another:

```{r,label=equalrearrangement}
identical(x, o[c(4,1,2,3)])
identical(o, x[c(2,3,4,1)])
```

One consequence of this is that the resulting Plackett-Luce support
functions will be the same:

```{r,showordervec}
(Sx <- ordervec2supp(x))
(So <- ordervec2supp(o))
```

Note carefully that the two support functions above can be
mathematically identical but not formally `identical()` because
neither the order of the brackets, nor the order of terms within a
bracket, is defined.  They look the same on my system but YMMV.  It is
possible that the two support functions might appear to be different
even though they are mathematically the same. We can verify that the
two are mathematically identical using package idiom `==`:

```{r}
Sx==So
```

## Skating

Let us consider the skating dataset and use `ordertrans()` to study it
(note that the skating dataset is analysed in more depth in
`skating.Rmd`).

```{r,label=lookatskating}
skating_table
```

We might ask how judges `J1` and `J2` compare to one another?  We need to
create vectors like `x` and `y` above:

```{r,makej1j2}
j1 <- skating_table[,1]  # column 1 is judge number 1
names(j1) <- rownames(skating_table)
j2 <- skating_table[,2]  # column 2 is judge number 2
names(j2) <- rownames(skating_table)
j1
j2
cbind(j1,j2)
```

In the above, see how objects `j1` and `j2` have identical names, in
the same order.  Observe that `hughes` is ranked 1 (that is, first) by
`J1`, and 4 (that is, 4th) by `J2`.  This makes it sensible to plot
`j1` against `j2`:

```{r j1vsj2,fig.cap="Judge 1 vs judge 2"}
par(pty='s')  # forces plot to be square
plot(j1,j2,asp=1,pty='s',xlim=c(0,25),ylim=c(0,25),pch=16,xlab='judge 1',ylab='judge 2')
abline(0,1)  # diagonal line
for(i in seq_along(j1)){text(j1[i],j2[i],names(j1)[i],pos=4,col='gray',cex=0.7)}
```

In figure \@ref(fig:j1vsj2), we see general agreement but differences
in detail.  For example, `hughes` is ranked first by judge 1 and
fourth by judge 2.  However, other problems are not so easy.  Suppose
we wish to compare the ranks according to likelihood with the ranks
according to some points system.

```{r,label=maxlikeskating,cache=TRUE}
mL <- maxp(skating)  # 'L' for Likelihood
mL
```

Note that in the above, the competitors' names are in alphabetical
order.  We first need to convert strengths to ranks:

```{r,label=strengthstoranks}
mL[] <- rank(-mL)  # minus because ranks orders from weak to strong
mL
```

(note that the names are in the same order as before, alphabetical).
In the above we see that `slutskya` ranks first, `hughes` second, and
so on.  Another way of ranking the skaters is to use a Borda-type
system: essentially the rowsums of the table (and, of course, the
_lowest_ score wins):

```{r,label=pointsskating,cache=TRUE}
mP <- rowSums(skating_table)  # 'P' for Points
mP[] <- rank(mP,ties='first') # positive sign here
mP
```

It is not at all obvious how to compare `mP` and `mL`.  For example,
we might be interested in `hegel`.  It takes some effort to find that
her likelihood rank is 20 and her Borda rank is 19.  Function
`ordertrans()` facilitates this:

```{r,label=ordertransexample}
ordertrans(mP,names(mL))  
```

See above how `ordertrans()` shows the points-based ranks but in
alphabetical order, to facilitate comparison with `mL`.  We can now
plot these against one another:

```{r,label=crapplot,fig.cap="points-based ranks vs likelihood ranks"}
plot(mL,ordertrans(mP,names(mL)))
```

However, figure \@ref(fig:crapplot) is a bit crude.  Function
`ordertransplot()` gives a more visually pleasing output, see figure
\@ref(fig:showoffordertransplot).

```{r,label=showoffordertransplot,fig.cap="points=based rank vs likelihood rank using `ordertransplot()`"}
ordertransplot(mL,mP,xlab="likelihood rank",ylab="Borda rank")
```

So now we may compare judge 1 against likelihood:

```{r transplotjudge1,fig.cap="Likelihood rank vs rank according to Judge 1"}
ordertransplot(mL,j1,xlab="likelihood rank",ylab="Judge 1 rank")
```

In figure \@ref(fig:transplotjudge1), looking at the lower-left
corner, we see (reading horizontally) that the likelihood method
placed Slutskya first, then Hughes second, then Kwan third; while
(reading vertically) judge 1 placed Hughes first, then Kwan, then
Slutskya.