### 2021-01-31 inst/rowing.Rmd
consistentify rowing dataset

---
title: "rowing: analysis of sculls in the 2016 Olympics"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

This short document analyses the `rowing` dataset: sculls in the 2016
olympics.  It is not really standalone and is here as part of a
consistent dataset generation mechanism used in the `hyper2` package.
See the `hyper2` vignette for a more formal discussion.  In Olympic
sculling, up to six individual competitors race a small boat called a
scull over a course of length 2\,km; the object is to cross the
finishing line first.  Note that actual timings are irrelevant, given
the model, as the sufficient statistic is the order in which
competitors cross the finishing line.  The 2016 Summer Olympics is a
case in point: the gold and silver medallists finished less than~5
milliseconds apart, corresponding to a lead of~$\sim
2.5\,\mathrm{cm}$.  The dataset is as follows;

```{r}
show <- function(filename){
  unlist(lapply(strsplit(readLines(filename)," "),paste,collapse=" "))
}					       
show("rowing.txt")
```

The first line shows that `fournier` came first, `cabrera` second, and
so on.  We can convert this dataset into a Plackett-Luce likelihood
function as follows:

```{r definerowing, cache=TRUE}

f <- function(filename){
  o <- strsplit(readLines(filename)," ")

  rowers <- sort(unique(unlist(o)))
  H <- hyper2(list(),0,pnames=rowers)

  for(v in o){
      v <- rev(v)
      for(i in seq_along(v)){
          H[v[i]] %<>% inc           # H[v[i]] %<>% `+`(1)
          H[v[seq_len(i)]] %<>% dec  # H[v[seq_len(i)]] %<>% `-`(1)
      }
  }
  return(H)
}

rowing <- f("rowing.txt")
pnames(rowing)
head(rowing)
```

Object `rowing` is a loglikelihood function on the strengths of the 32
competitors.  Our first step is to find the evaluate:

```{r maxpcalc,cache=TRUE}
rowing_maxp <- maxp(rowing)
rowing_maxp
```

and we may display it using a dot chart:

```{r showrowingmaxppie, echo=FALSE, fig.cap="dot chart of results in the 2016 Olympic Sculls"}
dotchart(rowing_maxp)
```

```{r showrowingmaxppielog, echo=FALSE, fig.cap="dot chart of results in the 2016 Olympic Sculls (log scale)"}
dotchart(log10(rowing_maxp))
```

Figure \@ref(fig:showrowingmaxppie) shows that many competitors have
very small strength and figure \@ref(fig:showrowingmaxppielog) shows
that many competitors have zero strength to numerical precision.
Looking at the original dataset we can see why: observe that `gambour`
came last in every race except for race number 7 (won by `boudina`).
In race 7, he beat only `yaklovlev`.  But we see that `yaklovlev` came
last in every race except race 20, where he beat only `gambour`.  Thus
the maximum likelihood estimate for `gambour + yaklovlev` would be
zero, because these two competitors always lose to everyone else.  It
is thus reasonable to eliminate `gambour` and `yaklovlev` from the
dataset. We may apply this reasoning recursively and end up with the
following dataset:

```{r}
show("rowing_minimal.txt")
```

We may then analyse this:

```{r rowingminimaldef}
rowing_minimal <- f("rowing_minimal.txt")
head(rowing_minimal)
pnames(rowing_minimal)
```

This is a much more manageable dataset, much easier to study and less
cluttered, as competitors with zero estimated strength are not
present. *However, it is incorrect as `drysdale` does not appear.
Something is wrong.*

```{r minimalmaxp,cache=TRUE}
(rowing_minimal_maxp <- maxp(rowing_minimal))
```

```{r maxpcalcminimal,fig.cap="Minimal dataset, maximum likelihood estimate of strength"}
dotchart(rowing_minimal_maxp)
```

### Package dataset {-}

Following lines create `rowing.rda`, residing in the `data/` directory
of the package.

```{r,label=saverowingingdataset}
save(rowing,rowing_minimal,rowing_maxp,rowing_minimal_maxp,file="rowing.rda")
```