### 2021-02-01 inst/eurovision.Rmd
inst/eurovision.Rmd

---
title: "Eurovision Song Contest: semifinal 1, 2009"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
abbreviated <- TRUE  # change to FALSE for full names
```

The file `eurovision.txt`, defined below, is copied from "Eurovision
Song Contest 2009," Wikipedia, accessed May 13, 2018.  It refers to
semi-final 1.  More documentation is given in `eurovision.Rd` [type
`help(euro2009)` at the `R` prompt].

First we specify the matrix as appearing in the Wikipedia page:

```{r, specifymatrix}
eurovision_table <- as.matrix(read.table("eurovision.txt"))
eurovision_table
```

Each row corresponds to a contestant and each column to a judge.  Note
that the matrix is not square: the last two columns correspond to
Germany and the UK, who were judges but not competitors.  I have
removed the first column from the Wikipedia table (which gave the
total of the points), replaced self-voting entries with `NA`, and
replaced blanks with `0`.  Thus the first row corresponds to votes
cast for `ME`, Montenegro.  We see that `BY` (Belarus) awarded them
three points, `AM` (Armenia) five points, and so on.  The first column
corresponds to the points awarded by Monetenegro.  We see that they
placed gave Belarus (`BY`) two points, Armenia (`AM`) four points, and
so on.  Their favourite was Bosnia and Herzegovina (`BA`) to whom they
gave twelve points.  The points system used was:

```{r,pointsforplaces}
points <- c(12,10,8,7,6,5,4,3,2,1)
```

Variable `points` gives the number of points awarded, under Eurovision
rules, to voters' first, second, third, etc choice.  Points for any
competitor are added and the winner is the competitor with the most
points.  However, in the `hyper2` model, the numerical values
themselves do not affect the likelihood function; only the order of
the voters' preferences matters.  The following R idiom translates
`wiki_matrix` into a form suitable for analysis with `hyper2`:

```{r,prefsetter}
preference <- eurovision_table*0  
for(i in seq_along(points)){ preference[eurovision_table == points[i]] <- i }

countries <- data.frame(
    fullname =
        c("Montenegro", "Czech rep", "Belgium", "Belarus", "Sweden",
          "Armenia", "Andorra", "Switzerland", "Turkey", "Israel",
          "Bulgaria", "Iceland", "Macedonia", "Romania", "Finland",
          "Portugal", "Malta", "Bosnia Herz", "Germany", "UK"),
    abbreviation = c("ME","CZ","BE", "BY", "SW", "AM", "AD", "CH",
                     "TR", "IL", "BG", "IS", "MK", "RO", "FI", "PT",
                     "MT", "BA", "DE", "UK")
)

if(abbreviated){ 
    jj <- countries$abbreviation
} else { 
    jj <- countries$fullname
}

competitors <- as.character(jj[1:18])
colnames(preference) <- jj      # voters; 20 countries (18 + DE + UK)
rownames(preference) <- competitors

rownames(eurovision_table) <- jj[1:18]
colnames(eurovision_table) <- jj
```

In the above, matrix `preference` records voters' first, second,
third, etc choice.  A zero entry means no points (nul punkte!) and
`NA` means that voter was forbidden from voting for that player
(countries cannot vote for themselves).  The competitors were the
first 18 countries (the last two countries,
Germany and the UK, voted but did not compete).

```{r}
preference
```
Now, take the first column of `preference`.  This represents the order
of preferences of Montenegro (`ME`).  Their favourite was [last row]
Bosnia & Herzegovina (`BA`), who they gave rank 1 to (that is, 12
points).  Their second favourite was Macedonia (`MK`), their third was
Turkey (`TR`), and so on.  They were not allowed to vote for themselves,
which is why the first row is `NA`.  So the order was: `BA` (first),
`MK` (second) , `TR` (third), `IS` (fourth), `RO` (fifth), `IL`
(sixth), `AM` (seventh), `FI` (eigthth), `BY` (ninth), `MT` (tenth).
The other countries (`CZ`, `BE`, `SW`, `CH`, `BG`, `PT`) came joint
last and did not receive any points from Montenegro.  Note the final
two rows corresponding to votes cast by Germany and the UK, who did
not compete and therefore cast votes for all competitors.

We need to convert matrix `preference` into a likelihood function,
here `eurovision`:

```{r,morestuff}
eurovision <- hyper2()

for(i in seq_len(ncol(preference))){   # cycle through the rows; each row is a voter
    d <- preference[,i,drop=TRUE]
	eurovision <- eurovision + ordervec2supp(d[!is.na(d)])
} # i loop closes
```

```{r,checknull,cache=TRUE}
eurovision_maxp <- maxp(eurovision)
eurovision_maxp
```

```{r,plotpieeurovision,fig.cap="Maximum likelihood strengths of competitors in the 2009 Eurovision song contest, semifinal 1"} 
pie(eurovision_maxp)
```

Now we can check null of equal strengths:

'''{r,label=equalptesteurovision,cache=TRUE}
equalp.test(eurovision)
```

Thus the difference is about 122 units of support, surely significant.

```{r,orderplot}
points_scored <- rowSums(eurovision_table,na.rm=TRUE)  # points
ox <- rank(-points_scored)
oy <- rank(-eurovision_maxp)   # m = MLE
oyp <- ordertrans(oy,names(ox))

par(pty='s') # square plot
plot(ox,oyp,asp=1,pty='s',xlim=c(0,18),ylim=c(0,18),pch=16,
xlab="official order",ylab="my order",main="Eurovision 2009")

par(xpd=TRUE) # allow country names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oyp[i],names(ox)[i],pos=4,col='gray') }
par(xpd=FALSE) # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

### Package dataset {-}

Following lines create `eurovision.rda`, residing in the `data/` directory of the package.

```{r,label=saveeurovisiondataset}
save(eurovision_table,eurovision_maxp,eurovision,file="eurovision.rda")
```