### 2021-02-02 inst/retain.Rmd
bugfix

---
title: "Keeping and discarding players"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

## Functions keep() and keep_flawed()

Function `keep_flawed()` can be confusing.  Originally, the package
included a function `keep()` but this used flawed logic, hence the
name of its replacement.  Here is a tiny example that should clarify
things.  We will use a subset of `skating_table`:

```{r}
jj <- skating_table[1:3,1:6]
rownames(jj) <- letters[1:3]
jj
```

Above, rows of object `jj` correspond to competitors and columns to
judges.  Object `jj` does not function as a rank table because the
entries are not sequential.

We could convert explicitly:

```{r}
a <- apply(jj,2,rank)
a
```

So each column is a permutation of `123`.  So, for example judge 1
ranked `a` first, `b` third, and `c` second.  We can coerce to a
`hyper2` object with `ordertable2supp()`:

```{r}
OT <- ordertable2supp(a)
OT
```

(function `ordertable2supp()` does this automatically, so
`ordertable2supp(jj)` would have given the same result).  Now suppose
we are interested only in the first two competitors, `a` and `b`,
which are competitors 1 and 2.

```{r}
keep_flawed(OT, c("a","b"))
```

Above, function `keep_flawed()` has effectively taken `OT`, set `c=0`,
and then discarded the `c^5` term.  This process, while natural from a
computing perspective, has the effect of replacing `(a+c)^-4` with
`a^-4`, which has no natural probabilistic interpretation: we are
ignoring `c`'s victories but nevertheless interpreting the joint
strength term on the denominator [viz `a+c`] as a *loss* for `a`.

What happens if we keep only `a`?

```{r}
keep_flawed(OT,"a")
```

the above support function is meaningless (?) as, having set `b=0` we
then treat `a+b` as simply `a` which sort of makes sense until we
realise that `(a+b)^-6` reduces to `a^-6`.  It is as though `a` has
wins only 3 trials out of 8, against a competitor of known zero
strength whose (impossible) wins we ignore.  Perhaps this has a
natural interpretation in probability, but if so I don't see it.