### 2021-02-04 inst/volleyball.Rmd
consistentify volleyball

---
title: "volleyball dataset"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

This short document discusses a dataset first presented by Hankin
(2010), although here only the first 52 (out of 4000) observations are
used.

```{r readcurlingtable}
volleyball_table <- as.matrix(read.table("volleyball.txt",header=TRUE))
nrow(volleyball_table)
head(volleyball_table)
```

The first row shows a game between a team comprising `p1`, `p4`, and
`p8` against a team comprising `p5` and `p6`.  Team `p1 p4 p8` won.
We may use function `volley()` to convert this to a likelihood
function:

```{r}
volleyball <- volley(volleyball_table)
(volleyball_maxp <- maxp(volleyball))
```

The original synthetic dataset was prepared using Zipf's law for the
players' strengths, so we may test the hypothesis that this is the
case; $H_0\colon p_i\propto i^{-1}$:

```{r}
zipf <- function(n){jj <- 1/(1:n); jj/sum(jj)}
zipf(9)
(null_support <- loglik(indep(zipf(9)),volleyball))
(alternative_support <- loglik(indep(volleyball_maxp),volleyball))
(Lambda <- 2*(alternative_support-null_support))
pchisq(Lambda,df=8,lower.tail=FALSE)
```

somewhat disappointingly rejecting the null with a $p$-value of about
4\% (and indeed--just---with a two units of support per degree of
freedom criterion).  However, it is not clear to me the extent to
which Wilks's theorem is applicable here [Wilks is an asymptotic
result; recall that we have only 52 observations here], nor whether
the support criterion is appropriate with 8 degrees of freedom.

### Package dataset {-}

Following lines create `volleyball.rda`, residing in the `data/` directory of the package.

```{r,label=savecurlingdataset}
save(volleyball_table,volleyball_maxp,volleyball,file="volleyball.rda")
```