### 2021-02-06 inst/powerboat.Rmd
consitentify powerboat

---
title: "powerboat"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

[powerboat](https://en.wikipedia.org/w/index.php?title=2018_F1_Powerboat_World_Championship&oldid=920386507)
racing is an important and prestigious sport.  Here I analyse a
dataset taken from the 2018 F1 Powerboat world championships.

```{r readpowerboattable}
powerboat_table <- read.table("powerboat.txt",header=TRUE)
```

We need some data cleaning first; specifically a zero entry is treated
as "did not start":

```{r}
powerboat_table[powerboat_table==0] <- "DNS"
powerboat_table[20:21,8] <- 0  # above does not apply to points
powerboat_table
```

Each row is a driver and each column a venue (apart from the final
column, which is points scored).  Thus we see from the first column
that, in Portugal, Torrente came first, Stark second, T. AlQamzi
second, and so on.

In this form we can easily convert to a support function:

```{r,definsupportb,cache=TRUE}
powerboat <- ordertable2supp(powerboat_table[,1:7])
head(powerboat)
(powerboat_maxp <- maxp(powerboat))
pie(powerboat_maxp)
par(pty="s")
points <- as.numeric(powerboat_table[,8])
names(points) <- rownames(powerboat_table)
points
points <- points[order(names(points))]
points
plot(powerboat_maxp,points*max(powerboat_maxp)/max(points),xlab="likelihood",ylab="points",pch=16)
abline(0,1)
```

### Package dataset

Following lines create `powerboat.rda`, residing in the `data/`
directory of the package.

```{r,label=savepowerboatdataset}
save(powerboat_table,powerboat,powerboat_maxp,file="powerboat.rda")
```