### 2021-02-06 inst/kka.Rmd
consistentify kka

---
title: "Karpov, Kasprov, Anand"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

This file defines dataset `kka` which is the observed scorelines
between the three players.  It is used to calculate likelihood
functions: `karpov_kasparov_anand`, `kka_3whites` and `kka_3draws`
This Rmd file supercedes files `inst/karpov_kasparov_anand.R` and
`inst/kka_3draws.R` and `inst/kka_array.R`.  Some more detail is given
in `man/karpov_kasparov_anand.Rd`.

```{r setupkka}
kka <- c(
## Kasparov vs Karpov
karpov_plays_white_beats_kasparov = 18,  # 12 on p1, 6 on p2
kasparov_plays_white_beats_karpov = 30, # 13 on p1, 17 on p2
kasparov_plays_white_losesto_karpov = 07,
karpov_plays_white_losesto_kasparov = 09,
karpov_plays_white_draws_kasparov = 11+13+14+17+14+03,
kasparov_plays_white_draws_karpov = 14+12+11+08+11+01,

## Kasparov vs Anand
kasparov_plays_white_beats_anand = 15,
anand_plays_white_beats_kasparov = 06,
anand_plays_white_losesto_kasparov = 11,
kasparov_plays_white_losesto_anand = 02,
kasparov_plays_white_draws_anand = 26,
anand_plays_white_draws_kasparov = 20,

## Karpov vs Anand
karpov_plays_white_beats_anand = 07,
anand_plays_white_beats_karpov = 18,
anand_plays_white_losesto_karpov = 05,
karpov_plays_white_losesto_anand = 13,
karpov_plays_white_draws_anand = 29,
anand_plays_white_draws_karpov = 20
)
kka
```

Quite a lot of information here, but for example we might ask what the
scoreline is for Karpov playing white against Kasparov:

```{r kkaexample}
(a1 <- c(
kka["karpov_plays_white_beats_kasparov"],
kka["karpov_plays_white_draws_kasparov"],
kka["karpov_plays_white_losesto_kasparov"]
))
```

If we were to ask instead what happens if Karpov plays black against
Kasparov this would be:

```{r kkaexample2}
(a2 <- c(
kka["kasparov_plays_white_beats_karpov"],
kka["kasparov_plays_white_draws_karpov"],
kka["kasparov_plays_white_losesto_karpov"]
))
```

To investigate the effect of playing white we might do this:

```{r anamwa}
M <- rbind(a1,rev(a2))
dimnames(M) <- list(karpov_plays=c("White","Black"),result=c("karpov wins","karpov draws","kaprov loses"))
M
fisher.test(M)
```

but we need a likelihood function.  Here I define likelihood function
`karpov_kasparov_anand` but to save typing we work with temporary
object `H` and rename it at the end:

```{r defineH}

results <- as.list(kka)
attach(results)
H <- hyper2()

## Kasparov vs Karpov
karpov_vs_kasparov <- c("Karpov","Kasparov","white","draw")  # all "players", real and imaginary

H <- H + trial(c("Karpov"  ,"white"), karpov_vs_kasparov, karpov_plays_white_beats_kasparov)
H <- H + trial(c("Kasparov","white"), karpov_vs_kasparov, kasparov_plays_white_beats_karpov)
H <- H + trial(  "Karpov"           , karpov_vs_kasparov, kasparov_plays_white_losesto_karpov)
H <- H + trial(  "Kasparov"         , karpov_vs_kasparov, karpov_plays_white_losesto_kasparov)
H <- H + trial(  "draw"             , karpov_vs_kasparov, karpov_plays_white_draws_kasparov)
H <- H + trial(  "draw"             , karpov_vs_kasparov, kasparov_plays_white_draws_karpov)

## Kasparov vs Anand
kasparov_vs_anand <- c("Kasparov","Anand","white","draw")
H <- H + trial(c("Kasparov","white"), kasparov_vs_anand, kasparov_plays_white_beats_anand)
H <- H + trial(c("Anand"   ,"white"), kasparov_vs_anand, anand_plays_white_beats_kasparov)
H <- H + trial(c("Kasparov"        ), kasparov_vs_anand, anand_plays_white_losesto_kasparov)
H <- H + trial(c("Anand"           ), kasparov_vs_anand, kasparov_plays_white_losesto_anand)
H <- H + trial(  "draw"             , kasparov_vs_anand, kasparov_plays_white_draws_anand)
H <- H + trial(  "draw"             , kasparov_vs_anand, anand_plays_white_draws_kasparov)

## Karpov vs Anand
karpov_vs_anand <- c("Karpov","Anand","white","draw")
H <- H + trial(c("Karpov","white"), karpov_vs_anand, karpov_plays_white_beats_anand)
H <- H + trial(c("Anand" ,"white"), karpov_vs_anand, anand_plays_white_beats_karpov)
H <- H + trial(  "Karpov"         , karpov_vs_anand, anand_plays_white_losesto_karpov)
H <- H + trial(  "Anand"          , karpov_vs_anand, karpov_plays_white_losesto_anand)
H <- H + trial(  "draw"           , karpov_vs_anand, karpov_plays_white_draws_anand)
H <- H + trial(  "draw"           , karpov_vs_anand, anand_plays_white_draws_karpov)

detach(results)

karpov_kasparov_anand <- H
karpov_kasparov_anand
(karpov_kasparov_anand_maxp <- maxp(karpov_kasparov_anand))
```

Test the hypothesis that all three players have the same strength; first do the free optimization:

```{r testall3}
samep.test(karpov_kasparov_anand,c("Karpov","Kasparov","Anand"))
```

and we reject that null.  Now test the hypothesis that playing white confers no advantage:

```{r whitenoadvantage,cache=TRUE}
specificp.test(karpov_kasparov_anand,"white",0)
```

We now allow each player to have a personalised draw monster.  The
resulting likelihood function is called `kka_3draws` but the idiom
below uses `H` to save typing.

```{r kka3draws}
library("hyper2")
H <- hyper2()

results <- as.list(kka)
attach(results)

karpov_vs_kasparov <- c("Karpov","Kasparov","Karpov_draw","Kasparov_draw","white")
draw1 <- c("Karpov_draw","Kasparov_draw")

H <- H + trial(c("Karpov"  ,"white"), karpov_vs_kasparov, karpov_plays_white_beats_kasparov)
H <- H + trial(c("Kasparov","white"), karpov_vs_kasparov, kasparov_plays_white_beats_karpov)
H <- H + trial(c("Karpov"  )        , karpov_vs_kasparov, kasparov_plays_white_losesto_karpov)
H <- H + trial(c("Kasparov")        , karpov_vs_kasparov, karpov_plays_white_losesto_kasparov)
H <- H + trial(draw1                , karpov_vs_kasparov, karpov_plays_white_draws_kasparov)
H <- H + trial(draw1                , karpov_vs_kasparov, kasparov_plays_white_draws_karpov)

## Kasparov vs Anand
kasparov_vs_anand <- c("Kasparov","Anand","Kasparov_draw","Anand_draw","white")
draw2 <- c("Kasparov_draw","Anand_draw")
H <- H + trial(c("Kasparov","white"), kasparov_vs_anand, kasparov_plays_white_beats_anand)
H <- H + trial(c("Anand"   ,"white"), kasparov_vs_anand, anand_plays_white_beats_kasparov)
H <- H + trial(c("Kasparov"  )      , kasparov_vs_anand, anand_plays_white_losesto_kasparov)
H <- H + trial(c("Anand")           , kasparov_vs_anand, kasparov_plays_white_losesto_anand)
H <- H + trial(draw2                , kasparov_vs_anand, kasparov_plays_white_draws_anand)
H <- H + trial(draw2                , kasparov_vs_anand, anand_plays_white_draws_kasparov)

## Karpov vs Anand
karpov_vs_anand <- c("Karpov","Anand","Karpov_draw","Anand_draw","white")
draw3 <- c("Karpov_draw","Anand_draw")

H <- H + trial(c("Karpov","white"), karpov_vs_anand, karpov_plays_white_beats_anand)
H <- H + trial(c("Anand" ,"white"), karpov_vs_anand, anand_plays_white_beats_karpov)
H <- H + trial(c("Karpov"        ), karpov_vs_anand, anand_plays_white_losesto_karpov)
H <- H + trial(c("Anand"         ), karpov_vs_anand, karpov_plays_white_losesto_anand)
H <- H + trial(draw3              , karpov_vs_anand, karpov_plays_white_draws_anand) 
H <- H + trial(draw3              , karpov_vs_anand, anand_plays_white_draws_karpov)

detach(results)

kka_3draws <- H
kka_3draws 
```

and one natural test is that the three draw monsters are of equal strength:

```{r threedrawsequal,cache=TRUE}
(kka_3draws_maxp <- maxp(kka_3draws))
pie(kka_3draws_maxp)
samep.test(kka_3draws,c("Karpov_draw","Kasparov_draw","Anand_draw"))
```

Now allow each player to have distinct white strength.  We create
`kka_3whites`, again using temporary variable `H` to save typing:

```{r kka3whitedef}
H <- hyper2()

results <- as.list(kka)
attach(results)

D <- "draw"

## First: Karpov vs Kasparov
karpov_plays_white_vs_kasparov <- c("Karpov","Kasparov","Karpov_white","draw"  )  # "players" 
kasparov_plays_white_vs_karpov <- c("Karpov","Kasparov","Kasparov_white","draw")
H <- H + trial(c("Karpov"  ,"Karpov_white"  ), karpov_plays_white_vs_kasparov, karpov_plays_white_beats_kasparov  ) # Karpov wins playing white
H <- H + trial(c("Kasparov","Kasparov_white"), kasparov_plays_white_vs_karpov, kasparov_plays_white_beats_karpov  ) # Kasparov wins playing white
H <- H + trial(c("Kasparov")                 , karpov_plays_white_vs_kasparov, karpov_plays_white_losesto_kasparov) # Kasparov wins playing black
H <- H + trial(c("Karpov"  )                 , kasparov_plays_white_vs_karpov, kasparov_plays_white_losesto_karpov) # Karpov wins playing black
H <- H + trial(D                             , karpov_plays_white_vs_kasparov, karpov_plays_white_draws_kasparov  ) # Karpov white, draws
H <- H + trial(D                             , kasparov_plays_white_vs_karpov, kasparov_plays_white_draws_karpov  ) # Kasparov white, draws

## Second: Karpov vs Anand
karpov_plays_white_vs_anand <- c("Karpov","Anand","Karpov_white","draw"  )
anand_plays_white_vs_karpov <- c("Karpov","Anand","Anand_white","draw")
H <- H + trial(c("Karpov","Karpov_white"), karpov_plays_white_vs_anand, karpov_plays_white_beats_anand  ) # Karpov wins playing white
H <- H + trial(c("Anand" ,"Anand_white" ), anand_plays_white_vs_karpov, anand_plays_white_beats_karpov  ) # Anand wins playing white
H <- H + trial(c("Anand"                ), karpov_plays_white_vs_anand, karpov_plays_white_losesto_anand) # Anand wins playing black
H <- H + trial(c("Karpov"               ), anand_plays_white_vs_karpov, anand_plays_white_losesto_karpov) # Karpov wins playing black
H <- H + trial(D                         , karpov_plays_white_vs_anand, karpov_plays_white_draws_anand  ) # Karpov white, draws
H <- H + trial(D                         , anand_plays_white_vs_karpov, anand_plays_white_draws_karpov  ) # Anand white, draws

## Third: Kasparov vs Anand
anand_plays_white_vs_kasparov <- c("Anand","Kasparov","Anand_white","draw"   )
kasparov_plays_white_vs_anand <- c("Anand","Kasparov","Kasparov_white","draw")
H <- H + trial(c("Kasparov","Kasparov_white"), kasparov_plays_white_vs_anand, kasparov_plays_white_beats_anand  ) # Kasparov wins playing white
H <- H + trial(c("Anand"   ,"Anand_white"   ), anand_plays_white_vs_kasparov, anand_plays_white_beats_kasparov  ) # Anand wins playing white
H <- H + trial(c("Anand"                    ), kasparov_plays_white_vs_anand, kasparov_plays_white_losesto_anand) # Anand wins playing black
H <- H + trial(c("Kasparov"                 ), anand_plays_white_vs_kasparov, anand_plays_white_losesto_kasparov) # Kasparov wins playing black
H <- H + trial(D                             , kasparov_plays_white_vs_anand, kasparov_plays_white_draws_anand  ) # Kasparov white, draws
H <- H + trial(D                             , anand_plays_white_vs_kasparov, anand_plays_white_draws_kasparov  ) # Anand white, draws

detach(results)

kka_3whites <- H
kka_3whites
```

Natural test for equality of the three white monsters's strengths:

```{r threewhitesequal,cache=TRUE}
(kka_3whites_maxp <- maxp(kka_3whites))
pie(kka_3whites_maxp)
samep.test(kka_3whites,c("Karpov_white","Kasparov_white","Anand_white"))
```
(fail to reject the null).

We now create a 3x3x3 array of results for the dataset.

```{r kkaarray}

library("hyper2")
library("abind")
attach(as.list(kka))

players <- c("Anand","Karpov","Kasparov")

plays_white_wins <- matrix(NA,3,3)
dimnames(plays_white_wins) <- list(plays_white_wins=players,plays_black_loses=players)
plays_white_wins["Anand"   ,"Karpov"  ] <- anand_plays_white_beats_karpov
plays_white_wins["Anand"   ,"Kasparov"] <- anand_plays_white_beats_kasparov
plays_white_wins["Karpov"  ,"Anand"   ] <- karpov_plays_white_beats_anand
plays_white_wins["Karpov"  ,"Kasparov"] <- karpov_plays_white_beats_kasparov
plays_white_wins["Kasparov","Anand"   ] <- kasparov_plays_white_beats_anand
plays_white_wins["Kasparov","Karpov"  ] <- kasparov_plays_white_beats_karpov

plays_white_draws  <- matrix(NA,3,3)
dimnames(plays_white_draws) <- list(plays_white_draws=players,plays_black_draws=players)
plays_white_draws["Anand"   ,"Karpov"  ] <- anand_plays_white_draws_karpov
plays_white_draws["Anand"   ,"Kasparov"] <- anand_plays_white_draws_kasparov
plays_white_draws["Karpov"  ,"Anand"   ] <- karpov_plays_white_draws_anand
plays_white_draws["Karpov"  ,"Kasparov"] <- karpov_plays_white_draws_kasparov
plays_white_draws["Kasparov","Anand"   ] <- kasparov_plays_white_draws_anand
plays_white_draws["Kasparov","Karpov"  ] <- kasparov_plays_white_draws_karpov

plays_white_loses <- matrix(NA,3,3)
dimnames(plays_white_loses) <- list(plays_white_loses=players,plays_black_wins=players)
plays_white_loses["Karpov"   ,"Anand"   ] <- karpov_plays_white_losesto_anand
plays_white_loses["Kasparov" ,"Anand"   ] <- kasparov_plays_white_losesto_anand
plays_white_loses["Anand"    ,"Karpov"  ] <- anand_plays_white_losesto_karpov
plays_white_loses["Kasparov" ,"Karpov"  ] <- kasparov_plays_white_losesto_karpov
plays_white_loses["Anand"    ,"Kasparov"] <- anand_plays_white_losesto_kasparov
plays_white_loses["Karpov"   ,"Kasparov"] <- karpov_plays_white_losesto_kasparov
detach(as.list(kka))

    plays_white_wins
    plays_white_draws
    plays_white_loses

kka_array <- abind(
    plays_white_wins,
    plays_white_draws,
    plays_white_loses,
    along=3)
dimnames(kka_array)[[3]] <- c("plays_white_wins", "plays_white_draws", "plays_white_loses")
kka_array
```

### Package dataset

Following lines create `kka.rda`, residing in the `data/` directory of
the package.

```{r,label=savecurlingdataset,cache=TRUE}
save(kka,karpov_kasparov_anand,kka_3draws,kka_3whites,kka_array,file="kka.rda")
```