### 2021-02-07 inst/table_tennis_serve.Rmd
consistentify table tennis

---
title: "The importance of serve in table tennis"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{table_tennis_serve}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# The importance of serve in table tennis.

(In this vignette, object `H` is created.  It is identical to
object `table_tennis` in the package; to see it type
`data(table_tennis)` at the command line, and to see the help
page, type `help(table_tennis)`).

Suppose we have three players, "A","B", and "C" who play table tennis.
We seek to quantify the effect of the serve.  The scoreline is as
follows:

* A vs B, A serves, 5-1
* A vs B, B serves, 1-3
* A vs C, A serves, 4-1
* A vs C, C serves, 1-2

(note that B does not play C).  This dataset is easily analysed with
the `hyper2` package although the model implicitly assumes that the
serve strength does not depend on who is serving.  The first step is
to define a `hyper2` object:

```{r,message=FALSE}
library("hyper2",quietly=TRUE)
(H <- hyper2(pnames=c("S","a","b","c")))
```

so the likelihood function is uniform, for the only term is raised to
the zeroth power (players' names are not capitalised in R idiom for
clarity).  Now we have to input the data, and we will go through the
four scorelines above, one by one.  Note that this approach implicitly
assumes independence of scores.

### A vs B, A serves, score 5-1:
```{r}
H[c("a","S")]     %<>% "+"(5  ) # a+S win 5 games
H["b"]            %<>% "+"(  1) # b wins 1 game
H[c("a","b","S")] %<>% "-"(5+1) # a+b+S play 1+5=6 games
H
```

The likelihood function is of the form
$\frac{\left(a+S\right)^5b}{\left(a+b+S\right)^6}$.  We can add the
other observations to $H$:

### A vs B, B serves, score 1-3:
```{r}
H[c("b","S")]     %<>% "+"(3  ) # b+S win 3 games
H["a"]            %<>% "+"(  1) # a wins 1 game 
H[c("a","b","S")] %<>% "-"(3+1) # a+b+S play 1+3=4 games
```

### A vs C, A serves, score 4-1:
```{r}
H[c("a","S")]     %<>% "+"(4  ) # a+S win 4 games
H["c"]            %<>% "+"(  1) # c wins 1 game 
H[c("a","c","S")] %<>% "-"(4+1) # a+c+S play 1+4=5 games
```

### A vs C, C serves, score 1-2:
```{r}
H["a"]            %<>% "+"(1  ) # a wins 1 game
H[c("c","S")]     %<>% "+"(  2) # c+S win 2 games
H[c("a","c","S")] %<>% "-"(1+2) # a+c+S play 1+2=3 games
```

## Likelihood function

For the entire dataset we have

```{r}
table_tennis <- H
table_tennis
```

Here $H$ is a likelihood function on $a,b,c,S$, algebraically
proportional to

\[
\frac{(S+a)^9(S+b)^3(S+c)^2a^2bc}{(S+a+b)^{10}(S+a+c)^8}
\]

# Maximum likelihood estimate for the strengths

We now find the unconstrained MLE:

```{r}
table_tennis_maxp <- maxp(table_tennis)
table_tennis_maxp
```

# Test hypothesis that serve strength S is zero

```{r testservezero}
specificp.test(table_tennis,"S",0.01)
```

Thus the tail area, pval, is about 0.022 which gives us a significant
result: the serve has nonzero strength.

## Assessment of equal player strength.

We now assess the hypothesis that the three (real) players are of zero
strength:  $H_E\colon p_A=p_B=p_C$.

## Assessment of B vs C

The maximum likelihood estimate for strengths allows us to assess what
might happen when B plays C (which was not observed).  For
convenience, here is the evaluate:

```{r}
maxp(table_tennis)
```

If B serves, we would have $B+S:C\simeq 0.14+0.45:0.15\equiv 59:15$,
or B wins with probability of about $0.78$.  If C serves we have
$B:C+S\simeq 0.14:0.16+0.45\equiv 14:61$, so B wins with a probability
of about $0.19$.

# Profile likelihood for serve strength

We can give a profile likelihood for $S$ by maximizing the likelihood
of $(S,p_a,p_b,p_c)$ subject to a particular value of $S$ (in addition
to the unit sum constraint).  The following function defines a profile
likelihood function for $S$:

```{r}
S_vals <- seq(from=0.02,to=0.85,len=30)  # possible values for S
plot(S_vals,profsupp(table_tennis,"S",S_vals))
abline(h=c(0,-2))
```

So we can see that the credible range is from about 0.07-0.8 (and in
particular excludes zero).

# Profile likelihood for B vs C

The profile likelihood technique can be applied to the question of the
strengths of players B and C, even though no competition between the
players was observed:

```{r,dpi=72}
proflike <- function(bc,give=FALSE){
  B <- bc[1]
  C <- bc[2]
  if(B+C>=1){return(NA)}

  objective <- function(S){     #

    jj <- c(S,A=1-(S+B+C),B,C)  # B,C fixed
    loglik(indep(jj),H)         # no minus: we use maximum=T in optimize()
  }

  maxlike_constrained <-  # single DoF is S [B,C given, A is fillup]
    optimize(             # single DoF -> use optimize() not optim()
        f = objective,         
        interval = c(0,1-(B+C)),
        maximum = TRUE
    )

  if(give){
    return(maxlike_constrained)
  } else{
    return(maxlike_constrained$objective)
  }
}

small <- 1e-5
n <- 50
p <- seq(from=small,to=1-small,length=n)
jj <- expand.grid(p,p)
support <- apply(jj,1,proflike)
support <- support-max(support,na.rm=TRUE)
supportx <- support
dim(supportx) <- c(n,n)

x <- jj[,1]+jj[,2]/2
y <- jj[,2]*sqrt(3)/2
plot(x,y,cex=(support+6)/12, 
   pch=16,asp=1,xlim=c(-0.1,1.1),ylim=c(-0.2,0.9),
   axes=FALSE,xlab="",ylab="",main="profile likelihood for B,C")
polygon(x=c(0,1/2,1),y=c(0,sqrt(3)/2,0))
text(0.07,-0.04,'A+S=1, B=0, C=0')
text(0.50,+0.90,'A+S=0, B=1, C=0')
text(0.93,-0.04,'A+S=0, B=0, C=1')
```

But what one is really interested in is the relative strength of
player B and player C.  We are indifferent between $p_B=0.1,p_C=0.2$
and $p_B=0.3,p_C=0.6$.  To this end we need the profile likelihood
curve of the log contrast, $X=\log\left(p_B/p_C\right)$.

```{r}
logcontrast <- apply(jj,1,function(x){log(x[1]/x[2])})
plot(logcontrast,support,xlim=c(-5,5),ylim=c(-4,0))
abline(h=-2)
```

From the graph, we can see that the two-units-of-support interval is
from about $-3$ to $+3$ which would translate into a probability of B
winning against C being in the interval
$\left(\frac{1}{1+e^{3}},\frac{1}{1+e^{-3}}\right)$, or about
$(0.047,0.953)$.  However, it is not clear if $p_B$ and $p_C$ are
meaningful in isolation, as a real table tennis point has to have a
server and the value of $S$ is itself uncertain.

### Package dataset

Following lines create `table_tennis.rda`, residing in the `data/`
directory of the package.

```{r,label=savetabletennisdataset}
save(table_tennis,file="table_tennis.rda")
```