### 2021-11-01 inst/pudding.Rmd
pudding

---
title: "Analysis of puddings following Davidson 1970"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
library("PlackettLuce")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document creates the `pudding` suite of R objects and
presents some preliminary analysis.  It follows Davidson 1970 and the
`PlackettLuce` package.

Here is the pudding result:

```{r}
pudding
```

Here is a method to convert the pudding matrix to a likelihood function:

```{r}

    numbers <- as.character(1:6)
    puddings <- paste("pudding",numbers,sep="_")
    tyers    <- paste("tie",numbers,sep="_")

`convert` <- function(M){

    H <- hyper2()
    I <- hyper2()
    for(i in seq_len(nrow(M))){
        r_ij <- M[i,3]
        w_ij <- M[i,4]
        w_ji <- M[i,5]
        t_ij <- M[i,6]
        H[c(puddings[c(M[i,1]       )]      )] %<>% inc(w_ij)
        H[c(puddings[c(       M[i,2])]      )] %<>% inc(w_ji)
        H[c(                           "tie")] %<>% inc(t_ij) 
        H[c(puddings[c(M[i,1],M[i,2])],"tie")] %<>% dec(r_ij)

        I[c(puddings[c(M[i,1]       )]      )] %<>% inc(w_ij)
        I[c(puddings[c(       M[i,2])]      )] %<>% inc(w_ji)
        I[c(tyers   [c(M[i,1],M[i,2])]      )] %<>% inc(t_ij)
        I[c(puddings[c(M[i,1],M[i,2])],tyers[c(M[i,1],M[i,2])])] %<>% dec(r_ij)    
    }

    return(list(H,I))
}
```

Now use it:

```{r calculatelikelihoodfunctionsandmaximize,cache=TRUE}
lf <- convert(pudding)
H1 <- lf[[1]]
H2 <- lf[[2]]
m1 <- maxp(H1)
m2 <- maxp(H2)
```

```{r}
m1
m2
```

We will analyse `m1` first:

```{r}
pie(m1)
```

First we use `H1` to test a null of equal pudding strengths:

```{r,m1test}
samep.test(H1,puddings)
```

so according to `samep.test()` the puddings are all the same strength.
Now some tests where the puddings' tying ability is allowed to vary:

```{r,m2test,cache=TRUE}
samep.test(H2,puddings)
samep.test(H2,tyers)
```