### 2021-11-04 inst/pudding.Rmd
work

---
title: "Analysis of pudding following Davidson 1970"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
library("PlackettLuce")
library("partitions")
options("digits" = 5)
showsome <- function(D){
jj <- round(D[c(1:4,nrow(D)),],4)
jj[4,] <- "..."
rownames(jj)[4] <- "..."
noquote(jj)
}
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document creates the `pudding` suite of R objects and
presents some preliminary analysis.  It follows Davidson 1970 and the
`PlackettLuce` package.

Here is the pudding result:

```{r}
pudding_table <- pudding  # standard hyper2 terminology
rownames(pudding_table) <- apply(allbinom(6,2),2,paste,collapse="v")
pudding_table
```

Here is a method to convert the pudding matrix to a likelihood function:

```{r,cache=TRUE}

    numbers <- as.character(1:6)
    puddings <- paste("pudding",numbers,sep="_")
    tyers    <- paste("tie",numbers,sep="_")

`convert` <- function(M){

    H <- hyper2()
    I <- hyper2()
    J <- hyper2()
    for(i in seq_len(nrow(M))){
        r_ij <- M[i,3]
        w_ij <- M[i,4]
        w_ji <- M[i,5]
        t_ij <- M[i,6]
        H[c(puddings[c(M[i,1]       )]      )] %<>% inc(w_ij)
        H[c(puddings[c(       M[i,2])]      )] %<>% inc(w_ji)
        H[c(                           "tie")] %<>% inc(t_ij) 
        H[c(puddings[c(M[i,1],M[i,2])],"tie")] %<>% dec(r_ij)

        I[c(puddings[c(M[i,1]       )]      )] %<>% inc(w_ij)
        I[c(puddings[c(       M[i,2])]      )] %<>% inc(w_ji)
        I[c(tyers   [c(M[i,1],M[i,2])]      )] %<>% inc(t_ij)
        I[c(puddings[c(M[i,1],M[i,2])],tyers[c(M[i,1],M[i,2])])] %<>% dec(r_ij)

        J[c(puddings[c(M[i,1]       )],"first")] %<>% inc(w_ij)
        J[c(puddings[c(       M[i,2])]        )] %<>% inc(w_ji)
        J[c(tyers   [c(M[i,1],M[i,2])]        )] %<>% inc(t_ij)
        J[c(puddings[c(M[i,1],M[i,2])],tyers[c(M[i,1],M[i,2])],"first")] %<>% dec(r_ij)
    }

    return(list(H,I,J))
}
```

Now use it:

```{r calculatelikelihoodfunctionsandmaximize,cache=TRUE}
lf <- convert(pudding_table)
pudding <- lf[[1]] # standard hyper2 terminology
pudding2 <- lf[[2]]
pudding3 <- lf[[3]]
pudding_maxp  <- maxp(pudding)
pudding2_maxp <- maxp(pudding2)
pudding3_maxp <- maxp(pudding3)
```

```{r}
pudding_maxp
pudding2_maxp
pudding3_maxp
```

# Pudding 1

We will analyse `pudding` first:

```{r}
summary(pudding)
pie(pudding_maxp)
```

First we use `pudding` to test a null of equal pudding strengths:

```{r,puddingtest,cache=TRUE}
samep.test(pudding,puddings)
```

so according to `samep.test()` the puddings are all the same strength.

# Pudding 2

Object `pudding2` includes a pudding-specific propensity to tie.  We
can test different nulls:

```{r,m2test,cache=TRUE}
samep.test(pudding2,puddings)
samep.test(pudding2,tyers)
```

# Pudding 3

Object `pudding3` includes a first-mover advantage:

```{r}
summary(pudding3)
pudding3_maxp
```

# Davidson's analysis

Davidson (1970) considers the following probability model:

\[
\operatorname{Prob}(i\succ j) = \frac{\pi_i}{\pi_i+\pi_j+\nu\sqrt{\pi_i\pi_j}}\\
\operatorname{Prob}(j\succ i) = \frac{\pi_j}{\pi_i+\pi_j+\nu\sqrt{\pi_i\pi_j}}\\
\operatorname{Prob}(i\sim  j) = \frac{\nu\sqrt{\pi_i\pi_j}}{\pi_i+\pi_j+\nu\sqrt{\pi_i\pi_j}}
\]

and gives the following estimate for the pudding data:

```{r}
davidson <- c(0.1388005,0.1729985, 0.1617420, 0.1653930, 0.1586805, 0.2023855, 0.7468147)
names(davidson) <- c(puddings,"tie")
davidson
par(pty='s')
plot(davidson[1:6],pudding_maxp[1:6]/(1-pudding_maxp[7]),asp=1,xlim=c(0.13,0.2),ylim=c(0.13,0.2),pch=16)
abline(0,1)
```

Although the probability model is different, we may compare the
probabilities of $i\succ j$, $j\succ i$, and $i\sim j$ for $i<j$ with
the two models.  This gives us a total of $3{6\choose 2}=45$
(nonindependent) probabilities and we may plot a scattergraph:

```{r,davidsonvshyper2,cache=TRUE}
ph <- pudding_maxp # ph == 'pudding hankin'; saves typing
pd <- davidson     # pd == 'pudding davidson'
print(ph)
print(pd)

`hankin` <- function(ph){
  ##  ph a vector of length n+1; first n entries are strengths, last one
  ##  propensity to tie
  n <- length(ph)-1
  m <- allbinom(n,2)
  out <- matrix(NA,nrow=ncol(m),ncol=3)
  colnames(out) <- c("i","j","tie")
  rownames(out) <- apply(m,2,paste,collapse="v")
  th <- ph[n+1]
  for(o in seq_len(ncol(m))){
      i <- m[1,o]
      j <- m[2,o]
      dh <- ph[i] + ph[j] + th    # dh == 'denominator hankin'
      out[o,1] <- ph[i]/dh
      out[o,2] <- ph[j]/dh
      out[o,3] <- th/dh
  }
  return(out)
}

`davidson` <- function(pd){

  ##  pd a vector of length n+1; first n entries are strengths, last one
  ##  nu
  n <- length(pd)-1  # how many puddings
  nu <- pd[n+1]  
  m <- allbinom(n,2)
  out <- matrix(NA,nrow=ncol(m),ncol=3)
  colnames(out) <- c("i","j","tie")
  rownames(out) <- apply(m,2,paste,collapse="v")
  nu <- pd[n+1]
  for(o in seq_len(ncol(m))){
      i <- m[1,o]
      j <- m[2,o]
      dd <- pd[i] + pd[j] + nu*sqrt(pd[i]*pd[j])  # dd == 'denominator davidson'
      out[o,1] <- pd[i]/dd
      out[o,2] <- pd[j]/dd
      out[o,3] <- nu*sqrt(pd[i]*pd[j])/dd
  }
  return(out)
}

plotter <- function(ph,pd,...){
  h <- hankin(ph)
  d <- davidson(pd)
  nh <- nrow(h)
  plot(c(h),c(d),pch=16,asp=1,col=c(rep("red",nh),rep("blue",nh),rep("green",nh)),...)
}
```

```{r}
print(ph)
print(hankin(ph))
print(pd)
print(davidson(pd))
jj <- c(2,5)/10
par(pty='s')
plotter(ph,pd,xlim=jj,ylim=jj)
grid()
```

Now calculate likelihoods:

```{r}
n <- pudding_table[,-(1:3)]
h <- hankin(ph)
d <- davidson(pd)
showsome(n)
showsome(h)
showsome(d)
sum(n*log(h))
sum(n*log(d))
(suppdiff <- sum(n*log(d))-sum(n*log(h)))
```

Above we see that Davidson's method has almost 3 units of support more
than RBT.  However, we note that Davidson's model has seven degrees of
freedom while RBT has only six (because of the the unit sum
constraint).  Here Wilks's theorem is not applicable because the two
models are not nested but even so the likelihood ratio is a meaningful
statistic.

```{r}
pchisq(suppdiff,df=1,lower.tail=FALSE)
```

# Appendix:  synthetic datasets

Above we see that, according to this model, there is no evidence to
suggest that the puddings' strengths, or their tying ability, differ.
But we can adduce a dataset for which the tying ability _does_ differ.
Consider this:

```{r, makem}
p <- pudding_table
p[1 ,4:6] <- c(10,10,37)  # 1v2
p[2 ,4:6] <- c(10,10,27)  # 1v3
p[4 ,4:6] <- c(10,10,24)  # 1v4
p[7 ,4:6] <- c(10,10,20)  # 1v5
p[11,4:6] <- c(10,10,31)  # 1v6
p
```

(above, we doctor the dataset so that any comparison involving pudding
1 has a high probability of tying).  Then:

```{r,tiepud,cache=TRUE}
Hp <- convert(p)[[2]]
mHp <- maxp(Hp)
mHpt <- mHp[7:12]  #just the tyers
```

visualise:

```{r}
mHpt
plot(mHpt)
```

(above we see that pudding 1 does indeed have a higher tie
probability).  Test the null of equal tying probabilities:

```{r,thptest,cache=TRUE}
(tHp <- samep.test(Hp,tyers))
```

Above analysis shows that we can reject the null that all puddings
have the same tying propensity.

## Test of Davidson vs Hankin

Although the likelihood for Davidson's model was higher than mine for
the original dataset, we can create a synthetic dataset which is a
random observation drawn from a reified Bradley-Terry probability
model.  I will then test the hypothesis that the data is in fact drawn
from an RBT distribution by calculating a likelihood ratio.

```{r,synthmaker}
set.seed(0)
f <- function(i){rmultinom(1,50,prob=h[i,])} # f(i) = sample from row i of h
synth <- pudding_table
synth[,-(1:3)] <- t(sapply(1:15,f))  # random sample
synth[,3] <- 50
rownames(synth) <- rownames(p)
colnames(synth) <- colnames(p)
synth
```

Above, `synth` is a dataset drawn from an RBT model with maximum
likelihood parameters obtained from the original dataset.  We now use
both methods.  First reified Bradley-Terry:

```{r,useboth,cache=TRUE}
(H_synth <- convert(synth)[[1]])
ph_synth <- maxp(H_synth)
```

```{r}
ph_synth
```

The above strengths are different from (but not too different from)
the value obtained from `maxp(H)` above.  Now analyse the same dataset
but using the method given in the `PlackettLuce` vignette:

```{r}
PlackettLuce_coef <- function(pudding_table){
i_wins <- data.frame(Winner = pudding_table$i, Loser = pudding_table$j)
j_wins <- data.frame(Winner = pudding_table$j, Loser = pudding_table$i)
  ties <- data.frame(Winner = asplit(pudding_table[c("i", "j")], 1),
                     Loser = rep(NA, 15))
R <- as.rankings(rbind(i_wins, j_wins, ties),
                 input = "orderings")
w <- unlist(pudding_table[c("w_ij", "w_ji", "t_ij")])
mod <- PlackettLuce(R, weights = w, npseudo = 0, maxit = 70)
return(coef(mod,log=FALSE))
}	
```

As a consistency check, try verifying the original result:

```{r,verifypl,cache=TRUE}
PlackettLuce_coef(pudding_table)
```

Above, we see values that match those in the `PlackettLuce` vignette.
Now apply the same method to the synthetic dataset `synth`:

```{r,calcpdsynth,cache=TRUE}
(pd_synth <- PlackettLuce_coef(synth))
```

Now we need to calculate the probabilities:

```{r}
showsome(h_synth <- hankin(ph_synth))
showsome(d_synth <- davidson(pd_synth))
```

And calculate a likelihood ratio for the two hypotheses, RBT and
Davidson:

```{r}
showsome(n <- synth[,-(1:3),])
(lh_synth <- sum(n*log(h_synth)))  # Hankin; RBT
(ld_synth <- sum(n*log(d_synth)))  # Davidson
lh_synth - ld_synth
```

So reified Bradley-Terry is better than Davidson, but (in this case at
least) not by much.  The difference is small because the puddings have
very similar strengths.  We can repeat the analysis but with puddings
of different strengths:

```{r}
dpudstrengths <- c(zipf(6)*0.9,tie=0.1)  # different pudding strengths
names(dpudstrengths) <- c(puddings,"tie")
dpudstrengths
```

Above we see that `pudding_1` is six times stronger than `pudding_6`
in accordance with Zipf's law.  Same analysis as before:

```{r}
showsome(h2 <- hankin(dpudstrengths))
f <- function(i){rmultinom(1,50,prob=h2[i,])} 
synth2 <- pudding_table
synth2[,-(1:3)] <- t(sapply(1:15,f))  # random sample
synth2[,3] <- 50
rownames(synth2) <- rownames(p)
colnames(synth2) <- colnames(p)
showsome(synth2)
```

OK so `synth2` is a different synthetic dataset but this time with
pudding strengths following Zipf.  As before, create a `hyper2`
object and find the evaluate:

```{r,synth2maxp,cache=TRUE}
H_synth2 <- convert(synth2)[[1]] # a hyper2 likelihood function
ph_synth2 <- maxp(H_synth2)
```

```{r}
ph_synth2
```

As a consistency check compare estimate with true value:

```{r}
ph_synth2-dpudstrengths  # difference is small
```

Now Davidson estimate using `synth2`:

```{r,synth2david,cache=TRUE}
(pd_synth2 <- PlackettLuce_coef(synth2))
```

```{r}
h_synth2 <- hankin(ph_synth2)
d_synth2 <- davidson(pd_synth2)
showsome(h_synth2)  # same as before
```

And calculate a support function:

```{r}
n <- synth2[,-(1:3),]  # as before, extract just the counts
(lh_synth2 <- sum(n*log(h_synth2)))
(ld_synth2 <- sum(n*log(d_synth2)))
lh_synth2 - ld_synth2
```

So this time the difference in support is larger, due to the puddings
having different strengths.

# References

* R. D. Davidson 1970. "On extending the Bradley-Terry model to accommodate ties in paired comparison experiments".  _Journal of the American Statistical Association_, Volume 65, Number 329, pp317--328

### Package dataset

Following lines create `pudding.rda`, residing in the `data/` directory of the package.

```{r,label=savecurlingdataset}
save(pudding_table,pudding,pudding_maxp,curling2_maxp,file="pudding.rda")
```