### 2021-11-06 inst/zermelo.Rmd

---
title: "Some thoughts on Zermelo's iteration process in the context of the hyper2 package"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document sets out a use-case for functions `zermelo()` and
`pairwise()`.  First I create `n=10` competitors with Zipf strengths:

```{r}
n <- 10
p <- zipf(n)
names(p) <- letters[seq_len(n)]
p
```

Now I create `M`, a matrix holding pairwise comparisons between the
competitors:

```{r}
M <- matrix(0,n,n)
rownames(M) <- names(p)
colnames(M) <- names(p)
index <- which(upper.tri(M),arr.ind=TRUE)
for(o in seq_len(nrow(index))){
  i <- index[o,1]
  j <- index[o,2]
  prob <- p[i]/(p[i] + p[j]) # Bradley-Terry
  nobs <- 80 
  i_wins <- rbinom(1,nobs,prob)
  j_wins <- nobs - i_wins
  M[i,j] <- i_wins
  M[j,i] <- j_wins
}
M
```

Above we see that `a` plays `b` a total of $48+32=80$ times, winning
$48$ and losing $32$.  The probability of `a` winning is
$\frac{p_1}{p_1+p_2}=\frac{1}{1+\frac{1}{2}}=\frac{2}{3}$ so we would
expect $80\cdot\frac{2}{3}=53.3$ victories.  We can estimate the
strengths of the competitors using either Zermelo's iteration scheme
or function `maxp()`.  First Zermelo:

```{r}
(pZ <- zermelo(M))
```

And now hyper2:

```{r,cache=TRUE}
(H <- pairwise(M))
(pH <- maxp(H))
```

We can compare the two using scattergraphs and likelihood:

```{r}
par(pty='s')
plot(pZ,pH,asp=1,log='xy')
abline(0,1)
```

```{r}
plot(log(pZ),log(abs(pH-pZ)/pZ))
abline(h=0)
```

```{r}
loglik(pZ,H)-loglik(pH,H)
```

Above we see that the two methods agree closely in both point and
likelihood.  Also, both evaluates are reasonably close to the true Zipf
strengths:

```{r}
par(pty='s')
plot(zipf(n),log(abs(zipf(n)-pZ)/zipf(n)),pch=16,col='black',log='x')
```

Now we can have a bit of fun with Zermelo:

```{r}
o <- zermelo(M,give=TRUE)
head(o)
l <- apply(o,1,loglik,H)
plot(log(abs(l-max(l))))
```