### 2022-04-16 inst/race3.Rmd
more work on order statistics in hyper3

---
title: "Races with hyper3 objects"
author: "Robin KS Hankin"
date: "10/04/2022"
output: html_document
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("disordR")
```

Consider a race in which there are six runners 1-6 but we happen to
know that three of the runners (1,2,3) are clones of strength $p_a$,
say, two of the runners (4,5) have strength $p_b$, and the final
runner (6) is of strength $p_c$.  The runners race and the finishing order is:

$$a\succ c\succ b\succ a\succ a \succ b$$

Thus the winner was $a$, second place for $b$, third for $a$, and so
on.  Alternatively we might say that $a$ came first, fourth, and
fifth; $b$ came third and sixth, and $c$ came second.  The
Plackett-Luce likelihood function for $p_a,p_b,p_c$ would be

$$
\frac{p_a}{3p_a+2p_b+p_c}\cdot
\frac{p_c}{2p_a+2p_b+p_c}\cdot
\frac{p_b}{2p_a+2p_b    }\cdot
\frac{p_a}{2p_a+ p_b    }\cdot
\frac{p_a}{ p_a+ p_b    }\cdot
\frac{p_b}{      p_b    },\qquad p_a+p_b+p_c=1
$$

The likelihood function is easily specified in package idiom:

```{r, label=definehyper3}
H <- ordervec2supp3(c("a","c","b","a","a","b"))
H
```

```{r, label=maximizeprobs, cache=TRUE}
mH  <-  maxp(H)
mH
```

We can test various hypotheses, for example that the competitors are of equal strength:

```{r testequality, cache=TRUE}
equalp.test(H)
```

showing that this small dataset does not allow one to say that the
competitors differ.  However, it is easy to generate synthetic order
statistics where one can reject the null of equality of strength:

```{r,biggerrace,cache=TRUE}
o <- rrace3(pn=c(a=20,b=11,c=7),ps=c(a=0.7,b=0.2,c=0.1))
o
H <- ordervec2supp3(o)
H
```

Above, we specify that there are 20 $a$'s, 11 $b$'s and 7 $c$'s with
strength $70\%$, $20\%$, and $10\%$ respectively.  Recall this means
that, in a head-to-head between $a$ and $b$ we would have $a$ winning
with probability $0.7/(0.7+0.2)=\frac{7}{9}$.

```{r,bigraceestimation,cache=TRUE}
mH <- maxp(H)
mH
pTRUE <- c(a=0.7,b=0.2,c=0.1)   # true values 
mH - pTRUE # residual
```

Above we see a reasonably close agreement between the true values and the estimate.

```{r,dothetest,cache=TRUE}
equalp.test(H)
```

Above we see strong evidence to reject the hypothesis that
$p_a=p_b=p_c=\frac{1}{3}$.  We may also test the hypothesis that the
strength vector is equal to its true value:

```{r,isittrue,cache=TRUE}
knownp.test(H,pTRUE)
```

And above we see that there is no evidence to reject the hypothesis
that the strengths are in fact the specified strengths in `pTRUE`.

# Some simple cases

The simplest nontrivial case is three competitors with strengths
$a,a,b$ and finishing ordr $a\succ b\succ a$.  The likelihood
function would be

$$
\frac{a}{2a+b}\cdot\frac{b}{a+b}
$$

In this case we know that $a+b=1$ so this is equal to
$\mathcal{L}=\mathcal{L}(a)=\frac{a(1-a)}{1+a}$.

$$
\frac{d\mathcal{L}}{da}=\frac{(1+a)(1-2a)-a(1-a)}{(1+a)^2}=
\frac{1-2a+a-2a^2 -a + a^2}{(1+a)^2}=
\frac{1-2a-a^2}{(1+a)^2}
$$

This will be zero at $\sqrt{2}\pm 1$; we also note that
$d^2\mathcal{L}/da^2=-4(1+a)^{-3}$, manifestly strictly negative for
$0\leq a\leq 1$: the root is a maximum.

```{r label=maxaba,cache=TRUE}
maxp(ordervec2supp3(c("a","b","a")))
```

above, we see close agreement with the theoretical value of
$(\sqrt{2}-1,1-\sqrt{2})\simeq (0.414,0.586)$

With four competitors we have five nontrivial order statistics which,
without loss of generality, are $(a,b,a,a)$ and $(a,a,b,a)$.
$(a,b,a,b)$, $(a,b,b,a)$, $(a,b,c,a)$. Taking $(a,b,a,b)$ as an
example we have

$$
\mathcal{L}(a)=
\frac{a}{2a+2b}\cdot
\frac{b}{ a+2b}\cdot
\frac{a}{ a+ b}\cdot
\frac{b}{    b}=\frac{a^2-a^3}{4-2a}
$$

the evaluate would be a root of $2a^2-7a+4$, or $(7-\sqrt{17})/4\simeq
0.72$:

```{r label=maxabab,cache=TRUE}
maxp(ordervec2supp3(c("a","b","a","b")))
```

If we allow non-finishers there is another nontrivial order statistic,
viz $\left(a,b,\left\lbrace a,b\right\rbrace\right)$:

# The constructors championship

Here we define  likelihood function for the constructors' championship.

```{r,dotheconst}
jj <- read.table("constructor_2021.txt",header=TRUE)
constructor_table_2021 <- jj[,-ncol(jj)]
constructor_table_2021[,1:10]
```

```{r,maketheconst,cache=TRUE}
constructor_2021 <- ordertable2supp3(constructor_table_2021)
```

Then

```{r,label=maxconstr,cache=TRUE}
mc <- maxp(constructor_2021)
mc
```

```{r}
pie(mc)
```