### 2022-04-21 inst/home_advantage.Rmd
more work

---
title: "Quantifying the home ground advantage"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{home_ground_advantage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

(takes about fifteen minutes to run without cache)

Here I analyse a small synthetic dataset for home ground advantage, then go on to discuss
a dataset presented by Agresti.

First load the package:
```{r,label=libload}
library(hyper2)
```

Now generate the dataset:

```{r,label=genthedata}
home_games_won <- matrix(c(
    NA, 16, 12, 11,
    19, NA, 19, 16,
    17, 12, NA, 11,
    11, 12, 12, NA),
    nrow=4,ncol=4,byrow=TRUE)

away_games_won <- matrix(c(
    NA, 05, 02, 02,
    02, NA, 10, 02,
    00, 04, NA, 07,
    02, 06, 04, NA),
    nrow=4,ncol=4,byrow=TRUE)

teams <- LETTERS[1:4]
dimnames(home_games_won) <- list("@home" = teams,"@away"=teams)
dimnames(away_games_won) <- list("@home" = teams,"@away"=teams)

home_games_won
away_games_won
```

Thus `home_games_won[1,2] == 16` means A played at home against B and
won 16 times; `home_games_won[2,1] == 19` means B played at home
against A and won 19 times.  Also, `away_games_won[1,2] == 5` means A
played away against B and won 5 times, and `away_games_won[2,1] == 2`
means B played away against A and won 2 times.  Alternatively, A
played B 16+19+5+2=45 times; A played at home 16+2=18 times and won 16
times and lost 2 times; and B played at home 19+5=24 times and won 19
times and lost 5 times.

We will create a hyper2 object with the teams and home ground advantage term:

```{r,label=makehyper2,cache=TRUE}
H <- home_away(home_games_won,away_games_won)
H
options("use_alabama" = TRUE) # to avoid the stupid wmmin bug
specificp.gt.test(H,"home",0)
```

We see strong evidence to support the contention that home advantage
is real.  Further, we may test the hypothesis that all the teams have
the same strength, after accounting for the home team advantage:

```{r,label=testhomegroundadvantageAgresti,cache=TRUE}
samep.test(H,teams)
```

Above, we see no evidence for a difference in team strengths.  Visually:

```{r,label=showvisH}
dotchart(maxp(H))
```

A profile likelihood diagram:

```{r,label=showprofilelikelihood, cache=TRUE}
home_strength <- seq(from=0.2,to=0.6,length=20)
plot(home_strength,profsupp(H,"home",home_strength),type="b",pch=16,
     xlab="home strength",ylab="support",main="profile likelihood for home advantage")
abline(h=c(0,-2))
```

The figure give a support interval between about 0.3 and 0.5 for home
ground advantage.

# Example in Agresti

Agresti considers a dataset of seven baseball teams who play one
another repeatedly.  Each game is played at the home ground of one
team, the other playing away.  The dataset is as follows:

```{r,label=makeAgrestimatrix}
baseball_table <- matrix(c(
NA,   4+3i, 4+2i, 4+3i, 6+1i, 4+2i, 6+0i,
3+3i, NA  , 4+2i, 4+3i, 6+0i, 6+1i, 4+3i,
2+5i, 4+3i, NA  , 2+4i, 4+3i, 4+2i, 6+0i,
3+3i, 5+1i, 2+5i, NA  , 4+3i, 4+2i, 6+1i,
5+1i, 2+5i, 3+3i, 4+2i, NA  , 5+2i, 6+0i,
2+5i, 3+3i, 3+4i, 4+3i, 4+2i, NA  , 2+4i,
2+5i, 1+5i, 1+6i, 2+4i, 1+6i, 3+4i, NA),7,7)
teams <- c("Miulwaukee", "Detroit", "Toronto", "New York", "Boston","Cleveland","Baltimore")
rownames(baseball_table) <- teams
colnames(baseball_table) <- teams
M <- baseball_table # saves typing
```

Above, we represent home team wins with the real part of the matrix,
the imaginary component being away wins.  Now process it:

```{r,label=makesuppfrommatrix}
baseball <- home_away(M)
baseball
```

Maximum likelihood:

```{r,label=calculatemaxpAgresti,cache=TRUE}
baseball_maxp <- maxp(baseball)
baseball_maxp
```

Visually:

```{r,label=showmaxpAgresti}
dotchart(baseball_maxp,pch=16)
pie(baseball_maxp)
```

Above, we see that the teams are all of roughly equal strength, and
the home advantage is small.  However, we may easily assess the null
that the home strength is zero:

```{r,analyseAgresti,cache=TRUE}
specificp.gt.test(baseball,"home",0)
```

```{r,Agrestiprofsupp,cache=TRUE}
home_strength <- seq(from=0.005,to=0.1,length=20)
plot(home_strength,profsupp(baseball,"home",home_strength),type="b",pch=16,
     xlab="home strength",ylab="support",main="profile likelihood for home advantage")
abline(h=c(0,-2))
```

Further, we may test the hypothesis that the baseball teams have the same strength:

```{r,label=testhomegroundadvantage,cache=TRUE}
samep.test(baseball,teams)
```

Above, we see no evidence for the teams having differential strengths.

### Some further thoughts on the home ground monster and Agresti's analysis.

Considering the simplest case of two teams, a and b, with results:

* $a$ at home, $b$ away, $a,b$ wins with probability $\frac{\lambda a,b}{\lambda a+b}$ (Agresti), or $\frac{a+H,b}{a+b+H}$ (me). 
* $a$ away, $b$ at home, $a,b$ wins with probability $\frac{a,\lambda b}{\lambda a+b}$ (Agresti), or $\frac{a,b+H}{a+b+H}$ (me).

Matching the probabilities for $a$ at home gives us $H=a(\lambda-1)$
and matching the probabilities for $a$ away gives $H=b(\lambda-1)$.
These cannot both be true unless $a=b$ or $\lambda=1$.  But we can, in principle,
test which one is a better model.

### Package dataset {-}

Following lines create `home_advantage.rda`, residing in the `data/`
directory of the package.

```{r,label=savesolingdataset}
save(baseball_table, baseball, baseball_maxp, 
  file="baseball.rda")
```