### 2022-04-24 inst/surfing.Rmd
more work on referee 2

---
title: "surfing"
output: bookdown::html_document2
bibliography: chess.bib
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

# Introduction

Surfing is a popular and growing activity in terms of both
participation and viewing [@warshaw2010].  Competitive surfing
involves surfers competing against one or two other surfers in heats
lasting 30-50 minutes [@booth1995].  Points are awarded on a 10-point
scale: each of five judges awards points to a surfer riding a
particular wave.  The World Surf League (WSL) is the main governing
body of surfing competitions [@wsl].  The WSL conducts a world tour in
which the best ranked 34 surfers compete; in addition, at each tour
venue, two "wildcard" surfers also enter the competition, who are
ignored here.  The 2019 WSL tour saw competitive events held at eleven
different surf locations.

The typical format is as follows.  Up to four surfers---the
competitors---are in the water simultaneously, watched by up to five
judges.  Subject to surfing etiquette, competitors are free to catch a
wave at their discretion within the heat's time window, typically
20-30 minutes.  The judges rate each wave ride and award points to the
surfer based on a range of subjective criteria such as stylish
execution and transition of manoeuvres, but also credit elements such
as the difficulty and novelty of the performance.  The top score and
bottom score among the five judges are removed and the remaining three
judges' scores are averaged to give the final score for the surfer for
that wave.  Each surfer's aggregate score is given by the average of
the two highest-scoring waves.  The winner is the surfer with the
highest aggregate score.

Such scoring systems are designed to account for the random nature of
wave quality while reflecting competitors' abilities fairly.
Differences in wave quality between successive days mean that direct
comparison of scores between one day and another are not informative
about competitors' abilities as they are strongly dependent on details
of wave quality at the time of the competition.  However, for a
particular heat, the order statistic---that is, which competitor
scored most highly, second highest, and third---is informative about
competitors' abilities: the wave environment is common to each surfer.

We note in passing that `wct8`, hosted at Lemoore had a different
format from the other venues in the championship, being held in an
artificial wave pool in which every wave was essentially identical (up
to a mirror reflection).  Statistically, there were seven heats of
either 2, 4, 8, or 16 surfers.  We show below how to incorporate the
information present in such observations with the remainder of the
championship using a consistent and intuitive statistical model.

# Previous related work

@farley2013 analyse surfing results using one-way ANOVA techniques and
report comparisons between the top 10 and bottom 10 surfers in the
2013 World Championship Tour and conclude that the top-ranked athletes
are more consistent than the lower-ranked group. They report that "To
date, only a limited number of research studies have reported on the
results of competitive surfing" [p44], and the analysis presented here
is hoped to change that.

Other statistical research into competitive surfing appears to focus
on the scoring of aerial manoeuvres compared with other manoeuvres in
events.  @lundgren2014, for example, report that aerial manoeuvres
scored higher than other stunts but had a lower completion rate, a
theme we return to below.

# Bradley-Terry

The Bradley-Terry model [@bradley1954] assigns non-negative strengths
$p_1,\ldots, p_n$ to each of $n$ competitors in such a way that the
probability of $i$ beating $j\neq i$ in pairwise competition is
$\frac{p_i}{p_i+p_j}$; it is conventional to normalize so that $\sum
p_i=1$.  Further, we use a generalization due to [@luce1959], in which
the probability of competitor~$i$ winning in a field of $\left\lbrace
1,\ldots, n\right\rbrace$ is $\frac{p_i}{p_1+\cdots +p_n}$.  Noting
that there is information in the whole of the finishing order, and not
just the first across the line, we can follow [@plackett1975] and
consider the runner-up to be the winner among the remaining
competitors, and so on down the finishing order. Without loss of
generality, if the order of finishing were $1,2,3,4,5$, then a
suitable Plackett-Luce likelihood function would be

\begin{equation}\label{competitors_1_to_5_likelihood}
\frac{p_1}{p_1+p_2+p_3+p_4+p_5}\cdot
\frac{p_2}{p_2+p_3+p_4+p_5}\cdot
\frac{p_3}{p_3+p_4+p_5}\cdot
\frac{p_4}{p_4+p_5}\cdot
\frac{p_5}{p_5}
\end{equation}

and this would be a forward ranking Plackett-Luce model in the
terminology of [@johnson2020].  We now use a technique due to Hankin
(2010, 2017) and introduce fictional (reified) entities whose nonzero
Bradley-Terry strength helps certain competitors or sets of
competitors under certain conditions.  The original example was the
home-ground advantage in football.  If players (teams) $1,2$ with
strengths $p_1,p_2$ compete, and if our observation were $a$ home wins
and $b$ away wins for team $1$, and $c$ home wins and $d$ away wins
for team~$2$, then a suitable likelihood function would be

\[
\left(\frac{p_1+p_H}{p_1+p_2+p_H}\right)^a
\left(\frac{p_1}{p_1+p_2+p_H}\right)^b
\left(\frac{p_2+p_H}{p_1+p_2+p_H}\right)^c
\left(\frac{p_2+p_H}{p_1+p_2+p_H}\right)^d,
\]

\noindent where $p_H$ is a quantification of the beneficial home
ground effect.  Similar techniques have been used to account for the
first-move advantage in chess, and here we apply it to assess the
Brazilian  preference for reef-breaking waves.

```{r}
surfing_table <- read.table("wct_table2019.txt",header=TRUE)
nationality <- function(surfers){surfing_table$nation[rownames(surfing_table) %in% surfers]}
```

```{r}
surfing_venuetypes <- read.table("venuetypes2019.txt",header=TRUE)
surfing_venuetypes
beachtype <- function(venue){surfing_venuetypes$beachtype[which(surfing_venuetypes$comp==venue)]}
beachtype("wct05")
beachtype("wct06")

options(width=120)
surfing_table
wildcards <- rownames(surfing_table)[-(1:36)]
wildcards
QS <- rownames(surfing_table)[c(23:33,35:36)]
QS
```

Some quick analysis of the wildcard surfers.  First, how did the wildcards perform?

```{r}
wilds <- as.matrix(surfing_table[-(1:36),2:12])
wilds[wilds == "INJ"] <- NA
wilds[wilds == "-"] <- NA
table(wilds)
```

Above we see that 10 wildcards placed "33" [meaning last place], 16
placed "17" [second to last] three placed "9" and two placed "5".
[the placings are treated as character strings which is why they are
not in numerical orer].  Next, how many venues did each wildcard
participate at?

```{r}
table(apply(wilds,1,function(x){sum(!is.na(x))}))
```

We see that 20 wildcards appeared at 1 venue, 4 at two venues, and one
(Willcox) at three venues.q

```{r}
o <- strsplit(readLines("wct2019.txt")," ")
jj <- unlist(lapply(o,paste,collapse=" "))
head(jj)
tail(jj)
```

The first line shows that, at one of the heats in `wct01`, `Colapinto`
came first, `Bailey` second, and `Wright` came third.  The last line
shows that, at `wct11`, `Ferreira` came first and `Medina` came
second.

Following idiom creates a table:

```{r}
table(unlist(lapply(strsplit(jj, " "),length)))
```

This shows that object `jj` has 312 lines of 2 surfers, 160 lines with
3 surfers, 2 lines with 4 surfers, 2 lines with 8 surfers, and one
line with 16 surfers.

We can convert the dataset into a Plackett-Luce likelihood
function but first have to remove the wildcards and also player
`Vries` whose maximum likelihood strength is zero.

```{r definesurfing, cache=TRUE}
remove <- c(wildcards, QS, "Vries")
wanted <- which(!(rownames(surfing_table) %in% remove)) # used at directlogplotofstrengths
surfing <- hyper2()
surfing_pairs <- hyper2()
for(v in o){
   r <- v[-1]  # first entry is venue
   numberretained <- sum(!(r %in% remove))
   if(numberretained > 1){
     r <- r[!(r %in% remove)]
     surfing <- surfing + race(r)
     if(numberretained == 2){
     r <- r[!(r %in% remove)]
     surfing_pairs <- surfing_pairs + race(r)
     }	      
    }
}	
summary(surfing)
summary(surfing_pairs)
```

```{r showsurfing_pairs}
summary(surfing_pairs)
```

Object `surfing` is a loglikelihood function on the strengths of the
competitors.  Our first step is to find the evaluate:

```{r maxpcalc,cache=TRUE}
surfing_maxp <- maxp(surfing)
surfing_maxp
sort(surfing_maxp)
```

Now compare `surfing` with `surfing_pairs`.  First calculate

```{r maxpcalcpairs,cache=TRUE}
surfing_pairs_maxp <- maxp(surfing_pairs)
```

then plot

```{r plotmaxppairs}
surfing_pairs_maxp
surfing_maxp
par(pty='s')
plot(surfing_pairs_maxp,surfing_maxp,asp=1,pch=16)
abline(0,1)
plot(surfing_pairs_maxp,surfing_maxp,asp=1,pch=16,log='xy',xlim=c(0.01,0.2),ylim=c(0.01,0.2))
abline(0,1)
plot(log(surfing_pairs_maxp/surfing_maxp),pch=16)
abline(0,0)
```

and the second step is to establish its consistency:

```{r consistencycheck,cache=TRUE}
consistency(surfing)
consistency(surfing_pairs)
```

```{r findhessiansurfing,cache=TRUE}
e_all <- eigen(hessian(surfing,indep(surfing_maxp)),T,T)$values
e_pairs <- eigen(hessian(surfing_pairs,indep(surfing_pairs_maxp)),T,T)$values
```

```{r showhessiansurfing}
e_all
e_pairs
f <- function(x){prod(sort(x)[-1])}
f(e_all)
f(e_pairs)
f(e_all)/f(e_pairs)
```

and we may display it using a dot chart:

```{r echo=FALSE}
`mydotchart` <- 
function (x, labels = NULL, groups = NULL, gdata = NULL, offset = 1/8, 
    ann = par("ann"), xaxt = par("xaxt"), frame.plot = TRUE, 
    log = "", cex = par("cex"), pt.cex = cex, pch = 21, gpch = 21, 
    bg = par("bg"), color = par("fg"), gcolor = par("fg"), lcolor = "gray", 
    xlim = range(x[is.finite(x)]), main = NULL, xlab = NULL, 
    ylab = NULL, ...) 
{
    opar <- par("mai", "mar", "mgp", "cex", "yaxs")
    on.exit(par(opar))
    par(cex = cex, yaxs = "i")
    if (!is.numeric(x)) 
        stop("'x' must be a numeric vector or matrix")
    n <- length(x)
    if (is.matrix(x)) {
        if (is.null(labels)) 
            labels <- rownames(x)
        if (is.null(labels)) 
            labels <- as.character(seq_len(nrow(x)))
        labels <- rep_len(labels, n)
        if (is.null(groups)) 
            groups <- col(x, as.factor = TRUE)
        glabels <- levels(groups)
    }
    else {
        if (is.null(labels)) 
            labels <- names(x)
        glabels <- if (!is.null(groups)) 
            levels(groups)
        if (!is.vector(x)) {
            warning("'x' is neither a vector nor a matrix: using as.numeric(x)")
            x <- as.numeric(x)
        }
    }
    linch <- if (!is.null(labels)) 
        max(strwidth(labels, "inch"), na.rm = TRUE)
    else 0
    if (is.null(glabels)) {
        ginch <- 0
        goffset <- 0
    }
    else {
        ginch <- max(strwidth(glabels, "inch"), na.rm = TRUE)
        goffset <- offset
    }
    nmai <- opar[["mai"]]
    if (ann) 
        nm.2 <- nmai[2L]
    if (!(is.null(labels) && is.null(glabels))) {
        yi <- if (is.null(ylab) || !ann) 
            0
        else offset
        nm.2 <- nmai[4L] + max(yi + linch + goffset, ginch) + 
            1/16
        if (nmai[2L] < nm.2) {
            nmai[2L] <- nm.2
            par(mai = nmai)
        }
    }
    if (is.null(groups)) {
        o <- seq_len(n)
        y <- o
        ylim <- c(0, n + 1)
    }
    else {
        o <- sort.list(as.numeric(groups), decreasing = TRUE)
        x <- x[o]
        groups <- groups[o]
        color <- rep_len(color, length(groups))[o]
        lcolor <- rep_len(lcolor, length(groups))[o]
        pch <- rep_len(pch, length(groups))[o]
        of.1 <- cumsum(c(0, diff(as.numeric(groups)) != 0))
        y <- seq_len(n) + 2 * of.1
        ylim <- range(0, y + 2)
    }
#    plot.window(xlim = xlim, ylim = ylim, log = log)
    lheight <- par("csi")
    if (!is.null(labels)) {
        loffset <- (linch + 0.1)/lheight
        mtext(labels[o], side = 2, line = loffset, at = y, adj = 0, 
            col = color, las = 2, cex = cex, ...)
    }
    abline(h = y, lty = "dotted", col = lcolor)
    points(x, y, pch = pch, col = color, bg = bg, cex = pt.cex/cex)
    if (!is.null(groups)) {
        gpos <- rev(cumsum(rev(tapply(groups, groups, length)) + 
            2) - 1)
        ginch <- max(strwidth(glabels, "inch"), na.rm = TRUE)
        goffset <- (max(linch + offset, ginch, na.rm = TRUE) + 
            1/16)/lheight
        mtext(glabels, side = 2, line = goffset, at = gpos, adj = 0, 
            col = gcolor, las = 2, cex = cex, ...)
        if (!is.null(gdata)) {
            abline(h = gpos, lty = "dotted")
            points(gdata, gpos, pch = gpch, col = gcolor, bg = bg, 
                cex = pt.cex/cex, ...)
        }
    }
    axis(1, xaxt = xaxt)
    if (frame.plot) 
        box()
    if (ann) {
        title(main = main, xlab = xlab, ...)
        mgp <- par("mgp")
        par(mgp = c(max(mgp[1], nm.2/lheight - 1.5), mgp[-1]))
        title(ylab = ylab, ...)
    }
    invisible()
}
```

```{r showsurfingmaxppie, echo=FALSE, fig.height=8,fig.cap="dot chart of result"}
dotchart(surfing_maxp)
dotchart(log10(surfing_maxp))
pie(surfing_maxp)
pdf(file="showsurfingmaxppie.pdf")
pie(surfing_maxp)
dev.off()
jj <- log(surfing_maxp)
kk <- log(surfing_pairs_maxp)
names(jj) <- NULL
names(kk) <- NULL
l <- c(-4.6, -1.5)
dotchart(log(surfing_maxp),pch=NA,xlim=l)
mydotchart(jj,col="red" ,pch=16,xlim=l,xaxt='n',frame.plot=FALSE)
mydotchart(kk,col="blue",pch=16,xlim=l,xaxt='n',frame.plot=FALSE)
```

Test the null of equal strengths:

```{r pval,cache=TRUE}
equalp.test(surfing)
```

```{r directlogplotofstrengths,cache=TRUE}
jjO <- surfing_table$points[wanted]
names(jjO) <- rownames(surfing_table)[wanted]
jjL <- surfing_maxp
jjO <- ordertrans(jjO)
jjL <- ordertrans(jjL)
plot(jjO,jjL,pch=16,log='xy',xlab='points',ylab='maximum likelihood strength')
fit <- lm(log10(jjL)~log10(jjO))
summary(fit)
for(i in seq_along(jjO)){
text(jjO[i],jjL[i],names(jjL)[i],col='gray',pos=4,cex=0.5)
}
abline(fit)

pdf(file="directlogplotofstrengths.pdf")
plot(jjO,jjL,pch=16,log='xy',xlab='points',ylab='maximum likelihood strength')
fit <- lm(log10(jjL)~log10(jjO))
for(i in seq_along(jjO)){
text(jjO[i],jjL[i],names(jjL)[i],col='gray',pos=4,cex=0.5)
}
abline(fit)
dev.off()
```

Now compare the finishing order based on points vs the finishing order
on likelhood:

```{r ordertransplotplot,fig.cap="Official ranks vs likelihood rank, WCT"}
rL <- sort(surfing_maxp,decreasing=TRUE)
rL[] <- seq_along(rL)    # likelihood-based rank
rO <- seq_len(nrow(surfing_table[wanted,]))
names(rO) <- rownames(surfing_table)[wanted]
rO
rL
sort(names(rO))
sort(names(rO))
sort(names(rO))==sort(names(rO))
ordertransplot(rO,rL,xlab="offical rank",ylab="likelihood rank",main="WCT 2019")
pdf(file="ordertransplotplot.pdf")
ordertransplot(rO,rL,xlab="offical rank",ylab="likelihood rank",main="WCT 2019")
dev.off()
```

```{r dosameptest, cache=TRUE}
samep.test(surfing, c("Ferreira","Medina"))
samep.test(surfing, c("Ferreira","Florence"))
samep.test(surfing, c("Florence","Medina"))
samep.test(surfing, c("Florence","Medina","Ferreira"))
samep.test(surfing, c("Medina","Wilson"))
```

# Medina's performance at the Surf Ranch

We will test the null that Medina's performance at the Surf Ranch wa
the same as his performance in the remainder of the tourament.

```{r label=medinaattheranchlike, cache=TRUE}
surfing_wct08     <- hyper2()
surfing_elsewhere <- hyper2()

for(v in o){
   venue <- v[1]
   r <- v[-1]  # first entry is venue

   r <- r[!(r %in% remove)]
   if(length(r)>1){
       if(venue == "wct08"){
           rmod <- r
           rmod[rmod == "Medina"] <- "Medina_wct08"
           surfing_wct08 <- surfing_wct08 + race(rmod)
       } else {
           surfing_elsewhere <- surfing_elsewhere + race(r)
       } 
   }
} # v loop closes	
both_together <- surfing_wct08 + surfing_elsewhere
```

Now test the null that Medina's performance is identical at wct08 and
elsewhere:

```{r, label=testmedinasranchperformance,cache=TRUE}
samep.test(both_together,c("Medina","Medina_wct08"))
```

# Brazilian null

One suggestion is that Brazilians do worse on reef break and better on
point and beach break beaches.  We now assess the null that Brazilians
perform equally well on all types of beaches.  To that end, we create
a reified surfer, "Brazilian_beachtype", whose strength helps
Brazilian surfers when surfing on their (putative) preferred
beachtype.

```{r surfing2, cache=TRUE}
surfing2 <- hyper2()
for(v in o){
   venue <- v[1]  # first entry is venue
   r <- v[-1] 
   numberretained <- sum(!(r %in% remove))
   if(numberretained > 1){
     r <- r[!(r %in% remove)]
     jj <- race(r)
     surfers <- pnames(jj)
     if(sum(nationality(surfers)=="BRA")==1){	
       if((beachtype(venue)=="point") | (beachtype(venue) == "beach")){
         jj %<>% pwa(surfers[nationality(surfers)=="BRA"],"Brazilian_beachtype")
       }
     }
     surfing2 %<>% sum(jj)
   }
}	
summary(surfing2)
```

```{r surfing2max, cache=TRUE}
(surfing2_maxp <- maxp(surfing2))
specificp.gt.test(surfing2,"Brazilian_beachtype",0)
```

```{r profsuppcalcsurfing2, cache=TRUE}
p <- seq(from=0.001,to=0.01,by=0.0005)
l <- profsupp(surfing2,"Brazilian_beachtype",p,relative=FALSE)
l0 <- loglik(indep(surfing_maxp),surfing)
```

```{r plotpthink}
plot(c(0,p),c(l0,l)-max(l),type='b',xlab="Brazilian beachtype",ylab="support")
```

```{r profsuppcalcsurfing2wide, cache=TRUE}
p_wide <- seq(from=0.002,to=0.04,by=0.002)
l_wide <- profsupp(surfing2,"Brazilian_beachtype",p_wide,relative=FALSE)
```

```{r plotwidep}
plot(c(0,p_wide),c(l0,l_wide)-max(l),type='b',xlab="Brazilian beachtype wide",ylab="support")
abline(h= -2,lty=2)
pdf(file="plotwidep.pdf")
plot(c(0,p_wide),c(l0,l_wide)-max(l),type='b',xlab="Brazilian beachtype wide",ylab="support")
abline(h= -2,lty=2)
dev.off()
```

## Generalized log-odds ratios

```{r allbrazsurf}
all_Brazilian_surfers <- rownames(surfing_table)[surfing_table$nation == "BRA"]
all_Brazilian_surfers
pnames(surfing2)
pnames(surfing2) %in% all_Brazilian_surfers
table(pnames(surfing2) %in% all_Brazilian_surfers)
Brazilian_surfers <- pnames(surfing2)[pnames(surfing2) %in% all_Brazilian_surfers]
Brazilian_surfers
```

```{r biglikefun,cache=TRUE}
f <- function(B,...){
  jj <- specificp.test(surfing2, "Brazilian_beachtype", B)
  mp <- jj$null_estimate
  n <- sum(names(mp) %in% all_Brazilian_surfers)
  GLO <- log(mp["Brazilian_beachtype"]^n/prod(mp[which(names(mp) %in% all_Brazilian_surfers)]))
  out <- c(B,GLO, support=jj$null_support)
  names(out) <- c("B","GLO","support")
  return(out)
}
```

```{r calcbrazodr,cache=TRUE}
braz <- c(0.0003,0.0005,0.0008, 0.001,0.0015,0.002,0.003,0.004,0.005,0.007,0.01,0.015,0.02,0.03,0.05)
M <- do.call("rbind",lapply(braz,f))
M[,3] <- M[,3]-max(M[,3])
M
```

```{r plotBraz}
plot(M[,2],M[,3]-max(M[,3]),type='b',xlab="Brazilian beachtype (log-odds)",ylab="support")
plot(M[,2],M[,3]-max(M[,3]),type='b',ylim=c(-0.3,0),xlab="Brazilian beachtype (log-odds)",ylab="support")
```

# compatriot selection

```{r calculate_compatriot_monster_LF,cache=TRUE}
scomp <- hyper2()
scomp3 <- hyper2()  # scomp3 is the likelihood function for just the races with 3 competitors, 2 of one nationality and one odd-man-out.
for(v in o){
   r <- v[-1]  # first entry is venue; "r" for "race"
   numberretained <- sum(!(r %in% remove))
   if(numberretained > 1){
     r <- r[!(r %in% remove)]
     if(numberretained ==3){
       nationalities <- c(nationality(r[1]),nationality(r[2]),nationality(r[3]))
       if(length(unique(nationalities)) == 2){
         if(sum(nationalities %in% nationality(r[1])) == 1){ # winner is the odd one out
           Hjj <- pwa(race(r),pwa=r[1])
           print("odd one out is number 1")
           print(100*surfing_maxp[r])
         } else if(sum(nationalities %in% nationality(r[2]) ) == 1){ # second best is the odd one out
           Hjj <- pwa(race(r),pwa=r[2])
           print("odd one out is number 2")
           print(100*surfing_maxp[r])
         } else if(sum(nationalities %in% nationality(r[3])) == 1){  # third best is the odd one out
	   Hjj <- pwa(race(r),pwa=r[3])
           print("odd one out is number 3")
           print(100*surfing_maxp[r])
         } else {
           stop("this cannot happen")
         }
         print(Hjj)
         print("---")
         scomp <- scomp + Hjj      # add compatriot-monster modified likelihood 
         scomp3 <- scomp3 + Hjj
       } else {                    # 3 in race, but no compatriot monster...
         scomp <- scomp + race(r)  # ... so use regular Plackett-Luce for 3-competitor race
       }
     } else {                    # if(numberretained==3) closes...
       scomp <- scomp + race(r)  # numberretained !=3; use regular Plackett-Luce
     }
  } # if(numberretained>1) closes
} # for(v in o) loop closes
```

```{r maximize_compatriot_monster,cache=TRUE}
maxs <- maxp(scomp)
maxs
```

```{r plotpiemaxs}
pie(maxs)
```

```{r testcompatriotmonster,cache=TRUE}
specificp.gt.test(scomp,"S",0,n=100,maxtry=100)
```

Now profile likelihood:

```{r profsuppcalcscompatriotmonster, cache=TRUE}
p <- c(0.0002,seq(from=0.001,to=0.02,len=10))
l <- profsupp(scomp,"S",p,relative=FALSE)
```

plot it:

```{r plotprofilescomp}  
plot(p,l,type='b',xlab="compatriot support",ylab="support",main="profile support for S, scomp")
```

```{r scomp3maxp, cache=TRUE}
scomp3
maxp(scomp3)
specificp.gt.test(scomp3,'S',0,n=100)
````

Now profile likelihood:

```{r profsuppcalcscomp3atriotmonster, cache=TRUE}
p <- c(0.0002,seq(from=0.001,to=0.02,len=10))
l <- profsupp(scomp3,"S",p,relative=FALSE,n=100)
```

plot it:

```{r plotprofilescomp3}  
plot(p,l,type='b',xlab="compatriot support",ylab="support",main="profile support for S, scomp3")
```

We can do a much more basic test on just the entries with three competitors comprising
two of one nationality and one of a different nationality:

```{r basicanalysis}
##     Dora Ferreira   Slater 
##   2.5741  12.5929   3.8933  (3)

##     Smith    Buchan Freestone 
##    5.2764    2.0552    1.8673  (1)

##   Slater   Andino Callinan 
##   3.8933   5.9190   1.8408  (3)

## Crisanto   Wilson   Ibelli 
##   2.1644   2.7100   2.2479  (2)

##     Moniz    Wright Freestone 
##    3.0250    3.6502    1.8673 (1)

## Carmichael  Freestone     Flores 
##     1.8855     1.8673     2.0480 (3)

##    Bourez    Wright Freestone 
##    1.9068    3.6502    1.8673 (1)

## Colapinto    Flores    Bourez 
##    2.9871    2.0480    1.9068  (1)

## Ferreira   Morais   Ibelli 
##  12.5929   1.0817   2.2479 (2)

## Freestone    Flores    Bourez 
##    1.8673    2.0480    1.9068  (1)

##     Dora Ferreira   Morais 
##   2.5741  12.5929   1.0817 (1)

M <- matrix(c(
  2.5741,12.5929,3.8933,3
, 5.2764, 2.0552,1.8673,1
, 3.8933, 5.9190,1.8408,3
, 2.1644, 2.7100,2.2479,2
, 3.0250, 3.6502,1.8673,1
, 1.8855, 1.8673,2.0480,3
, 1.9068, 3.6502,1.8673,1
, 2.9871, 2.0480,1.9068,1
,12.5929, 1.0817,2.2479,2
, 1.8673, 2.0480,1.9068,1
, 2.5741,12.5929,1.0817,1),byrow=TRUE, ncol=4)

for(i in seq_len(nrow(M))){ M[i,1:3] <-   M[i,1:3]/sum(M[i,1:3])}
print(M)  # rows 1-3 of M are strengths of the first, second, third
          # competitors.  Row 4 is the place of the odd-man-out.  Thus
          # on row 1, the odd-man-out came third, and on row 2, the
          # odd-man-out came first.

p <- M[cbind(1:11, M[,4])]
print(p) 
                                        ## element i of p is the probability of the odd-man-out
                                        ## winning, given the maximum likelihood estimate.

ww <-table(replicate(1e5,sum(rbinom(11,1,p))))
print(ww)
sum(ww[7:10])/sum(ww) # pvalue for observation of six wins by the odd-man-out
```

### Package dataset {-}

Following lines create `surfing.rda`, residing in the `data/` directory
of the package.

```{r,label=savesurfiningingdataset}
save(surfing,surfing_maxp,surfing_venuetypes,surfing_table,file="surfing.rda")
```

# references