### 2022-05-11 inst/carcinoma.Rmd
further work on issue #131

---
title: "Carcinoma"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

```{r loadlib}
carcinoma_table <- read.table("carcinoma.txt",header=TRUE)
carcinoma_table[,1:7] <- carcinoma_table[,1:7] > 0
wanted <- !(apply(carcinoma_table[,1:7],1,sum) %in% c(0,7))
carcinoma_table <- carcinoma_table[wanted,]
carcinoma_table
```

```{r makeggrllist,cache=TRUE}
W <- hyper2(pnames=LETTERS[1:7])
Wi <- list()

for(i in seq_len(nrow(carcinoma_table))){
      jj <- unlist(carcinoma_table[i,1:7])
      negative <- names(jj)[jj==0]
      positive <- names(jj)[jj==1]
      print(negative)
      print(positive)
      Wi[[i]] <- ggrl(W,positive,negative)
}
HL <- lsl(Wi,powers= carcinoma_table$n)
```

Now

```{r,label=calccloglikzipf,cache=TRUE}
z <- zipf(7)
names(z) <- LETTERS[1:7]
system.time(lz <- loglik_lsl(z,HL))
lz
```

```{r,label-calcloglikequalp,cache=TRUE}
e <- equalp(W)
system.time(le <- loglik_lsl(e,HL))
le
```

Agresti, table 13.1, p542.

Landis, J. R., and G. G. Koch. 1977. An application of hierarchical
kappa-type statistics in the assessment of majority agreement among
multiple observers. Biometrics 33: 363-374.

### Package dataset {-}

Following lines create `carcinoma.rda`, residing in the `data/`
directory of the package.

```{r}
carcinoma <- 5

```{r,label=savecarcinomadataset}
save(carcinoma_table,carcinoma, file="carcinoma.rda")
```