### 2022-05-12 inst/rock_paper_scissors_monster.Rmd
fixes issue #133

---
title: "Quantifying the rock-paper-scissors effect"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
abbreviated <- TRUE  # change to FALSE for full names
```

## Abstract

Here I use the `hyper2` package (Hankin 2017) to quantify the strength
of a circular rock-paper-scissors effect using the `hyper2` package,
improving on previous results (West and Hankin 2008) which merely
established its existence.  I use two datasets, a three-player
scoreline and a synthetic dataset referring to wear on a seven-tooth
gear wheel.

## Rock Paper Scissors

Consider the following dataset:

```{r,chesstable}
chess_table <- as.matrix(read.table("chess.txt"))
dimnames(chess_table) <- list(
    match=c("T_vs_A","A_vs_K","T_vs_K"),
    player=c("Topalov","Anand","Karpov"))
chess_table
```

This is documented in `chess.Rd` and shows that Topalov played Anand
$22+13=35$ times, winning 22 and losing 13 matches.  The corresponding
likelihood function is `chess`:

```{r,showchess}
chess <- saffy(chess_table)
chess
chess_maxp <- maxp(chess)
chess_maxp
```

(we refer to this as the "TAK" model).  We might ask whether there is
evidence for the players having differing strengths:

```{r equalptestchunk,cache=TRUE}
equalp.test(chess)
```

Thus we fail to reject $H_0\colon p_T=p_A=p_K$ under T-A-K as the
p-value is about 0.39.

## RPS in chess

The existence of a rock-paper-scissors relationship between the
players was discussed in Hankin (2010) and the null of no RPS
phenomena could be rejected using the Alymer test.

Here we improve on that work by introducing a monster, `M`, that
accounts for the rock-paper-scissors nature of the players.  The RPS
monster helps players that benefit from the RPS nature of the result.
The simplest nontrivial probability model is that all three real
players have the same strength: $p_T=p_A=p_K$, which we write $T=A=K$.
Under $T=A=K$, writing $p_T=p_A=p_K=\alpha$ we can use a binomial test
to calculate the p-value for the hypothesis that $p_M=0$:

```{r,binomtest}
binom.test(sum(diag(chess_table)),sum(chess_table,na.rm=TRUE),alternative="greater")
```

Estimating $p_M$ is also straightforward:

$$
Prob(\mathrm{procyclic}\ \mathrm{win})=\frac{\alpha+M}{2\alpha+M}
$$

(a "procyclic win" is a win that would be expected under a strong
monster).  Here we also have $3\alpha+M=1$, so to estimate the
strength of the monster we would solve

$$
\frac{\alpha+M}{2\alpha+M}=p,\qquad{3\alpha + M}=1
$$

for $\alpha$ and $M$, where $p$ is the probability of a procyclic win.
This gives us

$$
\alpha=\frac{1-p}{2-p},\qquad M=\frac{2p-1}{2-p}.
$$

If we use the maximum likelihood estimate value for a procyclic win of
$\frac{55}{88}=0.625$ we find that $\hat{M}=0.18$ approximately.
Further, we may express $p$, the probability of a pro-cyclic win, as
$p=\frac{1+2M}{2+M}$.  If we observe $a$ procyclic wins and $b$
countercyclic wins a likelihood function for $M$ would then be

$$
\mathcal{L}(M)= p^a(1-p)^b=\left(\frac{1+2M}{2+M}\right)^a\cdot
\left(\frac{1-M}{2+M}\right)^b
$$

and here we have $a=22+23+10=55$ and $b=13+12+8=33$ (a graph is given
below).  It is straightforward but not trivial to translate this into
`hyper3` idiom.  Noting that $3\alpha+M=1$ we define $\beta=3\alpha$
and then the unit sum constraint is $\beta+M=1$ and our probabilities
are

$$
\frac{\alpha+M;\alpha}{2\alpha+M   } =
\frac{\beta/3+M;\beta/3}{2\beta/3+M} =
\frac{1+2M;1-M}{2+M}
$$

(where the last expression is obtained using $\beta+M=1$).  We may set
up a log-likelihood function of $a,b$ using `hyper3` idiom as follows:

```{r,usehyper3}
TqAqK <- function(a,b){
  H3 <- hyper3(pnames=c("beta","M"))
  H3[c(beta=1/3,M=1)] %<>% inc(a  )
  H3[c(beta=1/3    )] %<>% inc(  b)
  H3[c(beta=2/3,M=1)] %<>% dec(a+b)
  return(H3)
}
H3 <- TqAqK(55,33)
loglik(c(M=0.1,beta=0.9),H3)
loglik(c(beta=0.9,M=0.1),H3)  #  NB: identical!
```

[above, we note that $\beta+M=3\alpha+M=1$, so the monster is the
fillup value with respect to $\beta$].  This is easy to maximize:

```{r,maximizeTqAqK}
optimize(function(M){loglik(c(M=M,beta=1-M),H3)},c(0.01,0.99),maximum=TRUE)
```

This gives maximum likelihood estimates of $M=0.18$ as would be
expected from the simple binomia model; the probability of a procyclic
win is about $0.625$.

We may relax the assumption of equal strength of the real
players by using a more sophisticated hyper2 construction:

```{r,setupH}
H <- hyper2()
H[c("Topalov",                 "M")] %<>% inc(22   )
H[c(          "Anand"             )] %<>% inc(   13)
H[c("Topalov","Anand",         "M")] %<>% dec(22+13)

H[c(          "Anand",         "M")] %<>% inc(   23)
H[c(                  "Karpov"    )] %<>% inc(   12)
H[c(          "Anand","Karpov","M")] %<>% dec(23+12)

H[c("Topalov"                     )] %<>% inc( 8   )
H[c(                  "Karpov","M")] %<>% inc(   10)
H[c("Topalov",        "Karpov","M")] %<>% dec( 8+10)
H
```

(in the above, a win by a player helped by the RPS effect has `M`
added to his strength).  We can call this the TAKM model, and find the
evaluate:

```{r}
maxp(H)
```

We see that, under TAKM, the monster has strength of about 0.16,
slightly less than the 0.18 obtained for $M$ under the equality
assumption $p_T=p_A=p_K$.  Further, under TAKM, the three "real"
players are again all of approximately equal strength, and we may test
this:

```{r,testequalTAKM,cache=TRUE}
samep.test(H,c("Topalov","Anand","Karpov"))
```

(note the estimated monster strengths of 0.16 and 0.18 agree with the
previous estimate).  Thus the p-value of about 0.77 indicates failure
to reject the null (compare 0.39 under TAK).  Now we can test
$H_0\colon p_M=0$ under TAKM:

```{r,specmtest,cache=TRUE}
specificp.gt.test(H,"M",0)
```

(in the test above, note that the players may appear in a different
order from previously).  The test shows that we may reject the
hypothesis that $p_M=0$ under TAKM with a p-value of $0.04$ (compare
0.012 from the binomial test).

### Profile likelihood

We may plot a profile likelihood curve for $M$:

```{r,label=plotproflik,cache=TRUE}
M <- seq(from=0.001,to=0.4,len=40)
L1 <- profsupp(H,"M",M)
L1 <- L1-max(L1)

L2 <- 55*log(1+2*M)+33*log(1-M)-88*log(2+M)
L2 <- L2-max(L2)

plot(M,L1,type="l",col="black",ylab="log likelihood",xlab="Monster strength")
points(M,L2,type="l",col="red")
abline(h=c(0,-2),lty=2)
legend("topright",col=c("black","red"),lty=1,legend=c("TAKM","T=A=K"))
```

Above we see support curves for RPS monster strength using TAKM and
$T=A=K$, showing very similar results.  Note that the T=A=K curve
could have been obtained more nicely with function `TqAqK` but I
haven't got round to changing it yet.

# Generalized RPS

West and Hankin (2008) consider a synthetic dataset referring to seven
teeth of a gear wheel.  Consecutive pairs are repeatedly compared and
the most worn tooth is nominated by independent judges.  Here a
dataset of the same form will be used, but the entries have been
changed for pedagogical reasons:

```{r,echo=FALSE}
t1 <- "t1"
t2 <- "t2"
t3 <- "t3"
t4 <- "t4"
t5 <- "t5"
t6 <- "t6"
t7 <- "t7"
M <- "M"
```

```{r,label=setuphw}
hw <- matrix(c(
    16,05,NA,NA,NA,NA,NA,
    NA,06,09,NA,NA,NA,NA,
    NA,NA,05,11,NA,NA,NA,
    NA,NA,NA,06,12,NA,NA,
    NA,NA,NA,NA,05,09,NA,
    NA,NA,NA,NA,NA,04,07,
    13,NA,NA,NA,NA,NA,05),
    7,7,byrow=TRUE)
colnames(hw) <- paste("t",1:7,sep="")
hw
```

Thus, considering the top row, we see that in a total of $16+5=21$
independent comparisons between `t1` and `t2`, tooth `t1` was judged
to be the more worn one 16 times, and `t2` was judged to be the more
worn one $5$ times (NB: in the above, I have changed the dataset in
order to make `t1` stronger than the other players while maintaining
the RPS phenomenon).  We may convert this dataset to a generalized
`hyper2` object.  First a standard Bradley-Terry:

```{r,label=HWnomonster}
HW <- hyper2()
rs <- -rowSums(hw,na.rm=TRUE)
cs <- colSums(hw,na.rm=TRUE)
HW[t1] <- cs[1]
HW[t2] <- cs[2]
HW[t3] <- cs[3]
HW[t4] <- cs[4]
HW[t5] <- cs[5]
HW[t6] <- cs[6]
HW[t7] <- cs[7]
HW[c(t1,t2)] <- rs[1]
HW[c(t2,t3)] <- rs[2]
HW[c(t3,t4)] <- rs[3]
HW[c(t4,t5)] <- rs[4]
HW[c(t5,t6)] <- rs[5]
HW[c(t6,t7)] <- rs[6]
HW[c(t7,t1)] <- rs[7]
HW
```

(likelihood function `HW` has no monster).  We can then analyse this:

```{r,label=testHW,cache=TRUE}
maxp(HW)
equalp.test(HW)
```

Thus according to this model we may reject the hypothesis of equal
strength.  We now incorporate the monster:

```{r,label=NWwithmonster}
HWM <- hyper2()
tt <- c("t1","t2","t3","t4","t5","t6","t7")
for(i in 1:6){
  HWM[c(tt[i]          )] <-  hw[i,i  ]
  HWM[c(      tt[i+1],M)] <-  hw[i,i+1]
  HWM[c(tt[i],tt[i+1],M)] <- rs[i]
}
HWM[c(tt[7]        )] <-  hw[7,7]
HWM[c(      tt[1],M)] <-  hw[7,1]
HWM[c(tt[7],tt[1],M)] <- rs[7]
HWM
```

Then the same analysis applies:

```{r,label=maxpHWM,cache=TRUE}
maxp(HWM)
```

and we can test the hypothesis that $p_M=0$:

```{r,label=specicifgtest,cache=TRUE}
specificp.gt.test(HWM,"M",0)
```

suggesting that we may reject the hypothesis of zero monster strength
with a p-value of about 4.7\%, even after accounting for the different
strengths of the teeth.

## References

* L. J. West and R. K. S. Hankin 2008.  "Exact tests for two-way
  contingency tables with structural zeros".  _Journal of Statistical
  Software_ 2008, volume 28 issue 11.
* R. K. S. Hankin 2010.  "A generalization of the Dirichlet
  distribution".  _Journal of Statistical Software_ 2010, volume 33
  issue 11.
* R. K. S. Hankin 2017.  "Partial rank data with the `hyper2` package:
  likelihood functions for generalized Bradley-Terry models".  _The R
  Journal_, volume 9, number 2.

### Package dataset

Following lines create `chess.rda`, residing in the `data/` directory of the package.

```{r,label=savechessdataset}
save(chess_table,chess,chess_maxp,file="chess.rda")
```