### 2022-05-16 inst/hepatitis.Rmd
fixes #130

---
title: "Hepatitis"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

```{r loadlib}
hepatitis_table <- read.table("hepatitis.txt",header=TRUE)
hepatitis_table[,1:3] <- hepatitis_table[,1:3] > 0
wanted <- !(apply(hepatitis_table[,1:3],1,sum) %in% c(0,3))
hepatitis_table <- hepatitis_table[wanted,]
hepatitis_table
```

```{r makeggrllist,cache=TRUE}
W <- hyper2(pnames=c("P","Q","E"))
Wi <- list()
for(i in seq_len(nrow(hepatitis_table))){
      jj <- unlist(hepatitis_table[i,1:3])
      negative <- names(jj)[jj==0]
      positive <- names(jj)[jj==1]
      Wi[[i]] <- ggrl(W,positive,negative)
}
hepatitis <- lsl(Wi,powers= hepatitis_table$OC)
```

```{r}
hepatitis
```

There are several natural BT strengths to consider.  First, we could
simply count how many times each clinician gives a positive diagnosis:

```{r,label=counthep,cache=TRUE}
hepatitis_count <- colSums(sweep(hepatitis_table[,1:3],1,hepatitis_table$OC,"*"))
names(hepatitis_count) <- colnames(hepatitis_table)[1:3]
hepatitis_count <- hepatitis_count/sum(hepatitis_count)
hepatitis_count
```

Secondly, Zipf:

```{r,label=calccloglikzipf,cache=TRUE}
z <- zipf(3)
names(z) <- c("P","Q","E")
z
```

Thirdly, equal probabilities:

```{r,label=calcloglikequalp,cache=TRUE}
e <- equalp(W)
e
```

To find the maximum likelihood estimate we would use:

```{r,label=calculatehepmaxl,cache=TRUE}
hepatitis_maxp <- maxp_lsl(hepatitis,startp=indep(hepatitis_count), control=list(trace=110))
```

```{r,label=calculateloglikformaxp, cache=TRUE}
hepatitis_maxp
```

With this, we are now in a position to compare these four points:

```{r,label=showthreemethods}
lz <- loglik_lsl(z,hepatitis) 
le <- loglik_lsl(e,hepatitis)
lm <- loglik_lsl(hepatitis_maxp,hepatitis)
lc <- loglik_lsl(hepatitis_count,hepatitis)
(results <- c(zipf=lz, equal=le, maxlike=lm, count=lc))
results <- results - max(results)
results
```

We can use Wilks here to assess the null of `hepatitis_count`.  We
have $-2\log(\Lambda/\Lambda_0)\sim\chi^2_6$, giving us a $p$-value of

```{r,label=calculatechisquaredpvalueforcount}
pchisq(-2*results[4],df=2,lower.tail=FALSE)
```

Agresti, table 13.1, p542.

### Package dataset {-}

Following lines create `hepatitis.rda`, residing in the `data/`
directory of the package.

```{r,label=savehepatitisdataset}
save(hepatitis_table, hepatitis, hepatitis_count, hepatitis_maxp, file="hepatitis.rda")
```