### 2022-05-21 inst/race3.Rmd
more algebra

---
title: "Races with hyper3 objects"
author: "Robin KS Hankin"
date: "10/04/2022"
output: html_document
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("disordR")
```

(takes about an hour to process)

Consider a race in which there are six runners 1-6 but we happen to
know that three of the runners (1,2,3) are clones of strength $p_a$,
say, two of the runners (4,5) have strength $p_b$, and the final
runner (6) is of strength $p_c$.  The runners race and the finishing order is:

$$a\succ c\succ b\succ a\succ a \succ b$$

Thus the winner was $a$, second place for $b$, third for $a$, and so
on.  Alternatively we might say that $a$ came first, fourth, and
fifth; $b$ came third and sixth, and $c$ came second.  The
Plackett-Luce likelihood function for $p_a,p_b,p_c$ would be

$$
\frac{p_a}{3p_a+2p_b+p_c}\cdot
\frac{p_c}{2p_a+2p_b+p_c}\cdot
\frac{p_b}{2p_a+2p_b    }\cdot
\frac{p_a}{2p_a+ p_b    }\cdot
\frac{p_a}{ p_a+ p_b    }\cdot
\frac{p_b}{      p_b    },\qquad p_a+p_b+p_c=1
$$

The likelihood function is easily specified in package idiom:

```{r, label=definehyper3}
H <- ordervec2supp3(c("a","c","b","a","a","b"))
H
```

```{r, label=maximizeprobs, cache=TRUE}
mH  <-  maxp(H)
mH
```

We can test various hypotheses, for example that the competitors are of equal strength:

```{r testequality, cache=TRUE}
equalp.test(H)
```

showing that this small dataset does not allow one to say that the
competitors differ.  However, it is easy to generate synthetic order
statistics where one can reject the null of equality of strength:

```{r,biggerrace,cache=TRUE}
o <- rrace3(pn=c(a=20,b=11,c=7),ps=c(a=0.7,b=0.2,c=0.1))
o
H <- ordervec2supp3(o)
H
```

Above, we specify that there are 20 $a$'s, 11 $b$'s and 7 $c$'s with
strength $70\%$, $20\%$, and $10\%$ respectively.  Recall this means
that, in a head-to-head between $a$ and $b$ we would have $a$ winning
with probability $0.7/(0.7+0.2)=\frac{7}{9}$.

```{r,bigraceestimation,cache=TRUE}
mH <- maxp(H)
mH
pTRUE <- c(a=0.7,b=0.2,c=0.1)   # true values 
mH - pTRUE # residual
```

Above we see a reasonably close agreement between the true values and the estimate.

```{r,dothetest,cache=TRUE}
equalp.test(H)
```

Above we see strong evidence to reject the hypothesis that
$p_a=p_b=p_c=\frac{1}{3}$.  We may also test the hypothesis that the
strength vector is equal to its true value:

```{r,isittrue,cache=TRUE}
knownp.test(H,pTRUE)
```

And above we see that there is no evidence to reject the hypothesis
that the strengths are in fact the specified strengths in `pTRUE`.

# Some simple cases: likelihood functions for order statistics with clones

Here I consider some order statistics with nontrivial maximum
likelihood Bradley-Terry strength.   The simplest nontrivial case is
three competitors with strengths $a,a,b$ and finishing order $a\succ
b\succ a$.  The likelihood function would be

$$
\frac{a}{2a+b}\cdot\frac{b}{a+b}
$$

In this case we know that $a+b=1$ so this is equal to
$\mathcal{L}=\mathcal{L}(a)=\frac{a(1-a)}{1+a}$.

$$
\frac{d\mathcal{L}}{da}=\frac{(1+a)(1-2a)-a(1-a)}{(1+a)^2}=
\frac{1-2a+a-2a^2 -a + a^2}{(1+a)^2}=
\frac{1-2a-a^2}{(1+a)^2}
$$

This will be zero at $\sqrt{2}\pm 1$; we also note that
$d^2\mathcal{L}/da^2=-4(1+a)^{-3}$, manifestly strictly negative for
$0\leq a\leq 1$: the root is a maximum.

```{r label=maxaba,cache=TRUE}
maxp(ordervec2supp3(c("a","b","a")))
```

above, we see close agreement with the theoretical value of
$(\sqrt{2}-1,1-\sqrt{2})\simeq (0.414,0.586)$

At the evaluate of $\hat{a}=\sqrt{2}-1$ we can show that
$\log\mathcal{l}(\hat{a})=\log\left(3-2\sqrt{2}\right)\simeq -1.76$.
Defining $\mathcal{S}=\log\mathcal{L}$ as the support [log-likelihood]
we have

$$\mathcal{S}=\mathcal{S}(a)=\log\left(\frac{a(1-a)}{1+a}\right)-\log\left(3-2\sqrt{2}\right)$$

as a standard support function which has a maximum value of zero.
Further, we can use either analytical or numerical methods to find a
credible interval.  With an $n$-unit of support criterion the
analytical solution to $\mathcal{S}(p)=-2$ is given by defining
$X=\log(3-2\sqrt{2})-n$ and solving $p(1-p)/(1+p)=X, or
$\left(1-X\pm\sqrt{1+4X+X^2}\right)/2$, the two roots being the lower
and upper limits of the credible interval.  Graphically:

```{r}
a <- seq(from=0,by=0.005,to=1)
S <- function(a){log(a*(1-a)/((1+a)*(3-2*sqrt(2))))}
plot(a,S(a),type='b')
abline(h=c(0,-2))
abline(v=c(0.02438102,0.9524271),col='red')
abline(v=sqrt(2)-1)
```

With four competitors we have five nontrivial order statistics which,
without loss of generality, are $(a,b,a,a)$ and $(a,a,b,a)$.
$(a,b,a,b)$, $(a,b,b,a)$, $(a,b,c,a)$. Taking $(a,b,a,b)$ as an
example we have

$$
\mathcal{L}(a)=
\frac{a}{2a+2b}\cdot
\frac{b}{ a+2b}\cdot
\frac{a}{ a+ b}\cdot
\frac{b}{    b}=\frac{a^2-a^3}{4-2a}
$$

the evaluate would be a root of $2a^2-7a+4$, or $(7-\sqrt{17})/4\simeq
0.72$:

```{r label=maxabab,cache=TRUE}
maxp(ordervec2supp3(c("a","b","a","b")))
```

If we allow non-finishers there is another nontrivial order statistic,
viz $\left(a,b,\left\lbrace a,b\right\rbrace\right)$:

$$
\mathcal{L}(a)=
\frac{a}{2a+2b}\cdot
\frac{b}{ a+2b}\propto\frac{a(1-a)}{2-a}
$$

(see how the likelihood function is actually simpler than for the
complete order statistic).  The evaluate would be $2-\sqrt{2}\simeq
0.586$:

```{r label=maxnonfinishers,cache=TRUE}
o <- ordervec2supp3(c("a","b"),nonfinishers=c("a","b"))
o
maxp(o)
```

Now consider

$$a\succ b\succ c\succ a$$

with likelihood function

$$
\mathcal{L} =
\frac{a}{2a+b+c}\cdot
\frac{b}{ a+b+c}\cdot
\frac{c}{ a  +c},\qquad a+b+c=1$$

It can be shown that this has a maximum if $a^3+4a^2+4a-1=0$.
Although cubic equations have analytical roots, the expression is
complicated, and equal to about `c(a=0.2056,b=0.5466,c=0.2478)`:

```{r,label=abca,cache=TRUE}
maxp(ordervec2supp3(c("a","b","c","a")))
```

# Fisher information for simple cases for aab

We have three possible finishing orders:

$b\succ a\succ a$, probability $\frac{b}{2a+b}=\frac{1-a}{1+a}$.  And
$\frac{\partial\log\mathcal{L}}{\partial a}=\frac{2}{p^2-1}$.

$a\succ b\succ a$, probability $\frac{a}{2a+b}\cdot\frac{b}{a+b}=\frac{a(1-a)}{1+a}$. And
$\frac{\partial\log\mathcal{L}}{\partial a}=\frac{a^2+2a-1}{a^3-a}$.

$a\succ a\succ b$, probability $\frac{a}{2a+b}\cdot\frac{a}{a+b}=\frac{a^2}{1+a}$.  And
$\frac{\partial\log\mathcal{L}}{\partial a}=\frac{a+2}{a^2+a}$.

Thus the Fisher information is

$$
\frac{1-a}{1+a}\cdot\left(\frac{2}{a^2-1}\right)^2
+
\frac{a(1-a)}{1+a}\cdot\left(\frac{a^2+2a-1}{a^3-a}\right)^2
+\frac{a^2}{1+a}\cdot\left(\frac{a+2}{a^2+a}\right)^2
$$

Simplifying,

$$\mathcal{I}(a)=\frac{1+16a+18a^2-4a^3-15a^4-7a^5-a^6}{a(1-a)(1+a)^3}$$

# The constructors' championship

Here we define  likelihood function for the constructors' championship.

```{r,dotheconst}
jj <- read.table("constructor_2021.txt",header=TRUE)
i <- seq(from=1,by=2,to=nrow(jj)) # extract every second entry for points
constructor_points_2021 <- jj[i,ncol(jj)]
names(constructor_points_2021) <- jj[i,1]

constructor_points_2021 <- sort(constructor_points_2021,dec=TRUE)
constructor_points_2021

constructor_table_2021 <- jj[,-ncol(jj)]
constructor_table_2021[,1:17]
constructor_2021_table <- constructor_table_2021
```

Looking at the first column we see that the constructors are Mercedes
("`Merc`"), Red Bull Racing Honda ("`RBRH`"), and so on.  Each
constructor fields two cars in each race.  Looking at the next column,
we see the order statistic for Bahrain (`BHR`).  We see that Mercedes
came first and third, Red Bull came second and fifth, Ferrari sixth
and eighth, and so on.

```{r,maketheconst,cache=TRUE}
constructor_2021 <- ordertable2supp3(constructor_table_2021)
```

Then

```{r,label=maxconstr,cache=TRUE}
constructor_2021_maxp <-  maxp(constructor_2021)
mc <- constructor_2021_maxp  # saves typing
```

```{r,label=drawaconstructorpie}
pie(mc)
dotchart(mc,pch=16)
ordertransplot(rank(-mc),rank(-constructor_points_2021),xlab="likelihood rank",ylab="points rank")
```

We may test the null of equal strengths:

```{r,label=testequalityconstructor,cache=TRUE}
equalp.test(constructor_2021)
```

and see that there is (very) strong evidence to reject this null.
Further, we may test the hypothesis that the big four (viz Mercedes,
Ferrari, Red Bull Racing Honda, and McLaren-Mercedes) are equal:

```{r,label=testsameconstructor,cache=TRUE}
samep.test(constructor_2021,c("Merc","Ferrari","RBRH","MM")) # ~10 mins
```

and fail to reject that null.  How about the mid-range constructors:

```{r,label=testsameconstructorlowscores,cache=TRUE}
samep.test(constructor_2021,c("ATH","ARRF","AR","AMM")) # ~ 10 mins
```

### Package dataset {-}

Following lines create `constructor.rda`, residing in the `data/`
directory of the package.

```{r,label=saveconstructordataset}
save(
constructor_2021_maxp,
constructor_2021,
constructor_2021_table,
file="constructor.rda")
```