### 2022-07-04 inst/carcinoma.Rmd

---
title: "Carcinoma"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

(takes maybe twenty minutes to run without cache.  The value of
`carcinoma_maxp` is calculated separately [discussed below], and that
takes a few days to run).

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This is a short analysis of the carcinoma dataset presented by
Agresti, table 13.1, p542.  Seven pathologists classified each of 118
slides on the presence or absence of carcinoma in the uterine cervix.

Here, the object of inference is the set of pathologists.  The
probability model is as follows.  To each pathologist $i,1\leq i \leq
7$, we assign a non-negative Bradley-Terry strength and require $\sum
p_i=1$.  Given that two pathologists $i$ and $j$ differ in their
diagnosis then the probability of $i$ diagnosing the presence, and $j$
diagnosing the absence, of carcinoma is given by

\begin{equation}\tag{1}
\operatorname{Prob}(i^+,j^-) = \frac{p_i}{p_i+p_j},\qquad
\operatorname{Prob}(i^-,j^+) = \frac{p_j}{p_i+p_j}
\end{equation}

(here superscripts $+$ and $-$ refer to a positive and negative
diagnosis respectively).  Operationally the way we implement this in
`hyper2` idiom is to posit an (unobservable) ranking on the
clinicians: we place the clinicians in order from most convinced of
the presence of carcinoma to the least convinced.  Thus a ranking of
`DACBEGF` means that clinician `D` was most convinced of the presence
of carcinoma and `F` least convinced.  Given this, and given a single
slide, our actual observation might be that `D` and `A` diagnosed
presence, and the others diagnosed absence, of carcinoma.  The
corresponding entry would be

```
A B C D E F G
T F F T F F F
```

We ask what rankings are consistent with this observation.  Any
ranking with `A` and `D` ahead of `BCEFG` would be admissible.  In
`hyper2` notation we would say

$$\left\lbrace A,D\right\rbrace\succ\left\lbrace B,C,E,F,G\right\rbrace$$

Any ranking where `A` and `D` occupy the first two ranks and
`B`,`C`,`E`,`F`,`G` ranks three to seven, would be admissible.  We
would have $2!5!=240$ rankings.  The probability of the actual ranking
being one of these is simply the sum of these 48 as the rankings are
mutually exclusive.

Each ranking has a Plackett-Luce likelihood function which is created
in `hyper2` idiom using the bespoke `race()` function.  Here, we use
generalized grouped rank likelihood (`ggrl()`) to create an `lsl`
object, as used by Hankin 2019 for the MasterChef analysis.  This
furnishes a Bradley-Terry model for the clinicians' strength that
recovers equation (1).

```{r loadlib}
carcinoma_table <- read.table("carcinoma.txt",header=TRUE)
carcinoma_table[,1:7] <- carcinoma_table[,1:7] > 0
carcinoma_table
```

```{r makeshorttable}
wanted <- !(apply(carcinoma_table[,1:7],1,sum) %in% c(0,7))
carcinoma_table_short <- carcinoma_table[wanted,]
```

```{r makeggrllist,cache=TRUE}
W <- hyper2(pnames=LETTERS[1:7])
Wi <- list()

for(i in seq_len(nrow(carcinoma_table_short))){
      jj <- unlist(carcinoma_table_short[i,1:7])
      negative <- names(jj)[jj==0]
      positive <- names(jj)[jj==1]
      Wi[[i]] <- ggrl(W,positive,negative)
}
carcinoma <- lsl(Wi,powers= carcinoma_table_short$n)
```

There are several natural BT strengths to consider.  First, we could
simply count how many times each clinician diagnoses carcinoma:

```{r,label=countcarcinoma,cache=TRUE}
carcinoma_count <- colSums(sweep(carcinoma_table_short[,1:7],1,carcinoma_table_short$n,"*"))
names(carcinoma_count) <- colnames(carcinoma_table_short)[1:7]
carcinoma_count <- carcinoma_count/sum(carcinoma_count)
carcinoma_count
lc <- loglik_lsl(carcinoma_count,carcinoma)
```

Secondly, Zipf:

```{r,label=calccloglikzipf,cache=TRUE}
z <- zipf(7)
names(z) <- LETTERS[1:7]
z
lz <- loglik_lsl(z,carcinoma)  # takes about 30 seconds to run
```

Thirdly, equal probabilities:

```{r,label=calcloglikequalp,cache=TRUE}
e <- equalp(W)
e
le <- loglik_lsl(e,carcinoma) # takes about 30 seconds to run
```

The maximum likelihood estimate is somewhat more demanding.  To find
the maximum likelihood estimate we would use:

`maxp_lsl(carcinoma,startp=indep(carcinoma_count), control=list(trace=100))`

but this takes many hours to run.  It gives the following:

```{r,label=calculateloglikformaxp, cache=TRUE}
carcinoma_maxp <- 
c(A = 0.118778759012282, B = 0.542873969223425, C = 0.00654044305270073, 
D = 0.00242693714895262, E = 0.202821434317779, F = 0.00130596252816986, 
G = 0.125252494716691)
carcinoma_maxp
lmax <- loglik_lsl(carcinoma_maxp,carcinoma)
```

[of course, if `carcinoma_maxp` is available, we would use this as a
starting point, viz
`maxp_lsl(carcinoma,startp=indep(carcinoma_maxp))`].  With this, we
are now in a position to compare these four points:

```{r,label=showthreemethods}
(results <- c(count=lc, zipf=lz, equal=le,maxp=lmax))
results <- results - max(results)
results
```

We can use Wilks here to assess the null of `carcinoma_count`.  We
have $-2\log(\Lambda/\Lambda_0)\sim\chi^2_6$, giving us a $p$-value of

```{r,label=calculatechisquaredpvalueforcount}
pchisq(-2*results[1],df=6,lower.tail=FALSE)
```

What is the difference between `carcinoma_count` and `carcinoma_maxp`?

```{r}
carcinoma_count
carcinoma_maxp
ordertransplot(carcinoma_count,carcinoma_maxp)
ordertransplot(log(carcinoma_count),log(carcinoma_maxp),plotlims=c(-8,0))
```

Agresti, table 13.1, p542.

Landis, J. R., and G. G. Koch. 1977. An application of hierarchical
kappa-type statistics in the assessment of majority agreement among
multiple observers. Biometrics 33: 363-374.

### Package dataset {-}

Following lines create `carcinoma.rda`, residing in the `data/`
directory of the package.

```{r,label=savecarcinomadataset}
save(carcinoma_table,carcinoma, carcinoma_count, carcinoma_maxp,file="carcinoma.rda")
```