### 2022-07-19 inst/global_liveability_ranking.Rmd
global liveability ranking work

---
title: "Global liveability ranking"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

[The global liveability ranking](https://en.wikipedia.org/wiki/Global_Liveability_Ranking)
shows where you would like to live.

```{r showglrtable}
readLines("global_liveability_ranking.txt")
```

First we will turn it into a `hyper2` object:

```{r readglrtable}
o <- readLines("global_liveability_ranking.txt") %>% strsplit(" ") %>% lapply(function(l){l[-1]})
H <- hyper2()
for(i in seq_along(o)){
  H <- H  + race(o[[i]])
}
```

And do some tests:

```{r testgrl, cache=TRUE}
mH <- maxp(H)
pie(mH)
equalp.test(H)
```

Now we tether some of the cities together:

```{r, tether}
elastic <- list(
Austria = c("Vienna"),
Denmark = c("Copenhagen"),
Switzerland = c("Zurich","Geneva"),
Canada = c("Calgary","Vancouver","Toronto"),
Germany = c("Frankfurt","Hamburg"),
Netherlands = c("Amsterdam"),
Japan = c("Osaka","Tokyo"),
Australia = c("Melbourne","Adelaide","Perth","Brisbane","Sydney"),
NZ = c("Auckland","Wellington"),
Finland = c("Helsinki")
)
e <- NULL
for(i in seq_along(elastic)){
  jj <- rep(i,length(elastic[[i]]))
  names(jj) <- elastic[[i]]
  e <- c(e,jj)
}
e 
```

Now create a `hyper3` object:

```{r,createhyper3func}
ML <- function(h){
   H3 <- hyper3()
   for(r in o){ H3 <- H3 + cheering3(r,e=e,help=rep(h,10))}
   return(H3)
}
f <- function(h){maxp(ML(h),give=TRUE,n=1)$likes}
```

And plot a profile likelihood

```{r,plliveable,cache=TRUE}
l <- seq(from=1,to=5,len=10)
L <- sapply(l,f)
```

and plot it:

```{r}
plot(log(l),L-max(L),type='b')
```