### 2022-07-29 inst/setpartitions_hyper2.Rmd
start work on issue # 148

---
title: "setpartitions_hyper2"
output: html_document
date: "2022-07-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("partitions")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here we will sum likelihoods across a sample space and verify that the total is constant.

```{r}
M <- apply(setparts(c(2,2,1)),2,function(x){letters[x]})
M
```

```{r}
f <- function(strength,help,vec){loglik(strength,cheering3(v=vec[1:4],e=c(a=1,b=2,c=3),help=help,vec[5]))}
help <- c(1.11,2.222,1.7345)
strength <- 1
```