### 2022-07-29 inst/setpartitions_hyper2.Rmd
slight reworking

---
title: "setpartitions_hyper2"
output: html_document
date: "2022-07-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("partitions")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here we will sum likelihoods across a sample space and verify that the
total is constant.

```{r}
M <- apply(setparts(c(2,2,1)),2,function(x){letters[x]})
M
```

```{r}
f <- function(strength,help,vec){
loglik(strength,cheering3(v=vec[1:4],e=c(a=1,b=2,c=3),help=help,vec[5]),log=FALSE)
}

tfun <- function(help,strength){
   out <- 0
   for(i in seq_len(ncol(M))){
      out <- out + f(strength,help,vec=M[i,])
   } 
   return(out)
}

print(tfun(help=c(1.11,2.222,1.7345), strength= c(0.1,0.3,0.6)))
print(tfun(help=c(1.,2.2,1.2), strength= c(0.3,0.2,0.5)))
```