### 2022-07-30 inst/setpartitions_hyper2.Rmd
multiset() not setparts()

---
title: "setpartitions_hyper2"
output: html_document
date: "2022-07-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("partitions")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here we will sum likelihoods across a sample space and verify that the
total is constant.

```{r}
M <- apply(multiset(c(1,1,2,2,3)),2,function(x){letters[x]})
noquote(M)
```

```{r}
tfun <- function(help,strength){
   out <- 0
     for(i in seq_len(ncol(M))){
     vec <- M[,i]
     out <- out +
       loglik(strength,cheering3(v=vec,e=c(a=1,b=2,c=3),help=help),log=FALSE)
   }
   return(out)
}

h <- c(1.11,2.222,1.7345)
print(tfun(help=(1:3)/10, strength= c(a=0.1,b=0.3,c=0.6)))
print(tfun(help=(7:5)/22, strength= c(a=0.3,b=0.2,c=0.5)))

```