### 2022-08-04 inst/mann_whitney_wilcoxon.Rmd
work on Mann-Whitney-Wilcox test (as a source for jss_hyper3.Rnw)

---
title: "Mann-Whitney-Wilcoxon test with `hyper3` objects"
author: "Robin KS Hankin"
date: "10/04/2022"
output: html_document
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

We do some work with Mann-Whitney test, using the chorioamnion dataset at `wilcox.test.Rd`:

```{r,label=defineAB}
x <- c(0.80, 0.83, 1.89, 1.04, 1.45, 1.38, 1.91, 1.64, 0.73, 1.46)
y <- c(1.15, 0.88, 0.90, 0.74, 1.21)
```

First do the Mann-Whitney-Wilcox test:

```{r,domwwtest}
wilcox.test(x,y)
```

Now use `hyper3`:

```{r,hyper3AB,cache=TRUE}
names(x) <- rep("x",length(x))
names(y) <- rep("y",length(y))
equalp.test(ordervec2supp3(names(sort(c(x,y)))))
```

Now some permutation tests under the null:

```{r,repeat_mww_H0,cache=TRUE}
n <- 10
mww_pval <- rep(NA,n)
for(i in seq_len(n)){
      jj <- sample(c(x,y))
      xstar <- jj[seq(from=1,length=length(x))]
      ystar <- jj[seq(from=length(x)+1,to=length(x)+length(y))]
      mww_pval[i] <- suppressWarnings(wilcox.test(xstar,ystar)$p.value)
}
```

```{r,histmww}
mww_pval
hist(mww_pval)
```

```{r,repeat_hyper3_H0,cache=TRUE}
n <- 10
hyper3_pval <- rep(NA,n)
for(i in seq_len(n)){
      jj <- sample(c(x,y))
      xstar <- jj[seq(from=1,length=length(x))]
      ystar <- jj[seq(from=length(x)+1,to=length(x)+length(y))]
      hyper3_pval[i] <- equalp.test(ordervec2supp3(args2ordervec(x=xstar,y=ystar)))$p.value
}
```

```{r,histhyper3}
hyper3_pval
hist(hyper3_pval)
```