### 2022-08-11 inst/jss_hyper3.Rnw
more rewriting and detail

% -*- mode: noweb; noweb-default-code-mode: R-mode; -*-
\documentclass[article]{jss}

\author{Robin K. S. Hankin\\Auckland University of Technology}
\title{Generalized Plackett-Luce Likelihoods}
\Plainauthor{Robin K. S. Hankin} %% comma-separated
\Plaintitle{Generalized Plackett-Luce Likelihoods}

%% an abstract and keywords
\Abstract{

  The \code{hyper2} package furnishes extensive idiom to work with
  extensions of the Bradley-Terry probability model such as
  Plackett-Luce likelihood functions, team strengths, and reified
  entities (monsters).  Here, I present a generalization of
  \code{hyper2} in which multiple entities are constrained to have
  identical Bradley-Terry strengths.  A new \proglang{S3} class
  \code{hyper3}, along with associated methods, is motivated and
  introduced.  Four datasets are analysed, each analysis furnishing
  new insight, and each highlighting different capabilities of the
  package.
}
\Keywords{Plackett-Luce, Bradley-Terry, Mann-Whitney}
\Plainkeywords{Plackett-Luce, Bradley-Terry, Mann-Whitney}

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Robin K. S. Hankin\\
  School of Engineering and Mathematical Sciences\\
  Auckland University of Technology\\
  Wellesley Street\\
  Auckland, New Zealand\\
  E-mail: \email{hankin.robin@gmail.com}\\
  URL: \url{https://academics.aut.ac.nz/robin.hankin}
}

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

<<setup, cache=FALSE,echo=FALSE>>=
library("knitr")
library("hyper2")
opts_chunk$set(cache=TRUE, autodep=TRUE)
@

\section{introduction}

The \code{hyper2} package~\citep{hankin2009,hankin2017} furnishes
computational support for generalized
Plackett-Luce~\citep{plackett1975} likelihood functions.  Our
preferred interpretation is a race (as in track and field atheletics):
given six competitors $1-6$, we ascribe to them nonnegative strengths
$p_1,\ldots, p_6$; the probability that $i$ beats $j$ is
$p_i/(p_i+p_j)$.  It is conventional to normalise so that the total
strength is unity, and to identify a competitor with his strength.
Given an order statistic, say $p_1\succ p_2\succ p_3\succ p_4$, the
Plackett-Luce likelihood function would be

\begin{equation}\label{PL_like}
  \frac{p_1}{p_1+p_2+p_3+p_4}\cdot
  \frac{p_2}{    p_2+p_3+p_4}\cdot
  \frac{p_3}{        p_3+p_4}\cdot
  \frac{p_4}{            p_4}.
\end{equation}

The \code{hyper2} package generalizes this likelihood function to
functions of ${\mathbf p}=(p_1,\ldots,p_n)$ with

\begin{equation}\label{hyper2likelihood}
\mathcal{L}\left(\mathbf{p}\right)=
\prod_{s\in \mathcal{O}}\left({\sum_{i\in s}}p_i\right)^{n_s}
\end{equation}

\noindent where~$\mathcal{O}$ is a set of observations and~$s$ a
subset of~$\left[n\right]=\left\{1,2,\ldots,n\right\}$; numbers~$n_s$
are integers which may be positive or negative.  It becomes possible
to analyse a wider range of observations such as incomplete order
statistics or teams; and the introduction of reified entities allows
one to consider draws~\citep{hankin2009}, noncompetitive tactics such
as collusion~\citep{hankin2020}, and the phenomenon of team cohesion
wherein the team becomes stronger than the sum of its parts.

However, consider a race in which there are six runners 1-6 but we
happen to know that three of the runners (1,2,3) are clones of
strength $p_a$, two of the runners (4,5) have strength $p_b$, and the
final runner (6) is of strength $p_c$.  We normalise so
$p_a+p_b+p_c=1$.  The runners race and the finishing order is:

$$a\succ c\succ b\succ a\succ a \succ b$$

Thus the winner was $a$, second place for $b$, third for $a$, and so
on.  Alternatively we might say that $a$ came first, fourth, and
fifth; $b$ came third and sixth, and $c$ came second.  The
Plackett-Luce likelihood function for $p_a,p_b,p_c$ would be

\begin{equation}\label{plackettluce}
\frac{p_a}{3p_a+2p_b+p_c}\cdot
\frac{p_c}{2p_a+2p_b+p_c}\cdot
\frac{p_b}{2p_a+2p_b    }\cdot
\frac{p_a}{2p_a+ p_b    }\cdot
\frac{p_a}{ p_a+ p_b    }\cdot
\frac{p_b}{      p_b    },\qquad p_a+p_b+p_c=1
\end{equation}

Existing \code{hyper2} formalism as described by \citet{hankin2017}
cannot represent this likelihood function, because
Equation~\ref{hyper2likelihood} uses sets as the indexing elements,
and in this case we need multisets\footnote{Note that the version of
\code{hyper2} presented by \citet{hankin2017} used integer-valued sets
together with a print method that used a complicated mapping system
from integers to competitor names.  Current methodology [following
commit \code{51a8b46}] is to use sets of character strings which
represent the competitors directly; this allows for easier combination
of observations including different competitors.}.  The declarations

\begin{verbatim}
typedef map<string, long double> weightedplayervector;
typedef map<weightedplayervector, long double> hyper3;
\end{verbatim}

show how the \proglang{map} class of the Standard Template Library is
used: \proglang{weightedplayervector} objects map strings to long
doubles (specifically, mapping player names to their multiplicities),
and objects of class \proglang{hyper3} are maps from a
\proglang{weightedplayervector} object to long doubles.  One advantage
of this is efficiency: Search, removal, and insertion operations have
logarithmic complexity.  As an example, the following \proglang{C++}
pseudo code would create a log-likelihood function for the first term
in Equation~\ref{plackettluce}:

\begin{verbatim}
weightedplayervector n,d;
n["a"] = 1; // n for numerator
d["a"] = 3; // d for denominator
d["b"] = 2;
d["c"] = 1;

hyper3 L;
L[n] = 1;
L[d] = -1;
\end{verbatim}

Above, object \code{L} is an object of class \code{hyper3}; it may be
evaluated at points in probability space [that is, a vector
\code{[a,b,c]} of nonnegative values with unit sum] using standard
\proglang{R} idiom wrapping \proglang{C++} back end.

The package includes an \proglang{S4} class \code{hyper3} for this
type of object.  Package idiom for creating such objects uses named
vectors:

<<>>=
H <- hyper3()
H[c(a=1)] <- 1
H[c(a=3,b=2,c=1)] <- -1
H
@

Above, we see object \code{H} is a log-likelihood function of the
players' strengths, which may be evaluated at specified points in
probability space.  If, for example, we wish to assess $H_1\colon
p_a=0.9,p_b=0.05,p_c=0.05$ and $H_2\colon p_a=0.01,p_b=0.01,p_c=0.98$,
we can evaluate these hypotheses using generic function
\code{loglik()}:

<<>>=
loglik(c(a=0.01,b=0.01,c=0.98),H)
loglik(c(a=0.9,b=0.05,c=0.05),H)
@

Thus we prefer $H_1$ over $H_2$ with about 3.5 units of support,
satisfying the standard two units of support
criterion~\citep{edwards1972}, and we conclude that our observation
[in this case, that one of the three clones of player $a$ beat the $b$
twins and the singleton $c$] furnishes strong support against $H_2$ in
favour of $H_1$.

The package includes many helper functions to work with order
statistics of this type.  Function \code{ordervec2supp3()}, for
example, can be used to generate a loglikelihood function for
Equation~\ref{PL_like}:

<<>>=
(H <- ordervec2supp3(c("a","c","b","a","a","b")))
@

We may find a maximum likelihood estimate for the players' strengths:

<<maxexamp,cache=TRUE>>=
(mH <- maxp(H))
@

We may use a selection of statistical tests on the order statistic,
modified from~\citep{hankin2017} to work with \code{hyper3} objects.
For example, we may assess the hypothesis that all three players are
of equal strength [viz $H_0\colon p_a=p_b=p_c=\frac{1}{3}$]:

<<testequality,cache=TRUE>>=
equalp.test(H)
@

showing, perhaps unsurprisingly, that this small dataset is consistent
with $H_0$.

\section{Some simple cases}

Here I consider some order statistics with nontrivial maximum
likelihood Bradley-Terry strength.  The simplest nontrivial case would
be three competitors with strengths $a,a,b$ and finishing order
$a\succ b\succ a$.  The likelihood function would be

\begin{equation}\label{aba}
\frac{a}{2a+b}\cdot\frac{b}{a+b}
\end{equation}

In this case we know that $a+b=1$ so this is equal to
$\mathcal{L}=\mathcal{L}(a)=\frac{a(1-a)}{1+a}$.

\begin{equation}\label{aba_mle}
\frac{d\mathcal{L}}{da}=\frac{(1+a)(1-2a)-a(1-a)}{(1+a)^2}=
\frac{1-2a+a-2a^2 -a + a^2}{(1+a)^2}=
\frac{1-2a-a^2}{(1+a)^2}
\end{equation}

This will be zero at $\sqrt{2}\pm 1$; we also note that
$d^2\mathcal{L}/da^2=-4(1+a)^{-3}$, manifestly strictly negative for
$0\leq a\leq 1$: the root is a maximum.

<<maxpaba,cache=TRUE>>=
maxp(ordervec2supp3(c("a","b","a")))
@

above, we see close agreement with the theoretical value of
$(\sqrt{2}-1,1-\sqrt{2})\simeq (0.414,0.586)$.  Observe that the
maximum likelihood estimate for $a$ is strictly less than 0.5, even
though the finishing order is symmetric.  Using
$\mathcal{L}(a)=\frac{a(1-a)}{1+a}$, we can show that
$\log\mathcal{L}(\hat{a})=\log\left(3-2\sqrt{2}\right)\simeq -1.76$,
where $\hat{a}=\sqrt{2}-1$ is the maximum likelihood estimate for $a$.
Defining $\mathcal{S}=\log\mathcal{L}$ as the support [log-likelihood]
we have

\begin{equation}
\mathcal{S}=\mathcal{S}(a)=\log\left(\frac{a(1-a)}{1+a}\right)-\log\left(3-2\sqrt{2}\right)
\end{equation}

as a standard support function which has a maximum value of zero when
evaluated at $\hat{a}=\sqrt{2}-1$.  For example, we can test the null
that $a=b=\frac{1}{2}$, the statement that the competitors have equal
Bradley-Terry strengths:

<<>>=
a <- 1/2 # null
(S_delta <- log( a*(1-a)/(1+a))- log(3-2*sqrt(2)))
@

Thus the additional support gained in moving from $a=\frac{1}{2}$ to
the evaluate of $a=\sqrt{2}-1$ is 0.029, rather small [as might be
expected given that we have only one rather uninformative observation,
and also given that the maximum likelihood estimate ($\simeq 0.41$) is
quite close to the null of $0.5$].  Nevertheless we can follow
\cite{edwards1972} and apply Wilks's theorem for a $p$-value:

<<>>=
pchisq(-2*S_delta,df=1,lower.tail=FALSE)
@

The $p$-value is about 0.81, exceeding 0.05; thus we have no strong
evidence to reject the null of $a=\frac{1}{2}$.  The observation is
informative, in the sense that we can find a credible interval for
$a$.  With an $n$-unit of support criterion the analytical solution to
$\mathcal{S}(p)=-n$ is given by defining $X=\log(3-2\sqrt{2})-n$ and
solving $p(1-p)/(1+p)=X$, or
$p_\pm=\left(1-X\pm\sqrt{1+4X+X^2}\right)/2$, the two roots being the
lower and upper limits of the credible interval;
Figure~\ref{ABA_likelihood}.

\begin{figure}[htbp]
  \begin{center}
<<figaba>>=
a <- seq(from=0,by=0.005,to=1)
S <- function(a){log(a*(1-a)/((1+a)*(3-2*sqrt(2))))}
plot(a,S(a),type='b')
abline(h=c(0,-2))
abline(v=c(0.02438102,0.9524271),col='red')
abline(v=sqrt(2)-1)
@
\caption{A\label{ABA_likelihood} support function for $p_a$ with observation $a\succ b\succ a$}
\end{center}
\end{figure}

\subsection{All results for AAB}

\begin{itemize}
\item $A\succ A\succ B\longrightarrow\frac{a^2}{1+a}$
\item $A\succ B\succ A\longrightarrow\frac{1-2a-a^2}{(1+a)^2}$
\item $B\succ A\succ A\longrightarrow\frac{1-a}{1+a}$
\end{itemize}

Likelihood functions for these order statistics are given in
Figure~\ref{ABAetc_likelihood}.  It can be shown that the Fisher
information for such observations is

\begin{equation}
\mathcal{I}(a)=\frac{1+16a+18a^2-4a^3-15a^4-7a^5-a^6}{a(1-a)(1+a)^3}
\end{equation}

which has a minimum at about $0.575$.

\begin{figure}[htbp]
  \begin{center}
<<figabbabb,echo=FALSE>>=
f_aab <- function(a){a^2/(1+a)}
f_aba <- function(a){a*(1-a)/(1+a)^2}
f_baa <- function(a){(1-a)/(1+a)}

p <- function(f,...){points(a,f(a)/max(f(a)),...)}
plot(0:1,0:1,xlab="p(a)",ylab="Likelihood",type="n")
p(f_aab,type="l",col="black")
p(f_aba,type="l",col="red")
p(f_baa,type="l",col="blue")
legend(x=0.85,y=0.7,col=c("black","red","blue"),lty=1,pch=NA,legend=c("AAB","ABA","BAA"))
abline(h=exp(-2),lty=2)
@
\caption{Likelihood \label{ABAetc_likelihood} functions for
observations $a\succ a\succ b$, $a\succ b\succ a$, $b\succ a\succ a$}
\end{center}
\end{figure}

\subsection{Nonfinishers}

If we allow non-finishers [that is, a subset of competitors who are
beaten by all the ranked competitors], there is another nontrivial
order statistic, viz $a\succ b\succ\left\lbrace a,b\right\rbrace$
[thus one of the two $a$'s won, one of the $b$'s came second, and one
of each of $a$ and $b$ failed to finish].

\begin{equation}
\mathcal{L}(a)=
\frac{a}{2a+2b}\cdot
\frac{b}{ a+2b}\propto\frac{a(1-a)}{2-a}
\end{equation}

(see how the likelihood function is actually simpler than for the
complete order statistic).  The evaluate would be $2-\sqrt{2}\simeq
0.586$:

<<cache=TRUE>>=
maxp(ordervec2supp3(c("a","b"),nonfinishers=c("a","b")))
@

\section{Mann-Whitney test}

The ideas presented above can easily be extended to arbitrarily large
numbers of competitors, although the analytical expressions tend to be
intractable and numerical methods must be used.  Given an order
statistic of the type considered above, the Mann-Whitney-Wilcoxon
test~\citep{mann1947,wilcoxon1945} assesses a null of identity of
underlying distributions.  Consider the chorioamnion dataset at
\code{wilcox.test.Rd}:

<<define_xy_wilcox>>=
x <- c(0.80, 0.83, 1.89, 1.04, 1.45, 1.38, 1.91, 1.64, 0.73, 1.46)
y <- c(1.15, 0.88, 0.90, 0.74, 1.21)
@

Here we see permeability constants of the human chorioamnion at term
(\code{x}) and between 12 to 26 weeks gestational age (\code{y}).  The
order statistic is straightforward to calculate:

<<hyper3osdef,cache=TRUE>>=
names(x) <- rep("x",length(x))
names(y) <- rep("y",length(y))
(os <- names(sort(c(x,y))))
@

Then object \code{os} is converted to a \code{hyper3} object, again
with \code{ordervec2supp3()}, which may be assessed using the Method
of Support:

<<hyper3xytest,cache=TRUE>>=
Hxy <- ordervec2supp3(os)
equalp.test(Hxy)
@

Above, we use generic function \code{equalp.test()} to test the null
that the permeability of the two groups both have Bradley-Terry
strength of $0.5$.  We see a p-value of about 0.09; compare 0.25 from
\code{wilcox.test()}.  However, observe that the \code{hyper3}
likelihood approach gives more information than Wilcoxon's analysis:
Firstly, we see that the maximum likelihood estimate for the
Bradley-Terry strength of \code{x} is about 0.24, considerably less
than the null of 0.5; further, we may plot a support curve for this
dataset, given in Figure~\ref{wilcox_likelihood}.

\begin{figure}[htbp]
  \begin{center}
<<plotwilcoxlike,echo=FALSE>>=
a <- seq(from=0.02,to=0.8,len=40)
L <- sapply(a,function(p){loglik(p,Hxy)})
plot(a,L-max(L),type='b')
abline(h=c(0,-2))
abline(v=c(0.24))
abline(v=c(0.5),lty=2)
@
\caption{A\label{wilcox_likelihood} support for the Bradley-Terry
strength $p_a$ of permeability at term.  The evaluate of 0.24 is
shown; and the two-units-of support credible interval, which does not
exclude $H_0\colon p_a=0.5$ (dotted line), is also shown}
\end{center}
\end{figure}

\section{Multiple competitors: javelin}

The ideas presented above may be extended to more than two types of
competitors.  Consider the following table, drawn from the men's
javelin, 2020 Olympics:

<<javelintable>>=
javelin_table
@

Thus Chopra threw 87.03m on his first throw, 87.58m on his second, and
so on.  No-throws, ignored here, are indicated with an \code{X}.  We
may convert this to a named vector with elements being the throw
distances, and names being the competitors, using
\code{attemptstable2supp3()}:

<<label=converttosupp3,cache=TRUE>>=
javelin_vector <- attemptstable2supp3(javelin_table,decreasing=TRUE,give.supp=FALSE)
javelin_vector
@

Above we see that Chopra threw the longest and second-longest throws
of 87.58m and 87.03 respectively; Vadlejch threw the third-longest
throw of 86.67m, and so on (\code{NA} entries correspond to
no-throws.)  The attempts table may be converted to a \code{hyper3}
object, again using function \code{attemptstable2supp3()} but this
time passing \code{give.supp=TRUE}:

<<dothething2,cache=TRUE>>=
javelin <- ordervec2supp3(v=names(javelin_vector)[!is.na(javelin_vector)])
@

Above, object \code{javelin} is a \code{hyper3} likelihood function,
so one has access to the standard likelihood-based methods, such as
finding and displaying the maximum likelihood estimate:

<<setdigits,echo=FALSE>>=
options(digits=3)
@

\begin{figure}[htbp]
  \begin{center}
<<testthejav,cache=TRUE>>=
(mj <- maxp(javelin))
dotchart(mj,pch=16)
@
\caption{Maximum likelihood estimate for javelin throwers' Bradley-Terry strengths}
\end{center}
\end{figure}

Above, we see that Vadlejch has the highest estimated Bradley-Terry
strength, but further analysis with \code{equalp.test()} reveals that
there is no strong evidence in the datast to reject the hypothesis of
equal competitive strength ($p=0.26$), or that Vadlejch has a strength
higher than the null value of $\frac{1}{8}$ ($p=0.1$).

A particularly attractive feature of this analysis is that the
Bradley-Terry strengths have direct operational significance: If two
competitors, say Vadlejch and Vesely, were to throw a javelin, then we
would estimate the probability that competitors, say Vadlejch would
throw further than Vesely at $\displaystyle p_{\mbox{\tiny
Vad}}/\left(p_{\mbox{\tiny Vad}} + p_{\mbox{\tiny Ves}}\right)\simeq
0.74$.  Indeed, from a training or selection perspective we might
follow \cite{hankin2017} and observe that log-contrasts
~\citep{ohagan2004} $\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny
Ves}}\right)$ appear to have approximately Gaussian likelihood
functions for observations of the type considered here.  Profile
log-likelihood curves for log-contrasts are easily produced by the
package, Figure~\ref{profliklogcont}.

<<havealook,echo=FALSE,cache=TRUE>>=
f <- function(s){
  jj <- specificp.test(javelin,"Vesely",s,n=2)
  p <- jj$null_estimate
 return(c(log(p[6]/p[7]),jj$null_support))
}
Ves <- seq(from=0.0199,to=0.33,len=16)
M <- sapply(Ves,f)
M[2,] <- M[2,]-max(M[2,])
rownames(M) <- c("logcontrast","support")
@

\begin{figure}[htbp]
\begin{center}
<<plottheloglikcont,echo=FALSE>>=
colnames(M) <-  as.character(Ves)
plot(t(M),type="b")
abline(h=c(0,-2))
abline(v=0,lty=2)
abline(v=log(0.32062833/0.11402735)) # these from mp
@
\caption{Profile \label{profliklogcont} likelihood for log-contrast
$\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny Ves}}\right)$.  Null
of $p_{\mbox{\tiny Vad}}=p_{\mbox{\tiny Ves}}$ indicated with a dotted
line, and two-units-of-support limit indicated with horizontal lines;
thus the null is not rejected}
\end{center}
\end{figure}

\section{Formula 1 motor racing: the constructors' championship}

In Formula 1 motor racing, the constructors' championship takes place
between {\em manufacturers} of racing cars (compare the drivers'
championship, which is between drivers).  Here, we consider the
constructor to be the object of inference.  Each constructor typically
fields two cars, each of which separately accumulates ranking-based
points at each venue.  Here we use generalized Plackett-Luce to assess
the contructors' performance.  The following table, included in the
\code{hyper2} package as a dataset, shows rankings for the first 9
venues of the 2021 season:

<<showconstructortable>>=
constructor_2021_table[,1:9]
@

Above, we see that Mercedes (``\code{Merc}'') came first and third at
Bahrain (\code{BHR}); and came second and retired at Emilia Romagna
(\code{EMI}); full details of the notation and conventions are given
in the package at \code{constructor.Rd}.  The identity of the driver
is viewed as inadmissible information and indeed may change during a
season.  Alternatively, we may regard the driver and the constructor
as a joint entity, with the constructor's ability to attract and
retain a skilled driver being part of the object of inference.  The
associated generalized Plackett-Luce \code{hyper3} object is easily
constructed using package idiom, in this case
\code{ordertable2supp3()}, and we may use this to assess the
Plackett-Luce strengths of the constructors:

<<maxpconstructor2021, cache=TRUE>>=
const2020 <- ordertable2supp3(constructor_2020_table)
const2021 <- ordertable2supp3(constructor_2021_table)
maxp(const2020,n=1)
maxp(const2021,n=1)
@

Above, we see the strength of Mercedes falling from about 0.38 in 2020
to less than 0.20 in 2021 and it is natural to wonder whether this can
be ascribed to random variation.  Observe that testing such a
hypothesis is complicated by the fact that constructors field multiple
cars, and also that constructors come and go, with two 2020 teams
dropping out between years and two joining.  Package idiom to test
this statistically would be to define a combined likelihood function
for both years, keeping track of the year:

<<definecombinedlikelihoodfunction,cache=TRUE>>=
H <- psubs(constructor_2020,"Merc","Merc2020") + psubs(constructor_2021,"Merc","Merc2021")
@

Above, we use generic function \code{psubs()} to change the name of
Mercedes from \code{Merc} to \code{Merc2020} and \code{Merc2021}
respectively.  Note the use of ``\code{+}" to represent addition of
log-likelihoods, corresponding to the assumption of conditional
independence of results.  The null would be simply that the strengths
of \code{Merc2020} and of \code{Merc2021} are identical.  Package
idiom would be to use generic function \code{samep.test()}:

<<usesamep,cache=TRUE>>=
samep.test(H,c("Merc2020","Merc2021"))
@

Above, we see strong evidence for a real decrease in the strength of
the Mercedes team from 2020 to 2021, with $p=0.002$.

\section{Conclusions and further work}

Plackett-Luce likelihood functions for rank datasets have been
generalized to impose identity of Bradley-Terry strengths for certain
groups; the preferred interpretation is a running race in which the
competitors are split into equivalence classes of clones.
Implementing this in \proglang{R} is accomplished via a \proglang{C++}
back-end making use of the \proglang{STL} ``map" class which offers
efficiency advantages, especially for large objects.

New likelihood functions for simple cases with three or four
competitors were presented, and extending to larger numbers furnishes
a generalization of the Mann-Whitney-Wilcoxon test that offers a
specific alternative (Bradley-Terry strength) with a clear operational
definition.  Further generalizations allow the analysis of more than
two groups, here applied to Olympic Javelin throw distances.
Generalized Plackett-Luce likelihood functions were used to assess the
constructors' championship and a reasonable null.  Specifically, the
hypothesis that the strength of the Mercedes team remained unchanged
between 2020 and 2021 was tested and rejected.

Further work might include a systematic comparison between
\code{hyper3} approach and the Mann-Whitney-Wilcoxon test, including
the characterisation of the power function of both tests.  Further,
the package could easily be extended to allow non-integer
multiplicities, which might prove useful in the context of reified
Bradley Terry techniques~\citep{hankin2020}.

\bibliography{chess}
\end{document}