### 2022-08-25 inst/collusion3.Rmd
try collusion with hyper3 objects

---
title: "Chess collusion with hyper3 objects"
author: "Robin KS Hankin"
output: html_document
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

We analyse results from Curacao 1962 using `hyper3` formalism using
the weighted Draw monster appraoch outlined in `kka.Rmd`:

$$
\left(\frac{\lambda p_1, D(p_1+p_2),p_2}{\lambda p_1+D(p_1+p_2)+p_2}\right)^{(a,b,c)}
\left(\frac{p_1, D(p_1+p_2),\lambda p_2}{p_1+D(p_1+p_2)+\lambda p_2}\right)^{(d,e,f)}
$$

\newcommand{\wld}[3]{\mathord{+}#1\,\mathord{=}#2\,\mathord{-}#3}

Above, $p_1$ plays $p_2$ a total of $a+b+c+d+e+f$ times; we have
$a+b+c$ matches with $p_1$ playing white and $p_2$ black, with $a$
wins, $b$ draws and $c$ losses [write $\wld{a}{b}{c}$], and $d+e+f$
matches with $p_1$ playing black and $p_2$ playing white with
$\wld{d}{e}{f}$.

```{r,label=loadjavelinandconvert}
cur_matches <- read.table("stockholm1962_matches.txt",header=FALSE)
head(cur_matches)
nrow(cur_matches)
```

Above, we see on the first line that Aaron (white) lost to Barcza
(black).  We will set up `hyper2` and `hyper3` objects corresponding
to likelihood functions for these observations.

```{r,label=nat}
jj <- read.table("stockholm1962.txt",header=FALSE)[,1:2]
(players <- jj$V1)
(nationalities <- jj$V2)
```

```{r,label=H2H3}
H2 <- hyper2()
H3 <- hyper3()
f <- function(lambda,D){
  for(i in seq_len(nrow(cur_matches))){
    white_player <- cur_matches[i,1]
    black_player <- cur_matches[i,2]
    result <- cur_matches[i,3]
    if(result == "1-0"){ # white player wins
       num <- lambda 
       names(num) <- white_player
    } else if(result == "0-1"){  # black player wins
       num <- 1
       names(num) <- black_player
    } else if (result == "1/2-1/2"){ # draw, duh
      num <- c(D,D)
      names(num) <- c(white_player,black_player)
    } else {
      stop("this cannot happen")
    }

    H3[num] %<>% inc()       
    den <- c(lambda+D,1+D)
    names(den) <- c(white_player,black_player)
    H3[den] %<>% dec()
  }
  return(H3)
}
```

Now use it

```{r,try11and1888,cache=TRUE}
H3 <- f(1.1,1.888)
```

```{r,showH3}
H3
summary(H3)
```