### 2022-09-01 inst/interzonal.Rmd

---
title: "Analysis of the 1962 world chess championship, Curacao"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This file creates likelihood functions `H` and `H_coll`, which appear in
the package as `interzonal` and `collusion` respectively.

```{r readtabs}
jj <- read.table("stockholm1962.txt",header=FALSE)
results <- as.matrix(jj[,-(1:2)])
players <- as.character(jj$V1)
nationality <- as.character(jj$V2)
rownames(results) <- players
colnames(results) <- players
points <- rowSums(results,na.rm=TRUE)
```

Now some comparison between `stockholm1962.txt` (`results`) and
`stockholm1962_matches.txt` (`results2`):

```{r makeH2,cache=TRUE}
H <- hyper2(pnames=c("white","draw",players))
H_coll <- hyper2(pnames=c("white","draw","collusion",players))

restab <- read.table("stockholm1962_matches.txt",header=FALSE)
stopifnot(all(unique(sort(c(as.character(restab$V1),as.character(restab$V2)))) == sort(players)))
results2 <- matrix(0,length(players),length(players))
rownames(results2) <- players
colnames(results2) <- players
diag(results2) <- NA

for(i in seq_len(nrow(restab))){
    white_player <- as.character(restab[i,1])
    black_player <- as.character(restab[i,2])
    match_result <- as.character(restab[i,3])

    nw <- which(players==white_player)
    nb <- which(players==black_player)

    if(match_result == "1-0"){ # white win
        results2[nw,nb] <- 2
        results2[nb,nw] <- 0

        H[c(white_player             ,"white"       )] %<>% inc()
        H[c(white_player,black_player,"white","draw")] %<>% dec()

        if(nationality[nw]=="USSR" & nationality[nb]=="USSR"){
            H_coll[c(white_player             ,"white"       )] %<>% inc()
            H_coll[c(white_player,black_player,"white","collusion")] %<>% dec()
        } else {
            H_coll[c(white_player             ,"white"       )] %<>% inc()
            H_coll[c(white_player,black_player,"white","draw")] %<>% dec()
        }
    } else if(match_result == "0-1"){ # black wins
        results2[nw,nb] <- 0
        results2[nb,nw] <- 2
        H[c(             black_player               )] %<>% inc()
        H[c(white_player,black_player,"white","draw")] %<>% dec()

        if(nationality[nw]=="USSR" & nationality[nb]=="USSR"){
            H_coll[c(             black_player                    )] %<>% inc()
            H_coll[c(white_player,black_player,"white","collusion")] %<>% dec()
        } else {  # collusion not playing
            H_coll[c(             black_player               )] %<>% inc()
            H_coll[c(white_player,black_player,"white","draw")] %<>% dec()
        }

    } else if (match_result == "1/2-1/2"){ # draw
        results2[nw,nb] <- 1
        results2[nb,nw] <- 1
        H[c(                                  "draw")] %<>% inc()
        H[c(white_player,black_player,"white","draw")] %<>% dec()

        if(nationality[nw]=="USSR" & nationality[nb]=="USSR"){
            H_coll[c(                                  "collusion")] %<>% inc()
            H_coll[c(white_player,black_player,"white","collusion")] %<>% dec()
        } else {  # collusion not playing
            H_coll[c(                                  "draw")] %<>% inc()
            H_coll[c(white_player,black_player,"white","draw")] %<>% dec()
        }
    } else {
        stop("not possible")
    }
}  # i loop closes
```

So, `results` [from the square matrix] and `results2` [from the
3-column dataframe] should match:

```{r checksquareandtab}
stopifnot(all(results[!is.na(results)] == results2[!is.na(results2)]))
```

Now some maximum likelihood estimation.

```{r,hmaxpcalc,cache=TRUE}
H_maxp <- maxp(H)
H_maxp
```

```{r ml1,cache=TRUE}
small <- 0.01

maxlike_free <- maxp(H_coll,startp=c(small,small,small/2,rep(small,22)),give=TRUE)
print("l")
dput(maxlike_free$value)
```

```{r ml2,cache=TRUE}
jj <- maxlike_free$par
maxlike_free <- maxp(H_coll,startp=jj,give=TRUE)
dput(maxlike_free$value)
```

```{r}
H_coll_maxp <- maxlike_free$value
```

```{r dosometests,cache=TRUE}
samep.test(H_coll,c("draw","collusion"))
```

### Package dataset {-}

Following lines create `interzonal.rda`, residing in the `data/`
directory of the package.

```{r,label=saveinterzonaldataset}
interzonal <- H
interzonal_maxp <- H_maxp
collusion <- H_coll
collusion_maxp <- H_coll_maxp
save(interzonal,collusion,interzonal_maxp,collusion_maxp,
file="interzonal.rda")
```

# References

C. C. Moul and J. V. C. Nye 2009. "Did the Soviets collude? A
 statistical analysis of championship chess 1940-1978.  Journal of
 Economic Behaviour and Organization 70(2009) 10-21