### 2022-09-25 inst/surfing_jsa_rebuttal_minorcomments.tex
start work on #166

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

\section*{MS596 ``Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods": response to minor comments}

Below, we reproduce the two reviewers' minor comments; our replies to
the issues are in \textcolor{blue}{blue}.  We have indicated changes
to the manuscript where appropriate.

Short story: each of the  comments is perfectly reasonable and we have 
changed the manuscript as suggested.

\subsection*{placeholder comment}

Foo should be bar

\textcolor{blue}{generic placeholder response}

\end{document}