### 2022-10-18 inst/salad.Rmd
salad

---
title: "salad"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
library("prefmod")
library("PlackettLuce")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document shows how to apply `hyper2` idiom to the `salad`
dataset of the `prefmod` package.  From `salad.Rd`:

```
The dataset contains the rankings of four salad dressings concerning
tartness by 32 judges, with values ranging from 1 (most tart) to 4
(least tart).
```

```{r showsalad}
head(salad)
nrow(salad)
```

From row 3, for exampl,e we see that salad `A` was ranked as second
most tart, `B` the most tart, `C` the third most tart, and `D` the
least tart of the four.  We may process this using two methods,
explicit and slick.  First, explicit:

```{r useexplicitmethod,cache=TRUE}
H1 <- hyper2()
for(i in seq_len(nrow(salad))){
   H1 <-  H1 + ordervec2supp(as.matrix(salad)[i,])
}
H1
```

And second, slick:

```{r useslickmethod,cache=TRUE}
(H2 <- ordertable2supp(t(salad)))
```

Just to check:

```{r checkequalityofsalads}
H1 == H2
```

```{r equalsaladtart, cache=TRUE}
equalp.test(H1)
```

The null estimate agrees to six places of decimals with that
presented by Turner:

```{r usepladmm, cache=TRUE}
standardPL_PlackettLuce <- PlackettLuce(salad, npseudo = 0)
(p1 <- itempar(standardPL_PlackettLuce))
(p2 <- maxp(H1))
p1-p2
```

# References

Turner HL, van Etten J, Firth D, Kosmidis I (2020).  "Modelling
Rankings in R: The PlackettLuce Package."  _Computational Statistics_,
35, 1027–1057. doi:10.1007/s00180-020-00959-3,
\url{https://doi.org/10.1007/s00180-020-00959-3}

Yildiz, Ilkay, Jennifer Dy, Deniz Erdogmus, Jayashree Kalpathy-Cramer,
Susan Ostmo, J. Peter Campbell, Michael F. Chiang, and Stratis
Ioannidis. 2020. "Fast and Accurate Ranking Regression."  In
Proceedings of the Twenty Third International Conference on Artificial
Intelligence and Statistics, edited by Silvia Chiappa and Roberto
Calandra, 108:77–88. Proceedings of Machine Learning
Research. \url{http://proceedings.mlr.press/v108/yildiz20a.html}