### 2022-11-21 inst/ternaryplot_hyper2.Rmd
last part of issue #187

---
title: "ternary_hyper2"
author: "Robin Hankin"
date: "2022-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ternary plots for hyper2 samples

Here I show how to use the excellent `Ternary` package to generate triangular plots for `hyper2` likelihood functions with size 3.  First load the `Ternary` and `hyper2` packages:

```{r loadlibraries}
library("hyper2")
library("Ternary")
```

Now generate a synthetic dataset.  Here we will sample from the posterior of a uniform prior with a likelihood function specified by the `chess` object (this takes quite a long time):

```{r generatedata, cache=TRUE}
set.seed(0)
xx <- rp(1000,chess)
```

Using the `TernaryPlot()` and `TernaryPoints()` functions is straightforward:
```{r plotdata}
jj <- pnames(chess)
TernaryPlot(atip=jj[1],btip=jj[2],ctip=jj[3])
TernaryPoints(xx,cex=0.5)
TernaryPoints(chess_maxp,pch=16,col="red")
```

Above we see black circles being samples from the posterior and the red point is the maximum likelihood estimate for the three players' strengths.

We can adapt this technique to higher-dimensional `hyper2` objects.