### 2022-12-01 inst/skeleton.Rmd
nice analysis of skeleton

---
title: "skeleton"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
library("igraph")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

The [skeleton](https://en.wikipedia.org/wiki/Skeleton_(sport)) is a
winter sliding sport.  This short document discusses a dataset taken
from the [2018 Winter
Olympics](https://en.wikipedia.org/wiki/Skeleton_at_the_2018_Winter_Olympics_%E2%80%93_Men%27s)
and analyses the order statistic for the competitors.

```{r readskeletontable}
(skeleton_table <- read.table("skeleton2022.txt",header=TRUE))
```

Thus the top line shows that Grotheer came first in runs 1-3 and sixth
in run 4.  The first column shows the results for run 1: we see that
Grotheer came first, Jungk came fifth, and so on.  There is in fact no
dominated set: Timmings [place 25] beats Enzie [place 19] (in run 4),
Enzie [19] beats Crumpton [22] (run 1), Crumpton [17] beats Wyatt [18]
(2), Wyatt [16] beats Maier [15] (1), Maier [11] beats Bagnis [12]
(4), Bagnis [10] beasts Seunggi [11] (1), Seunggi [9] beats T. Dukurs
[10] (3 and 4), Dukurs [7] beats Tretyakov [8] (2), Tretyakov [2]
beats Wengang [3] (1), and Wengang beats everybody (4).

We can *try* to automate the above cycle as follows.  The first step
is to convert `skeleton_table` to a numeric matrix, which proves to be
a very awkward process

NB: `data.matrix()` fucks things up:

```{r,label=datamatrixfucksthingsup}
head(data.matrix(skeleton_table))
```

Above, we see that Grotheer apparently came 17th in run 4, clearly
some weird infelicity.  OK, moving swiftly on, can we coerce our
dataframe to a numeric matrix?  Yes, we can, but it's klunky:

```{r,showextremerank}
st <- matrix(as.numeric(c(as.matrix(skeleton_table))),ncol=4)
rownames(st) <- rownames(skeleton_table)
colnames(st) <- colnames(skeleton_table)
st
st[is.na(st)] <- 999
```

Taking Maier as an example (a mid-range competitor) we want to
identify competitors who beat Maier at least once, and competitors who
Maier beats at least once (these groups are not necessarily distinct)

```{r}
bm1 <- st[,1] < st["Maier",1]  # beats Maier in run 1
bm2 <- st[,2] < st["Maier",2]  # beats Maier in run 2
bm3 <- st[,3] < st["Maier",3]  # beats Maier in run 3
bm4 <- st[,4] < st["Maier",4]  # beats Maier in run 4
bm1 | bm2 | bm3 | bm4
st[bm1 | bm2 | bm3 | bm4,]
```

Above we see all competitors who beat `Maier` at least once.  Who is
the worst of these, as measured by his worst performance?  This would
be `Weston`, who ranked 14 in run 1, compared to `Maier` who ranked
15.  Idiomatically:

```{r}
jj <- st[bm1 | bm2 | bm3 | bm4,]
which(jj==max(jj),arr.ind=TRUE)
rownames(jj)[which(jj==max(jj),arr.ind=TRUE)[1]]
```

We can automate this process as follows:

```{r}
wf <- function(p){
 beats <- function(run){st[,run] < st[p,run]}
 jj <- st[apply(sapply(seq_len(ncol(st)),beats),1,any),]
 rownames(jj)[which(jj==max(jj),arr.ind=TRUE)[1]]
}
wf("Maier")
```

To check:

```{r}
wf("Zheng")
```

Above we see that `Dukurs_M` beats `Zheng` in run 7, but `Dukurs_M`
ranked only 13 in run 4.

```{r}
loop <- c("Grotheer")
for(i in 1:20){loop <- c(loop,wf(loop[length(loop)]))}
loop
```

Above we see an endless repetition of `Blaser` -> `Timmings` -> `Blaser`.

```{r}
loop <- c("Jungk")
for(i in 1:20){loop <- c(loop,wf(loop[length(loop)]))}
loop
```

```{r}
loop <- c("Wengang")
for(i in 1:20){loop <- c(loop,wf(loop[length(loop)]))}
loop
```

```{r}
forw <- function(start){
  loop <- start
  while(all(table(loop)==1)){ loop <- c(loop,wf(loop[length(loop)])) }
  return(loop)
}
forw("Wengang")
```

We can do the process in reverse:

```{r}
fw <- function(p){
 beatenby <- function(run){st[,run] > st[p,run]}
 jj <- st[apply(sapply(seq_len(ncol(st)),beatenby),1,any),]
 rownames(jj)[which(jj==min(jj),arr.ind=TRUE)[1]]
}
fw("Maier")
```

Above, we see that `Dukurs_M` is the best (ranking 3 in run 2) of
competitors who lose to `Maier` [11 vs 13 in run 4].

```{r}
back <- function(start){
  loop <- start
  while(all(table(loop)==1)){ loop <- c(loop,fw(loop[length(loop)])) }
  return(loop)
}
back("Wengang")
```

```{r}
gd <- graph(c(1,2, 2,3, 1,2, 2,4, 1,4, 5,5, 3,6))
plot(gd)
path <- function(a){c(t(cbind(a[-length(a)],a[-1])))}

gpath <- NULL
jjf <- sapply(rownames(st),forw)
jjb <- sapply(rownames(st),back)
for(i in seq_along(jjf)){gpath <- c(gpath,path(jjf[[i]]))}
for(i in seq_along(jjb)){gpath <- c(gpath,rev(path(jjb[[i]])))}
gpath <- substr(gpath,1,3)
plot(graph(gpath[1:20]))
plot(graph(gpath[1:40]))
plot(graph(gpath[1:80]))
```

# Rank table analysis {-}

We can present the result table in another way:

```{r showranktable}
wikitable_to_ranktable(skeleton_table)
```

Above we see, in the first column, that Grotheer came first ("`c1`")
in runs 1-3 and Wengang came first in run 4.  From the first column we
see that in run 1, Grotheer came fist, Tretyakov came second, Wengang
came third, and so on.

```{r,label=processskeleton,cache=TRUE}
skeleton <- ordertable2supp(skeleton_table)
skeleton_maxp <- maxp(skeleton)
skeleton_maxp
pie(skeleton_maxp)
```

And we can test the null that the competitors all have the same
Plackett-Luce strengths:

```{r,label=testskeleton,cache=TRUE}
equalp.test(skeleton)
```

We see very strong evidence for different competitor strengths, at
$p=2\times 10^{-33}$.  How strong is the evidence that Grotheer is
better than Jungk?

```{r,label=testgrotheerandjungk,cache=TRUE}
samep.test(skeleton,c("Grotheer","Jungk"))
```

### Package dataset {-}

Following lines create `skeleton.rda`, which is not part of the CRAN version of the package.

```{r,label=savecurlingdataset}
save(skeleton_table,skeleton,skeleton_maxp,file="skeleton.rda")
```