### 2023-01-05 inst/multiset_hyper2.Rmd
clarifications and rewording

---
title: "The `cheering3()` probability model for order statistics: a
verification with the `multiset()` function of the `partitions` package"
output: html_document
date: "2022-07-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("partitions")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here we will sum probabilities across a sample space and verify that
the total is constant.  In theory the sum of the probabilities of all
members of the sample space is 1, but in practice the likelihood
calculation includes an indeterminate constant which means that all we
can say is that the total probability is not a function of the
parameters of the distribution.

```{r defineM}
M <- apply(partitions::multiset(c(1,1,2,2,3)),2,function(x){letters[x]})
colnames(M) <- rep("",ncol(M))   # tidier and more compact appearance
M <- noquote(M)
M
```

Above, we use `multiset()` to enumerate the sample space of five
competitors comprising two twins denoted `a`, two twins denoted `b`,
and a singleton `c`.  This document discusses the issue of whether the
twins are identical twins and hence indistinguishable.  Matrix `M` has
${5\choose2\,2\,1}=30$ columns, one per possible result.  We will
calculate the likelihood of each column and sum them to get the
probability of observing any one of the 30 order statistics [the
observations are mutually exclusive, so
$\operatorname{Prob}\left(x_1\cup\cdots\cup x_n\right) =
\operatorname{Prob}\left(x_1\right)+\cdots
+\operatorname{Prob}\left(x_n\right)$]:

```{r definetfun}
tfun <- function(help,strength,M){
  out <- 0
  for(i in seq_len(ncol(M))){
     vec <- M[,i]
     out <- out +
       loglik(strength,cheering3(v=vec,e=c(a=1,b=2,c=3),help=help),log=FALSE)
  }
  return(out)
}
```

```{r showtisconstant,cache=TRUE}
print(tfun(help=(1:3)/10, strength= c(a=0.1,b=0.3,c=0.6),M))
print(tfun(help=(7:5)/22, strength= c(a=0.3,b=0.2,c=0.5),M))
```

Above we see that the sum of the likelihoods is constant.  The value
turns out to be $\frac{1}{1!2!2!}=0.25$.  This is because the
likelihood function (being Plackett-Luce) considers the observation to
be one of the $5!=120$ ways of ordering five distinct competitors.
Matrix `M` has only $\frac{5!}{2!2!1!}=30$ columns, each one of which
corresponds to four different observations, all of which have the same
probability.  Much of the confusion arises from the practice of
identifying a competitor with his strength; we write "$p_a$'' [or
sometimes just "$a$''] to signify both a particular competitor, and
his Bradley-Terry strength.

To explain the factor of $1/4$ discussed above we need some new
notation.  We have five competitors who I will call $\mathbf{do}$,
$\mathbf{re}$, $\mathbf{me}$, $\mathbf{fa}$, and $\mathbf{sol}$.  Now
$p_X$ means "the Bradley Terry strength of $X$''.  The setup here is
that $\mathbf{do}$ and $\mathbf{re}$ are twins, and that $\mathbf{me}$
and $\mathbf{fa}$ are twins (and $\mathbf{sol}$ is a singleton).  We
have that the Bradley-Terry strength of $\mathbf{do}$ and
$\mathbf{re}$ is `a`, the strength of $\mathbf{me}$ and $\mathbf{fa}$
is `b`, and the strength of $\mathbf{sol}$ is `c`.

Consider, for example, the column of `M` reading `a c b b a`.  In
theory, this means the competitor who came first had strength `a`, the
competitor who came second had strength `c`.  So the order statistic
is one of the $2!2!1!=4$ following orders:

$$\mathbf{do}\succ\mathbf{sol}\succ\mathbf{me}\succ\mathbf{fa}\succ\mathbf{re}$$
$$\mathbf{do}\succ\mathbf{sol}\succ\mathbf{fa}\succ\mathbf{me}\succ\mathbf{re}$$
$$\mathbf{re}\succ\mathbf{sol}\succ\mathbf{me}\succ\mathbf{fa}\succ\mathbf{do}$$
$$\mathbf{re}\succ\mathbf{sol}\succ\mathbf{fa}\succ\mathbf{me}\succ\mathbf{do}$$

Above we see that $\mathbf{do}$ and $\mathbf{re}$ come first and last
(or last and first), and $\mathbf{me}$ and $\mathbf{fa}$ come third
and fourth (or fourth and third), while $\mathbf{sol}$ comes second.
All four possibilities have the same probability and all four are
admissible order statistics for the five solfeggio syllables.

But the `hyper2` package [specifically the `loglik()` and
`cheering3()` functions] consider `a c b b a` to mean something
different.  It considers it to be an observation of

$$\mathbf{do}\succ\mathbf{sol}\succ\mathbf{me}\succ\mathbf{fa}\succ\mathbf{re}$$

followed by an identification of $\mathbf{do}$ with (BT strength) `a`,
$\mathbf{sol}$ with strength `c`, and so on.  Any of the four
possibilities above would be equally admissible, and they all have the
same probability.  So to get "real'' probabilities [real in the sense
that summing the probabilities over the columns in matrix `M`], we
have to multiply by four to include all the admissible order
statistics.

The difference between these two interpretations does not matter from
a likelihood perspective as likelihood is defined only up to a
multiplicative constant.

## A more exacting numerical test

```{r,exactingtest,cache=TRUE}
rs <- rp(100,dirichlet(c(a=1,b=1,c=1))) # random strengths
colnames(rs) <- letters[1:3]
head(rs)
f <- function(s){
  jj <- s
  names(jj) <- letters[1:3]
  tfun(help=1:3,strength=jj,M=M)
}
x <- apply(rs,1,f)
```

```{r,plotdifferences}
hist(x-mean(x))
```

Above we see very tight clustering about the mean value of 0.25