### 2023-01-05 inst/salad.Rmd
use samep.test()

---
title: "salad"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
library("prefmod")
library("PlackettLuce")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document shows how to apply `hyper2` idiom to the `salad`
dataset of the `prefmod` package.  From `salad.Rd`:

```
The dataset contains the rankings of four salad dressings concerning
tartness by 32 judges, with values ranging from 1 (most tart) to 4
(least tart).
```

```{r showsalad}
head(salad)
nrow(salad)
```

From row 3, for example, we see that salad `A` was ranked as second
most tart, `B` the most tart, `C` the third most tart, and `D` the
least tart of the four.  We may process this using two methods,
explicit and slick.  First, explicit:

```{r useexplicitmethod,cache=TRUE}
H1 <- hyper2()
for(i in seq_len(nrow(salad))){
   H1 <-  H1 + ordervec2supp(as.matrix(salad)[i,])
}
H1
```

And second, a slick method.  We take the transpose of `salad` (things
are clearer if we name the rows):

```{r}
rownames(salad) <-
   paste("j",formatC(seq_len(nrow(salad)),width=2,format="d",flag="0"),sep="")
ts <- t(salad)
ts
```

Above, we see that object `ts` is an order table.  It is just like the
formula 1 result table `F1_table_2016`, except that venues are judges
and drivers are the salads [which are "competing" for the "who has the
most tart dressing" prize].  Converting this to a support function is
accomplished with `ordertable2supp()`:

```{r useslickmethod,cache=TRUE}
(H2 <- ordertable2supp(ts))
```

Just to check:

```{r checkequalityofsalads}
H1 == H2
```

```{r equalsaladtart, cache=TRUE}
equalp.test(H1)
```

The null estimate agrees to six places of decimals with that
presented by Turner:

```{r usepladmm, cache=TRUE}
standardPL_PlackettLuce <- PlackettLuce(salad, npseudo = 0)
(p1 <- itempar(standardPL_PlackettLuce))
(p2 <- maxp(H1))
p1-p2
```

We may use `hyper2` to test some hypotheses.  Firstly, $H_0\colon
p_B=\frac{1}{4}$, which we test with `specificp.test()`:

```{r specificptestB,cache=TRUE}
specificp.test(H2,"B",1/4)
```

Above we see a $p$-value of about $10^{-7}$, clearly significant.
However, we may also test the hypothesis that $p_B$ is less than the
second strongest, $p_C$:

```{r sameptestBC,cache=TRUE}
samep.test(H2,c("B","C"))
```

We see a $p$-value of about $7\times 10^{-4}$, much higher than the
value from `specificp.test()` of $H_0\colon p_B=\frac{1}{4}$.  This is
because `samep.test()` allows one to change the value of $p_C$ and we
would expect a less significant result.

# References

Turner HL, van Etten J, Firth D, Kosmidis I (2020).  "Modelling
Rankings in R: The PlackettLuce Package."  _Computational Statistics_,
35, 1027–1057. doi:10.1007/s00180-020-00959-3,
\url{https://doi.org/10.1007/s00180-020-00959-3}

Yildiz, Ilkay, Jennifer Dy, Deniz Erdogmus, Jayashree Kalpathy-Cramer,
Susan Ostmo, J. Peter Campbell, Michael F. Chiang, and Stratis
Ioannidis. 2020. "Fast and Accurate Ranking Regression."  In
Proceedings of the Twenty Third International Conference on Artificial
Intelligence and Statistics, edited by Silvia Chiappa and Roberto
Calandra, 108:77–88. Proceedings of Machine Learning
Research. \url{http://proceedings.mlr.press/v108/yildiz20a.html}