### 2023-01-08 inst/plackett_luce_monster.Rmd

---
title: "A generalization of Plackett-Luce likelihoods to incorporate non-independence of competitors"
output: html_document
date: '2022-05-16'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("partitions")
library("pdftools")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here I modify standard Plackett-Luce likelihood functions to include
an element of non-independence among the competitors.  There seem to
be a few distinct statistical models for nonindependence:

* Runner $i$ might wish to support $j$.  If runner $i$ finishes, once
  his rank is established he might be able to help $j$ in some way.
  He might cheer his still-running friend, spurring him to run faster
  and augmenting his competitive strength.

* In general, we can consider $n$ ordered pairs $(i_r,j_r)$, $1\leq
  r\leq n$, with runner $i_1$ helping runner $j_1$, runner $i_2$
  helping $j_2$, and so on.  One simple case would be equivalence
  classes of $[n]=\{1,2,\ldots,n\}$ with each equivalence class
  comprising mutually supporting runners.  In this model, with a pair
  of mutually supporting runners only the higher-placed competitor can
  affect the other.  Each equivalence class might have a different
  supportive proclivity.

* Runner $i$ might hate runner $j$.  If $i$ finishes, he might expend
  his spare energy hindering $j$ (by throwing stones or booing, for
  example).  Note that equivalence classes of mutual hatred might
  exist.

* Also, the help or hindrance might have a finite lifetime.  Under one
  model, if $i$ finishes then he can stay helping (or hindering) $j$
  wherever $j$ places.  Alternatively, suppose $i$ places $n$th.  Then
  perhaps he can affect $j$ only place $n+1$, and after a competitor
  finishes in place $n+1$, then $i$ can exert no further influence.

* The help or hindrance discussed above might be contingent on
  placing: perhaps only the winnner (that is, first-placing
  competitor) would be able to influence other competitors.  Or maybe
  competitors with any placing could influence other competitors.

* Runners $i$ and $j$ might elect to cross the finishing line holding
  hands, with the apparent tie broken randomly.  This would be a
  symmetrical relationship.  In general, we could consider
  hand-holding equivalence classes.

* We might have a resource allocation problem, with runners having a
  certain [possibly unknown] amount of "strength" but also the ability
  to expend this strength on objectives other than ranking in the
  race.  A runner might be able to expend just enough energy to remain
  slightly ahead of a rival, reserving energy for other purposes.  If
  runner $i$ wishes to beat runner $j$, and has sufficient resources
  to do so, then runners $i$ and $j$ will finish consecutively with
  $i$ ahead of $j$.

# Start simple: three runners

The simplest nontrivial case is three runners, $\left\lbrace
1,2,3\right\rbrace$ where runner 1 helps runner 2 with multiplier
$\lambda$ [and that is the only non PL interaction].  Note that this
is _not_ a symmetric interaction: 1 helps 2, but 2 does not help 1.
The model is that 1 helps 2 if and only if 1 has finished and 2 has
not.

123: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{\lambda p_2}{\lambda p_2+p_3}$

132: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{p_3}{\lambda p_2+p_3}$

213: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{p_1}{p_1+p_3}$

231: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{p_3}{p_1+p_3}$

312: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_1}{p_1+p_2}$

321: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_2}{p_1+p_2}$

(Above, the probabilities clearly sum to 1).  Next simplest: 1 and 2
mutually support each other with $\lambda$:

123: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{\lambda p_2}{\lambda p_2+p_3}$

132: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{p_3}{\lambda p_2+p_3}$

213: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{\lambda p_1}{p_1+p_3}$

231: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{p_3}{\lambda p_1+p_3}$

312: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_1}{p_1+p_2}$

321: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_2}{p_1+p_2}$

Now we retain the mutual support between 1 and 2 but allow an
asymmetry: 1 helps 2 with $\lambda$ as before, but now in addition, 2
helps 1 with $\mu$:

123: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{\lambda p_2}{\lambda p_2+p_3}$

132: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{p_3}{\lambda p_2+p_3}$

213: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{\mu p_1}{\mu p_1+p_3}$

231: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{p_3}{\mu p_1+p_3}$

312: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_1}{p_1+p_2}$

321: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_2}{p_1+p_2}$

Now how about 1 must beat 2.  He does not care whether he comes first
or second, only that he must beat 2.  Player 1 has sufficient energy
to win if he wants, but is subject to some sort of resource constraint
(money, perhaps) and wishes to expend only enough money to beat 2.  So
conventional P-L likelihoods are modified for 1, but only if 2 is
_still running_.

123: $\frac{\lambda p_1}{\lambda p_1+p_2+p_3}\cdot\frac{p_2}{p_2+p_3}$

132: $\frac{\lambda p_1}{\lambda p_1+p_2+p_3}\cdot\frac{p_3}{p_2+p_3}$

213: $\frac{p_2}{\lambda p_1+p_2+p_3}\cdot\frac{p_1}{p_1+p_3}$

231: $\frac{p_2}{\lambda p_1+p_2+p_3}\cdot\frac{p_3}{p_1+p_3}$

312: $\frac{p_3}{\lambda p_1+p_2+p_3}\cdot\frac{\lambda p_1}{\lambda p_1+p_2}$

321: $\frac{p_3}{\lambda p_1+p_2+p_3}\cdot\frac{p_2}{\lambda p_1+p_2}$

## Four runners

Now we have four runners $\left\lbrace 1,2,3,4\right\rbrace$, again
with the only interaction being runner 1 helping runner 2.  Thus, in a
situation where $p_1$ has finished and $p_2$ is still running, we have
$p_2\longrightarrow\lambda p_2$ [although the $\lambda$ cancels if 2
is the only competitor still running, that is, if *all* the remaining
competitors are helped]

1234: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_3+p_4}\cdot\frac{p_3}{p_3+p_4}$

1243: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_3+p_4}\cdot\frac{p_4}{p_3+p_4}$

1324: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_3}{\lambda p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_4}$

1342: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_3}{\lambda p_2+p_3+p_4}\cdot\frac{p_4}{\lambda p_2+p_4}$

1324: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_3}{\lambda p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_4}$

1423: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_4}{p_1+\lambda p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_3}$

1432: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_4}{p_1+\lambda p_2+p_3+p_4}\cdot\frac{p_3}{\lambda p_2+p_3}$

2xxx: P-L [no help given by 2 to 1]

3124: $\frac{p_3}{p_1+p_2+p_3+p_4}\cdot\frac{p_1}{p_1+p_2+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_4}$

3142: $\frac{p_3}{p_1+p_2+p_3+p_4}\cdot\frac{p_1}{p_1+p_2+p_4}\cdot\frac{p_4}{\lambda p_2+p_4}$

32xx: P-L

3412: P-L [1 helps 2 but when 1 crosses the finishing line, only 2 is
still running; there are no non-helped competitors]

4123: $\frac{p_4}{p_1+p_2+p_3+p_4}\cdot\frac{p_1}{p_1+p_2+p_3}\cdot\frac{\lambda p_2}{\lambda p_2+p_3}$

4132: $\frac{p_4}{p_1+p_2+p_3+p_4}\cdot\frac{p_1}{p_1+p_2+p_3}\cdot\frac{p_3}{\lambda p_2+p_3}$

42xx: P-L

43xx: P-L

Now 1 and 2 mutually support one another by $\lambda$.  Here I will
put the last term in, to show explicitly that the $\lambda$ term
remains in the expressions for numerator and denominator.

1234: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda
p_2+p_3+p_4}\cdot\frac{p_3}{p_3+p_4}\cdot\frac{p_4}{p_4}$

1243: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda
p_2+p_3+p_4}\cdot\frac{p_4}{p_3+p_4}\cdot\frac{p_3}{p_3}$

1324: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_3}{\lambda
p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda
p_2+p_4}\cdot\frac{p_4}{p_4}$

1342: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_3}{\lambda
p_2+p_3+p_4}\cdot\frac{p_4}{\lambda p_2+p_4}\cdot\frac{\lambda
p_2}{\lambda p_2}$

1423: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_4}{\lambda
p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda
p_2+p_3}\cdot\frac{p_3}{p_3}$

1432: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_4}{\lambda
p_2+p_3+p_4}\cdot\frac{p_3}{\lambda p_2+p_3}\cdot\frac{\lambda
p_2}{\lambda p_2}$

2134: $\frac{p_2}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_1}{\lambda
p_1+p_3+p_4}\cdot\frac{p_3}{p_3+p_4}\cdot\frac{p_4}{p_4}$

2143: $\frac{p_2}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_1}{\lambda
p_1+p_3+p_4}\cdot\frac{p_4}{p_3+p_4}\cdot\frac{p_3}{p_3}$

2314: $\frac{p_2}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_3}{\lambda
p_1+p_3+p_4}\cdot\frac{\lambda p_1}{\lambda
p_1+p_4}\cdot\frac{p_4}{p_4}$

2341: $\frac{p_2}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_3}{\lambda
p_1+p_3+p_4}\cdot\frac{p_4}{\lambda p_1+p_4}\cdot\frac{\lambda
p_1}{\lambda p_1}$

We will calculate a Plackett-Luce likelihood function for four runners
$\left\lbrace 1,2,3,4\right\rbrace$ but with a reified Bradley-Terry
monster that corresponds to having subsets of mutual friends who
support one another.  For example, we might have runners 1 and 2 being
friends, and runners 3 and 4 being friends: we would represent this as
$\left\lbrace\left\lbrace 1,2\right\rbrace,\left\lbrace
3,4\right\rbrace\right\rbrace$.

The runner metaphor is not the only one.  We can apply the resulting
likelihood functions to other cases including the red bus-blue bus
problem (see `inst/red_bus_blue_bus.Rmd` for more discussion).

First do three runners, and suppose the equivalence classes are
$\left\lbrace\left\lbrace 1,2\right\rbrace,\left\lbrace
3\right\rbrace\right\rbrace$.

 123: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{p_2+M}{p_2+p_3+M}\qquad\frac{p_1}{p_1+p_2+p_3}\cdot\frac{\lambda p_2}{\lambda p_2+p_3}$

 132: $\frac{p_1}{p_1+p_3+p_2}\cdot\frac{p_3}{p_3+p_2+M}\qquad\frac{p_1}{p_1+p_3+p_2}\cdot\frac{p_3}{p_3+\lambda p_2}$

 213: $\frac{p_2}{p_2+p_1+p_3}\cdot\frac{p_1+M}{p_1+p_3+M}\qquad\frac{p_2}{p_2+p_1+p_3}\cdot\frac{\lambda p_1}{\lambda p_1+p_3}$

 231: $\frac{p_2}{p_2+p_3+p_1}\cdot\frac{p_3}{p_3+p_1+M}\qquad\frac{p_2}{p_2+p_1+p_3}\cdot\frac{\lambda p_1}{p_3+\lambda p_1}$

 312: $\frac{p_3}{p_3+p_1+p_2}\cdot\frac{p_1}{p_2+p_1}\qquad\frac{p_3}{p_3+p_1+p_2}\cdot\frac{\lambda p_1}{p_3+p_1}$

 321: $\frac{p_3}{p_3+p_2+p_1}\cdot\frac{p_2}{p_2+p_1}\qquad\frac{p_3}{p_3+p_2+p_1}\cdot\frac{p_2}{p_2+p_1}$

Now four runners with classes $\left\lbrace\left\lbrace
1,2\right\rbrace,\left\lbrace 3,4\right\rbrace\right\rbrace$.

 1234: $\frac{1}{1+2+3+4}\cdot\frac{2+M}{2+3+4+M}\cdot\frac{3}{3+4}$

 1243: $\frac{1}{1+2+4+3}\cdot\frac{2+M}{2+4+3+M}\cdot\frac{4}{4+3}$

 1324: $\frac{1}{1+3+2+4}\cdot\frac{3}{3+2+4+M}\cdot\frac{2}{2+4+M}$

 1342: $\frac{1}{1+3+4+2}\cdot\frac{3}{3+4+2+M}\cdot\frac{4+M}{2+4+M}$

 1423: $\frac{1}{1+4+2+3}\cdot\frac{4}{4+2+3+M}\cdot\frac{2}{2+3}$

 1432: $\frac{1}{1+4+3+2}\cdot\frac{4}{4+3+2+M}\cdot\frac{3}{2+3}$

 2134: $\frac{2}{2+1+3+4}\cdot\frac{1+M}{1+3+4+M}\cdot\frac{3}{3+4}$

 2143: $\frac{2}{2+1+4+3}\cdot\frac{1+M}{1+4+3+M}\cdot\frac{4}{4+3}$

 2314: $\frac{2}{2+3+1+4}\cdot\frac{3}{3+1+4+M}\cdot\frac{1}{1+4+M}$

 2341: $\frac{2}{2+3+4+1}\cdot\frac{3}{3+4+1+M}\cdot\frac{4+M}{1+4+M}$

 2413: $\frac{2}{2+4+1+3}\cdot\frac{4}{4+1+3+M}\cdot\frac{1}{1+3}$

 2431: $\frac{2}{2+4+3+1}\cdot\frac{4}{4+3+1+M}\cdot\frac{3}{1+3}$

etc.

Now 123 tethered, 45 tethered:

12345: $\frac{1}{1+2+3+4+5}\cdot\frac{2+M}{2+3+4+5+M}\cdot\frac{3+M}{3+4+5+M}\cdot\frac{4}{4+5}$

12354: $\frac{1}{1+2+3+5+4}\cdot\frac{2+M}{2+3+5+4+M}\cdot\frac{3+M}{3+5+4+M}\cdot\frac{5}{5+4}$

12435: $\frac{1}{1+2+3+5+4}\cdot\frac{2+M}{2+3+5+4+M}\cdot\frac{4}{3+5+4+M}\cdot\frac{3}{3+5+M}$

12453: $\frac{1}{1+2+3+5+4}\cdot\frac{2+M}{2+3+5+4+M}\cdot\frac{4}{3+5+4+M}\cdot\frac{5+M}{5+3+M}$

12534: $\frac{1}{1+2+3+5+4}\cdot\frac{2+M}{2+5+3+4+M}\cdot\frac{5}{3+5+4+M}\cdot\frac{3}{3+4+M}$

12543: $\frac{1}{1+2+3+5+4}\cdot\frac{2+M}{2+5+4+3+M}\cdot\frac{5}{3+5+4+M}\cdot\frac{4+M}{4+3+M}$

Might be better to consider ${5\choose 2\,3}$ arrangements of
$aaabb\ldots bbaaa$.  FOLLOWING IS NOT RIGHT: THE PROBABILITY MODEL IS
NOT PRECISELY DEFINED.

aaabb: $\frac{a}{a+a+a+b+b}\cdot\frac{a+M}{a+a+b+b+M}\cdot\frac{a+M}{a+b+b+M}\cdot\frac{b}{b+b+M}$

aabab: $\frac{a}{a+a+a+b+b}\cdot\frac{a+M}{a+a+b+b+M}\cdot\frac{b}{a+b+b+M}\cdot\frac{a}{a+b+M}$

aabba: $\frac{a}{a+a+a+b+b}\cdot\frac{a+M}{a+a+b+b+M}\cdot\frac{b}{a+b+b+M}\cdot\frac{b+M}{b+a+M}$

abaab: $\frac{a}{a+a+a+b+b}\cdot\frac{b}{a+a+b+b+M}\cdot\frac{a}{a+a+b+M}\cdot\frac{a+M}{a+b+M}$

ababa: $\frac{a}{a+a+a+b+b}\cdot\frac{b}{a+a+b+b+M}\cdot\frac{a}{a+a+b+M}\cdot\frac{b}{b+a+M}$

abbaa: $\frac{a}{a+a+a+b+b}\cdot\frac{b}{a+a+b+b+M}\cdot\frac{b+M}{b+a+a+M}\cdot\frac{a}{a+a+M}$

baaab: $\frac{b}{a+a+a+b+b}\cdot\frac{a}{a+a+a+b+M}\cdot\frac{a+M}{a+a+b+M}\cdot\frac{a+M}{a+b+M}$

baaba: $\frac{b}{a+a+a+b+b}\cdot\frac{a}{a+a+a+b+M}\cdot\frac{a+M}{a+a+b+M}\cdot\frac{b}{a+b+M}$

babaa: $\frac{b}{a+a+a+b+b}\cdot\frac{a}{a+a+a+b+M}\cdot\frac{b}{b+a+a+M}\cdot\frac{a}{a+a}$

bbaaa: $\frac{b}{a+a+a+b+b}\cdot\frac{b+M}{a+a+a+b+M}\cdot\frac{a}{a+a+a}\cdot\frac{a}{a+a}$

```{r,label=definefunctionf}
f <- function (a,b,M){
return(
  factorial(3)*factorial(2)*(
   a/(a+a+a+b+b)*(a+M)/(a+a+b+b+M)*(a+M)/(a+b+b+M)*(b  )/(b+b+M) +
   a/(a+a+a+b+b)*(a+M)/(a+a+b+b+M)*(b  )/(a+b+b+M)*(a  )/(a+b+M) +
   a/(a+a+a+b+b)*(a+M)/(a+a+b+b+M)*(b  )/(a+b+b+M)*(b+M)/(b+a+M) +
   a/(a+a+a+b+b)*(b  )/(a+a+b+b+M)*(a  )/(a+a+b+M)*(a+M)/(a+b+M) +
   a/(a+a+a+b+b)*(b  )/(a+a+b+b+M)*(a  )/(a+a+b+M)*(b  )/(b+a+M) +
   a/(a+a+a+b+b)*(b  )/(a+a+b+b+M)*(b+M)/(b+a+a+M)*(a  )/(a+a+M) +
   b/(a+a+a+b+b)*(a  )/(a+a+a+b+M)*(a+M)/(a+a+b+M)*(a+M)/(a+b+M) +
   b/(a+a+a+b+b)*(a  )/(a+a+a+b+M)*(a+M)/(a+a+b+M)*(b  )/(a+b+M) +
   b/(a+a+a+b+b)*(a  )/(a+a+a+b+M)*(b  )/(b+a+a  )*(a  )/(a+a  ) +
   b/(a+a+a+b+b)*(b+M)/(a+a+a+b+M)*(a  )/(a+a+a  )*(a  )/(a+a  )
   ) 
   )
}
```

```{r}
f(4.4,3.3, 0.2)*12
f(4.4,3.3, 0.0)*12
```

```{tikz, fig.cap = "Partial probability tree for five competitors $a-e$ with $a$ supporting $b$, hyper3 approach", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2,3); 
\draw (0,0) -- (2,2);
\draw (0,0) -- (2,1);
\draw (0,0) -- (2,0);
\draw (0,0) -- (2,-1);
\node at (0,2.5) (eq1) {$a\longrightarrow b$};

\node at (1.5,2.5) {$a$};
\node at (1.5,1.7) {$b$};
\node at (1.5,1.0) {$c$};
\node at (1.5,0.2) {$d$};
\node at (1.5,-0.5) {$e$};

\node at (3, 3)  {$\frac{a}{a+b+c+d+e}$};
\node at (3, 2)  {$\frac{b}{a+b+c+d+e}$};
\node at (3, 1)  {$\frac{c}{a+b+c+d+e}$};
\node at (3, 0)  {$\frac{d}{a+b+c+d+e}$};
\node at (3,-1)  {$\frac{e}{a+b+c+d+e}$};

\fill  (4, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill  (4, 2) circle[radius=2pt];  % terminal node
\fill  (4, 1) circle[radius=2pt];  % terminal node
\fill  (4, 0) circle[radius=2pt];  % d finishes; paths abce
\fill  (4,-1) circle[radius=2pt];  % terminal node

\draw (4,3) -- (5,5); 
\draw (4,3) -- (5,4); 
\draw (4,3) -- (5,3); 
\draw (4,3) -- (5,2);

\node at (4.5,4.6) {$b$};
\node at (4.5,3.7) {$c$};
\node at (4.5,3.2) {$d$};
\node at (4.5,2.7) {$e$};

\newcommand{\lb}{\lambda b }
\newcommand{\lc}{\lambda c }
\newcommand{\lee}{\lambda e }
\node at (6, 5)  {$\frac{\lb}{\lb+c+d+e}$};
\node at (6, 4)  {$\frac{c}{\lb+c+d+e}$};
\node at (6, 3)  {$\frac{d  }{\lb+c+d+e}$};
\node at (6, 2)  {$\frac{e  }{\lb+c+d+e}$};

\fill (7.4, 5) circle[radius=2pt];  % ab finishes; paths cde
\draw (7.4,5) -- (8.4,6);
\draw (7.4,5) -- (8.4,5);
\draw (7.4,5) -- (8.4,4);

\node at (7.9,5.7) {$c$};
\node at (7.9,5.2) {$d$};
\node at (7.9,4.7) {$e$};
\node at (9.3, 6) {$\frac{c}{c+d+e}$};
\node at (9.3, 5) {$\frac{d  }{c+d+e}$};
\node at (9.3, 4) {$\frac{e  }{c+d+e}$};

\fill (10.2, 6) circle[radius=2pt];  % abc finishes; paths de
\draw (10.2, 6) -- (11.2,7);
\node at (10.7,6.8) {$d$};

\node at (11.8, 7) {$\frac{d}{d+e}$};

\draw (10.2, 6) -- (11.2,6.2);
\node at (11.8, 6.2) {$\frac{e}{d+e}$};
\node at (10.7,6.3) {$e$};

\fill (10.2, 5) circle[radius=2pt];
\draw (10.2, 5) -- (11.2,5.3);
\node at (11.8, 5.3) {$\frac{c}{c+e}$};
\node at (10.7,5.3) {$c$};

\draw (10.2, 5) -- (11.2,4.7);
\node at (11.8, 4.7) {$\frac{e}{c+e}$};
\node at (10.7,4.7) {$e$};

\fill (10.2, 4) circle[radius=2pt];
\draw (10.2, 4) -- (11.2,3.9);
\node at (11.8, 3.9) {$\frac{c}{c+e}$};
\node at (10.7,4.1) {$c$};

\draw (10.2, 4) -- (11.2,3.1);
\node at (11.8, 3.1) {$\frac{e}{c+e}$};
\node at (10.7,3.7) {$e$};

\draw (4, 0) -- (5,1);
\node at (6, 1)  {$\frac{a}{a+b+c+e}$};

\draw (4, 0) -- (5,0);
\node at (6, 0)  {$\frac{b}{a+b+c+e}$};

\draw (4, 0) -- (5,-1);
\node at (6, -1)  {$\frac{c}{a+b+c+e}$};

\draw (4, 0) -- (5,-2);
\node at (6, -2)  {$\frac{e}{a+b+c+e}$};

\node at (4.5,0.7) {$a$};
\node at (4.5,0.2) {$b$};
\node at (4.5,-0.3) {$c$};
\node at (4.5,-0.8) {$e$};

\fill (7.4, -2) circle[radius=2pt];  % de finishes; paths abc
\draw (7.4, -2) -- (8.4,-1);
\draw (7.4, -2) -- (8.4,-2);
\draw (7.4, -2) -- (8.4,-3);

\node at (7.9,-1.2) {$a$};
\node at (7.9,-1.8) {$b$};
\node at (7.9,-2.3) {$c$};

\node at (9.3, -1) {$\frac{a}{a+b+c}$};
\node at (9.3, -2) {$\frac{b}{a+b+c}$};
\node at (9.3, -3) {$\frac{c}{a+b+c}$};

\fill (10.2, -1) circle[radius=2pt];  % dea finishes; paths bc
\fill (10.2, -2) circle[radius=2pt];  % deb finishes; paths ac
\fill (10.2, -3) circle[radius=2pt];  % dec finishes; paths ab

\draw (10.2, -1) -- (11.2,-0);
\node at (10.6,-0.4) {$b$};
\node at (10.6,-0.8) {$c$};

\node at (11.8, 0) {$\frac{\lb}{\lb+c}$};
\draw (10.2, -1) -- (11.2,-1);

\node at (11.8, -1) {$\frac{c}{\lb+c}$};

\draw (10.2, -2) -- (11.2,-1.6);
\node at (11.8, -1.6) {$\frac{a}{a+c}$};

\node at (10.6,-1.7) {$a$};
\node at (10.6,-2) {$c$};

\draw (10.2, -2) -- (11.2,-2.2);
\node at (11.8, -2.2) {$\frac{c}{a+c}$};

\draw (10.2, -3) -- (11.2,-2.9);
\node at (11.8, -2.9) {$\frac{a}{a+b}$};

\draw (10.2, -3) -- (11.2,-3.4);
\node at (11.8, -3.4) {$\frac{b}{a+b}$};

\node at (10.6,-2.8) {$a$};
\node at (10.6,-3.3) {$b$};

\fill (7.4, 1) circle[radius=2pt];  % da finishes; paths bce

\draw (7.4, 1) -- (8.4,2);
\draw (7.4, 1) -- (8.4,1);
\draw (7.4, 1) -- (8.4,0);

\node at (7.9,1.7) {$b$};
\node at (7.9,1.2) {$c$};
\node at (7.9,0.7) {$e$};

\node at (9.3, 2) {$\frac{\lb}{\lb+c+e}$};
\node at (9.3, 1) {$\frac{  c}{\lb+c+e}$};
\node at (9.3, 0) {$\frac{e  }{\lb+c+e}$};

\fill (10.2, 1) circle[radius=2pt];  % da finishes; paths bce
\draw (10.2, 1) -- (11.2,2);
\node at (11.8, 2) {$\frac{\lb}{\lb+e}$};
\draw (10.2, 1) -- (11.2,1);
\node at (11.8, 1) {$\frac{e}{\lb+e}$};

\end{tikzpicture}
```

```{tikz, fig.cap = "Partial probability tree for five competitors $a-e$ with mutually supporting  group $abc$, hyper3 approach.  If any competitor in the set {$a,b,c$} finishes, they lend their strength to the other competitors in the set who are still running", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2,3); 
\draw (0,0) -- (2,2);
\draw (0,0) -- (2,1);
\draw (0,0) -- (2,0);
\draw (0,0) -- (2,-1);
\node at (0,2.5) (eq1) {$\underbrace{\left\lbrace a,b,c\right\rbrace}_{\lambda}$};

\node at (1.5,2.5) {$a$};
\node at (1.5,1.7) {$b$};
\node at (1.5,1.0) {$c$};
\node at (1.5,0.2) {$d$};
\node at (1.5,-0.5) {$e$};

\node at (3, 3)  {$\frac{a}{a+b+c+d+e}$};
\node at (3, 2)  {$\frac{b}{a+b+c+d+e}$};
\node at (3, 1)  {$\frac{c}{a+b+c+d+e}$};
\node at (3, 0)  {$\frac{d}{a+b+c+d+e}$};
\node at (3,-1)  {$\frac{e}{a+b+c+d+e}$};

\fill  (4, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill  (4, 2) circle[radius=2pt];  % terminal node
\fill  (4, 1) circle[radius=2pt];  % terminal node
\fill  (4, 0) circle[radius=2pt];  % d finishes; paths abce
\fill  (4,-1) circle[radius=2pt];  % terminal node

\draw (4,3) -- (5,5); 
\draw (4,3) -- (5,4); 
\draw (4,3) -- (5,3); 
\draw (4,3) -- (5,2);

\node at (4.5,4.6) {$b$};
\node at (4.5,3.7) {$c$};
\node at (4.5,3.2) {$d$};
\node at (4.5,2.7) {$e$};

\newcommand{\la}{\lambda a }
\newcommand{\lb}{\lambda b }
\newcommand{\lc}{\lambda c }
\newcommand{\lee}{\lambda e }
 \node at (6, 5)  {$\frac{\lb}{\lb+\lc+d+e}$};
 \node at (6, 4)  {$\frac{\lc}{\lb+\lc+d+e}$};
 \node at (6, 3)  {$\frac{d  }{\lb+\lc+d+e}$};
 \node at (6, 2)  {$\frac{e  }{\lb+\lc+d+e}$};

 \fill (7.4, 5) circle[radius=2pt];  % ab finishes; paths cde
 \draw (7.4,5) -- (8.4,6);
 \draw (7.4,5) -- (8.4,5);
 \draw (7.4,5) -- (8.4,4);

 \node at (7.9,5.7) {$c$};
 \node at (7.9,5.2) {$d$};
 \node at (7.9,4.7) {$e$};
 \node at (9.3, 6) {$\frac{\lc}{\lc+d+e}$};
 \node at (9.3, 5) {$\frac{d  }{\lc+d+e}$};
 \node at (9.3, 4) {$\frac{e  }{\lc+d+e}$};

 \fill (10.2, 6) circle[radius=2pt];  % abc finishes; paths de
 \draw (10.2, 6) -- (11.2,7);
 \node at (10.7,6.8) {$d$};

 \node at (11.8, 7) {$\frac{d}{d+e}$};

 \draw (10.2, 6) -- (11.2,6.2);
 \node at (11.8, 6.2) {$\frac{e}{d+e}$};
 \node at (10.7,6.3) {$e$};

 \fill (10.2, 5) circle[radius=2pt];
 \draw (10.2, 5) -- (11.2,5.3);
 \node at (11.8, 5.3) {$\frac{\lc}{\lc+e}$};
 \node at (10.7,5.3) {$c$};

 \draw (10.2, 5) -- (11.2,4.7);
 \node at (11.8, 4.7) {$\frac{e}{\lc+e}$};
 \node at (10.7,4.7) {$e$};

 \fill (10.2, 4) circle[radius=2pt];
 \draw (10.2, 4) -- (11.2,3.9);
 \node at (11.8, 3.9) {$\frac{\lc}{\lc+e}$};
 \node at (10.7,4.1) {$c$};

 \draw (10.2, 4) -- (11.2,3.1);
 \node at (11.8, 3.1) {$\frac{e}{\lc+e}$};
 \node at (10.7,3.7) {$e$};

 \draw (4, 0) -- (5,1);
 \node at (6, 1)  {$\frac{a}{a+b+c+e}$};

 \draw (4, 0) -- (5,0);
 \node at (6, 0)  {$\frac{b}{a+b+c+e}$};

 \draw (4, 0) -- (5,-1);
 \node at (6, -1)  {$\frac{c}{a+b+c+e}$};

 \draw (4, 0) -- (5,-2);
 \node at (6, -2)  {$\frac{e}{a+b+c+e}$};

 \node at (4.5,0.7) {$a$};
 \node at (4.5,0.2) {$b$};
 \node at (4.5,-0.3) {$c$};
 \node at (4.5,-0.8) {$e$};

 \fill (7.4, -2) circle[radius=2pt];  % de finishes; paths abc
 \draw (7.4, -2) -- (8.4,-1);
 \draw (7.4, -2) -- (8.4,-2);
 \draw (7.4, -2) -- (8.4,-3);

 \node at (7.9,-1.2) {$a$};
 \node at (7.9,-1.8) {$b$};
 \node at (7.9,-2.3) {$c$};

 \node at (9.3, -1) {$\frac{a}{a+b+c}$};
 \node at (9.3, -2) {$\frac{b}{a+b+c}$};
 \node at (9.3, -3) {$\frac{c}{a+b+c}$};

 \fill (10.2, -1) circle[radius=2pt];  % dea finishes; paths bc
 \fill (10.2, -2) circle[radius=2pt];  % deb finishes; paths ac
 \fill (10.2, -3) circle[radius=2pt];  % dec finishes; paths ab

 \draw (10.2, -1) -- (11.2,-0);
 \node at (10.6,-0.4) {$b$};
 \node at (10.6,-0.8) {$c$};

 \node at (11.8, 0) {$\frac{\lb}{\lb+\lc}$};
 \draw (10.2, -1) -- (11.2,-1);

 \node at (11.8, -1) {$\frac{c}{\lb+\lc}$};

 \draw (10.2, -2) -- (11.2,-1.6);
 \node at (11.8, -1.6) {$\frac{\la}{\la+\lc}$};

 \node at (10.6,-1.7) {$a$};
 \node at (10.6,-2) {$c$};

 \draw (10.2, -2) -- (11.2,-2.2);
 \node at (11.8, -2.2) {$\frac{\lc}{\la+\lc}$};

 \draw (10.2, -3) -- (11.2,-2.9);
 \node at (11.8, -2.9) {$\frac{\la}{\la+\lb}$};

 \draw (10.2, -3) -- (11.2,-3.4);
 \node at (11.8, -3.4) {$\frac{\lb}{\la+\lb}$};

 \node at (10.6,-2.8) {$a$};
 \node at (10.6,-3.3) {$b$};

 \fill (7.4, 1) circle[radius=2pt];  % da finishes; paths bce

 \draw (7.4, 1) -- (8.4,2);
 \draw (7.4, 1) -- (8.4,1);
 \draw (7.4, 1) -- (8.4,0);

 \node at (7.9,1.7) {$b$};
 \node at (7.9,1.2) {$c$};
 \node at (7.9,0.7) {$e$};

 \node at (9.3, 2) {$\frac{\lb}{\lb+\lc+e}$};
 \node at (9.3, 1) {$\frac{\lc}{\lb+\lc+e}$};
 \node at (9.3, 0) {$\frac{e  }{\lb+\lc+e}$};

 \fill (10.2, 1) circle[radius=2pt];  % da finishes; paths bce
 \draw (10.2, 1) -- (11.2,2);
 \node at (11.8, 2) {$\frac{\lb}{\lb+e}$};
 \draw (10.2, 1) -- (11.2,1);
 \node at (11.8, 1) {$\frac{e}{\lb+e}$};

\end{tikzpicture}
```

```{tikz, fig.cap = "Partial probability tree for five competitors $a$-$e$ with mutually supporting subsets $\\left\\lbrace a,b,c\\right\\rbrace$ [with support term $\\lambda$] and $(de)$ [with support term $\\mu$], hyper3 approach", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2,3); 
\draw (0,0) -- (2,2);
\draw (0,0) -- (2,1);
\draw (0,0) -- (2,0);
\draw (0,0) -- (2,-1);
\node at (0,2.5) (eq1) {$\underbrace{\left\lbrace a,b,c\right\rbrace}_{\lambda}\underbrace{\left\lbrace d,e\right\rbrace}_{\mu}$};

\node at (1.5,2.5) {$a$};
\node at (1.5,1.7) {$b$};
\node at (1.5,1.0) {$c$};
\node at (1.5,0.2) {$d$};
\node at (1.5,-0.5) {$e$};

\node at (3, 3)  {$\frac{a}{a+b+c+d+e}$};
\node at (3, 2)  {$\frac{b}{a+b+c+d+e}$};
\node at (3, 1)  {$\frac{c}{a+b+c+d+e}$};
\node at (3, 0)  {$\frac{d}{a+b+c+d+e}$};
\node at (3,-1)  {$\frac{e}{a+b+c+d+e}$};

\fill  (4, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill  (4, 2) circle[radius=2pt];  % terminal node
\fill  (4, 1) circle[radius=2pt];  % terminal node
\fill  (4, 0) circle[radius=2pt];  % d finishes; paths abce
\fill  (4,-1) circle[radius=2pt];  % terminal node

\draw (4,3) -- (5,5); 
\draw (4,3) -- (5,4); 
\draw (4,3) -- (5,3); 
\draw (4,3) -- (5,2);

\node at (4.5,4.6) {$b$};
\node at (4.5,3.7) {$c$};
\node at (4.5,3.2) {$d$};
\node at (4.5,2.7) {$e$};

\newcommand{\la}[1]{\lambda{#1}}
\newcommand{\ld}[1]{\mu    {#1}}
\node at (6, 5)  {$\frac{\la{b}}{\la{b}+\la{c}+d+e}$};
\node at (6, 4)  {$\frac{\la{c}}{\la{b}+\la{c}+d+e}$};
\node at (6, 3)  {$\frac{d     }{\la{b}+\la{c}+d+e}$};
\node at (6, 2)  {$\frac{e     }{\la{b}+\la{c}+d+e}$};

\fill (7.4, 5) circle[radius=2pt];  % ab finishes; pahts cde
\draw (7.4,5) -- (8.4,6);
\draw (7.4,5) -- (8.4,5);
\draw (7.4,5) -- (8.4,4);

\node at (7.9,5.7) {$c$};
\node at (7.9,5.2) {$d$};
\node at (7.9,4.7) {$e$};
\node at (9.3, 6) {$\frac{\la{c}}{\la{c}+d+e}$};
\node at (9.3, 5) {$\frac{d  }{\la{c}+d+e}$};
\node at (9.3, 4) {$\frac{e  }{\la{c}+d+e}$};

\fill (10.2, 6) circle[radius=2pt];  % abc finishes; paths de
\draw (10.2, 6) -- (11.2,7);
\node at (10.7,6.8) {$d$};

\node at (11.8, 7) {$\frac{d}{d+e}$};

\draw (10.2, 6) -- (11.2,6.2);
\node at (11.8, 6.2) {$\frac{e}{d+e}$};
\node at (10.7,6.3) {$e$};

\fill (10.2, 5) circle[radius=2pt];
\draw (10.2, 5) -- (11.2,5.3);
\node at (11.8, 5.3) {$\frac{\la{c}}{\la{c}+\ld{e}}$};
\node at (10.7,5.3) {$c$};

\draw (10.2, 5) -- (11.2,4.7);
\node at (11.8, 4.7) {$\frac{\ld{e}}{\la{c}+\ld{e}}$};
\node at (10.7,4.7) {$e$};

\fill (10.2, 4) circle[radius=2pt];
\draw (10.2, 4) -- (11.2,3.9);
\node at (11.8, 3.9) {$\frac{\la{c}}{\la{c}+\ld{e}}$};
\node at (10.7,4.1) {$c$};

\draw (10.2, 4) -- (11.2,3.1);
\node at (11.8, 3.1) {$\frac{\ld{e}}{\la{c}+\ld{e}}$};
\node at (10.7,3.7) {$e$};

\draw (4, 0) -- (5,1);
\node at (6, 1)  {$\frac{a}{a+b+c+\ld{e}}$};

\draw (4, 0) -- (5,0);
\node at (6, 0)  {$\frac{b}{a+b+c+\ld{e}}$};

\draw (4, 0) -- (5,-1);
\node at (6, -1)  {$\frac{c}{a+b+c+\ld{e}}$};

\draw (4, 0) -- (5,-2);
\node at (6, -2)  {$\frac{\ld{e}}{a+b+c+\ld{e}}$};

\node at (4.5,0.7) {$a$};
\node at (4.5,0.2) {$b$};
\node at (4.5,-0.3) {$c$};
\node at (4.5,-0.8) {$e$};

\fill (7.4, -2) circle[radius=2pt];  % de finishes; paths abc
\draw (7.4, -2) -- (8.4,-1);
\draw (7.4, -2) -- (8.4,-2);
\draw (7.4, -2) -- (8.4,-3);

\node at (7.9,-1.2) {$a$};
\node at (7.9,-1.8) {$b$};
\node at (7.9,-2.3) {$c$};

\node at (9.3, -1) {$\frac{a}{a+b+c}$};
\node at (9.3, -2) {$\frac{b}{a+b+c}$};
\node at (9.3, -3) {$\frac{c}{a+b+c}$};

\fill (10.2, -1) circle[radius=2pt];  % dea finishes; paths bc
\fill (10.2, -2) circle[radius=2pt];  % deb finishes; paths ac
\fill (10.2, -3) circle[radius=2pt];  % dec finishes; paths ab

\draw (10.2, -1) -- (11.2,-0);
\node at (10.6,-0.4) {$b$};
\node at (10.6,-0.8) {$c$};

\node at (11.8, 0) {$\frac{\la{b}}{\la{b}+\la{c}}$};
\draw (10.2, -1) -- (11.2,-1);

\node at (11.8, -1) {$\frac{\la{c}}{\la{b}+\la{c}}$};

\draw (10.2, -2) -- (11.2,-1.6);
\node at (11.8, -1.6) {$\frac{\la{a}}{\la{a}+\la{c}}$};

\node at (10.6,-1.7) {$a$};
\node at (10.6,-2) {$c$};

\draw (10.2, -2) -- (11.2,-2.2);
\node at (11.8, -2.2) {$\frac{\la{c}}{\la{a}+\la{c}}$};

\draw (10.2, -3) -- (11.2,-2.9);
\node at (11.8, -2.9) {$\frac{\la{a}}{\la{a}+\la{b}}$};

\draw (10.2, -3) -- (11.2,-3.4);
\node at (11.8, -3.4) {$\frac{\la{b}}{\la{a}+\la{b}}$};

\node at (10.6,-2.8) {$a$};
\node at (10.6,-3.3) {$b$};

\fill (7.4, 1) circle[radius=2pt];  % da finishes; paths bce

\draw (7.4, 1) -- (8.4,2);
\draw (7.4, 1) -- (8.4,1);
\draw (7.4, 1) -- (8.4,0);

\node at (7.9,1.7) {$b$};
\node at (7.9,1.2) {$c$};
\node at (7.9,0.7) {$e$};

\node at (9.3, 2) {$\frac{\la{b}}{\la{b}+\la{c}+e}$};
\node at (9.3, 1) {$\frac{\la{c}}{\la{b}+\la{c}+e}$};
\node at (9.3, 0) {$\frac{e     }{\la{b}+\la{c}+e}$};

\fill (10.2, 1) circle[radius=2pt];  % da finishes; paths bce
\draw (10.2, 1) -- (11.2,2);
\node at (11.8, 2) {$\frac{\la{b}}{\la{b}+\ld{e}}$};
\draw (10.2, 1) -- (11.2,1);
\node at (11.8, 1) {$\frac{\ld{e}}{\la{b}+e}$};

\end{tikzpicture}
```

# Package idiom

We can investigate red bus-blue bus phenomenon (as discussed, in a
slightly different context, in `inst/red_bus_blue_bus.Rmd`).  Here, we
consider a person who is given the choice of five transport methods:

* `C`, car
* `T`, train
* `RB` a red bus
* `BB` a blue bus
* `W` walking

Now, he does not really care what colour the bus is.  If we ask him to
rank his options, it is highly probable that he will put `RB` and `BB`
consecutively (because they are essentially indistinguishable).  Can
we quantify the strength of this effect?  To do this, we define a
bespoke function `RB_BB_LF()` which returns a `hyper3` log-likelihood
function corresponding to repeated observations of our commuter's
reported ranks for the five options:

```{r,defineredbuslf}
`RB_BB_LF` <- function(s){
    ec <- c(C=1,T=2,RB=3,BB=3,W=4) # equivalence classes
    h <- c(1,1,s,1)                # strength of support
    (
        cheering3(v=c("RB","BB","C" ,"T","W"),e=ec,h=h)*3 + 
        cheering3(v=c("BB","RB","T" ,"C","W"),e=ec,h=h)*2 + 
        cheering3(v=c("T" ,"BB","RB","C","W"),e=ec,h=h)*2 + 
        cheering3(v=c("W" ,"BB","RB","T","C"),e=ec,h=h)*4 + 
        cheering3(v=c("C" ,"RB","BB","W","T"),e=ec,h=h)*4 + 
        cheering3(v=c("BB","C" ,"RB","T","W"),e=ec,h=h)*3
    )
}
```

Above, we see from the function body that he reported `RB,BB,C,T,W`
three times [first row], `BB,RB,T,C,W` twice [second row], and so on;
perhaps his ranking depends on the weather or how tired he is on any
given day.  Observe that in almost every case he ranks `RB` and `BB`
consecutively.  Function `RB_BB_LF()` takes argument `s` that
quantifies the perceived similarity between `RB` and `BB`.  For
example:

```{r,examplerbbblf,cache=TRUE}
(H <- RB_BB_LF(s=1.8888))
(mH <- maxp(H,n=1))
```

Now to find a profile likelihood function for `s`:

```{r,label=proflikeredbus,cache=TRUE}
o <- function(s){maxp(RB_BB_LF(s),give=TRUE,n=1)$likes} # optimand
s <- exp(seq(from=log(1.3),to=log(47),len=17)) # putative similarity measures
L <- sapply(s,o)
L <- L-max(L)
```

We can plot these:

```{r,plotproflike}
plot(s,L,type="b")
abline(h=c(0,-2))
abline(v=1)
plot(log(s),L,type="b")
abline(h=c(0,-2))
abline(v=0)
```

And formally maximize the likelihood:

```{r,label=maxlikesim,cache=TRUE}
(osup <- optimize(o,c(6,10),maximum=TRUE))
```

So a likelihood ratio test of the null that $S=1$ would be:

```{r,label=LRT_RB,cache=TRUE}
(suppdiff <- o(osup$maximum) - o(1))
```

Easily satisfying Edwards's two-units-of-support criterion; Wilks gives us an asymptotic $p$-value:

```{r,label=RBBBpval}
pchisq(suppdiff*2,df=1,lower.tail=FALSE)
```

# University ranking analysis

Here we use a dataset of university rankings, `timesData.csv`, taken
from
`https://github.com/arnaudbenard/university-ranking/blob/master/timesData.csv`.

```{r}
a <- read.table("timesData.csv",sep=",", header=TRUE)
wanted <- c("California Institute of Technology", "Harvard University", 
"Massachusetts Institute of Technology", "Princeton University", 
"Stanford University", "University of Cambridge", "University of Oxford")
names(wanted) <- c("cal","harv","mass","prin","stan","cam","ox")

a <- a[a$university_name %in% wanted,]
a <- cbind(a,"top7rank"=0)
for(y in unique(a$year)){
    a[a$year==y,"top7rank"] <- order(
     			  as.numeric(a[a$year==y,"world_rank"]) + 
                                     a[a$year==y,"research"]/1e6,
decreasing=TRUE)}
a <- a[,c("top7rank","university_name","year")]
a <- reshape(a,idvar="university_name",timevar="year",direction="wide")
for(i in seq_len(nrow(a))){
   a$university_name[i] <- names(which(wanted == a$university_name[i]))
}
rownames(a) <- a$university_name
a <- a[,-1]
colnames(a) <- paste("Y",2011:2016,sep="")
a
H <- ordertable2supp(a)
```

```{r,label=testuniv,cache=TRUE}
equalp.test(H)
samep.test(H,c("ox","cam"))
```

Start to use `hyper3` idiom:

```{r,label=defineoxcam}
H3 <- function(oxcam){
  out <- hyper3()
  for(i in seq_len(ncol(a))){
    jj <- rep("",nrow(a))
    jj[a[,i]] <- rownames(a)
    out <- out + cheering3(v=jj,e=c(ox=1,cam=1,prin=2, stan=3, mass=4, harv=5, cal=6), help=c(oxcam,1,1,1,1,1))
  }
  return(out)
}
```

```{r,label=calculateoxcam,cache=TRUE}
o <- function(oxcam){maxp(H3(oxcam),give=TRUE,n=1)$likes}
oc <- exp(seq(from=log(0.5),to=log(5),len=15))
L <- sapply(oc,o)
L <- L - max(L)
```

```{r,label=plotoxcam}
plot(log(oc),L,type="b")
abline(v=0)
```

# Five nations championship

The five nations rugby championship was held from 1910 to 1999 and
file `five_nations.txt` shows the order statistic for England (E),
Scotland (S), Ireland (I), Wales (W), and France (F).

`https://en.wikipedia.org/wiki/Six_Nations_Championship`

Here is `hyper2` analysis:

```{r,load5nationsdata,cache=TRUE}
a <- as.matrix(read.table("five_nations.txt",header=FALSE))
head(a)
H <- hyper2()
for(i in seq_len(nrow(a))){
  H <- H + race(a[i,-1])
} 
mH <- maxp(H)
pie(mH)
equalp.test(H)
```

Now use `hyper3` to see whether teams do better following a win:

```{r,rugby3,cache=TRUE}
rugby <- function(lambda){
  H3a <- hyper3()
  for(i in seq(from=2,to=nrow(a))){
        last_year_winner <- a[i-1,2] # e.g. "W" or "E"
        H3a <- H3a + ordervec2supp3a(a[i,-1],nonfinishers=NULL,helped=last_year_winner,lambda=lambda)
  }
  return(H3a)
}
rugby(1.888)
rugby(1.111111)
```

```{r,rugby3max, cache=TRUE}
maxp(rugby(1.8),n=1,give=TRUE)
maxp(rugby(1.9),n=1,give=TRUE)
```