### 2023-01-15 inst/NBA.Rmd
fixes #73

---
title: "An analysis of a single basketball game"
output: html_document
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

File `NBA.txt` contains a point-by-point analysis of a basketball
match between the Cleveland Cavaliers and the Golden State Warriors on
13 June 2017, data taken from
[here](https://www.espn.com/nba/playbyplay?gameId=400954514).

```{r readNBAtable}
NBA_table  <- read.csv("NBA.txt",header=TRUE)
head(NBA_table)
```

Each row corresponds to a point scored.  The first column is the time
of the score, the second is the number of points scored, the third
shows which team had possession at the start of play, and the fourth
shows which team scored.  The other columns show the players.  Table
entries show whether or not that particular player was on the pitch
when the point was scored.  Note that the two "ghost" players
represent the effect of having possession.

First some basics:

```{r,fisherbasic}
(TT <- table(NBA_table[,3:4]))
fisher.test(TT,alternative="greater")
```

Thus the identity of the scoring team is not independent of the
identity of the possessing team (as one might expect).  
We can convert `NBA_table` to a likelihood function:

```{r,label=definesupportb,cache=TRUE}
allplayers <- as.matrix(NBA_table[,5:23])
NBA <- hyper2()
NBA2 <- hyper2()

## players 1-9 Cleveland
## players 10-19 (sic) Warriors:
C_onpitch <- allplayers[,1:9]
W_onpitch <- allplayers[,10:19]

possession <- NBA_table[,3]
scored     <- NBA_table[,4]

for(i in seq_len(nrow(NBA_table))){
  if(scored[i] == 'W'){  # Warriors score a point
    onpitch_scored  <- names(which(W_onpitch[i,]))
    onpitch_noscore <- names(which(C_onpitch[i,]))
  } else if(scored[i] == "C"){ # Cleveland score a point
    onpitch_scored  <- names(which(C_onpitch[i,]))
    onpitch_noscore <- names(which(W_onpitch[i,]))
  } else {
    stop("scorint team must be either W or C")
  }
  if(possession[i] == 'W'){
    onpitch_scored1 <- c(onpitch_scored,"possession")
    onpitch_scored2 <- c(onpitch_scored,"W_possession")
  } else if (possession[i] == 'C'){
    onpitch_scored1 <- c(onpitch_scored,"possession")
    onpitch_scored2 <- c(onpitch_scored,"C_possession")
  } else {
    stop("possession must be either W or C")
  }

  NBA[onpitch_scored1] %<>% inc()
  NBA[c(onpitch_scored1,onpitch_noscore)] %<>% dec()

  NBA2[onpitch_scored2] %<>% inc()
  NBA2[c(onpitch_scored,onpitch_noscore)] %<>% dec()
}

NBA2_maxp <- maxp(NBA2)

NBA2_maxp %<>% ordertrans(c(colnames(NBA_table)[-(1:4)],"C_possession","W_possession"))
dotchart(NBA2_maxp,col=c(rep("black",9),rep("red",10),rep("blue",2)),pch=16)

NBA_maxp_precomputed  <- c(
                                        # precomputed; very very computer
                                        # intensive but far higher likelihood
                                        # than the result of maxp(NBA):

Barnes    = 0.165210241492775,    Curry      = 1.11314508611164e-06,
Durant    = 0.00538619616133955,  Green      = 0.0110738194130432,
Iguodala  = 1.0005765074368e-06,  Irving     = 0.0198810257932021,
James     = 1.00000582238398e-06, Jefferson  = 0.00990927936847305,
Korver    = 0.00843806151505123,  Livingston = 0.075108213128189,
Love      = 0.00175976502052894,  McCaw      = 0.0119305671842558,
Pachiuila = 0.0496281731046277,   possession = 0.479366441714251,
Shumpert  = 0.0666196978946451,   Smith      = 1.0008827880116e-06,
Thompson  = 0.00159564073991825,  Thompson.1 = 0.00135198197477646,
West      = 0.0927348818403838,   Williams   = 1.89904433611776e-06)

NBA_maxp <- maxp(NBA)
NBA_maxp %<>% ordertrans(c(colnames(NBA_table)[-(1:4)],"possession"))

dotchart(NBA_maxp,col=c(rep("black",9),rep("red",10),rep("blue",1)),pch=16)
```

# Statistical analysis

```{r,label=statsanal,cache=TRUE}
samep.test(NBA2,c("C_possession","W_possession"))
```

We may thus assume that Cleveland having possession is the same
strength as the Warriors having possession, and from now on we work
with `NBA` rather than `NBA2`.

```{r,label=doCWtest, cache=TRUE}
(C_allstrengths <- sort(NBA_maxp[01:9],decreasing=TRUE))
(W_allstrengths <- sort(NBA_maxp[10:19],decreasing=TRUE))
samep.test(NBA, names(C_allstrengths))
samep.test(NBA, names(W_allstrengths))
```

Now, what is the probability of C scoring both with and without
possession?  We may condition on the estimated player and possession
strengths, and assume that only the 5 strongest players on each team
play.

```{r}
C_strongestteam <- sum(C_allstrengths[1:5])
W_strongestteam <- sum(W_allstrengths[1:5])
possession <- 0.2155786 # taken from samep.test() above

# probability of C scoring without possession:
(C_strongestteam           )/(C_strongestteam + W_strongestteam + possession)

# probability of C scoring with possession:
(C_strongestteam+possession)/(C_strongestteam + W_strongestteam + possession)
```

### Package dataset

Following lines create `NBA.rda`, residing in the `data/` directory of the package.

```{r,label=saveNBAdataset}
NBA_maxp <- NBA_maxp_precomputed
save(NBA_table,NBA,NBA_maxp,file="NBA.rda")
```