### 2023-01-18 inst/monster_vs_lambda.Rmd
monster v lambda

---
title: "Monster or Lambda?  A likelihood analysis"
author: "Robin KS Hankin"
output: html_document
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

(takes about an hour to process without cache)

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here I consider pairwise comparisons with a home-ground advantage and
wish to compare two likelihood functions.  First, the monster parametrization:

$$
\mathcal{L_M}\left(p_1,p_2,M\right)=
\left(
\frac{p_1+M;p_2}{p_1+p_2+M}
\right)^{a;b}\cdot
\left(
\frac{p_1;p_2+M}{p_1+p_2+M}
\right)^{c;d}
$$

[here the first expression is for $p_1$ at home and $p_2$ away; the
second is $p_1$ away and $p_2$ at home; $a$ is number of home wins for
$p_1$, $b$ number of away wins for $p_2$, $c$ number of away wins for
$p_1$ and $d$ number of home wins for $p_2$].   Secondly the monster
parameterization:

$$
\mathcal{L_\lambda}\left(p_1,p_2,\lambda\right)=
\left(
\frac{\lambda p_1;p_2}{\lambda p_1+p_2}
\right)^{a;b}\cdot
\left(
\frac{p_1;\lambda p_2}{p_1+\lambda p_2}
\right)^{c;d}
$$

A suitable generalized function would be

$$
\mathcal{L}\left(p_1,p_2,\lambda,M\right)=
\left(
\frac{\lambda p_1+M;p_2}{\lambda p_1+p_2+M}
\right)^{a;b}\cdot
\left(
\frac{p_1;\lambda p_2+M}{p_1+\lambda p_2+M}
\right)^{c;d}
$$

First we need to define a slight generalization of `home_away3()`:

```{r}
`ha3` <- function(home_games_won,away_games_won,lambda){

    if(is.complex(home_games_won)){
        if(missing(lambda)){lambda <- away_games_won}
        away_games_won <- Im(home_games_won)
        home_games_won <- Re(home_games_won) 
    }

    teams <- rownames(home_games_won)
    stopifnot(identical(teams,colnames(home_games_won)))
    stopifnot(identical(teams,rownames(away_games_won)))
    stopifnot(identical(teams,colnames(away_games_won)))

    H <- hyper3(pnames=c(teams,"M","lambda"))

    for(i in seq_len(nrow(home_games_won))){
        for(j in seq_len(ncol(home_games_won))){
            if(i != j){  # diagonal means a team plays itself, meaningless.
                ## First deal with home_games:
                game_winner <- teams[i]
                game_loser  <- teams[j]
                no_of_matches <- home_games_won[i,j] # won with home strength helping

                jj <- c(lambda,1)
                names(jj) <- c(game_winner,"M")
                H[jj] %<>% inc(no_of_matches)  # won with home strength helping

                jj <- c(lambda,1,1)
                names(jj) <- c(game_winner,game_loser,"M")
                H[jj] %<>% dec(no_of_matches)

                ## now away games:
                no_of_matches <- away_games_won[i,j] # won without the benefit of home strength
                jj <- 1
                names(jj) <- game_winner
                H[jj] %<>% inc(no_of_matches)  # won without home ground helping...

                jj <- c(1,lambda,1)
                names(jj) <- c(game_winner,game_loser,"M")
                H[jj] %<>% dec(no_of_matches)  # home strength nevertheless on denominator

            } # if(i != j) closes
        } # j loop closes
    } # i loop closes
    return(H)
}
```

(maybe one day I will include this in the `hyper2` package).  Copying
from `home_advantage.Rmd`:

```{r label=definehomeawaymatrix}
D  <- matrix(c(
    NA, 16, 12, 11, 19, NA, 19, 16,
    17, 12, NA, 11, 11, 12, 12, NA),
    nrow=4,ncol=4,byrow=TRUE) + 1i* matrix(c(
    NA, 05, 02, 02, 02, NA, 10, 02,
    00, 04, NA, 07, 02, 06, 04, NA),
    nrow=4,ncol=4,byrow=TRUE)

teams <- LETTERS[1:4]
dimnames(D) <- list("@home" = teams,"@away"=teams)
D
```

```{r,tryha3,cache=TRUE}
H <- ha3(D,lambda=1.88)
s <- specificp.gt.test(H,"M",0,n=1)  # NB lambda = 1.88
s
```

```{r}
s$null_support
```

Above but in one line:

```{r,label=trytwovalues,cache=TRUE}
ll <- function(lambda){specificp.gt.test(ha3(D,lambda=lambda),"M",0,n=1)$null_support}
ll(1.22)
ll(1.88)
```

```{r,label=calclam,cache=TRUE}
lam <- seq(from=2.5,to=6,len=15)
supp <- sapply(lam,ll)
```

```{r}
plot(lam,supp-max(supp),type='b')
```