### 2023-01-26 inst/monster_vs_lambda.Rmd
reorganisation: start simple

---
title: "Monster or Lambda?  A likelihood analysis"
author: "Robin KS Hankin"
output: html_document
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

(takes about an hour to process without cache)

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Here I consider pairwise comparisons with a home-ground advantage and
wish to compare two likelihood functions.  First, the monster parametrization:

$$
\mathcal{L_M}\left(p_1,p_2,M\right)=
\left(
\frac{p_1+M;p_2}{p_1+p_2+M}
\right)^{a;b}\cdot
\left(
\frac{p_1;p_2+M}{p_1+p_2+M}
\right)^{c;d}
$$

[here the first expression is for $p_1$ at home and $p_2$ away; the
second is $p_1$ away and $p_2$ at home; $a$ is number of home wins for
$p_1$, $b$ number of away wins for $p_2$, $c$ number of away wins for
$p_1$ and $d$ number of home wins for $p_2$].   Secondly the monster
parameterization:

$$
\mathcal{L_\lambda}\left(p_1,p_2,\lambda\right)=
\left(
\frac{\lambda p_1;p_2}{\lambda p_1+p_2}
\right)^{a;b}\cdot
\left(
\frac{p_1;\lambda p_2}{p_1+\lambda p_2}
\right)^{c;d}
$$

A suitable generalized function would be

$$
\mathcal{L}\left(p_1,p_2,\lambda,M\right)=
\left(
\frac{\lambda p_1+M;p_2}{\lambda p_1+p_2+M}
\right)^{a;b}\cdot
\left(
\frac{p_1;\lambda p_2+M}{p_1+\lambda p_2+M}
\right)^{c;d}
$$

## Start simple: two teams

```{r,echo=FALSE}
jj <- matrix(c(NA, 1+2i ,4+5i, NA),2,2,byrow=TRUE)
teams <- LETTERS[1:2]
dimnames(jj) <- list("@home" = teams,"@away"=teams)
dimnames(jj) <- list("@home" = teams,"@away"=teams)
print(jj)
```

Above we see that A played at home 3 times, winning 1 and losing 2
[row 1 column 2 is $1+2i$]; B played at home 9 times, winning 4 and
losing 5 [row 2 column 1 is $4+5i$].  In its most general form, the
$2\times 2$ result matrix would be

$$
\begin{bmatrix}
-&\alpha+\beta i \\
\gamma + \delta i&-\\
\end{bmatrix}
$$

and the log-likelihood function would be, algebraically,

$$
\mathcal{S}=\log\mathcal{L}=
\log\left(\frac{A^\delta\cdot(A + M)^\alpha\cdot B^\beta
(B + M)^\gamma
}{
(A + B + M)^{\alpha+\beta+\gamma+\delta}
}\right)
=
\delta\log A + \alpha\log(1-B) + \beta\log B + \gamma\log(1-A)
$$

where $M$ is the strength of the home monster `home` and $A+B+M=1$, so
we understand that $M=1-A-B$.

$$
\frac{\partial\mathcal{S}}{\partial A}=\frac{\delta}{A}-\frac{\gamma}{1-A}\qquad
\frac{\partial\mathcal{S}}{\partial B}=\frac{\beta }{B}-\frac{\alpha}{1-A}
$$

At a maximum, therefore,

$$
\hat{A}=\frac{\delta}{\delta + \gamma}\qquad
\hat{B}=\frac{\beta }{\alpha  + \beta}
$$

Just to check, we evaluate the Hessian matrix and find

$$H=-
\left(\begin{array}{cc} 
\frac{\partial^2\mathcal{S}}{\partial A\partial A} & 
\frac{\partial^2\mathcal{S}}{\partial A\partial B} \\
\frac{\partial^2\mathcal{S}}{\partial B\partial A} & 
\frac{\partial^2\mathcal{S}}{\partial b\partial B} 
\end{array}\right)
=
\left(\begin{array}{cc}
\delta A^{-2} + \gamma(1-A)^{-2}&0\\
0&\beta B^{-2} + \alpha(1-B)^{-2}
\end{array}\right)
$$

which is clearly strictly positive-definite if $0<A,B<1$ and
$\alpha+\beta>0$, $\gamma+\delta>0$.

We will calculate numerical estimates using the
monster-parameterization for a simple case:

```{r,label=try2x2}
M <- matrix(c(
    NA, 5+1i, 9+2i, NA),
    nrow=2,ncol=2,byrow=TRUE)
teams <- LETTERS[seq_len(2)]
dimnames(M) <- list("@home" = teams,"@away"=teams)
dimnames(M) <- list("@home" = teams,"@away"=teams)
M
(Hsimp <- home_away(M))
maxp(Hsimp)
```

Compare the theoretical value of

```{r}
theoretical <- c(A=2/(2+9),B=1/(1+5))
fillup(theoretical,Hsimp)
```

showing good agreement.  We note that the model predicts a probability
of $\hat{A}/(\hat{A}+\hat{B})=\frac{12}{23}\simeq 0.52$ for $A$
winning in a match between $A$ and $B$ where the home team advantage
is not operating (on neutral ground, for example).

Now the $\lambda$ parameterization.  The likelihood function for the
same dataset is

$$
\mathcal{S}=\log\mathcal{L}=
\log\left(
\frac{(\lambda A)^\alpha B^\beta}{(\lambda A+B)^{\alpha+\beta}}
\cdot
\frac{A^\gamma (\lambda B)^\delta}{(A+\lambda B)^{\gamma+\delta}}
\right)$$

This is equal to $(\alpha+\delta)\log\lambda +(\alpha+\gamma)\log
A+(\beta+\delta)\log B -(\alpha+\beta)\log(\lambda A+B)
-(\gamma+\delta)\log(A+\lambda B)$ and even after simplifying [we may
assume $A+B=1$] the result does not appear to have a simple analytical
solution; we must proceed numerically.  To do so,

we need to define a
slight generalization of `home_away3()`:

```{r defineha3}
`ha3` <- function(home_games_won,away_games_won,lambda){  # based on home_away3()

    if(is.complex(home_games_won)){
        if(missing(lambda)){lambda <- away_games_won}
        away_games_won <- Im(home_games_won)
        home_games_won <- Re(home_games_won) 
    }

    teams <- rownames(home_games_won)
    stopifnot(identical(teams,colnames(home_games_won)))
    stopifnot(identical(teams,rownames(away_games_won)))
    stopifnot(identical(teams,colnames(away_games_won)))

    H <- hyper3(pnames=c(teams,"lambda","M"))

    for(i in seq_len(nrow(home_games_won))){
        for(j in seq_len(ncol(home_games_won))){
            if(i != j){  
                home_team <- teams[i]
                away_team <- teams[j]

		home_wins <- home_games_won[i,j]
		away_wins <- away_games_won[i,j] 
                no_of_matches <- home_wins + away_wins

                ## home wins:
                jj <- c(lambda,1)
                names(jj) <- c(home_team,"M")
                H[jj] %<>% inc(home_wins)

                ## away wins:
                jj <- 1
                names(jj) <- away_team
                H[jj] %<>% inc(away_wins)

                ## denominator
                jj <- c(lambda,1,1)
                names(jj) <- c(home_team,away_team,"M")
                H[jj] %<>% dec(no_of_matches)
            } # if(i != j) closes
        } # j loop closes
    } # i loop closes
    return(H)
}
```

(maybe one day I will include function `ha3()` in the `hyper2`
package).

Now a more complicated example, a
modified version of the 4x4 matrix on `home_advantage.Rmd`:

```{r label=definehomeawaymatrix}
home_games_won <- matrix(c(
    NA, 16, 12, 11,
    19, NA, 19, 16,
    17, 12, NA, 11,
    11, 12, 12, NA),
    nrow=4,ncol=4,byrow=TRUE)

away_games_won <- matrix(c(
    NA, 05, 02, 02,
     9, NA, 10, 02,
     3, 04, NA, 07,
     8, 06, 04, NA),
    nrow=4,ncol=4,byrow=TRUE)

teams <- LETTERS[1:4]
dimnames(home_games_won) <- list("@home" = teams,"@away"=teams)
dimnames(away_games_won) <- list("@home" = teams,"@away"=teams)

D <- home_games_won + 1i*away_games_won
D
```

Above we see, from row 1 column 2 [`36+10i`], that A played at home
against B a total of $36+10=46$ times, winning $36$ and losing $10$;
From row 2 column 1 [`11+10i`] we see that B played at home against A
a total of 21 times, winning 11 and losing 10.  We can use the Monster
likelihood function and test $H_0\colon M=0$:

```{r,tryha3,cache=TRUE}
H <- ha3(D,lambda=1.01)
print(H)
```

```{r,maxha3,cache=TRUE}
maxp(H)
```

```{r,testha3,cache=TRUE}
s <- specificp.gt.test(H,"M",0,n=10)  # NB lambda = 1.88
s
```

We see strong evidence that $M>0$.

```{r}
s$null_support
```

We want to plot a profile log-likelihood function, that is, a
likelihood function corresponding to the evaluate constrained to
having a particular value of $M$:

```{r,label=marginallambdaprofilelike,cache=TRUE}
write(date(),file="marginallambdaprofilelike",append=FALSE)
f <- function(lambda){
 write(date(),file="marginallambdaprofilelike",append=TRUE)  # so we can check its progress
 return(maxp(ha3(D,lambda=lambda),n=1,give=TRUE)$value)
}

lam <- seq(from=0.1,to=4.2,len=5)
supp_M_eq_0 <- sapply(lam,f)
```

```{r,plotprofilelikem0}
plot(lam,supp_M_eq_0,type="b",xlab="lambda")
plot(lam,supp_M_eq_0-max(supp_M_eq_0),type="b",xlab="lambda")
grid()
abline(h=c(0,-2))
```

Now enforce the constraint $\lambda=1$ [and allow $M$ to vary],
thereby reducing the likelihood function from a `hyper3` to a `hyper2`
object:

```{r,tryiconprofsupp,cache=TRUE}
(H <- home_away(D))  # observe that the monster is 'home'
home <- seq(from=0.15,to=0.4,len=15)
supp_lambda_eq_0 <- profsupp(H,"home",home,relative=FALSE)
plot(home,supp_lambda_eq_0 - max(supp_lambda_eq_0),type='b')
abline(h=c(0,-2))
```

```{r,showmaxplam1,cache=TRUE}
maxp(H)
```

```{r,label=plotordinarysupport}
plot(lam,supp_M_eq_0-max(supp_M_eq_0),type='b',main='Profile likelihood with M=0')
grid()
plot(home,supp_lambda_eq_0 - max(supp_lambda_eq_0),type='b')
```

```{r,label=plotbothsupports}
both <- c(supp_lambda_eq_0,NA,supp_M_eq_0)
plot(both-max(both,na.rm=TRUE),type='b',col=c(rep("black",length(supp_lambda_eq_0)),"green",rep("red",length(supp_M_eq_0))))
abline(h=c(0,-2))
```

Support difference:

```{r,label=supportdifference}
max(supp_M_eq_0)-max(supp_lambda_eq_0)
```

exceeding Edwards's two-units-of-support criterion by a comfortable
margin.  Note that both have equal numbers of degrees of freedom.

# Create a new dataset by direct simulation

```{r,syntheticdatapp}
set.seed(0)
P <- matrix(0,4,4)
teams <- LETTERS[1:4]
dimnames(P) <- list("@home" = teams,"@away"=teams)
n <- 90 # number of matches between a pair 
M <- 0.15  # home monster
strengths <- (4:1)/10
strengths <- strengths/sum(strengths)
names(strengths) <- teams
strengths
for(home_team in 1:4){
  for(away_team in 1:4){
    if(home_team != away_team){
      prob_home_win <- (strengths[home_team] + M)/(strengths[home_team] + strengths[away_team] + M)
      home_wins <- rbinom(1,n,prob_home_win)
      home_losses <- n - home_wins
      P[home_team,away_team] <- home_wins + 1i*home_losses
    }
  } 
}
diag(P) <- NA
P
HP <- home_away(P)
```

```{r,label=showlike}
HP
```

```{r,label=maxpmonstereq,cache=TRUE}
mH <- maxp(HP,n=3,give=T)
mH
```

```{r proflikesynthcc,cache=TRUE}
home <- seq(from=0.12,to=0.16,len=15)
supp_lambda_eq_1 <- profsupp(HP,"home",home,relative=FALSE)
```

```{r label=matchershower}
jjH <- ha3(P,lambda=1) # should match HP above
jjH
as.hyper3(HP) - jjH
```

```{r plotsynthprofa}
plot(home,supp_lambda_eq_1,type='b')
plot(home,supp_lambda_eq_1 - max(supp_lambda_eq_1),type='b')
abline(h=c(0,-2))
```

```{r showresult,cache=TRUE}
specificp.gt.test(ha3(P,lambda=1.4),"M",0)
specificp.gt.test(ha3(P,lambda=1.4),"M",1e-4)
```

```{r marginalsynthcale,cache=TRUE}
f <- function(lambda){
   o <- ha3(P,lambda=lambda)
   return(specificp.gt.test(o,"M",0)$null_support)
}

lam <- seq(from=1.9,to=2.4,len=5)
supp_M_eq_0 <- sapply(lam,f)
```

```{r plotsynthma}
plot(lam,supp_M_eq_0,type='b')
plot(lam,supp_M_eq_0-max(supp_M_eq_0),type='b')
abline(h=c(0,-2))
```

```{r plotbothtogether}
jj <- c(supp_M_eq_0,NA,supp_lambda_eq_1)
plot(jj-max(jj,na.rm=TRUE),type="b")
abline(h=c(0,-2))
```