### 2023-01-27 inst/home_advantage3.Rmd
consistentification

---
title: "Quantifying the home advantage with `hyper3` objects"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{hyper3}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

$$\mathcal{L}\left(p_1,\ldots, p_n\right)=\prod_j\left(\sum_{i=1}^n w^i_jp_i\right)^{n_j}.$$

As a motivating example, suppose two teams (players) with
Bradley-Terry strengths $p_1,p_2$ play football where we quantify the
home-ground advantage with a term $\lambda$.  If $p_1$ plays at home
$a+b$ times with $a$ wins and $b$ losses, and plays away [so $p_2$ is
at home] $c+d$ times with $c$ wins and $d$ losses, then a sensible
likelihood function might be

$$\mathcal{L}(p_1,p_2,\lambda;A,B,C,D)=
\left(\frac{\lambda p_1}{\lambda p_1 + p_2}\right)^{A}
\left(\frac{p_2        }{\lambda p_1 + p_2}\right)^{B}
\left(\frac{p_1        }{p_1 + \lambda p_2}\right)^{C}
\left(\frac{\lambda p_2}{p_1 + \lambda p_2}\right)^{D}.
$$

where we understand that $p_1+p_2=1$, $p_1,p_2,\lambda\geqslant 0$.
Elementary techniques allow us to identify a maximum and we find

$$
\hat{p_1}=\frac{\sqrt{AC}}{\sqrt{AC}+\sqrt{BD}}\qquad
\hat{p_2}=\frac{\sqrt{BD}}{\sqrt{AC}+\sqrt{BD}}\qquad
\hat{\lambda}=\sqrt{\frac{AD}{BC}}
$$

```{r label=defineABCD}
A <- 9
B <- 2
C <- 3
D <- 5
```

$$
\hat{p_1}=\frac{\sqrt{27}}{\sqrt{27} + \sqrt{10}}\simeq 0.62\qquad
\hat{p_2}=\frac{\sqrt{15}}{\sqrt{27} + \sqrt{10}}\simeq 0.38\qquad
\hat{\lambda}=\sqrt{\frac{AD}{BC}}\simeq 2.73$$

Keeping $A=x,B=x,C=x,D=xxxxxx$ and $\lambda=2.73$ [assumed for the moment
to be known, estimation techniques are discussed later], appropriate
package idiom might be to use bespoke function `home_away3()`:

```{r,label=usehomeaway3a}
library("hyper2",quietly=TRUE)
M <- matrix(c(NA, A+B*1i ,D+C*1i, NA),2,2,byrow=TRUE)
teams <- c("p1","p2")
dimnames(M) <- list("@home" = teams,"@away"=teams)
dimnames(M) <- list("@home" = teams,"@away"=teams)
print(M)
(H <- home_away3(M,lambda=2.73))
```

Above, obect `M` has real parts being home wins and imaginary parts
being away wins.

We may estimate $p_1$ and $p_2$ using maximum likelihood, `maxp()` in
package idiom:

```{r,label=findtheevaluate3,cache=TRUE}
H
maxp(H)
```

Further, we can test whether the players are in fact of equal strength:

```{r,label=testequalitynow,cache=TRUE}
equalp.test(H)
```

Showing convincingly that we may reject the null that $p_1=p_2$ and
assert that the players do in fact differ in strength, at least with
this value of $\lambda$.  Observe that a naive analysis would have
$p_1$ winning $w=A+C=`r A+C`$ games and losing $l=B+D=`r B+D`$ games,
from which we would obtain $\hat{p_1}=`r A+C`/`r A+B+C+D`$; a
likelihood ratio for $H_0\colon p_1=0.5$ would be

$$\frac
{
  {{w+l}\choose {w\, l}}\left(\frac{w}{w+l}\right)^w\left(\frac{l}{w+l}\right)^l
}{
  {{w+l}\choose {w\, l}}\left(\frac{1}{2}\right)^w\left(\frac{1}{2}\right)^l
}=\frac{2^{w+l}w^wl^l}{(w+l)^{w+l}}\simeq 1.94,
$$

not significant.

Now, how to estimate $\lambda$?

```{r,label=estimatelambda,cache=TRUE}
f <- function(lambda){maxp(home_away3(M,lambda=lambda),give=TRUE)$value}
f(2)
f(2.1)
lam <- seq(from=1,to=10,len=17)
Supp <- sapply(lam,f)
```

```{r,label=plotsup}
plot(log(lam),Supp-max(Supp),type="b")
```

We can proceed in two ways.  The flaw in this analysis is that it is
conditional on the estimated value of $\lambda$ [which is not too bad,
IMO].  Secondly

```{r,label=loopysuppcalc,cache=TRUE}
probs <- seq(from=0.1,to=0.8,len=17)
jj <- as.matrix(expand.grid(lam,probs))
S <- rep(NA,nrow(jj))
for(i in seq_len(nrow(jj))){
    lambda <- jj[i,1]
    p1 <- jj[i,2]
    p <- c(p1,1-p1)
    names(p) <- c("p1","p2")
   S[i] <- loglik(p=p,H=home_away3(M,lambda=lambda),log=TRUE)
}
S <- matrix(S,length(lam),length(probs))
```

```{r}
contour(lam,probs,S,levels=seq(from=-20,to=-10,by=0.5),xlab="lambda",ylab="p1")
abline(h=0.5)
```