### 2023-01-27 inst/hyper3.Rmd

---
title: "The weighted hyperdirichlet distribution: the `hyper3` class"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{hyper3}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

$$\mathcal{L}\left(p_1,\ldots, p_n\right)=\prod_j\left(\sum_{i=1}^n w^i_jp_i\right)^{n_j}.$$

As a motivating example, suppose two teams (players) with
Bradley-Terry strengths $p_1,p_2$ play football where we quantify the
home-ground advantage with a term $\lambda$.  If $p_1$ plays at home
$a+b$ times with $a$ wins and $b$ losses, and plays away [so $p_2$ is
at home] $c+d$ times with $c$ wins and $d$ losses, then a sensible
likelihood function might be

$$\mathcal{L}(p_1,p_2,\lambda;A,B,C,D)=
\left(\frac{\lambda p_1}{\lambda p_1 + p_2}\right)^{A}
\left(\frac{p_2        }{\lambda p_1 + p_2}\right)^{B}
\left(\frac{p_1        }{p_1 + \lambda p_2}\right)^{C}
\left(\frac{\lambda p_2}{p_1 + \lambda p_2}\right)^{D}.
$$

where we understand that $p_1+p_2=1$, $p_1,p_2,\lambda\geqslant 0$.
Elementary techniques allow us to identify a maximum and we find

$$
\hat{p_1}=\frac{\sqrt{AC}}{\sqrt{AC}+\sqrt{BD}}\qquad
\hat{p_2}=\frac{\sqrt{BD}}{\sqrt{AC}+\sqrt{BD}}\qquad
\hat{\lambda}=\sqrt{\frac{AD}{BC}}
$$

If $A=x,B=x,C=x,D=x$ and $\lambda=1.3$ [assumed for the moment to be
known], appropriate package idiom might be:

```{r,label=usenamedvectorsforH}
library("hyper2",quietly=TRUE)
A <- 6 ; B <- 2 ; C <- 7 ; D <- 1
H <- hyper3()
H[c(p1=1.3)]      %<>% inc(A) 
H[c(p2=1)]        %<>% inc(B) 
H[c(p1=1.3,p2=1)] %<>% dec(A+B)
H[c(p1=1)]        %<>% inc(C) 
H[c(p2=1.3)]      %<>% inc(D) 
H[c(p1=1,p2=1.3)] %<>% dec(C+D)
H
```

Bespoke function `home_away3()` achieves the same result with a more
convenient input:

```{r,label=usehomeaway3}
M <- matrix(c(NA, 6+2i ,1+7i, NA),2,2,byrow=TRUE)
teams <- c("p1","p2")
dimnames(M) <- list("@home" = teams,"@away"=teams)
dimnames(M) <- list("@home" = teams,"@away"=teams)
print(M)
home_away3(M,lambda=1.3)
```

Above, obect `M` has real parts being home wins and imaginary parts
being away wins.

We may estimate $p_1$ and $p_2$ using maximum likelihood, `maxp()` in
package idiom:

```{r,label=findtheevaluate,cache=TRUE}
maxp(H)
```

Further, we can test whether the players are in fact of equal strength:

```{r,label=testequality,cache=TRUE}
equalp.test(H)
```

Showing convincingly that we may reject the null that $p_1=p_2$ and
assert that the players do in fact differ in strength.

Another motivating example might come from Plackett-Luce likelihood
functions for order statistics.  Suppose we have three players with
nonnegative strengths $p_\mathrm{M},p_\mathrm{RB},p_\mathrm{F}$, with
unit sum.  These players might be conceputalised as F1 constructors
(here Mercedes (M), Red Bull (RB), and Ferrari (F) for the sake of
argument).  Each constructor fields two cars and the order statistic
might be

\newcommand{\pr}{p_\mathrm{RB}}
\newcommand{\pm}{p_\mathrm{M}}
\newcommand{\pf}{p_\mathrm{F}}

$$
RB\succ M\succ F\succ F\succ RB\succ M$$

indicating that the finishing order was Red Bull first, Mercedes
second, Ferrari third, and so on (this was in fact observed at the
2021 Emilia Romagna Grand Prix). The Placket-Luce likelihood function
would be

$$
\frac{\pr}{\pr+\pm+\pf+\pf+\pr+\pm}\cdot
\frac{\pm}{    \pm+\pf+\pf+\pr+\pm}\cdot
\frac{\pf}{        \pf+\pf+\pr+\pm}\cdot
\frac{\pf}{            \pf+\pr+\pm}\cdot
\frac{\pr}{                \pr+\pm}\cdot
\frac{\pm}{                    \pm}
$$

or, in a form more suitable for the package,

$$
\frac{\pr}{2\pr + 2\pm + 2\pf}\cdot
\frac{\pm}{ \pr + 2\pm + 2\pf}\cdot
\frac{\pf}{ \pr +  \pm + 2\pf}\cdot
\frac{\pf}{ \pr +  \pm  + \pf}\cdot
\frac{\pr}{ \pr  + \pm       }$$

We could proceed as follows:

```{r,label=workouta}
a <- hyper3()
a[c(R=1)] %<>% inc
a[c(R=2,M=2,F=2)] %<>% dec
a[c(M=1)] %<>% inc
a[c(R=1,M=2,F=2)] %<>% dec
a[c(F=1)] %<>% inc
a[c(R=1,M=1,F=2)] %<>% dec
a[c(F=1)] %<>% inc
a[c(R=1,M=1,F=1)] %<>% dec
a[c(R=1)] %<>% inc
a[c(R=1,M=1    )] %<>% dec
a
```

Or maybe more logically

```{r,label=usebetteridiomforb}
b <- hyper3()
b["R"] %<>% inc ; b[c("R","M","F","F","R","M")] %<>% dec
b["M"] %<>% inc ; b[c(    "M","F","F","R","M")] %<>% dec
b["F"] %<>% inc ; b[c(        "F","F","R","M")] %<>% dec
b["F"] %<>% inc ; b[c(            "F","R","M")] %<>% dec
b["R"] %<>% inc ; b[c(                "R","M")] %<>% dec

a==b
```

However, bespoke function `ordervec2supp3()` performs all this in one
step:

```{r, label=showbespoke}
ordervec2supp3(c("RB","M","F","F","RB","M"))
```

And then

```{r, label=maxpa,cache=TRUE}
ma <- maxp(a)
ma
```

Evidence for different strengths?

```{r,aretheydifferent,cache=TRUE}
equalp.test(a)
```

No evidence agains the null of equal strengths. Further:

```{r,label=showgradissmall}
gradient(a,ma)  # should be small at the evaluate
```