### 2023-01-30 inst/home_advantage3.Rmd
rewrite

---
title: "Quantifying the home advantage with `hyper3` objects"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{hyper3}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

$$\mathcal{L}\left(p_1,\ldots, p_n\right)=\prod_j\left(\sum_{i=1}^n w^i_jp_i\right)^{n_j}.$$

As a motivating example, suppose two teams (players) with
Bradley-Terry strengths $p_1,p_2$ play football where we quantify the
home-ground advantage with a term $\lambda$.  If $p_1$ plays at home
$a+b$ times with $a$ wins and $b$ losses, and plays away [so $p_2$ is
at home] $c+d$ times with $c$ wins and $d$ losses, then a sensible
likelihood function might be

$$\mathcal{L}(p_1,p_2,\lambda;A,B,C,D)=
\underbrace{
\overbrace{\left(\frac{\lambda p_1}{\lambda p_1 + p_2}\right)^{A}}^{\mbox{$A$ home wins for $p_1$}}\quad
\overbrace{\left(\frac{p_2        }{\lambda p_1 + p_2}\right)^{B}}^{\mbox{$B$ away wins for $p_2$}}
}_\mbox{$p_1$ plays at home to $p_2$ $A+B$ times}\quad
\underbrace{
\overbrace{\left(\frac{\lambda p_2}{p_1 + \lambda p_2}\right)^{C}}^{\mbox{$C$ home wins for $p_2$}}\quad
\overbrace{\left(\frac{        p_1}{p_1 + \lambda p_2}\right)^{D}}^{\mbox{$D$ away wins for $p_1$}}
}_\mbox{$p_2$ plays at home to $p_1$ $C+D$ times}
$$

where we understand that $p_1+p_2=1$, $p_1,p_2,\lambda\geqslant 0$.
Elementary techniques allow us to identify a maximum and we find

$$
\hat{p_1}=\frac{\sqrt{AD}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{p_2}=\frac{\sqrt{AD}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{\lambda}=\sqrt{\frac{AC}{BD}}
$$

Note that the evaluate is unchanged if $A,B,C,D$ are all multiplied by
a positive constant.

## Aside: `hyper2` analysis of the same scoreline

Now the likelihood function would be

$$\mathcal{L}(p_1,p_2,\lambda;A,B,C,D)=
\underbrace{
\overbrace{\left(\frac{p_1+M}{p_1 + p_2+M}\right)^{A}}^{\mbox{$A$ home wins for $p_1$}}\quad
\overbrace{\left(\frac{p_2  }{p_1 + p_2+M}\right)^{B}}^{\mbox{$B$ away wins for $p_2$}}
}_\mbox{$p_1$ plays at home to $p_2$ $A+B$ times}\quad
\underbrace{
\overbrace{\left(\frac{p_2+M}{p_1 + p_2+M}\right)^{C}}^{\mbox{$C$ home wins for $p_2$}}\quad
\overbrace{\left(\frac{p_1  }{p_1 + p_2+M}\right)^{D}}^{\mbox{$D$ away wins for $p_1$}}
}_\mbox{$p_2$ plays at home to $p_1$ $C+D$ times}
$$

The maximum would be

$$
\hat{p_1}=\frac{C}{C+D}\qquad
\hat{p_2}=\frac{A}{A+B}\qquad
M=\frac{AD-BC}{(A+B)(B+C)}
$$

although it is not clear to me what the _constrained_ optimal point
[that is, requiring $M\geq 0$] is.

## Numerical analysis

```{r label=defineABCD}
A <- 9
B <- 5
C <- 7
D <- 3

(p1hat <- sqrt(A*D)/(sqrt(A*D) + sqrt(B*C)))
(p2hat <- sqrt(B*C)/(sqrt(A*D) + sqrt(B*C)))

(lam_hat <- sqrt((A*C)/(B*D)))
```

$$
\hat{p_1}=\frac{\sqrt{21}}{\sqrt{21} + \sqrt{35}}\simeq 0.47\qquad
\hat{p_2}=\frac{\sqrt{63}}{\sqrt{21} + \sqrt{35}}\simeq 0.53\qquad
\hat{\lambda}=\sqrt{\frac{AD}{BC}}=2.05.
$$

Keeping $A=9,B=5, C=7,D=3$, and $\lambda=2.05$ [assumed for the moment
to be known, estimation techniques are discussed later], we can
represent this information in a standard format:

```{r,label=usehomeaway3a}
library("hyper2",quietly=TRUE)
M <- matrix(c(NA, A+B*1i ,C+D*1i, NA),2,2,byrow=TRUE)
teams <- c("p1","p2")
dimnames(M) <- list("@home" = teams,"@away"=teams)
dimnames(M) <- list("@home" = teams,"@away"=teams)
print(M)
```

Above, object `M` has real parts being home wins and imaginary parts
being away wins.  Appropriate package idiom to specify a likelihood
function might be to use bespoke function `home_away3()`:

```{r,label=ha3calc,cache=TRUE}
(H <- home_away3(M,lambda=2.05))
```

Keeping $A=9,B=5, C=7,D=3$, and $\lambda=2.05$ [assumed for the moment
to be known, estimation techniques are discussed later], we may
estimate $p_1$ and $p_2$ using maximum likelihood, `maxp()` in package
idiom:

```{r,label=findtheevaluateh3,cache=TRUE}
H
maxp(H)
```

Further, we can test whether the players are in fact of equal strength:

```{r,label=testequalitynow,cache=TRUE}
equalp.test(H)
```

Showing convincingly that we may reject the null that $p_1=p_2$ and
assert that the players do in fact differ in strength, at least with
this value of $\lambda$.  Observe that a naive analysis would have
$p_1$ winning $w=A+C=`r A+C`$ games and losing $l=B+D=`r B+D`$ games,
from which we would obtain $\hat{p_1}=`r A+C`/`r A+B+C+D`$; a
likelihood ratio for $H_0\colon p_1=0.5$ would be

$$\frac
{
  {{w+l}\choose {w\, l}}\left(\frac{w}{w+l}\right)^w\left(\frac{l}{w+l}\right)^l
}{
  {{w+l}\choose {w\, l}}\left(\frac{1}{2}\right)^w\left(\frac{1}{2}\right)^l
}=\frac{2^{w+l}w^wl^l}{(w+l)^{w+l}}\simeq 1.94,
$$

not significant.

Now, how to estimate $\lambda$?

```{r,label=estimatelambda3,cache=TRUE}
f <- function(lambda){maxp(home_away3(M,lambda=lambda),give=TRUE)$value}
f(2)
f(2.1)
lam <- seq(from=1,to=10,len=17)
Supp <- sapply(lam,f)
```

```{r,label=plotsup}
plot(log(lam),Supp-max(Supp),type="b")
```

Or even

```{r,label=optimlambda,cache=TRUE}
op <- optimize(f,interval=c(2,3),maximum=TRUE)
op
```

We can proceed in two ways.  Firstly we can use $\hat{\lambda}$ directly:

```{r,ept3,cache=TRUE}
equalp.test(home_away3(M,lambda=op$maximum))
```

The flaw in this analysis is that it is conditional on the estimated
value of $\lambda$ [which is not too bad, IMO].  Secondly we can
perform a fully 2D analysis, which allows for the fact that the
evaluate might have a different value of $\lambda$ from the
$\hat{\lambda}$:

```{r,label=loopysuppcalc,cache=TRUE}
probs <- seq(from=0.1,to=0.8,len=17)
jj <- as.matrix(expand.grid(lam,probs))
S <- rep(NA,nrow(jj))
for(i in seq_len(nrow(jj))){
    lambda <- jj[i,1]
    p1 <- jj[i,2]
    p <- c(p1,1-p1)
    names(p) <- c("p1","p2")
   S[i] <- loglik(p=p,H=home_away3(M,lambda=lambda),log=TRUE)
}
S <- matrix(S,length(lam),length(probs))
```

```{r}
contour(lam,probs,S,levels=seq(from=-20,to=-10,by=0.5),xlab="lambda",ylab="p1")
abline(h=0.5)
```