### 2023-02-08 inst/home_advantage3.Rmd
ternary plot

---
title: "Quantifying the home advantage with `hyper3` objects"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{hyper3}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

$$\mathcal{L}\left(p_1,\ldots, p_n\right)=\prod_j\left(\sum_{i=1}^n w^i_jp_i\right)^{n_j}.$$

As a motivating example, suppose two teams (players) with
Bradley-Terry strengths $p_1,p_2$ play football where we quantify the
home-ground advantage with a term $\lambda$.  If $p_1$ plays at home
$a+b$ times with $a$ wins and $b$ losses, and plays away [so $p_2$ is
at home] $c+d$ times with $c$ wins and $d$ losses, then a sensible
likelihood function might be

$$\mathcal{L}(p_1,p_2,\lambda;A,B,C,D)=
\underbrace{
\overbrace{\left(\frac{\lambda p_1}{\lambda p_1 + p_2}\right)^{A}}^{\mbox{$A$ home wins for $p_1$}}\quad
\overbrace{\left(\frac{p_2        }{\lambda p_1 + p_2}\right)^{B}}^{\mbox{$B$ away wins for $p_2$}}
}_\mbox{$p_1$ plays at home to $p_2$ $A+B$ times}\quad
\underbrace{
\overbrace{\left(\frac{\lambda p_2}{p_1 + \lambda p_2}\right)^{C}}^{\mbox{$C$ home wins for $p_2$}}\quad
\overbrace{\left(\frac{        p_1}{p_1 + \lambda p_2}\right)^{D}}^{\mbox{$D$ away wins for $p_1$}}
}_\mbox{$p_2$ plays at home to $p_1$ $C+D$ times}
$$

where we understand that $p_1+p_2=1$, $p_1,p_2,\lambda\geqslant 0$.
Elementary techniques allow us to identify a maximum and we find

$$
\hat{p_1}=\frac{\sqrt{AD}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{p_2}=\frac{\sqrt{BC}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{\lambda}=\sqrt{\frac{AC}{BD}}
$$

Note that the evaluate is unchanged if $A,B,C,D$ are all multiplied by
a positive constant.

## Aside: `hyper2` analysis of the same scoreline

Now the likelihood function would be

$$\mathcal{L}(p_1,p_2,\lambda;A,B,C,D)=
\underbrace{
\overbrace{\left(\frac{p_1+M}{p_1 + p_2+M}\right)^{A}}^{\mbox{$A$ home wins for $p_1$}}\quad
\overbrace{\left(\frac{p_2  }{p_1 + p_2+M}\right)^{B}}^{\mbox{$B$ away wins for $p_2$}}
}_\mbox{$p_1$ plays at home to $p_2$ $A+B$ times}\quad
\underbrace{
\overbrace{\left(\frac{p_2+M}{p_1 + p_2+M}\right)^{C}}^{\mbox{$C$ home wins for $p_2$}}\quad
\overbrace{\left(\frac{p_1  }{p_1 + p_2+M}\right)^{D}}^{\mbox{$D$ away wins for $p_1$}}
}_\mbox{$p_2$ plays at home to $p_1$ $C+D$ times}
$$

The maximum would be

$$
\hat{p_1}=\frac{D}{C+D}\qquad
\hat{p_2}=\frac{B}{A+B}\qquad
\hat{M}=\frac{AC-BD}{(A+B)(C+D)}
$$

although it is not clear to me what the _constrained_ optimal point
[that is, requiring $M\geq 0$] is.

## Numerical analysis

```{r label=defineABCD}
A <- 9
B <- 5
C <- 7
D <- 3

(p1hat <- sqrt(A*D)/(sqrt(A*D) + sqrt(B*C)))
(p2hat <- sqrt(B*C)/(sqrt(A*D) + sqrt(B*C)))

(lam_hat <- sqrt((A*C)/(B*D)))
```

$$
\hat{p_1}=\frac{\sqrt{21}}{\sqrt{21} + \sqrt{35}}\simeq 0.47\qquad
\hat{p_2}=\frac{\sqrt{63}}{\sqrt{21} + \sqrt{35}}\simeq 0.53\qquad
\hat{\lambda}=\sqrt{\frac{AC}{BD}}=2.05.
$$

Moving to the monster formulation we have

$$
\hat{p_1}=\frac{3}{7+3} = 0.3\qquad
\hat{p_2}=\frac{5}{5+9}\simeq  0.36\qquad
\hat{M}=\frac{63-21}{14\cdot 10} = 0.3
$$

Keeping $A=9,B=5, C=7,D=3$, and $\lambda=2.05$ [assumed for the moment
to be known, estimation techniques are discussed later], we can
represent this information in a standard format:

```{r,label=usehomeaway3a}
library("hyper2",quietly=TRUE)
M <- matrix(c(NA, A+B*1i ,C+D*1i, NA),2,2,byrow=TRUE)
teams <- c("p1","p2")
dimnames(M) <- list("@home" = teams,"@away"=teams)
dimnames(M) <- list("@home" = teams,"@away"=teams)
print(M)
```

Above, object `M` has real parts being home wins and imaginary parts
being away wins.  Appropriate package idiom to specify a likelihood
function might be to use bespoke function `home_away3()`:

```{r,label=ha3calc,cache=TRUE}
(H <- home_away3(M,lambda=2.05))
```

Keeping $A=9,B=5, C=7,D=3$, and $\lambda=2.05$ [assumed for the moment
to be known, estimation techniques are discussed later], we may
estimate $p_1$ and $p_2$ using maximum likelihood, `maxp()` in package
idiom:

```{r,label=findtheevaluateh3,cache=TRUE}
H
maxp(H)
```

Further, we can test whether the players are in fact of equal strength:

```{r,label=testequalitynow,cache=TRUE}
equalp.test(H)
```

Showing convincingly that we may reject the null that $p_1=p_2$ and
assert that the players do in fact differ in strength, at least with
this value of $\lambda$.  Observe that a naive analysis would have
$p_1$ winning $w=A+C=`r A+C`$ games and losing $l=B+D=`r B+D`$ games,
from which we would obtain $\hat{p_1}=`r A+C`/`r A+B+C+D`$; a
likelihood ratio for $H_0\colon p_1=0.5$ would be

$$\frac
{
  {{w+l}\choose {w\, l}}\left(\frac{w}{w+l}\right)^w\left(\frac{l}{w+l}\right)^l
}{
  {{w+l}\choose {w\, l}}\left(\frac{1}{2}\right)^w\left(\frac{1}{2}\right)^l
}=\frac{2^{w+l}w^wl^l}{(w+l)^{w+l}}\simeq 1.94,
$$

not significant.

Now, how to estimate $\lambda$?

```{r,label=estimatelambda3,cache=TRUE}
f <- function(lambda){maxp(home_away3(M,lambda=lambda),give=TRUE)$value}
f(2)
f(2.1)
lam <- seq(from=1,to=10,len=17)
Supp <- sapply(lam,f)
```

```{r,label=plotsup}
plot(log(lam),Supp-max(Supp),type="b")
```

Or even

```{r,label=optimlambda,cache=TRUE}
op <- optimize(f,interval=c(2,3),maximum=TRUE)
op
```

We can proceed in two ways.  Firstly we can use $\hat{\lambda}$ directly:

```{r,ept3,cache=TRUE}
equalp.test(home_away3(M,lambda=op$maximum))
```

The flaw in this analysis is that it is conditional on the estimated
value of $\lambda$ [which is not too bad, IMO].  Secondly we can
perform a fully 2D analysis, which allows for the fact that the
evaluate might have a different value of $\lambda$ from the
$\hat{\lambda}$:

```{r,label=loopysuppcalc,cache=TRUE}
probs <- seq(from=0.1,to=0.8,len=17)
jj <- as.matrix(expand.grid(lam,probs))
S <- rep(NA,nrow(jj))
for(i in seq_len(nrow(jj))){
    lambda <- jj[i,1]
    p1 <- jj[i,2]
    p <- c(p1,1-p1)
    names(p) <- c("p1","p2")
   S[i] <- loglik(p=p,H=home_away3(M,lambda=lambda),log=TRUE)
}
S <- matrix(S,length(lam),length(probs))
```

```{r}
contour(lam,probs,S,levels=seq(from=-20,to=-10,by=0.5),xlab="lambda",ylab="p1")
abline(h=0.5)
```

#  Ternary representation

```{r,label=useternary}
library(Ternary)

par(mar = rep(0.2, 4))
TernaryPlot(alab = "p1", blab = "p2", clab = "M")

FunctionToContour <- function(p1,p2,M) {
pmax(-20,log(M+p1)*A + log(p2)*B - log(M+p1+p2)*(A+B+C+D) + log(p2+M)*C + log(p1)*D )
}

# Compute and plot colour tiles
values <- TernaryPointValues(FunctionToContour, resolution = 24L)
ColourTernary(values)

# Add contour lines
TernaryContour(FunctionToContour, resolution = 36L)
```

# Appendix

Here is a transcript of a mathematica session which establishes the
estimates above.  First the `hyper2` likelihood:

```
rhankin@rhrpi4:~ $ wolfram
Mathematica 12.2.0 Kernel for Linux ARM (32-bit)
Copyright 1988-2021 Wolfram Research, Inc.

In[1]:= L = Log[M+p1]*A +Log[p2]*B- Log[M+p1+p2]*(A+B+C+D) + Log[p2+M]*C + Log[p1]*D

Out[1]= D Log[p1] + A Log[M + p1] + B Log[p2] + C Log[M + p2] - (A + B + C + D) Log[M + p1 + p2]

In[2]:= LL = L /. {p2 -> 1-p1-M}

Out[2]= C Log[1 - p1] + B Log[1 - M - p1] + D Log[p1] + A Log[M + p1]

In[3]:= dp = D[LL,p1]//FullSimplify

            C      D         B          A
Out[3]=  ------- + -- + ----------- + ------
         -1 + p1   p1   -1 + M + p1   M + p1

In[4]:= dM = D[LL,M]//FullSimplify

              B          A
Out[4]= ----------- + ------
         -1 + M + p1   M + p1

In[5]:= Solve[{dp==0,dM==0},{p1,M}]

                   D            A C - B D
Out[5]=  {{p1 -> -----, M -> ---------------}}
                 C + D       (A + B) (C + D)

In[6]:= p2 = 1-p1-M /. {%} //FullSimplify

             B
Out[6]=  {{-----}}
           A + B

```

Now the `hyper3` likelihood:

```
rhankin@rhrpi4:~ $ wolfram
Mathematica 12.2.0 Kernel for Linux ARM (32-bit)
Copyright 1988-2021 Wolfram Research, Inc.

In[1]:= L = A*Log[lambda*p1] + B*Log[p2] -(A+B)*Log[lambda*p1 + p2] + C*Log[lambda*p2] + D*Log[p1] -(C+D)*Log[p1 + lambda*p2]

Out[1]= D Log[p1] + A Log[lambda p1] + B Log[p2] + C Log[lambda p2] - (A + B) Log[lambda p1 + p2] - (C + D) Log[p1 + lambda p2]

In[2]:= LL = L /. {p2 -> 1-p1}

Out[2]= B Log[1 - p1] + C Log[lambda (1 - p1)] + D Log[p1] + A Log[lambda p1] - (C + D) Log[lambda (1 - p1) + p1] -

>    (A + B) Log[1 - p1 + lambda p1]

In[3]:= dp1 = D[LL,p1] //FullSimplify

         B + C    A + D   (A + B) (-1 + lambda)    (C + D) (-1 + lambda)
Out[3]= ------- + ----- - --------------------- + -----------------------
        -1 + p1    p1     1 + (-1 + lambda) p1    lambda + p1 - lambda p1

In[4]:= dlam = D[LL,lambda]//FullSimplify

          A        C           (A + B) p1           (C + D) (-1 + p1)
Out[4]= ------ + ------ - -------------------- + -----------------------
        lambda   lambda   1 + (-1 + lambda) p1   lambda + p1 - lambda p1

In[5]:= Solve[{dp1==0,dlam==0},{p1,lambda}]//FullSimplify

[snip]

In[6]:= %[[1]]

                        1                     Sqrt[A] Sqrt[C]
Out[6]= {p1 -> -------------------, lambda -> ---------------}
                   Sqrt[B] Sqrt[C]            Sqrt[B] Sqrt[D]
               1 + ---------------
                   Sqrt[A] Sqrt[D]

```