### 2023-04-17 inst/ijmm_rebuttal.tex
start on rebuttal as per issue #197

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

\section*{An objective assessment of point scoring systems in Formula 1 motor racing: rebuttal}

Below, I reproduce the reviewers comments; my replies to the issues
are in \textcolor{blue}{blue}.  I have indicated changes to the
manuscript where appropriate.

Short story: I have accommodated all the comments with rewording and
clarification.  The resulting document is, I believe, stronger and
more scholarly than before and I recommend it to you.

\subsection*{Detailed response to comments}

Conclusion in Introduction: The final paragraph of section 1
(Introduction) seems to need a conclusion.

Points systems similar to that of Formula 1 are common in other racing
sports; all have the common feature of translating ranks into points
which combine additively to generate an overall ranking.  Further, we
see points systems used in the wider context of competitive situations
such as the Eurovision Song Contest and many other such group
tournaments.  \mbox{Therefore, ...}

\textcolor{blue}{Response to comment here}\\ \\

Addition of literature review of Zipf (1949)

\textcolor{blue}{Response to comment here}\\ \\

Many of our readers are unfamiliar with this work.  Thus, a short
passage will help in understanding the significance of the inclusion
of Zipf in the manuscript.

\textcolor{blue}{Response to comment here}\\ \\

The charts need to be larger for easier reading.  Please place figures
in full-page format as appendices. Share the actual charts rather than
pictures of the charts. Doing so will allow us to retain image quality
as figures are enlarged.

\textcolor{blue}{Response to comment here}\\ \\

Replace Pie Charts with Pareto Charts

\textcolor{blue}{Response to comment here}\\ \\

The pie charts are difficult to distinguish between components. For
example, in Figure 1 for 2017, the reviewers could not see a
difference between Bottas and Vettel.  The Pareto Chart style would
improve that aspect.

\textcolor{blue}{Response to comment here}\\ \\

\section*{Rankings in Section 2.1}

Place the comparisons of the rankings systems in a table.  Readers will
be able to make comparisons more easily.

\textcolor{blue}{Response to comment here}\\ \\

Table 2 and Figures 4-6 Consistency

Please use WTA or Top1 consistently in these items.

\textcolor{blue}{Response to comment here}\\ \\

Section 2.2 Numerical Results

Explain how the WTA system maximizes the probability of the complete
order statistic.

\textcolor{blue}{Response to comment here}\\ \\

Comment for Practitioners

Please include a summary of how the points system could benefit the
managers of Formula 1 and other series in the conclusion.

\textcolor{blue}{Response to comment here}\\ \\

\end{document}