### 2023-05-02 inst/simplified_likelihood.Rmd
simplified likelihood

---
title: "simplified_likelihood"
author: "Robin Hankin"
date: "2023-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Consider a race between $n+2$ competitors of strength $p_1$, $p_2$, and $n$ clones, all of strength $p=1-p_1-p_2$.  Our observation is that $p_1$ places $n_1$, and $p_2$ places $n_2>n_1$.  The order statistic is thus

$$
\underbrace{p\succ\cdots\succ p}_{\mbox{$n_1-1$ terms}}\succ p_1\succ\underbrace{p\succ\cdots\succ p}_{n_2-n_1-1}\succ p_2\succ\underbrace{p\succ\cdots\succ p}_{n-n_2+2}
$$

The Plackett-Luce likelihood function would be as follows

$$
\begin{eqnarray}
&{}&
\underbrace{
\frac{p}{p_1+p_2+np}\cdot
\frac{p}{p_1+p_2+(n-1)p}
\cdots
\frac{p}{p_1+p_2+(n-n_1+2)p}
}_{n_1-1}\cdot\\&{}&
\frac{p_1}{p_1+p_2+(n-n_1+1)p}\cdot\\&{}&
\underbrace{
\frac{p}{p_2+(n-n_1+1)p}\cdot
\frac{p}{p_2+(n-n_1)p}\cdots
\frac{p}{p_2+(n-n_2+3)p}
}_{n_2-n_1-1}\cdot\\&{}&
\frac{p_2}{p_2+(n-n_2+2)p}\cdot\\&{}&
\underbrace{
\frac{p}{(n-n_2+2)p}\cdot\frac{p}{(n-n_2)p}
\cdots\frac{p}{2p}\cdot\frac{p}{p}
}_{n-n_2+2}
\end{eqnarray}
$$

Observing that the fifth line may be neglected, R idiom for this would be as follows

```{r}
library("hyper2")
simplike <- function(n1,n2,n){
  H <- hyper3()
  H[c(p =1)] <- n2-2
  H[c(p1=1)] <- 1
  H[c(p2=1)] <- 1
  for(i in (n-n1+1):n)     { H[c(p1=1,p2=1,p=i)]  %<>% dec }
  for(i in (n-n2+2):(n-n1+1)){ H[c(p2=1,   p=i)]  %<>% dec }
  return(H)
} 
simplike(2,6,11)
simplike(1,6,11)
simplike(1,2,11)
```