### 2023-05-08 inst/jss_4382_rebuttal.tex
more work on issue #153

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

\section*{JSS4832: Generalized Plackett-Luce Likelihoods by Robin
K. S. Hankin}

Below, the reviewer's comments are in black, and the replies to the
issues are in \textcolor{blue}{blue}.  I have indicated changes to
the manuscript where appropriate.

Short story: I have accommodated all the comments with rewording and
clarification.

\subsection*{Reviewer's comments, 15 April 2023}

Dear Robin K.S. Hankin,

We have reached a decision regarding your submission to Journal of
Statistical Software, \verb+#4832+, ``Generalized Plackett-Luce
Likelihoods".

Unfortunately we have to tell you that we decided to reject in its current state.

However, we encourage you to resubmit the article after detailed
revision. Please check the journal management system for any
additional feedback:
https://www.jstatsoft.org/index.php/jss/authorDashboard/submission/4832. We
expect you to provide a point-by-point response to this feedback in
your revision.

Your sincerely,

The editorial team

The manuscript describes functionality of the hyper2 package for
Fitting and testing variants of the Placket-Luce ranking model.

1.1 Recommendation

The manuscript may be publishable in the journal after revisions have been
made.

1.2 General comments

G1 The introduction should provide an overview of the contents of the
manuscript. \\  \textcolor{blue}{New sentence added to manuscript:\\ \\
Here I consider such generalized Plackett-Luce likelihood functions,
and give an exact analysis of several simple cases.  I then show how
this class of likelihood functions may be applied to a range of
inference problems involving order statistics.  Illustrative examples,
drawn Formula 1 motor racing, and track-and-field athletics, are
given.}\\ \\

In addition, the section headlines could be made more
informative. Currently, they are:

2. ``Some simple cases"
3. ``Mann-Whitney test"
4. ``Multiple competitors: javelin"
5. ``Formula 1 motor racing: the constructors' championship"

Maybe the headlines could provide more of a summary of what the
section contains.  \textcolor{blue}{The headings are indeed somewhat
  uninformative and are now more in line with the remainder of the
  manuscript.  I have changed what was section 4 to a {\em sub}section
  with heading ``A generalization of the Mann-Whitney test using
  generalized Plackett-Luce likelihood'' which I think is in keeping
  with this comment.  The headings are now
  \begin{itemize}
  \item Exact analytical solution for some simple generalized
    Plackett-Luce likelihood functions
  \item An alternative to the Mann-Whitney test using generalized
    Plackett-Luce likelihood
  \item A generalization of the Mann-Whitney test using
    generalized Plackett-Luce likelihood (now a subsection)
  \item Formula 1 motor racing: the constructors' championship
\end{itemize}}
\rule{0mm}{10mm}

G2 There should be a comparison with other software.  Although the
author states in the introduction that with hyper2 ``It becomes
possible to analyse a wider range of likelihood functions than
standard Plackett-Luce", it is not clear to me which of the features
described later on are unique to the \verb+hyper2+ package and which
can also be obtained from other packages (like, e.g., PlackettLuce).

\textcolor{blue}{Although the hyper2 package has functionality not
  present in other implementations, the current manuscript focuses on
  the new generalization represented in the package by the hyper3
  class (that is, generalized Plackett-Luce likelihoods for order
  statistics with clones of identical Bradle-Terry strength).  Turner
  et. al. (2020), which I have cited, does a good job of setting out
  the functionality of hyper2 not present in other packages.  I think
  that an extended discussion of hyper2 unique functionality would be
  a distraction from the focus of the manuscript.  But I agree that a
  short statement documenting the unique aspects of the hyper2 package
  would be useful to JSS readers, and I have added a short paragraph
  to this effect.  Actually (in the spirit of full disclosure), I have
  added quite a lot of machinery to the package since Turner's 2020
  review and this gives me a welcome opportunity to mention some of
  it.}

G3 The manuscript should comment on the timing of the functions. For the
simple example on P. 4, the maxp(H) call takes more than 8 seconds on
my computer. This makes me wonder how timings would be for realistic
examples.

\textcolor{blue}{Any optimization technique becomes slow when used
  over a high-dimensional space and {\tt maxp()} is no exception.  The
  examples chosen were well within its capabilities.  In addition, the
  default implementation of {\tt maxp()} in the package uses a *very*
  conservative set of assumptions, for example the default is to use
  $n=10$ slightly differing start points and perform (by default)
  $n-10$ independent optimizations, to check for
  convergence.  I have come to realise that this is absurdly
  over-engineered for the examples considered in the manuscript.  If
  one is sure that the convergence is guaranteed (as it is in the
  examples considered here), $n=1$ is perfectly adequate.  For
  example, using the case given on P4:
  {\tt\\ \\
> (system.time(mH <- maxp(H,n=1)))\\
   user  system elapsed \\
  0.355   0.003   0.357 \\
> maxp(H,n=1)\\
         a          b          c \\
0.21324090 0.08724824 0.69951086 \\
  }\\ \\
  The latest package version on github includes a discussion of this
  issue and I have added a brief discussion (``function {\tt maxp()}
  uses standard optimization techniques to locate the evaluate; it has
  access to first derivatives of the log-likelihood and as such has
  rapid convergence, if its objective function is reasonably
  smooth").}\\ \\

1.3 Specific comments\\ S1 I'm not claiming any expertise in English
language, but occasionally, I found the manuscript difficult to
understand. For example, in the abstract it says: "The package
. . . furnishes idiom" Couldn't that simply be: "The package
. . . provides functionality"?

\textcolor{blue}{The original motivation for the package occurred when
  I had a large complicated set of Bradley-Terry observations which I
  wanted to translate into a likelihood function.  The R code became
  extremely long and unweildy, and was almost impossible to check.  It
  also became very slow to run.  What I wanted was a nice way to enter
  this kind of observations, in such a way as to maintain both
  computational efficiency [hence the use of the {\tt STL} map class]
  {\bf and} human readability.  I think that the phrase ``furnishes
  idiom" expresses this reasonably clearly.  But on reflection it does
  sound a bit odd and I have rephrased the abstract accordingly.}\\ \\

On P. 13, it says "Here we use generalized Plackett-Luce to assess the
constructors' performance". Isn't there a "model" missing?
(Plackett-Luce is just a name and cannot be used).
\textcolor{blue}{The original text was too terse and I have
  reworded.}\\ \\

S2 Figure 2: For readability, maybe use xlab = expression(p[a]) and maybe
instead of a legend, write AAB, ABA, BAA next to the line in the color
of that line.
\textcolor{blue}{Response here}\\ \\

S3 P. 7: "https:github.com/RobinHankin/hyper2" should be
"https://github.com/RobinHankin/hyper2".
\textcolor{blue}{Response here}\\ \\

S4 P. 7: "Wolf" should be "Wolfe".
\textcolor{blue}{Response here}\\ \\

S5 P. 15, maxp(const2020, n=1): What does n do?
\textcolor{blue}{Response here}\\ \\

S6 P. 16, "we see strong evidence for a real decrease in the strength": I'd tone
that down. The model is almost certainly wrong, so it does not measure
the "real decrease." Also, power is unknown, so it is difficult to say what
"evidence" there is against H0. (See, e.g., Gelman, Regression an other
stories, Chapter 4: Time to unlearn what you thought you knew about
statistics. https://statmodeling.stat.columbia.edu/2022/01/27/regression-
and-other-stories-free-pdf/)
\textcolor{blue}{Response here}\\ \\

S7 The references are lacking their DOIs.
\textcolor{blue}{Response here}\\ \\

2 Review of software

The package includes many functions with a similar name. E.g., ordervec2supp,
ordervec2supp3, ordervec2supp3a, ordertable2supp, ordertable2supp3. Maybe
there could be some overview of which to use for what purpose.

\textcolor{blue}{Response here}\\ \\

Section editor comments:

The author proposes an extension of the alreadty existing hyper2 package.
This new S3 class allows for some extended likelihood specifications.

\textcolor{blue}{Response here}\\ \\

Even though this is just a code snippet and the hyper2 package was already
published elsewhere, the author should provide some introductory elaborations
on the Plackett-Luce model. Along these lines it would also be helpful to give
a quick overview of the hyper2 package, and mention other related packages. As it
is now it is difficult for a reader who's not super familiar with PL-models and
hyper2, respectively, to follow the paper. Anyway, I find the runners example
as motivating example very illustrative, and actually like the examples throughout
the paper.

\textcolor{blue}{Response here}\\ \\

Some more aesthetic effort should be made in the code file (e.g., providing
comments). The code looks like a simple purl() extraction.

\textcolor{blue}{Response here}\\ \\

\textcolor{blue}{Summary: this referee's insightful and
  constructive comments have led to a much improved submission.}

\end{document}