### 2023-05-10 README.Rmd
typo

---
title: "The hyper 2 package"
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<img src="man/figures/hyper2.png" width = "150" align="right" />

<!-- badges: start -->
[![Total Downloads](https://cranlogs.r-pkg.org/badges/grand-total/hyper2)](https://CRAN.R-project.org/package=hyper2)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/hyper2)](https://cran.r-project.org/package=hyper2)
<!-- badges: end -->

# Overview

The `hyper2` package gives a suite of methods for creating and
analysing likelihood functions generated from a generalization of the
Bradley-Terry probability model.  The package uses the `STL` map class
of `C++` for efficiency; it conforms to `disordR` discipline.

# Installation

The released version of `hyper2` on [CRAN](https://CRAN.R-project.org)
may be installed with:

```{r, message=FALSE}
# install.packages("hyper2")  # uncomment this to install the package
library("hyper2")
```

# Further information

For more detail, see the package vignette; type

`vignette("hyper2")`

at the R command line.

# Citation

To cite the `hyper2` package in publications, please use

Robin K. S. Hankin 2017.  "Partial rank data with the `hyper2`
package: Likelihood functions for generalized Bradley-Terry models".
_The R Journal_, 9(2):429--439