### 2023-05-22 inst/simplified_likelihood.Rmd
correct some stupid mistakes

---
title: "simplified_likelihood"
author: "Robin Hankin"
date: "2023-05-02"
output: html_document
---

```{r setup, include=FALSE}
{ # this line vandalises this file: nonfinishers are ignored and should not be!
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

Consider a race between $n+2$ competitors of strength $p_1$, $p_2$,
and $n$ clones, all of strength $p=1-p_1-p_2$.  Our observation is
that $p_1$ places $n_1$, and $p_2$ places $n_2>n_1$.  The order
statistic is thus

$$
\underbrace{p\succ\cdots\succ p}_{\mbox{$n_1-1$ terms}}\succ p_1\succ\underbrace{p\succ\cdots\succ p}_{n_2-n_1-1}\succ p_2\succ\underbrace{p\succ\cdots\succ p}_{n-n_2+2}
$$

The Plackett-Luce likelihood function would be as follows

$$
\begin{eqnarray}
&{}&
\underbrace{
\frac{p}{p_1+p_2+np}\cdot
\frac{p}{p_1+p_2+(n-1)p}
\cdots
\frac{p}{p_1+p_2+(n-n_1+2)p}
}_{n_1-1}\cdot\\&{}&
\frac{p_1}{p_1+p_2+(n-n_1+1)p}\cdot\\&{}&
\underbrace{
\frac{p}{p_2+(n-n_1+1)p}\cdot
\frac{p}{p_2+(n-n_1)p}\cdots
\frac{p}{p_2+(n-n_2+3)p}
}_{n_2-n_1-1}\cdot\\&{}&
\frac{p_2}{p_2+(n-n_2+2)p}\cdot\\&{}&
\underbrace{
\frac{p}{(n-n_2+2)p}\cdot\frac{p}{(n-n_2)p}
\cdots\frac{p}{2p}\cdot\frac{p}{p}
}_{n-n_2+2}
\end{eqnarray}
$$

```{r}
named_drivers <- function(n,...){ # twodrivers(20, ham=3, bottas=5)  # NB repeats allowed!
   a <- list(...)
   stopifnot(all(table(as.vector(unlist(a)))==1)) # stops two players having the same result
   s <- rep("other",n)
   for(i in seq_along(a)){
     s[a[[i]]] <- names(a)[i]   # ham=3 -> s[3] <- "ham"
   }
   return(ordervec2supp3(s))  
}
```

```{r}
(H <- named_drivers(20,ham=3,bott=5))
```

Thus `H` is a likelihood function on the assumption that the other
drivers are all clones of identical strength.  Maximizing this:

```{r readtable}
a <- read.table("formula1_2019.txt")[1:2,1:21]
a
f <- function(i){
  jj <- as.list(a[,i])
  suppressWarnings(jj <- as.numeric(jj))
  names(jj) <- rownames(a)
  jj[!is.na(jj)]   # NA entries ignored - THESE SHOULD BE NONFINISHERS
}

H <- hyper3()
for(i in seq_len(ncol(a))){
  jj <- f(i)
  if(all(!is.na(as.numeric(unlist(jj))))){
  }
}
H
```

```{r maxh,cache=TRUE}
maxp(H)
```
Thus we narrowly fail to reject the null at $\alpha = 0.05$.  Now
compare with the actual 2019 season:

```{r actual2019,cache=TRUE}
F1_2019 <- ordertable2supp(read.table("formula1_2019.txt",header=TRUE)[,1:21])
samep.test(F1_2019,c("Hamilton","Bottas"))
```

We see a much smaller p-value when using the complete dataset.  It is
not clear whether we would expect a larger or a smaller p-value here,
but it is not surprising to see a difference.  On the simplified case,
the estimated strength of `other` is about 0.016 (under the
alternative); but in the full case, most of the other drivers have
strength considerably more thn this.  I do not have a good way of
thinking about this at this time.  One other feature is that the
simplified likelihood results in different MLEs:

* simplified: $\widehat{p_\mbox{Ham}} = 0.49$, $\widehat{p_\mbox{Bot}}=0.31$
* full: $\widehat{p_\mbox{Ham}} = 0.38$, $\widehat{p_\mbox{Bot}}=0.13$

Consider the log-contrasts, $\log(p_1/p_2)$ giving