### 2023-10-29 inst/eurodance.Rmd
label each chunk

---
title: "Eurovision **Dance** Contest 2008"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
abbreviated <- TRUE  # change to FALSE for full names
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

The file `eurodance2008.txt`, used below, is copied from
"Eurovision Dance Contest 2008", Wikipedia, accessed March 31, 2021.
First we specify the matrix as appearing in the Wikipedia page:

```{r, specifymatrix}
eurodance_table <- as.matrix(read.table("eurodance2008.txt"))
eurodance_table
o <- eurodance_table # saves typing
```

Each row corresponds to a contestant and each column to a judge.  The
points system used was:

```{r,pointsforplaces}
points_voters <- c(12,10,8,7,6,5,4,3,2,1)
points_jury   <- 4*points_voters
ov <- o[,-(1:2)] # ='o' but just the voters: exclude total and jury
pref <- ov*0  # retains NA
for(i in seq_along(points_voters)){pref[ov == points_voters[i]] <- i}
pref
```

Now we treat each column as an order statistic:

```{r,label=eachcolisorder,cache=TRUE}
eurodance <- hyper2()
for(i in seq_len(ncol(pref))){
	x <- pref[,i]
	eurodance %<>% `+`(ordervec2supp(x[!is.na(x)]))
}  # i loop closes
eurodance_maxp <- maxp(eurodance)
eurodance_maxp
pie(eurodance_maxp)
```

Note the dominance of Poland.  We verify consistency:

```{r,label=verifyconsistency,cache=TRUE}
consistency(eurodance)
```

Now we consider the jury's vote:

```{r,label=considerjury}
(jj <- o[,2])
jj <- sort(jj,decreasing=TRUE)
jj[jj>0] <- seq_len(sum(jj>0))
jury <- jj
jury
```

Now the question is whether the partial rank conferred by the jury is
consistent with that of the voters.  The permutation test technique
used in `cook.Rmd` is not appropriate here as there are $14!/4!\simeq
3.6\times 10^9$ permutations to consider; we will use a randomised
sampling method.  Below, we use a resampling technique: 4000
randomised jury observations are created, and the likelihood [under
the Plackett-Luce likelihood function for the order statistics of the
nationals' voters] is calculated.  The $p$-value is then the
probability of observing the actual observation or an observation more
unlikely under this likelihood function; it is operationally identical
to a simulated Fisher's exact test.

```{r,label=simfish,cache=TRUE}
n <- 4000
LL <- rep(0,n)
for(i in seq_len(n)){
	jjstar <- sample(jury)
	names(jjstar) <- names(jury)
	LL[i] <- loglik(indep(eurodance_maxp),ordervec2supp(jjstar),log=FALSE)
}
```

```{r,label=comparelogliks}
plot(sort(LL),log="y")
p <- loglik(indep(eurodance_maxp),ordervec2supp(jury),log=FALSE)
abline(h=p)
```

```{r,label=sumll}
sum(LL[LL<p])/sum(LL)
```

The $p$-value of about 2.4\% shows that there is strong evidence to
suggest that the expert jury's ordering differs from that of the
national popular vote.

### Package dataset {-}

Following lines create `eurodance.rda`, residing in the `data/` directory of the package.

```{r,label=saveeurodancedataset}
save(eurodance_table,eurodance_maxp,eurodance,file="eurodance.rda")
```