### 2023-10-30 inst/fatigue.Rmd
label all chunks

---
title: "fatigue"
author: "Robin Hankin"
date: "16/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

## R Markdown

Analysis of  the 1973 interzonal competition

```{r readdata}
jj <- read.table("interzonal1973_petropolis.txt",header=TRUE,skip=2)
a <- data.frame(day=jj$day + 31*(jj$month=="August"),white=jj$white,black=jj$black,result=jj$result,round=jj$round,moves=jj$moves)

plot(sort(unique(a$day)),type='b',pch=c(1,1,16))
tail(a)
```

```{r,label=define_H,cache=TRUE}
players <-  unique(sort(c(a$white,a$black)))

wm <- "wm" 
dm <- "dm"
rm <- "rm"

tab <- matrix(0,length(players),3)
colnames(tab) <- c("won","lost","drawn")
rownames(tab) <- players

H <- hyper2()
for(i in seq_len(nrow(a))){
  white_player <- a$white[i]
  black_player <- a$black[i]
  result <- a$result[i]
  H[c(white_player,wm,black_player,dm)] %<>% dec()
  if(result == "1-0"){ # white win, black loss
    H[c(white_player,wm)] %<>% inc()
    tab[white_player,] <- tab[white_player,] + c(1,0,0)
    tab[black_player,] <- tab[black_player,] + c(0,1,0)
  } else if(result == "0-1"){ # white loss, black win
    H[black_player] %<>% inc()
    tab[white_player,] <- tab[white_player,] + c(0,1,0)
    tab[black_player,] <- tab[black_player,] + c(1,0,0)
  } else if(result == "1/2-1/2"){  # draw
    H[dm] %<>% inc()
    tab[white_player,] <- tab[white_player,] + c(0,0,1)
    tab[black_player,] <- tab[black_player,] + c(0,0,1)

  } else {
    stop("this cannot happen")
  }
}
```

```{r,label=calculatemaxp,cache=TRUE}
mH <- maxp(H,n=100)
```

```{r,testwhitezero,cache=TRUE}
specificp.gt.test(H,"wm",0)
```

```{r,summaryofH}
head(H)
pnames(H)

pie(mH)
dotchart(mH)
mH
summary(H)
```

```{r maketabs}
tab <- cbind(tab,played=rowSums(tab))
tab <- cbind(tab,points=tab[,1] + tab[,3]/2)
tab
```

```{r dostudentttest}
table(a$result)
hist(a$moves)
plot(moves~as.factor(result),data=a)
result
moves_drawn_games <- a$moves[a$result == "1/2-1/2"]
moves_won_games   <- a$moves[a$result != "1/2-1/2"]
t.test(moves_drawn_games,moves_won_games)
```

```{r makeo}
head(a,10)
o <- matrix(0,max(a[,5]),length(players))
dimnames(o) <- list(round=paste("r",seq_len(nrow(o)),sep=""),player=players)

for(i in seq_len(nrow(a))){
	o[a[i,5],which(players==a[i,2])] %<>% `+`(1)
	o[a[i,5],which(players==a[i,3])] %<>% `+`(1)
}
o
```

Now focus on pairs of successive rounds: 1,2; 4,5; 7,8; 10,11; 13,14; 16,17

```{r justsuccessive}
unlist(a[2,])
`fatigued` <- function(i){
  x <- unlist(a[i,])
  day <- as.numeric(a[i,1])
  white_player <- a[i,2]
  black_player <- a[i,3]
  result <- a[i,4]
  round <- as.numeric(a[i,5])
  out <- c(white_player_fatigued=FALSE,black_player_fatigued=FALSE)

  if(round%%3 == 1){return(out)}  # yesterday was a break day
  out <- c(white_player_fatigued=FALSE,black_player_fatigued=FALSE)
  b <- a[a$round==round-1,]  # b = yesterday's schedule
  yesterdays_white_players <- unique(sort(c(b[,2])))
  yesterdays_black_players <- unique(sort(c(b[,3])))

  if(white_player %in% yesterdays_white_players){  # he played white yesterday...
    if(b[which(white_player==b[,2]),4] !="1/2-1/2"){  # ...and didn't draw...
      out["white_player_fatigued"] <- TRUE # ... the white player is fatigued
    }
  }

  if(white_player %in% yesterdays_black_players){  # he played black yesterday...
    if(b[which(white_player==b[,3]),4] !="1/2-1/2"){  # ...and didn't draw...
      out["white_player_fatigued"] <- TRUE # ... the white player is fatigued
    }
  }

  if(black_player %in% yesterdays_white_players){  # he played white yesterday...
    if(b[which(black_player==b[,2]),4] !="1/2-1/2"){  # ...and didn't draw...
      out["black_player_fatigued"] <- TRUE # ... the black player is fatigued
    }
  }

  if(black_player %in% yesterdays_black_players){  # he played black yesterday...
    if(b[which(white_player==b[,3]),4] !="1/2-1/2"){  # ... and didn't draw...
      out["black_player_fatigued"] <- TRUE # ... the white player is fatigued
    }
  }

  return(out)
}

rm(H)
Hf <- hyper2() # H fatigue
for(i in seq_len(nrow(a))){
  white_player <- a[i,2]
  black_player <- a[i,3]
  result <- a[i,4]

  jj <- fatigued(i)
  white_player_fatigued <- jj[1]
  black_player_fatigued <- jj[2]

  if(white_player_fatigued & black_player_fatigued){  # both fatigued: no fatigue monster
    Hf[c(white_player,wm,black_player,dm)] %<>% dec()
    if(result == "1-0"){ # white win, black loss
      Hf[c(white_player,wm)] %<>% inc()
    } else if(result == "0-1"){ # white loss, black win
      Hf[c(black_player)] %<>% inc()
    } else if(result == "1/2-1/2"){ # draw
      Hf[dm] %<>% inc()
    } else {
      stop("result must be 0-1, 1-0, or 1/2-1/2")
    }
  } else if (white_player_fatigued & !black_player_fatigued){
    Hf[c(white_player,wm,black_player,dm,rm)] %<>% dec()
    if(result == "1-0"){ # white win, black loss
      Hf[c(white_player,wm)] %<>% inc()
    } else if(result == "0-1"){ # white loss, black win
      Hf[c(black_player,rm)] %<>% inc()
    } else if(result == "1/2-1/2"){ # draw
      Hf[dm] %<>% inc()
    } else {
      stop("result must be 0-1, 1-0, or 1/2-1/2")
    }
  } else if (!white_player_fatigued & black_player_fatigued){
    Hf[c(white_player,wm,black_player,dm,rm)] %<>% dec()
    if(result == "1-0"){ # white win, black loss
      Hf[c(white_player,wm,rm)] %<>% inc()
    } else if(result == "0-1"){ # white loss, black win
      Hf[c(black_player)] %<>% inc()
    } else if(result == "1/2-1/2"){ # draw
      Hf[dm] %<>% inc()
    } else {
      stop("result must be 0-1, 1-0, or 1/2-1/2")
    }
  } else if (!white_player_fatigued & !black_player_fatigued){ # neither fatigued
    Hf[c(white_player,wm,black_player,dm)] %<>% dec()
    if(result == "1-0"){ # white win, black loss
      Hf[c(white_player,wm)] %<>% inc()
    } else if(result == "0-1"){ # white loss, black win
      Hf[c(black_player)] %<>% inc()
    } else if(result == "1/2-1/2"){ # draw
      Hf[dm] %<>% inc()
    } else {
      stop("result must be 0-1, 1-0, or 1/2-1/2")
    }
  } else {
    stop("This is a logical impossibility")
  }
}
```

now play with it:

```{r,calchfmax,cache=TRUE}
Hf_maxp <- maxp(Hf)
```

```{r showsum}
summary(Hf)
pnames(Hf)
Hf_maxp
pie(Hf_maxp)
```

```{r maketable}
jj <- rep(0,nrow(a))
M <- data.frame(
white_player=rep(NA,nrow(a)),
black_player=rep(NA,nrow(a)),
white_player_fatigued=FALSE,
black_player_fatigued=FALSE
)
for(i in seq_len(nrow(a))){
	M[i,1] <- a[i,2]	
	M[i,2] <- a[i,3]	
	jj <- fatigued(i)
	M[i,3] <- jj[1]
	M[i,4] <- jj[2]
}
colnames(M) <- c("white","black","white_fatigued","black_fatigued")
tail(M)
table(M[,3])
table(M[,4])
```