### 2023-10-30 inst/T20.Rmd
name all chunks and cache the slow ones

---
title: "The impact of winning the coin toss on the Indian Premier League"
author: "Tanvi Chandel and Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: |
  %\VignetteIndexEntry{T20}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

# The Indian Premier League

This script creates and analyses hyper2 object `T20` which is a
likelihood function for the strengths of the competitors in the Indian
Premier League.

```{r kickoff}
library("hyper2",quietly=TRUE)
T20_table <- read.table("T20.txt",header=TRUE)
head(T20_table)
nrow(T20_table)
```

Dataframe `T20_table` has one row for each T20 IPL match in the
period 2008-2017 with the exception of three no-result matches and
seven tied matches, which were removed.  Object `T20` is a
likelihood function for the strengths of the 13 teams, and
`T20_toss` is a likelihood function that also includes a toss
strength term.

some details are given in the package help file
`T20.Rd` (type `?T20` at the R prompt).

## One-way univariate analysis

What is the probability of a team winning the match, given that they
won the toss?

```{r onewayanal}
x <- table(T20_table$toss_winner==T20_table$match_winner)
x
binom.test(x,alternative="less")$p.value
```

Binomial test against a null of $p=0.5$ is thus not significant: there
is no evidence that winning the toss increases one's probability of
winning the match.  Now we consider the decision (to bat first or to
field first) made by the toss winner:

```{r showtable}
table(T20_table$toss_decision)
```

a clear majority elect to field first. We can now ask what the
probability of the batting team winning the match is:

```{r pbtw}
attach(T20_table)
bat_first_wins <- 
((toss_winner==match_winner)&(toss_decision=='bat'  )) | 
((toss_winner!=match_winner)&(toss_decision=='field'))
sum(bat_first_wins)
field_first_wins <- 
((toss_winner==match_winner)&(toss_decision=='field'  )) | 
((toss_winner!=match_winner)&(toss_decision=='bat'))
sum(field_first_wins)
detach(T20_table)
```

as a consistency check we have

```{r conscheck}
sum(field_first_wins) + sum(bat_first_wins)
nrow(T20_table)
```
agreeing as expected.

## Two-way contingency tables

We can do some contingency tables:

```{r twowaycontables}
M <- table(
decision           = T20_table$toss_decision,
won_toss_won_match = T20_table$toss_winner==T20_table$match_winner)
dimnames(M) <- list(
decision=c('bat first','field first'),
won_match =c(FALSE,TRUE))
M
```

We may reject homogeneity of proportion (Fisher's exact test, $p= `r round(fisher.test(M)$p.value,3)`$).  Recalling that a slight majority
of toss winners elect to field first (361 out of 633), we see that
this is a consistent result.

# Strengths of the individual teams

We can now consider the matches as Bernoulli trials and try to infer
the teams' strengths.  The first step is to calculate a likelihood
function for the strengths:

```{r indstrength}
attach(lapply(T20_table[, ], as.vector))  # vector access to team1, etc
T20 <- hyper2()
for(i in seq_along(team1)){
   T20[match_winner[i]] %<>% inc
   T20[c(team1[i],team2[i])] %<>% dec

}		       
T20
```

and maximization is straightforward:

```{r maxisstraight,cache=TRUE}
T20_maxp <- maxp(T20)
dotchart(T20_maxp)
```

So we can test the null that all the teams have the same strength:

```{r nullallteams}
a1 <- maxp(T20,give=TRUE)$value        # likelihood at the evaluate
a2 <- loglik(indep(equalp(T20)),T20)  # likelihood at p1=p2=...=p14
a1-a2
```

and the above would correspond to a support $\Lambda$ of about 14.5,
or a likelihood ratio of about $e^\Lambda\simeq 2\times 10^6$.  This
is not significant by Edwards's criterion of two units of support per
degree of freedom (we have 13df here).  Alternatively, we might
observe that the statistic $2\Lambda$ is in the tail region of its
asymptotic distribution $\chi^2_{13}$ given by Wilks, with a p-value
of

```{r wilkstest}
pchisq(2*14.5,df=13,lower.tail=FALSE)
```

showing a significant result although frankly Edwards's criterion is
somewhat flaky in this context.  Wilks's theorem is only asymptotic
and it is not clear that the actual sampling distribution is close to
its asymptotic limit of $\chi^2$.  However, it turns out that the
asymptotic approximation is good in this particular case:
`inst/T20_analysis.Rmd` presents some analysis (it takes a very
long time to run which is why it is not included in this vignette).

# The effect of the toss

We can now attempt to see if the toss makes a difference:

```{r tossdiff}
T20_toss  <- hyper2(pnames=c('toss',levels(T20_table$team1)))
for(i in seq_along(team1)){
      players <- c(team1[i],team2[i])
      if(toss_winner[i] == match_winner[i]){ # win toss, win match
         T20_toss[c('toss',match_winner[i])] %<>% inc
      } else { # win toss, lose match
         T20_toss[match_winner[i]] %<>% inc
      }			 
   T20_toss[c('toss',players)] %<>% dec
   T20_toss
}			  
```
and again we can maximize the likelihood:

```{r maxlike,cache=TRUE}
T20_maxp_toss <- maxp(T20_toss)
dotchart(T20_maxp_toss)
```

Above, note the small estimated strength of the toss.

### Package dataset

The following lines create `T20.rda`, residing in the `data/`
directory of the package.

```{r,label=saveT20dataset}
save(T20_table,T20,T20_maxp,T20_toss,file="T20.rda")
```