### 2023-10-30 inst/gradient.Rmd

---
title: "Numerical verification of function `gradientn()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2",quietly=TRUE)
set.seed(0)
```

We will numerically verify the `gradientn()` function for two `hyper2`
objects.

## Chess dataset

The chess dataset is as follows:

```{r showchess}
chess
```

We will define two gradient functions: `chess_gradient_theoretical()`,
which calculates the partial derivatives using straightforward algebra
(specifically elementary calculus, $\partial(x+y)^n/\partial
x=n(x+y)^{n-1}$), and `chess_gradient_theoretical()`, which uses `C` code:

```{r chessgrad}
chess_gradient_theoretical <- function(x){
    Topalov <- x[1]
    Anand   <- x[2]
    Karpov  <- x[3]

    out <- c(
        Topalov =  30/Topalov -35/(Topalov+Anand) -18/(Topalov+Karpov),
        Anand   = -35/(Topalov+Anand) +36/Anand -35/(Anand+Karpov),
        Karpov  = -18/(Topalov+Karpov) -35/(Anand+Karpov) +22/Karpov
    )

    names(out) <- pnames(chess)
    out

}
chess_gradient_numerical  <- function(x){gradientn(chess,x)}
```

So in the above we have two ways of calculating the gardient: one
theoretical and one numerical.  Do they agree?

```{r pchessagree}
p_chess <- 
	c(
		Topalov = 0.5,
		Anand   = 0.3,
		Karpov  = 0.2
	)
chess_gradient_theoretical(p_chess) - chess_gradient_numerical(p_chess)
```

Yes, they agree

# Icons

The `icons` dataset is:

```{r showicons}
icons
```

We can do the same thing but the algebra is more involved:

```{r samethingalgebra}
icons_gradient_theoretical <- function(x){
    NB <- x[1]
    L <- x[2]
    PB <- x[3]
    THC <- x[4]
    OA <- x[5]
    WAIS <- x[6]

	out <- c(
		NB = (
			+32/NB
            -20/(NB + L + THC + OA)
            -15/(NB + L + THC + WAIS)
            -09/(NB + L + OA + WAIS)
            -18/(NB + PB + THC + OA)
            -18/(NB + PB + THC + WAIS)
            -08/(NB + PB + OA + WAIS)
        ),
        L = (
            -20/(NB + L + THC + OA)
            -15/(NB + L + THC + WAIS)
            -09/(NB + L + OA + WAIS)
            +24/L
            -11/(L + PB + THC + OA)
            -16/(L + PB + THC + WAIS)
            -18/(L + PB + OA + WAIS)
        ),
        PB = (
            -18/(NB + PB + THC + OA)
            -18/(NB + PB + THC + WAIS)
            -08/(NB + PB + OA + WAIS)
            -11/(L + PB + THC + OA)
            -16/(L + PB + THC + WAIS)
            -18/(L + PB + OA + WAIS)
            +30/PB
        ),
        THC = (
            -20/(NB + L + THC + OA)
            -15/(NB + L + THC + WAIS)
            -18/(NB + PB + THC + OA)
            -18/(NB + PB + THC + WAIS)
            -11/(L + PB + THC + OA)
            -16/(L + PB + THC + WAIS)
            +24/THC
        ),
        OA = (
            -20/(NB + L + THC + OA)
            -09/(NB + L + OA + WAIS)
            -18/(NB + PB + THC + OA)
            -08/(NB + PB + OA + WAIS)
            -11/(L + PB + THC + OA)
            -18/(L + PB + OA + WAIS)
            +14/OA
        ),
        WAIS = (
            -15/(NB + L + THC + WAIS)
            -09/(NB + L + OA + WAIS)
            -18/(NB + PB + THC + WAIS)
            -08/(NB + PB + OA + WAIS)
            -16/(L + PB + THC + WAIS)
            -18/(L + PB + OA + WAIS)
            +9/WAIS
        )
    )
	names(out) <- pnames(icons)
	return(out)
}

icons_gradient_numerical  <- function(x){gradientn(icons,x)}
```

Then

```{r isicons}
p_icons <- c(NB=0.3, L=0.1, PB=0.2, THC=0.15, OA=0.15, WAIS=0.1)
```

```{r comparegradientwithwithout}
icons_gradient_theoretical(p_icons)
icons_gradient_theoretical(p_icons) - icons_gradient_numerical(p_icons)
```

also agreeing.

## Difference between `gradientn()` and `gradient()`

Taking the chess dataset we have

```{r lookatchess}
rbind(indep(p_chess), gradient(chess,indep(p_chess)))
rbind(p_chess,gradientn(chess,p_chess))
```

Indeed, because the powers of `chess` sum to zero, we can test the
fact that the derivative in the direction parallel to `p_chess` is
zero:

```{r isderivzero}
sum(powers(chess))
sum(p_chess*gradientn(chess,p_chess))
```

which is correct to numerical precision.  Further,

```{r seesumpower}
sum(powers(icons))
sum(p_icons*gradientn(icons,p_icons))
```

## Case study: icons

Consider the `icons` likelihood function; we wonder whether the
loglikelihood function has a well-defined maximum.

```{r ishessian}
(M <- hessian(icons))
is_ok_hessian(M)
```

This would suggest that the maximum is defined; we may ask how sharp
it is:

```{r isnothessian}
is_ok_hessian(M,give=TRUE)
```

suggesting a reasonably sharp constrained maximum point.  We now
demostrate that the Hessian matrix of second derivatives is
numerically correct.  First the unconstrained problem, working with
the first five elements.  The first step is to look at the likelihood,
the gradient, and the Hessian, all evaluated at a certain point:

```{r evaluatehessian}
(p <- indep(equalp(icons))) # NB not the evaluate
(G <- gradient(icons,p))
(M <- hessian(icons,p,border=FALSE))
eigen(M,TRUE,TRUE)$values
```

```{r smallpert}
dp <- rnorm(5)*1e-4        # small perturbation
dL <- loglik(p+dp,icons) - loglik(p,icons)  # delta loglikelihood (exact)

dL1 <- sum(gradient(icons,p)*dp)    # first order approximation
dL2 <- dL1 + t(dp) %*% M %*% dp/2   # second order; should be quad.form(dp,M)

c(dL,dL1,dL2)

c(first_order_error=dL-dL1,second_order_error=dL-dL2)

```

In the above, see how the first order approximation is pretty good: it
is of ${\mathcal O}(f''(x)\delta x^2)$ which here would be about
$1000\times 10^{-8}=10^{-5}$ which is pretty much what we observe [the
1000 is the typical size of the elements of the Hessian].  The
second-order approximation is better, having a smaller error ... but
not that much better.  I would expect the error to be ${\mathcal
O}(f'''(x)\delta x^3)$ but maybe the third order derivatives are
larger than one might expect.  We can perform a similar analysis but
using bordered Hessians:

```{r useborderedhess}

dpc <- fillup(dp,0)   # constrained delta p

MC <- hessian(icons,p,border=TRUE)
dLC1 <- sum(gradientn(icons,fillup(p))*dpc)    # first order approximation
dLC2 <- dLC1 + t(dpc) %*% MC %*% dpc/2     # second order

c(dL,dLC1,dLC2)

c(first_order_error=dL-dLC1,second_order_error=dL-dLC2)

```

showing similar results.