### 2023-10-30 inst/skeleton.Rmd

---
title: "skeleton"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
library("igraph")
options("digits" = 5)
```

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

The [skeleton](https://en.wikipedia.org/wiki/Skeleton_(sport)) is a
winter sliding sport.  This short document discusses a dataset taken
from the [2018 Winter
Olympics](https://en.wikipedia.org/wiki/Skeleton_at_the_2018_Winter_Olympics_%E2%80%93_Men%27s)
and analyses the order statistic for the competitors.

```{r readskeletontable}
(skeleton_table <- read.table("skeleton2022.txt",header=TRUE))
```

Thus the top line shows that Grotheer came first in runs 1-3 and sixth
in run 4.  The first column shows the results for run 1: we see that
Grotheer came first, Jungk came fifth, and so on.  There is in fact no
dominated set: Timmings [place 25] beats Enzie [place 19] (in run 4),
Enzie [19] beats Crumpton [22] (run 1), Crumpton [17] beats Wyatt [18]
(2), Wyatt [16] beats Maier [15] (1), Maier [11] beats Bagnis [12]
(4), Bagnis [10] beasts Seunggi [11] (1), Seunggi [9] beats T. Dukurs
[10] (3 and 4), Dukurs [7] beats Tretyakov [8] (2), Tretyakov [2]
beats Wengang [3] (1), and Wengang beats everybody (4).

We can *try* to automate the above cycle as follows.  The first step
is to convert `skeleton_table` to a numeric matrix, which proves to be
a very awkward process.  NB: `data.matrix()` buggers things up:

```{r,label=datamatrixfucksthingsup}
head(data.matrix(skeleton_table))
```

Above, we see that Grotheer apparently came 17th in run 4, clearly
some weird infelicity.  OK, how to coerce our dataframe to a numeric
matrix?  Possible but _extremely_  klunky:

```{r,showextremerank}
st <- matrix(as.numeric(c(as.matrix(skeleton_table))),ncol=4)
rownames(st) <- rownames(skeleton_table)
colnames(st) <- colnames(skeleton_table)
st
st[is.na(st)] <- 999
```

Taking Maier as an example (a mid-range competitor) we want to
identify competitors who Maier beats at least once:

```{r identifycompetitors}
bm1 <- st[,1] > st["Maier",1]  # lost to Maier in run 1
bm2 <- st[,2] > st["Maier",2]  # lost to Maier in run 2
bm3 <- st[,3] > st["Maier",3]  # lost to Maier in run 3
bm4 <- st[,4] > st["Maier",4]  # lsot to Maier in run 4
st[bm1 | bm2 | bm3 | bm4,]
```

Above we see a subset of rows of  `skeleton_table` corresponding to
competitors who lost to Maier.

```{r losttomaier}
rownames(st[bm1 | bm2 | bm3 | bm4,])
jj <- cbind(won="Maier",lost=rownames(st[bm1 | bm2 | bm3 | bm4,]))
noquote(jj)
plot(graph(c(t(jj))))
```

Above we see a directed graph, the arrow means "beats", so for example
Maier beats Timmings (2 o'clock).  We can functionify the calculation
and define `beats()`:

```{r definebeats}
beats <- function(p){   # returns everyone whom competitor 'p' beats
  bm1 <- st[,1] > st[p,1]  
  bm2 <- st[,2] > st[p,2]  
  bm3 <- st[,3] > st[p,3] 
  bm4 <- st[,4] > st[p,4] 
  jj <- cbind(p,rownames(st[bm1 | bm2 | bm3 | bm4,]))
  return(c(t(jj)))
}
bb <- c()
for(i in rownames(st)){bb <- c(bb,beats(i))}
graph(bb)
plot(graph(bb))
```

Above we see a directed graph.  The aim was to detect dominated sets,
that is, a subset of competitors who are always beaten by everyone
outside the group.  But although there is no such group in
`skeleton_table`, this fact is not immmediately obvious from the
graph.  We can however consider just the top and bottom few
competitors and see if that makes a difference.

```{r excludemidrange}
(st2 <- st[-(4:19),])  # exclude a bunch of mid-range competitors
```

Above, it is pretty clear that there is a dominated set of everyone
except the top 3.

```{r justtop3}
bb <- c()
for(i in rownames(st2)){bb <- c(bb,beats(i))}
graph(bb)
plot(graph(bb),layout = layout.fruchterman.reingold)
```

Above we see that the existence of the dominated set is not obvious
without close scrutiny.  Observing that Grotheer, Jungk, and Wengang
dominate the other competitors, we can see that these three have
arrows pointing to each other within the group, but noone outside the
group has an arrow pointing to any member of the set.  But this is not
obvious from inspection.

# Rank table analysis {-}

We can present the result table in another way:

```{r showranktable}
wikitable_to_ranktable(skeleton_table)
```

Above we see, in the first column, that Grotheer came first ("`c1`")
in runs 1-3 and Wengang came first in run 4.  From the first column we
see that in run 1, Grotheer came fist, Tretyakov came second, Wengang
came third, and so on.

```{r,label=processskeleton,cache=TRUE}
skeleton <- ordertable2supp(skeleton_table)
skeleton_maxp <- maxp(skeleton)
skeleton_maxp
pie(skeleton_maxp)
dotchart(skeleton_maxp)
dotchart(log(skeleton_maxp))
```

And we can test the null that the competitors all have the same
Plackett-Luce strengths:

```{r,label=testskeleton,cache=TRUE}
equalp.test(skeleton)
```

We see very strong evidence for different competitor strengths, at
$p=2\times 10^{-33}$.  How strong is the evidence that Grotheer is
better than Jungk?

```{r,label=testgrotheerandjungk,cache=TRUE}
samep.test(skeleton,c("Grotheer","Jungk"))
```

Not very strong.

```{r,label=dgdtert,cache=TRUE}
samep.test(skeleton,c("Grotheer","Jungk"))$p.value
```

```{r,label=dgdtkkkert,cache=TRUE}
p <- rownames(st)
pvals <- seq_along(p)
for(i in seq(from=2,to=length(p))){ pvals[i] <- samep.test(skeleton,c(p[1],p[i]))$p.value }
names(pvals) <- p
pvals
```

(NB the above chunk takes about an hour to run, even with `n=1`).

```{r showpvals}
plot(pvals,log='y',pch=16)
for(i in seq_along(pvals)){text(i,pvals[i],names(pvals)[i],pos=4,col='gray')}
```

### Package dataset {-}

Following lines create `skeleton.rda`, which is not part of the CRAN version of the package.

```{r,label=saveskeletondataset}
save(skeleton_table,skeleton,skeleton_maxp,file="skeleton.rda")
```