### 2023-10-30 inst/rowing.Rmd

---
title: "rowing: analysis of sculls in the 2016 Olympics"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document analyses the `rowing` dataset: sculls in the 2016
olympics.  It is not really standalone and is here as part of a
consistent dataset generation mechanism used in the `hyper2` package.
See the `hyper2` vignette for a more formal discussion.  In Olympic
sculling, up to six individual competitors race a small boat called a
scull over a course of length $2\,\mathrm{km}$; the object is to cross
the finishing line first.  Note that actual timings are irrelevant,
given the model, as the sufficient statistic is the order in which
competitors cross the finishing line.  The 2016 Summer Olympics is a
case in point: the gold and silver medallists finished less than $\sim
5$ milliseconds apart, corresponding to a lead of $\sim
2.5\,\mathrm{cm}$.  The package dataset is file `rowing.txt`, here are
the first three lines:

```{r}
readLines("rowing.txt")[1:3]
```

We can process this file using standard R functions.

```{r}
rowing_table <- sapply(readLines("rowing.txt"),strsplit," ")
names(rowing_table) <- NULL  # names are pointless here
rowing_table[1:3]
```

The first line shows that `fournier` came first, `cabrera` second, and
so on.  We can convert this dataset into a Plackett-Luce likelihood
function as follows:

```{r definerowing, cache=TRUE}
rowing <- Reduce("+",sapply(rowing_table,race))
summary(rowing)
```

Object `rowing` is a loglikelihood function on the strengths of the 32
competitors.  Our first step is to find the evaluate:

```{r maxpcalc,cache=TRUE}
rowing_maxp <- maxp(rowing)
rowing_maxp
```

and we may display it using a dot chart:

```{r showrowingmaxppie, echo=FALSE, fig.cap="dot chart of results in the 2016 Olympic Sculls"}
dotchart(rowing_maxp)
```

```{r showrowingmaxppielog, echo=FALSE, fig.cap="dot chart of results in the 2016 Olympic Sculls (log scale)"}
dotchart(log10(rowing_maxp))
```

```{r conscheck, cache=TRUE,fig.cap="consistency check for rowing function"}
consistency(rowing)
```

Figure \@ref(fig:showrowingmaxppie) shows that many competitors have
very small strength and figure \@ref(fig:showrowingmaxppielog) shows
that many competitors have zero strength to numerical precision.
Looking at the original dataset we can see why: observe that `gambour`
came last in every race except for race number 7 (won by `boudina`).
In race 7, he beat only `yaklovlev`.  But we see that `yaklovlev` came
last in every race except race 20, where he beat only `gambour`.  Thus
the maximum likelihood estimate for `gambour + yaklovlev` would be
zero, because these two competitors always lose to everyone else.  It
is thus reasonable to eliminate `gambour` and `yaklovlev` from the
dataset. We may apply this reasoning recursively and end up with the
following dataset:

```{r}
show("rowing_minimal.txt")
```

(also, we have removed `drysdale`, as he won everything and his ML
strength would be 1).  We may then analyse this:

```{r rowingminimaldef}
rowing_minimal_table <- sapply(readLines("rowing_minimal.txt"),strsplit," ")
names(rowing_minimal_table) <- NULL  
rowing_minimal  <- Reduce("+",sapply(rowing_minimal_table,race))
summary(rowing_minimal)
```

This is a much more manageable dataset, much easier to study and less
cluttered, as competitors with zero estimated strength are not
present.

```{r minimalmaxp,cache=TRUE}
rowing_minimal_maxp <- maxp(rowing_minimal)
rowing_minimal_maxp
```

```{r maxpcalcminimal,fig.cap="Minimal dataset, maximum likelihood estimate of strength"}
dotchart(rowing_minimal_maxp)
dotchart(log10(rowing_minimal_maxp))
```

### Package dataset {-}

Following lines create `rowing.rda`, residing in the `data/` directory
of the package.

```{r,label=saverowingingdataset}
save(rowing,rowing_minimal,rowing_maxp,rowing_minimal_maxp,rowing_table,rowing_minimal_table,file="rowing.rda")
```

## References {-}