### 2023-10-31 inst/jester.Rmd

---
title: "The Jester joke recommender system"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This file creates objects `jester` and `maxjest`, which are datasets
available in the `hyper2` package, and documented at `jester.Rd`.
This file takes quite a long time to run.  Goldberg et al present a
dataset in which respondents rated a number of jokes.  Here, I analyse
a small portion of this dataset using the `hyper2` package.  This
document is intended to illustrate an extremely challenging
application of the `hyper2` package and (without cache) takes a long
time to process.  Goldberg's dataset has 24938 lines, one per
respondent, and 101 columns, one per joke (the first column shows the
number of jokes rated by each respondent); here I use 150 lines and 99
jokes (the 100th joke was not funny).

```{r,label=kickoff}
library("hyper2",quietly=TRUE)
a <- read.csv("jester-data-3.csv",head=FALSE) # File is 150 lines only
a <- as.matrix(a[,-c(1,100)])
a[a==99] <- NA
colnames(a) <- paste("joke",sprintf("%02d",seq_len(ncol(a))),sep="")
a[1,]
```

Row 1 of `a` is displayed (most entries are `NA`, signifying that
respondent 1 did not rank that particular joke).  It shows that the
first respondent rated joke 5 at -1.65, joke 7 at -0.78, and so on.
We can perform some summary statistics of `a`:

```{r,label=plothowmanyjokes}
plot(rowSums(!is.na(a)),xlab="respondent")
```

The above plot shows how many jokes each of the 100 respondents rated.

```{r,label=plothowmanyrespondents}
plot(colSums(!is.na(a)),xlab="joke index")
```

The above shows how many respondents rated each joke.  It would make
sense to remove the jokes that were not rated:

```{r,label=removeunratedjokes}
dim(a)
a <- a[,colSums(!is.na(a))>1]
dim(a)
```

showing that 91 jokes were rated by at least one respondent.  We need
to transform the dataset:

```{r,label=transformthedataset}
f <- function(x){
    x <- x[!is.na(x)]
    x[order(x,decreasing=TRUE)] <- seq_along(x)
    return(x)
}
f(a[1,])
```

Thus we see that this respondent rated `joke08` to be the funniest, having rank 1.

```{r,label=calculate_jester,cache=TRUE}
jester <- hyper2()
system.time(for(i in seq_len(nrow(a))){jester <- jester + rank_likelihood(f(a[i,]))})
```

now

```{r,label=findmaxp,cache=TRUE}
system.time(jester_maxp <- maxp(jester,n=1))
```

and

```{r,showjester_maxp}
jester_maxp
plot(jester_maxp)
```

```{r,label=equalptestdoer,cache=TRUE}
equalp.test(jester,startp=indep(jester_maxp))
```

## Reference

Eigentaste: A Constant Time Collaborative Filtering Algorithm. Ken
Goldberg, Theresa Roeder, Dhruv Gupta, and Chris Perkins. _Information
Retrieval_, 4(2), 133-151. July 2001.

## Create a table like formula 1 results table

```{r makejestertable}
jester_table <- a
for(i in seq_len(nrow(jester_table))){
  x <- jester_table[i,]
  x[!is.na(x)] <- rank(-x[!is.na(x)],ties.method="first")
  jester_table[i,] <- x
}
jester_table <- t(jester_table)
colnames(jester_table) <-paste("resp",seq_len(ncol(jester_table)),sep="_")
jester_table[1:6,1:10]
```

Above we see that respondent 1 ranked joke 8 as the funniest and joke 7 as the 15th funniest.
Comparing with `formula1_2022.txt`, for example, we see that respondents correspond to venues and jokes correspond to drivers.  We have to be a little bit careful because `NA` means "not rated", not "did not finish" as in the Formula 1 datasets.

```{r meanjokes}
plot(rowMeans(jester_table,na.rm=TRUE))
abline(v=17)
```

### Package dataset {-}

Following lines create `jester.rda`, residing in the `data/` directory of the package.

```{r,label=savejesterdataset}
save(jester,jester_table,jester_maxp,file="jester.rda")
```

 ### References