### 2023-10-31 inst/distance_matrix_idea.Rmd

---
title: "icons: some experimental results"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
library("ape")
library("phangorn")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This file presents some _speculative_ experimental work which is under
development.  Do not confuse this file with `icons.Rmd`, which creates
the `icons` objects in the package.

```{r,label=showicons}
icons
icons_maxp <- sort(icons_maxp,decreasing=TRUE)  # sort in decreasing order
icons_maxp
```

Recall `samep.test()`:

```{r,showsameptest,cache=TRUE}
samep.test(icons,c("NB","THC"))
```

I will calculate all pairwise distances between the icons, as measured
by the support difference [penultimate line of the output above, named
`statistic` in the object for consistency with other tests in R]

```{r,label=calculateM,cache=TRUE}
M <- matrix(0,6,6)
p <- names(icons_maxp)
rownames(M) <- p
colnames(M) <- p
for(i in seq_len(6)){
  for(j in seq_len(6)){
    if(i != j){M[i,j] <- samep.test(icons,p[c(i,j)])$statistic}
  }
}
```

```{r roundm}
M
(M <- round(M*100))
```

Observations: the triangle inequality is broken, consider `NB`-`L`
(81), `L`-`PB` (38) and `PB`-`NB` (8)

```{r plotphangorn}
plot(phangorn::upgma(as.dist(M)))
```

### References {-}