### 2023-10-31 inst/global_liveability_ranking.Rmd

---
title: "Global liveability ranking"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a;
takes about an hour to process without cache.

_This analysis is essentially without value, see the very end for why_

[The global liveability
ranking](https://en.wikipedia.org/wiki/Global_Liveability_Ranking)
shows the nicest places to live.

```{r showglrtable}
(glr <- readLines("global_liveability_ranking.txt"))
```

First we will turn it into a `hyper2` object:

```{r readglrtable}
(o <- glr  %>% strsplit(" ") %>% lapply(function(l){l[-1]}))
```

Will remove Helsinki, Brisbane, and Hamburg on account of their having
very low rank scores (if we leave them in, the Hessian becomes
non-positive-definite at the evaluate).

```{r,removeHelsinki_and_Hamburg}
o %<>%  lapply(function(x){x[!(x %in% c("Hamburg","Helsinki","Brisbane"))]})
o
```

```{r,turnintohyper2,cache=TRUE}
H <- hyper2()
for(i in seq_along(o)){
  H <- H  + race(o[[i]])
}
```

And do some tests:

```{r testgrl, cache=TRUE}
(mH <- maxp(H))
pie(mH)
equalp.test(H)
```

```{r,calculatemoreinfo,cache=TRUE}
jjm <- maxp(H,n=1,give=TRUE,hessian=TRUE)
```

```{r}
jjm$hessian
diag(jjm$hessian)
eigen(jjm$hessian)$values
```

So the Hessian is positive-definite [all the eigenvalues are
positive].  Also, the evaluate is robust, as indicated by
`consistency()`:

```{r,doconsis,cache=TRUE}
consistency(H)
```

Now we tether some of the cities together:

```{r, tether}
elastic <- list(
Austria     = c("Vienna"),
Denmark     = c("Copenhagen"),
Switzerland = c("Zurich","Geneva"),
Canada      = c("Calgary","Vancouver","Toronto"),
Germany     = c("Frankfurt","Hamburg"),
Netherlands = c("Amsterdam"),
Japan       = c("Osaka","Tokyo"),
Australia   = c("Melbourne","Adelaide","Perth","Brisbane","Sydney"),
NZ          = c("Auckland","Wellington"),
Finland     = c("Helsinki")
)
```

Now create a `hyper3` object:

```{r,createhyper3func}
ML <- function(h){
   H3 <- hyper3()
   for(r in o){ H3 <- H3 + cheering3(r,e=list2nv(elastic),help=rep(h,10))}
   return(H3)
}
f <- function(h){maxp(ML(h),give=TRUE,n=1)$likes}
```

And plot a profile likelihood

```{r,plliveable,cache=TRUE}
l <- seq(from=1,to=5,len=10)
L <- sapply(l,f)
```

and plot it:

```{r}
plot(log(l),L-max(L),type='b')
```

# A coarser division

We will try dividing the countries into continents:

```{r,label=elasticcontinent}
elastic_continent <- list(
Europe      = c("Vienna","Copenhagen","Zurich","Geneva","Frankfurt","Hamburg","Amsterdam","Helsinki"),
America     = c("Calgary","Vancouver","Toronto"),
Asia        = c("Osaka","Tokyo"),
Australasia = c("Melbourne","Adelaide","Perth","Brisbane","Sydney","Auckland","Wellington")
)

ec <- NULL
for(i in seq_along(elastic_continent)){
  jj <- rep(i,length(elastic_continent[[i]]))
  names(jj) <- elastic_continent[[i]]
  ec <- c(ec,jj)
}
ec 
MLc <- function(h){
   H3 <- hyper3()
   for(r in o){ H3 <- H3 + cheering3(r,e=ec,help=rep(h,4))}
   return(H3)
}
fc <- function(h){maxp(ML(h),give=TRUE,n=1)$likes}
```

```{r,evaluatefc,cache=TRUE}
l <- seq(from=0.8,to=3,len=10)
L <- sapply(l,fc)
```

```{r,plotfc}
plot(l,L-max(L),type="b")
abline(v=1)
```

## Country-level analysis

We need to replace each city with its country:

```{r,label=replacecity}
elastic
(jj <- list2nv(elastic,FALSE))
o_countries <- lapply(o, function(x){jj[names(jj) %in% x]})
(o_countries <- lapply(o_countries,as.vector))
```

We see that NZ and Australia are dominated so we should remove them:

```{r}
(o_countries <- lapply(o_countries,function(x){x[!(x %in% c("NZ","Australia","Japan"))]}))
```

But hang on, if we continue to remove dominated countries we
eventually remove every country that is not Austria.  So that was a big
fat waste of time.

## References {-]