### 2024-02-23 inst/very_simplified_likelihood.Rmd
include Stirling

---
title: "Very simplified likelihood"
author: "Robin Hankin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("pracma")
library("magrittr")
```

```{r hexsticker, out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

Consider a race between a competitor of Bradley-Terry strength $a$ and
$n$ competitors each of strength $b$; we require $a+b=1$.  NB we have
$n+1$ competitors in total, 1 of strength $a$ and $n$ of strength $b$.
An observation is indexed by $r$, the number of $b$ clones finishing
ahead of $a$, so $0\leqslant r\leqslant n$.  The initial field
strength is $a+nb=1+b(n-1)$.

Suppose our bro comes first (zero $b$'s ahead of him):

\[
\mathcal{L}(a)=\frac{a}{a+nb}=\frac{1-b}{1+(n-1)b}
\]

Comes second, one $b$ ahead of him:

\[
\mathcal{L}(a)=\frac{b}{a+nb}\cdot\frac{a}{a+(n-1)b}
\]

Third, two ahead of him:

\[
\mathcal{L}(a)=
\frac{b}{a+ n   b}\cdot
\frac{b}{a+(n-1)b}\cdot
\frac{a}{a+(n-2)b}
\]

Fourth, three ahead:

\[
\mathcal{L}(a)=
\frac{b}{a+ n   b}\cdot
\frac{b}{a+(n-1)b}\cdot
\frac{b}{a+(n-2)b}\cdot
\frac{a}{a+(n-3)b}
\]

Placing after $r<n$ of the $n$ clones of strength $b$ finish:

\[
\mathcal{L}(a)=
\frac{b^ra}{
(a+nb)(a+(n-1)b)(a+(n-2)b)\cdots(a+(n-r)b)}
=
\frac{B-1}{(B+n-1)(B+n-2)\cdots(B+n-r-1)}
\]

where $B=1/b$ [also using the fact that $a+b=1$].

This is

\[
\mathcal{L}_{r,n}(a)=\frac{(B-1)(B+n-r-2)!}{(B+n-1)!},\qquad B=1/(1-a)
\]

Then, using Stirling [viz $n!\sim n^ne^{-n}\sqrt{2\pi n}$] we get

\[
\log(B-1) + ((B+n-r-2)\log(B+n-r-2)-(B+n-r-2)+\frac{\log(B+n-r-2)}{2})
          - ((B+n  -1)\log(B+n  -1)-(B+n  -1)+\frac{\log(B+n  -1)}{2})
\]

If you give Sage this:

```
var('B n r')
pretty_print_default(True)
X=log(B-1) + (
+ ((B+n-r-2)*log(B+n-r-2)-(B+n-r-2)+log(B+n-r-2)/2)    
- ((B+n  -1)*log(B+n  -1)-(B+n  -1)+log(B+n  -1)/2)
)
X.taylor(n,Infinity,2)
```

It returns this:

\[
-(r+1)\log n
-\frac{(2B-3)r-r^2+2B-2}{2n}
-\frac{3(2B-3)r^2-6B^2-6(B^2-3B+2)r+12B-5}{12n^2}
+\log(B-1)+\mathcal{O}(n^{-3})
\]

After simplifying and taking an asymptotic expansion we find

\[
\mathcal{S}=\log\mathcal{L}_{r,n}(a)=\log(B-1)-B\frac{r+1}{n} + \frac{B^2-B(5+r^2)/2}{n^2}
+ K + \mathcal{O}\left(n^{-3}\right)
\]

(where $K$ is an arbitrary constant).  If we then neglect terms of
order $n^{-2}$ and above, viz $\log(B-1)-B\frac{r+1}{n}$, observing
that $\partial^2\mathcal{S}/\partial B^2=-(B-1)^{-2}<0$, the evaluate
would be $\hat{a}=\frac{n}{n+r+1}$ [although this seems to be a poor
approximation to the exact evaluate]

## Placing last

Placing last is finishing  after $n=r$ clones finish:

\[
n=1\longrightarrow\mathcal{L}=b\]

\[
n=2\longrightarrow\mathcal{L}=\frac{b}{a+2b}\cdot\frac{b}{a+b}=\frac{b^2}{1+b}
=\frac{1}{B(B+1)}
\]

\[
n=3\longrightarrow\mathcal{L}=\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}=\frac{b^3}{(1+2b)(1+b)}=\frac{1}{B(B+1)(B+2)}
\]

\[
n=4\longrightarrow\mathcal{L}=\frac{b}{a+4b}\cdot\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}=
\frac{b^4}{(1+3b)(1+2b)(1+b)}=
\frac{1}{B(B+1)(B+2)(B+3)}
\]

For $n=4$ we have equivalently

\[
\frac{1}{(B+3)(B+2)(B+1)B}=\frac{(B-1)!}{(B+3)!}
\]

In general we would have

\[
\frac{1}{(B+n-1)(B+n-2)\cdots(B+1)B}=
\frac{(B-1)!}{(B+n-1)!}
\]

(denominator has $n+1$ terms).  In R:

```{r,label=definefsingle}
f_single <- function(a,r,n,log=FALSE){  # not vectorised
	B <- 1/(1-a)
	if(log){
		out <- log(B-1) - lgamma(B+n) + lgamma(B+n-r-1)
	} else {  # not-log 
		out <- (B-1)/prod(B+(n-r-1):(n-1))
	}	
   return(out)
}

f_vec_a <- function(a,r,n, log=FALSE){  # vectorised in 'a' but not in 'r'
	sapply(a,function(a){f_single(a,r=r,n=n,log=log)})
}

f_vec_arn <- function(a,r,n,log=FALSE){
	## vectorized in 'a'; treats 'r' and 'n' as
	## vectors of independent observations

	M <- cbind(r,n)

	if(log){
		out <- 0
		for(i in seq_len(nrow(M))){out <- out + f_vec_a(a,r=M[i,1],n=M[i,2],log=TRUE)}
	} else {
		out <- 1
		for(i in seq_len(nrow(M))){out <- out * f_vec_a(a,r=M[i,1],n=M[i,2],log=FALSE)}
	}
	return(out)
}

fapprox <- function(a,r,n,log=FALSE){
	if(log){
		return(log(B-1)-B*(r+1)/n)
	} else {
		return(B-1 - (r+1)*exp(B)/n)
	}
}
```

```{r label=defdiff}
fdash <- function(a,r,n,log=TRUE){
	B <- 1/(1-a)
	out <- (1/(B-1) - psigamma(B+n) + psigamma(B+n-r-1))*B^2
	if(!log){out <- out * f_vec_arn(a,r,n,log=FALSE)}
	return(out)
}
```

Now verify `fdash()`:

```{r label=verifydiff}
a <- 0.34
d <- 1e-3
r <- 4
n <- 10
c(
  numerical  = (f_single(a+d/2,r,n,log=TRUE)-f_single(a-d/2,r,n,log=TRUE))/d,
  analytical = fdash(a,r,n)
)

c(
  numerical  = (f_single(a+d/2,r,n,log=FALSE)-f_single(a-d/2,r,n,log=FALSE))/d,
  analytical = fdash(a,r,n,log=FALSE)
)
```

Now use the derivative to find the maximum likelihood point

```{r,label=defmle}
MLE <- function(r,n,give=FALSE){
	d <- 1e-6
	out <- optimize(f_vec_arn,c(d,1-d),r=r,n=n,log=TRUE,maximum=TRUE)
	if(!give){out <- out$maximum}
	return(out)
}

MLEa <- function(r,n,give=FALSE){
	out <- uniroot(f=function(a){fdash(a,r,n)},interval=c(0.01,0.99))
	if(!give){out <- out$root}
	return(out)
}

rbind(
c(MLE(1,9),MLEa(1,9)),
c(MLE(3,9),MLEa(3,9)),
c(MLE(3,7),MLEa(3,7))
)
```

```{r usestirling}
rbind(
c(MLE(10 , 100),100 / 111),
c(MLE(100,1000),1000/1101),
c(MLE(500,5000),5000/5501)
)

rbind(
c(MLEa(10 , 100),100 / 111),
c(MLEa(100,1000),1000/1101),
c(MLEa(500,5000),5000/5501)
)
```

Now verification

```{r,label=verifytwo}
a <- 0.23423434
b <- 1-a
B <- 1/b
n <- 9
r <- 3
c(
way1 = b^3*a/((a+9*b) * (a+8*b) * (a+7*b) * (a+6*b)),
way2 = (B-1)/prod(B+(5:8)),
way3 = f_vec_arn(a,r,n,log=FALSE)
)
```

```{r,label=makejj}
out <- list()
for(n in 1:10){
  out[[n]]  <- rep(NA,n+1)
  for(r in 0:n){
    if(r==0){
      jj <- 1
    } else if(r==n){n
      jj <- 0
    } else {
      jj <- MLE(r,n)
    }  
    out[[n]][r+1] <- jj
  }
} 
out
```

```{r,label=plotjj}
plotterp <- function(...){
plot(NA,xlim=c(1,10),ylim=c(0,1),type="n",xlab="n",ylab=expression(hat(a)))
for(n in 1:10){
  for(r in 0:n){   points(n,out[[n]][r+1],pch=16) }
  for(r in 0:n){   text(n+0.2,out[[n]][r+1],r,cex=0.5,col='gray') }
}
}
plotterp()
pdf(file="dotprobs.pdf")
plotterp()
dev.off()
```

## Note on the differences between probability and likelihood

The likelihood calculations above can be confusing becuase likelihood
is not the same as probability.

Consider, for example, $n=2$ so we have 3 competitors.

probability of focal competitor coming first [i.e. $r=0$]:

\[\frac{a}{a+2b}=\frac{1-b}{1+b}\]

Now what is the probability of coming second, that is, $r=1$?

My first thought was:

\[
\mbox{prob of coming second}   =
\frac{b}{a+2b}\cdot\frac{a}{a+b}=\frac{b(1-b)}{1+b}\qquad\mbox{WRONG}
\]

but this is incorrect: there are two clones of (strength) $b$, and the
likelihood arguments above do not distinguish between their finishing
order.  We can name the competitors $a$, $b_1$ and $b_2$ and specify
that $b_1$ and $b_2$ both have strength $b$.  Then "coming second"
means that the finishing order was $b_1\succ a\succ b_2$ or $b_2\succ
a\succ b_1$.  Noting that these events are disjoint, the probability
would be

\[\mbox{prob of coming second}   =
\operatorname{P}(b_1\succ a\succ b_2)      +
\operatorname{P}(b_2\succ a\succ b_1)      =
\frac{b}{a+2b}\cdot\frac{a}{a+b} +
\frac{b}{a+2b}\cdot\frac{a}{a+b} = \frac{2(1-b)}{1+b}
\]

Similarly for the focal competitor coming last we need to sum the
probabilities of finishing orders $b_1\succ b_2\succ a$ and $b_2\succ
b_1\succ a$:

\[\mbox{prob of coming third}=
\operatorname{P}(b_1\succ b_2\succ a) +
\operatorname{P}(b_2\succ b_1\succ a) =
\frac{b}{a+2b}\cdot\frac{b}{a+b}+
\frac{b}{a+2b}\cdot\frac{b}{a+b}=
\frac{2b^2}{1+b}.
\]

Above we see a factor of two difference between the probabilities and
the likelihoods presented previously.  Note that there is no such
factor for the probability of the focal competitor finishing first:

\[
\mbox{prob of coming first}      =
\operatorname{P}(a\succ b_1\succ b_2)      +
\operatorname{P}(a\succ b_2\succ b_1)      =
\frac{b}{a+2b}\cdot\frac{b}{b+b} +
\frac{b}{a+2b}\cdot\frac{b}{b+b} = \frac{1-b}{1+b}
\]

Now we see that the probabilities sum to

\[
\frac{1-b}{1+b} + \frac{2b^2}{1+b} + \frac{2(1-b)}{1+b}=1
\]

Now try $r=3$ [i.e. 4 competitors].  Taking things one step at a time:

\[
\mbox{prob of coming first} =
3!\cdot\frac{a}{a+3b}\cdot\frac{b}{3b}\cdot\frac{b}{2b}\cdot\frac{b}{b}=
\frac{1-b}{1+2b}
\]

\[
\mbox{prob of coming second} =
3!\cdot\frac{b}{a+3b}\cdot\frac{a}{a+2b}\cdot\frac{b}{2b}\cdot\frac{b}{b}=
3\cdot\frac{b(1-b)}{(1+2b)(1+b)}
\]

\[
\mbox{prob of coming third} =
3!\cdot\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{a}{a+b}\cdot\frac{b}{b}
=
6\cdot\frac{b^2(1-b)}{(1+2b)(1+b)}
\]
\[
\mbox{prob of coming fourth (last)} =
3!\cdot\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}\cdot\frac{a}{a}
=
6\cdot\frac{b^3}{(1+2b)(1+b)}
\]

The sum would be $[(1+b)(1-b) + 3b(1-b) + 6b^2(1-b) +
6b^3]/[(1+b)(1+2b)]=1$.

Now consider the general case.  Again, $n$ clones of strength $b$ and
one of strength $a=1-b$, $n+1$ competitors.

\[
\mbox{prob of coming first} =
n!\cdot\frac{b}{a+nb}\cdot\frac{b}{nb}\cdot\frac{b}{(n-1)b}\cdots\frac{b}{b}
=
\frac{b}{1+(n-1)b}
\]

\[
\mbox{prob of coming second} =
n!\cdot\frac{b}{a+nb}\cdot\frac{a}{a+(n-1)b}\cdot\frac{b}{(n-1)b}\cdot\frac{b}{(n-2)b}\cdots\frac{b}{2b}\cdot\frac{b}{b}
=n\cdot X
\]

## Visuals

Now some likelihood visuals.  Suppose $r=4,n=8$:

```{r,label=likevis,cache=F}
plotter <- function(...){
a <- seq(from=0,to=1,by=0.01)
n <- 7
jj <- f_vec_arn(a,3,n,log=FALSE)
plot(a,jj/max(jj,na.rm=TRUE),type='n')
grid()
for(r in seq(from=0,to=n)){
  y <- f_vec_a(a,r,n,log=FALSE)
  if(r==0){y[length(y)] <- 1}  
  y <- y/max(y,na.rm=TRUE)
  if(r>0){y[length(y)] <- 0}
  points(a,y,type='l',lwd=4,col=rainbow(n+1)[r+1])
  segments(x0=r/(n),y0=0.9,y1=1)
  text(x=r/(n),y=0.8,r)
}
}
plotter()
pdf(file="ninelikes.pdf")
plotter()
dev.off()
```

```{r usehyper3,cache=TRUE}
n <- 7
a <- seq(from=0,to=1,by=0.01)
fhyper3 <- function(r,n){
  out <- hyper3()
  out['a'] <- 1
  out['b'] <- r
  for(i in (n-r):n){
    out[c(a=1,b=i)] %<>% dec
  }
  return(out)
}

plot(NA,xlim=c(0,1),ylim=c(0,1),type='n')
M <- cbind(a=a,b=1-a)
for(r in 0:7){
  y <- loglik(M,fhyper3(r,n),log=FALSE)
  y <- y/max(y, na.rm=TRUE)
points(M[,1],y,type="l",lwd=8)
}
```

# Formula 1.

```{r maxvscheco}
a <- read.table("formula1_2023.txt",header=TRUE)
a <- a[,seq_len(ncol(a)-1)]
d_ocon  <- as.numeric(as.matrix(a)["Ocon",])
d_gasly <- as.numeric(as.matrix(a)["Gasly",])
d_ocon [is.na(d_ocon )] <- 22
d_gasly[is.na(d_gasly)] <- 22
d_ocon
d_gasly
MLE(d_ocon ,22)
MLE(d_gasly,22)
```

two sample test:

```{r 2samptest}
both <- c(d_ocon,d_gasly)
bothmax <- MLE(both,22)

d_ocon_max  <- MLE(d_ocon ,22)
d_gasly_max <- MLE(d_gasly,22)
f_vec_arn(bothmax,d_ocon ,22,log=TRUE)
f_vec_arn(bothmax,d_gasly,22,log=TRUE)
f_vec_arn(d_ocon_max ,d_ocon ,22,log=TRUE)
f_vec_arn(d_gasly_max,d_gasly,22,log=TRUE)

Lambda <- (
   (f_vec_arn(d_gasly_max  ,d_ocon,22,log=TRUE)+f_vec_arn(d_gasly_max  ,d_gasly,22,log=TRUE)) -
   (f_vec_arn(bothmax      ,d_ocon,22,log=TRUE)+f_vec_arn(bothmax      ,d_gasly,22,log=TRUE))
)
Lambda
pchisq(-2*Lambda,df=1,lower.tail=FALSE)
```

## Mathematical Olympiad

```{r readoly}
sametest <- function(d1,d2,n){
  both <- c(d1,d2)
  bothmax <- MLE(both,n)

  d1max <- MLE(d1,n)
  d2max <- MLE(d2,n)
  Lambda <- (
  (f_vec_arn(d2max  ,d1,n,log=TRUE)+f_vec_arn(d2max  ,d2,n,log=TRUE)) -
  (f_vec_arn(bothmax,d1,n,log=TRUE)+f_vec_arn(bothmax,d2,n,log=TRUE))
)
  pchisq(-2*Lambda,df=1,lower.tail=FALSE)
}

a <- as.matrix(read.table("olympiad.txt"))
nrow(a)
d1 <- a["AUS",]
d2 <- a["NZL",]
d1
d2
sametest(d1,d2,100)
a["USA",]
a["CHN",]
sametest(a["USA",],a["CHN",],100)
sametest(a["USA",],a["CHN",],5)
```

## Parkrun

```{r readresults}
results <- read.table("parkrun_results.txt",header=TRUE)
results
```

Above we see three parkrun results.  In parkrun number 148, I placed
229 out of a field of 318.

```{r definelikelihood}

L <- function(a,results){
   out <- a*0
   for(i in seq_len(nrow(results))){
   jj <- f_vec_a(a,r=results$place[i]-1,results$runners[i]-1,log=TRUE)
   out <- out + jj
   }
   out[is.infinite(out)] <- NA
   return(out)
}
```

```{r calculatelikes,cache=F}
a <- seq(from=0.25,to=0.65,by=0.01)
La <- f_vec_arn(a,r=results$place-1,results$runners-1,log=TRUE)
La <- La - max(La,na.rm=TRUE)
```

```{r plotlike}
plot(a,La,ylab='support')
abline(h=c(0,-2))
```

```{r optimsinglelike}
optimize(function(a){L(a,results)},c(0.2,0.4),maximum=TRUE)
```

```{r calculatelikes1,cache=F}
a <- seq(from=0.005,to=0.75,by=0.01)
La <- f_vec_arn(a,r=results$place[1]-1,results$runners[1]-1,log=TRUE)
La <- La - max(La,na.rm=TRUE)
plot(a,La,type='b',ylab='fishy')
abline(h=c(0,-2))
```

```{r splitresults}
supp2 <- function(vec,place,runners){
  a_stirling <- vec[1]
  a_auckland <- vec[2]
  out <- 0
  for(i in 1:3){
    out <- out + f_vec_a(a_stirling,r=place[i]-1,runners[i]-1,log=TRUE)
  }
  for(i in 4:10){
    out <- out + f_vec_a(a_auckland,r=place[i]-1,runners[i]-1,log=TRUE)
  }
  return(out)
}
```

```{r,label=calcbothlikes,cache=TRUE}
x <- seq(from=0.1,to=0.8,by=0.01)
jj <- as.matrix(expand.grid(x,x))
L <- apply(jj,1,supp2,results$place,results$runners)
L <- L - max(L,na.rm=TRUE)
L <- matrix(L,length(x),length(x))
```

```{r,label=plotbothlikes}
par(pty='s')
contour(x,x,L,asp=1,xlim=range(x),ylim=range(x),levels=-(0:5))
abline(0,1)
```

```{r, label=onedimopt}
f <- function(a){supp2(c(a,a),results$place,results$runners)}
a <- seq(from=0.2,to=0.6,by=0.01)
alike <- sapply(a,f)
plot(a,alike-max(alike,na.rm=TRUE),type='b',ylab='blong')
abline(h=c(0,-2))
optimize(f,c(0.1,0.6),maximum=TRUE)
```

```{r, label=twodimopt}
f2 <- function(v){supp2(v,results$place,results$runners)}
optim(par=c(0.6,0.3),f2,control=list(fnscale=-1))
```

```{r,label=showchisquareasymptotic}
7380.314-7381.496
pchisq(2*(7381.496-7380.314),df=1,lower.tail=FALSE)
```

## Education

The focal competitor

```{r focaleducation}
rank <- c( 9, 7, 2, 3,2 ,8 ,5 ,4 ,9 )
class_size <- c(12, 17,23, 9,13,14,13,12,15)
course <- c("rings and modules","group theory","calculus","linear algebra",
"differential equations","topology","special relativity","fluid mechanics","Lie algebra")
category=c("pure","pure","applied","applied","applied","pure","applied","applied","pure")
data.frame(course,category,rank,class_size)
```

```{r seppurefromapplied}
a <- seq(from=0.2,by=0.01,to=0.8)
wp <- category=='pure'
wa <- category=='applied'
```

Pure first

```{r dopurefurst}
Lpure <- f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)
Lpure <- Lpure - max(Lpure,na.rm=TRUE)
plot(a,Lpure,ylab='pure strength')
optimize(function(a){f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
optimize(function(a){f_vec_arn(a,rank[wa]-1,class_size[wa]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
```

```{r definesupp2ed}
supp2_ed <- function(vec,place1,place2,runners1,runners2){
  out <- 0
  M1 <- cbind(place1,runners1)
  M2 <- cbind(place2,runners2)
  for(i in seq_len(nrow(M1))){
    out <- out + f_vec_arn(vec[1],r=M1[i,1]-1,n=M1[i,2]-1,log=TRUE)
  }
  for(i in seq_len(nrow(M2))){
    out <- out + f_vec_arn(vec[2],r=M2[i,1]-1,n=M2[i,2]-1,log=TRUE)
  }
  return(out)
}
```

```{r,label=fishbish,cache=TRUE}
x <- seq(from=0.1,to=0.95,by=0.01)
jj <- as.matrix(expand.grid(x,x))
L <- apply(jj,1,supp2_ed,rank[wp]-1,rank[wa]-1,class_size[wp]-1,class_size[wa]-1)
L <- L - max(L,na.rm=TRUE)
L <- matrix(L,length(x),length(x))
```

```{r,label=contourstrength}
par(pty='s')
contour(x,x,L,asp=1,xlim=range(x),ylim=range(x),levels=-(0:5),
xlab='pure strength',ylab='applied strength')
abline(0,1)
```

```{r,label=defsupp2}
jj1 <- 
optim(
   c(.4,.8),
   function(vec){
      supp2_ed(vec,rank[wp]-1,rank[wa]-1,class_size[wp]-1,class_size[wa]-1)},
   control=list(fnscale = -1)
)
jj1
```

```{r,label=findmaxlike}
jj2 <-
optimize(function(v){supp2_ed(c(v,v),rank[wp]-1,rank[wa]-1,class_size[wp]-1,class_size[wa]-1)},
c(0.1,0.9),maximum=TRUE)
jj2
```

```{r,label=shochisw}
pchisq(2*(jj1$value-jj2$objective),lower.tail=FALSE,df=1)
```