### 2024-03-08 inst/vsl_masterchef.Rmd
work on very simplified likelihood

---
title: "Very simplified likelihood: masterchef"
author: "Robin Hankin"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("pracma")
library("magrittr")
```

```{r hexsticker, out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

This follows on from `vsl_teams.Rmd` and `vsl_partial_ranks.Rmd`.  But
the application of those ideas to _MasterChef_ is flawed: as the
competition progresses, Brent and Laura face increasingly skilled
competitors, for the weakest ones have been eliminated.  And it is not
clear to what extent that observation interferes with the assumtion of
constant non-focal competitor strength.

```{r,label=define_f_range_setup,include=FALSE}
f_single_old <- function(a,r,n){  # not vectorised
	B <- 1/(1-a)
    (B-1)/prod(B+(n-r-1):(n-1)) * exp( lfactorial(n)-lfactorial(n-r))
}

f_single <- function(a,r,n){
	B <- 1/(1-a)
    exp(log(B-1)-lgamma(B+n)+lgamma(B+n-r-1)+lfactorial(n)-lfactorial(n-r))
}
f_vec_a <- function(a,r,n){  # vectorised in 'a' but not in 'r'
	sapply(a,function(a){f_single(a,r=r,n=n)})
}

```

```{r label=show_f_range}
L_range <- function(a,possible,n){ # r1=4 -> came fifth [four b clones ahead]  
    out <- a*0
    for(i in (possible)){
        out <- out + f_vec_a(a,i-1,n-1) }
    return(out)
}
```

Two teams of size $m$:

```{r}
`L_win` <- function(a,m){
    b <- 1-a
    ((m-2)*b+1)/((2*m-2)*b+1)}

`L_lose` <- function(a,m){1-L_win(a,m)}
```

# Laura in  Australian MasterChef series 6

* week 1
  +  top 3 out of 25 but not the winner: (2:3)/24

* week 2
  +  not top 3, not bottom 3: (4:20)/23
  +  team lose (11 v 11)
  +  not bottom 6: (1:5)/11

* week 3
  +  top 3 out of 21: (1:3)/21
  +  team win (10 v 10)

* week 4
  +  not top 3, not bottom 3:  (4:16)/19
  +  team win (9 v 9)

* week 5
  +  not top 3, not bottom 4:  (4:13)/17
  +  4 teams of 4; Laura came in second team.

* week 6
  + bottom 3, not eliminated: (13:14)/15
  + two teams; Laura came in top team

* week 7
  + top3: (1:3) /14 
  + 2 teams of 6, losing team

* week 8
  + bottom 2, not eliminated, (13/15)
  + 2 teams of 5, losing team

* week 9
  + not bottom 5 out of 11 (1:6)/11
  + two teams of 5 (again); losing team
  + of the 5 in the losing team, came 4th

* week 10
  + top 3 out of 9
  + 4 teams of 2, Laura came in top team

* week 11
  + top 3 out of 7

* Finals
  + bottom 3 out of 6     (4:6)/6
  + not eliminated out of 3
  + top 3 out of 5   (1:3)/5
  + top out of 4  (1)/4
  + top 2 out of 3 (1:2)/3
  + not the winner

# Brent in  Australian MasterChef series 6

* Week 1
  + Bottom 9 out of 24 (16:24)/24
  + elimination out of 9, not in bottom 3 (1:6)/9

* Week 2
  + out of 23, not in top 3, not in bottom 3 (4:20)/23
  + team win; two teams of 11
  + out of 11, in the bottom 6 (6:11)/11
  + out of 6, not in bottom 3 (1:3)/6

* week 3
  + out of 21, not bottom 3, not top 3 (4:18)/21
  + bottom 3 out of 18 (16:18)/18
  + two teams of 10, losing team
  + out of 10, bottom 6 (5:10)/10
  + out of 6, in top 3 (1:3)/6

* week 4
  + out of 19, in top 3, (1:3)/19
  + out of 18, in bottom 3 (16:18)/18
  + out of 3, did not get eliminated (1:2)/3

* Week 5
  + out of 17, top 3 (1:3)/17
  + four teams of 4, in losing team
  + out of 4, did not get eliminaed (1:3)/4

* week 6
  + out of 15, not in bottom 3, not in top 3 (4:12)/15
  + two teams of 7, in losing team
  + out of 7, not in bottom 3 (1:4)/7

* week 7
  + out of 13, not in top 3 not in bottom 3 (4:10)/13
  + two teams of 6, losing team
  + out of 6, bottom 2 (5:6)/6
  + out of 2, not eliminated (1)/2

* week 8 
  + (2:4)/11
  + 2 teams of 5, winning team

* week 9
  + bottom 5, (7:11)/11
  + out of 5, not bottom 3 (1:2)/5
  + 2 teams of 5, losing team
  + out of 5, not bottom 2 (1:3)/5

* week 10
  + out of 9, not in top 3 or bottom 4 (4:5)/9
  + 4 teams of 2, in third or fourth team (?)
  + out of 4, not eliminted (1:3)/4

* week 11
  + out of 7, bottom 4 (4:7)/7
  + out of 4, not bottom 2 (1:2)/4

* Finals
  + out of 6, top 3 (1:3)/6
  + out of 5, bottom 2 (4:5)/5
  + out of 2, not eliminated (1)/2
  + out of 4, (2:3)/4
  + out of 3, not eliminated (1:2)/3
  + out of 3, not eliminatged (1:2)/3
  + overall winner (1)/3

```{r label=defLauraLike}
L_laura <- function(a){
    b <- 1-a
    out <- a*0 + 1
    out <- out * L_range(a,2:3,24)    # week 1
    out <- out * L_range(a,4:20,24)   # week 2
    out <- out * L_lose(a,11)         # week 2
    out <- out * L_range(a,1:5,11)    # week 2
    out <- out * L_range(a,1:3,21)    # week 3
    out <- out * L_win(a,10)          # week 3
    out <- out * L_range(a,4:16,19)   # week 4
    out <- out * L_win(a,9)           # week 4
    out <- out * 4*b*(1+2*b)/
                 (1+14*b)/(1+10*b)    # week 5
    out <- out * L_range(a,14:15,16)  # week 6
    out <- out * L_win(a,7)           # week 6
    out <- out * L_range(a,1:3,13)    # week 7
    out <- out * L_lose(a,6)          # week 7
    out <- out * L_range(a, 10, 11)   # week 8
    out <- out * L_lose(a,5)          # week 8
    out <- out * L_range(a,1:6,11)    # week 9
    out <- out * L_lose(a,5)          # week 9
    out <- out * L_range(a,4,5)       # week 9
    out <- out * L_range(a,1:3,9)     # week 10
    out <- out * 1/(7-6*a)            # week 10
    out <- out * L_range(a,1:3,7)     #  week 11
    out <- out * L_range(a,4:6,6)     # finals
    out <- out * L_range(a,1:2,3)     # finals
    out <- out * L_range(a,1:3,5)     # finals
    out <- out * L_range(a,1,4)       # finals
    out <- out * L_range(a,1:2,3)     # finals
    out <- out * L_range(a,2,2)       # finals
    return(out)
}
```
```{r label=optimizeLauras,cache=TRUE}
evaluate <- optimize(L_laura,c(0.1,0.9),maximum=TRUE)$max
small <- 1e-4
second <- (2*log(L_laura(evaluate))-log(L_laura(evaluate-small))-log(L_laura(evaluate+small)))/small^2
evaluate
second
f <- function(x){x*(1-x)}
(2*log(f(evaluate))-log(f(evaluate-small))-log(f(evaluate+small)))/small^2
```

```{r label=plotlaurassupport}
a <- seq(from=0,to=1,by=0.01)
L <- L_laura(a)
L <- L/max(L,na.rm=TRUE)
plot(a,L)
abline(v=evaluate)
plot(a,log(L),ylim=c(-5,0))
abline(h= -2)
abline(v=evaluate)
```