### 2024-03-08 inst/vsl_partial_ranks.Rmd

 ---
title: "Very simplified likelihood: partial ranks"
author: "Robin Hankin"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("pracma")
library("magrittr")
library("knitr")
```

```{r hexsticker, out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

This follows on from `very_simplified_likelihood.Rmd`.
Consider a race between a competitor of Bradley-Terry strength $a$ and
$n$ competitors each of strength $b$; we require $a+b=1$.  NB we have
$n+1$ competitors in total, 1 of strength $a$ and $n$ of strength $b$.
An observation is indexed by $r$, the number of $b$ clones finishing
ahead of $a$, so $0\leqslant r\leqslant n$.  The initial field
strength is $a+nb=1+b(n-1)$.

```{r,label=definefsingle}
f_single <- function(a,r,n){
	B <- 1/(1-a)
    exp(log(B-1)-lgamma(B+n)+lgamma(B+n-r-1)+lfactorial(n)-lfactorial(n-r))
}
f_vec_a <- function(a,r,n){  # vectorised in 'a' but not in 'r'
	sapply(a,function(a){f_single(a,r=r,n=n)})
}

f_range <- function(a,possible,n){ # r1=4 -> came fifth [four b clones ahead]  
    out <- a*0
    for(i in (possible)){
        out <- out + f_vec_a(a,i-1,n-1) }
    return(out)
}
```

verify verify verify

```{r verifyverifyverify}
f_range(0.2323423,1:10,10)-1
```

Suppose there are 30 in the class, we know our bro came between 5th
and 29th:

```{r plot4tto9th}
a <- seq(from=0,by=0.1,to=1)
L <- f_range(a,possible=5:29,30)
plot(a,L,type='b')
plot(a,log(L),type='b')
abline(h=-2)
```

What I want to do now is compare students with different grades:

```{r}
M <- t(matrix(c(
    "4-9","10-14", "50",
    "5-10","1-4", "23",
    "23-40","23-40", "100",
    "1-3","5-9", "100"),3,4))
rownames(M) <- c("Algebra","Calculus","Geometry","Topology")
colnames(M) <- c("Alice range","Bob range", "class size")
kable(M,caption='asdf')
```

```{r}
alice <- t(matrix(c(
    4,9,50,
    5,10,23,
    23,40,100,
    1,3,100),nrow=3))
dimnames(alice) <- list(topic=rownames(M),alice_results=c("low","high","class"))

bob <- t(matrix(c(
    30,34,50,
    17,19,23,
    70,90,100,
    66,88,100),nrow=3))
dimnames(bob) <- list(topic=rownames(M),bob_results=c("low","high","class"))
```

```{r showalicebob}
alice
bob
```

```{r definemakelike}
makelike <- function(a,M){
    L <- a*0 + 1
    for(i in seq_len(nrow(M))){
        L <- L * f_range(a,M[i,1]:M[i,2],M[i,3])
    }
    return(L)
}
```

```{r plotalicebobsep}
s <- seq(from=0,to=1,by=0.02)
La <- makelike(s,alice)
Lb <- makelike(s,bob)
plot(s,La/max(La,na.rm=TRUE))
plot(s,Lb/max(Lb,na.rm=TRUE))
```

```{r deflikeboth}
Likeboth <- function(a){makelike(a,alice)*makelike(a,bob)}
plot(s,Likeboth(s),col='red',pch=16)
optimize(Likeboth,c(0.4,0.6),maximum=TRUE)
```

```{r contouralicebob,cache=TRUE}
s_alice <- seq(from=0.1,to=0.95,by=0.01)
s_bob   <- seq(from=0.1,to=0.95,by=0.01)
like_sep <- function(v){log(makelike(v[1],alice)) + log(makelike(v[2],bob))}
L <- apply(as.matrix(expand.grid(s_alice,s_bob)),1,like_sep)
L <- L - max(L)
```

```{r contourbothalicebob}
par(pty='s')
contour(s_alice,s_bob,
        matrix(L,length(s_alice),length(s_bob)),
        asp=1,levels = -(0:7)
        )
abline(0,1)

```

```{r dobothoptims,cache=TRUE}
H_alt <- optim(c(0.8,0.4),like_sep,control=list(fnscale=-1))
H_null <- optim(c(0.8,0.4),function(v){
    v[2] <- v[1]
    like_sep(v)},control=list(fnscale=-1))
pchisq(2*(H_alt$value-H_null$value),df=1,lower.tail=FALSE)
```