### 2024-03-11 inst/formula1_start_vs_finish.Rmd
more work on formula 1 start vs finish position

---
title: "formula1_start_vs_finish.Rmd"
output: html_document
date: "2024-03-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

Formula 1 grand prix racing, 2023 season.

```{r loadfinish}
setwd("~/rstudio/hyper2/inst")
finish <- read.table("formula1_2023.txt",header=TRUE)
finish <- finish[,seq_len(ncol(finish)-1)]
finish[,1:5]
```

```{r loadgrid}
grid <- read.table("formula1_starting_2023.txt",header=TRUE)
grid[,1:5]
```

```{r definemakeH3}
`makeH3` <- function(i,lambda){
  suppressWarnings(a <- setNames(as.numeric(finish[,1]),rownames(finish)))
  suppressWarnings(b <- setNames(as.numeric(grid  [,1]),rownames(finish)))
  wanted <- !is.na(b)
  a <- a[wanted]
  b <- b[wanted]
  a[is.na(a)] <- 0 # did not finish
  H <- ordervec2supp(a)
  for(i in seq_along(b)){H[names(b)[i]] <- as.weight(lambda^(i-1))}
  return(H)
}

likelam <- function(lambda){
  H <- hyper3()
  for(i in seq_len(ncol(finish))){H <- H + makeH3(1,lambda=lambda)}
  return(H)
}
```

```{r evaluatelamrange,cache=TRUE}

f <- function(lambda){
  H <- likelam(lambda)
  loglik(maxp(H,n=4),H)
}
f(0.9)
f(0.901)
```

```{r evalam,cache=TRUE}
lam <- seq(from=0.9,to=1.05,len=5)
like <- sapply(lam,f)
```

```{r plotlikelam}
plot(lam,like,type='b')
```