### 2024-03-22 inst/hankin-ejs.Rmd
more for git to ignore

---
title: "Recreation of figures and results in the manuscript"
author: "Robin Hankin"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("pracma")
library("magrittr")
```

This terse document reproduces the results presented in the associated
manuscript.

## Asymptotic likelihood expression {-}

```
var('B n r')
pretty_print_default(True)
X=log(B-1) + (
+ ((B+n-r-2)*log(B+n-r-2)-(B+n-r-2)+log(B+n-r-2)/2)
- ((B+n  -1)*log(B+n  -1)-(B+n  -1)+log(B+n  -1)/2)
)
X.taylor(n,Infinity,2)
```

It returns this:

\[
-(r+1)\log n
-\frac{(2B-3)r-r^2+2B-2}{2n}
-\frac{3(2B-3)r^2-6B^2-6(B^2-3B+2)r+12B-5}{12n^2}
+\log(B-1)+\mathcal{O}(n^{-3})
\]

Note that if we use instead $n~\sim n^ne^{-n}\sqrt{2\pi
n}\left(1+\frac{1}{12n}\right)$, then only terms of
$\mathcal{O}(n^{-2})$ and above are changed; the first-order terms and
indeed the zeroth order terms are unaffected.  Anyway, after
simplifying and taking an asymptotic expansion to order $n^{-1}$, we
find, to first order,

\[
\mathcal{S}=\log\mathcal{L}_{r,n}(a)=\log(B-1)-B\frac{r+1}{n}
+ K + \mathcal{O}\left(n^{-2}\right)
\]

## Definition of likelihood and log-likelihood in R {-}

```{r,label=definefsingle}
f_single <- function(a,r,n,log=FALSE){  # not vectorised
	B <- 1/(1-a)
	if(log){
		out <- log(B-1) - lgamma(B+n) + lgamma(B+n-r-1)
	} else {  # not-log
		out <- (B-1)/prod(B+(n-r-1):(n-1))
	}
   return(out)
}

f_vec_a <- function(a,r,n, log=FALSE){  # vectorised in 'a' but not in 'r'
	sapply(a,function(a){f_single(a,r=r,n=n,log=log)})
}

f_vec_arn <- function(a,r,n,log=FALSE){
	## vectorized in 'a'; treats 'r' and 'n' as
	## vectors of independent observations

	M <- cbind(r,n)

	if(log){
		out <- 0
		for(i in seq_len(nrow(M))){out <- out + f_vec_a(a,r=M[i,1],n=M[i,2],log=TRUE)}
	} else {
		out <- 1
		for(i in seq_len(nrow(M))){out <- out * f_vec_a(a,r=M[i,1],n=M[i,2],log=FALSE)}
	}
	return(out)
}

fapprox <- function(a,r,n,log=FALSE){
	if(log){
		return(log(B-1)-B*(r+1)/n)
	} else {
		return(B-1 - (r+1)*exp(B)/n)
	}
}
```

```{r label=defdiff}
fdash <- function(a,r,n,log=TRUE){
	B <- 1/(1-a)
	out <- (1/(B-1) - psigamma(B+n) + psigamma(B+n-r-1))*B^2
	if(!log){out <- out * f_vec_arn(a,r,n,log=FALSE)}
	return(out)
}
```

Now verify `fdash()`:

```{r label=verifydiff}
a <- 0.34
d <- 1e-3
r <- 4
n <- 10
c(
  numerical  = (f_single(a+d/2,r,n,log=TRUE)-f_single(a-d/2,r,n,log=TRUE))/d,
  analytical = fdash(a,r,n)
)

c(
  numerical  = (f_single(a+d/2,r,n,log=FALSE)-f_single(a-d/2,r,n,log=FALSE))/d,
  analytical = fdash(a,r,n,log=FALSE)
)
```

## Maximum likelihood estimation

Use numerical optimization [`optim()`] to find the maximum likelihood
point [R function `MLE()`] and, as a check, use calculus `uniroot()`
to find the point of zero derivative; R function `MLEa()`] to find the
maximum likelihood point:

```{r,label=defmle}
MLE <- function(r,n,give=FALSE){
	d <- 1e-6
	out <- optimize(f_vec_arn,c(d,1-d),r=r,n=n,log=TRUE,maximum=TRUE)
	if(!give){out <- out$maximum}
	return(out)
}

MLEa <- function(r,n,give=FALSE,small=1e-4){
	out <- uniroot(f=function(a){fdash(a,r,n)},interval=c(small,1-small))
	if(!give){out <- out$root}
	return(out)
}

showdiff <- function(x,y){c(way1=x,way2=y,diff=x-y,logratio=log(x/y))}

rbind(
showdiff(MLE(1,9),MLEa(1,9)),
showdiff(MLE(3,9),MLEa(3,9)),
showdiff(MLE(3,7),MLEa(3,7))
)
```

```{r usestirling}
rbind(
showdiff(MLE(10, 100), 100 / 111),
showdiff(MLE(10, 500), 500 / 511),
showdiff(MLE(10,1000),1000 /1011),
showdiff(MLE(10,5000),5000 /5011)
)
```

```{r usestirlinga}
rbind(
showdiff(MLEa(10, 100), 100 / 111),
showdiff(MLEa(10, 500), 500 / 511),
showdiff(MLEa(10,1000),1000 /1011),
showdiff(MLEa(10,5000),5000 /5011)
)

```

Now verification

```{r,label=verifytwo}
a <- 0.23423434
b <- 1-a
B <- 1/b
n <- 9
r <- 3
c(
way1 = b^3*a/((a+9*b) * (a+8*b) * (a+7*b) * (a+6*b)),
way2 = (B-1)/prod(B+(5:8)),
way3 = f_vec_arn(a,r,n,log=FALSE)
)
```

## Produce Maximum Likelihood diagram, Figure 1 {-}

```{r,label=makejj}
out <- list()
for(n in 1:10){
  out[[n]]  <- rep(NA,n+1)
  for(r in 0:n){
    if(r==0){
      jj <- 1
    } else if(r==n){n
      jj <- 0
    } else {
      jj <- MLE(r,n)
    }
    out[[n]][r+1] <- jj
  }
}
out
```

```{r,label=plotjj}
plotterp <- function(...){
plot(NA,xlim=c(1,10),ylim=c(0,1),type="n",xlab="n",ylab=expression(hat(a)))
for(n in 1:10){
  for(r in 0:n){   points(n,out[[n]][r+1],pch=16) }
  for(r in 0:n){   text(n+0.2,out[[n]][r+1],r,cex=0.5,col='gray') }
}
}
plotterp()
pdf(file="dotprobs.pdf")
plotterp()
dev.off()
```

## Visuals {-}

Produce the rainbow-coloured likelihood plot

```{r,label=likevis,cache=F}
plotter <- function(...){
a <- seq(from=0,to=1,by=0.01)
n <- 8
jj <- f_vec_arn(a,3,n,log=FALSE)
plot(a,jj/max(jj,na.rm=TRUE),type='n',xlab=expression(a),ylab="likelihood")
grid()
rain <- rainbow(n+1)
for(r in seq(from=0,to=n)){
  y <- f_vec_a(a,r,n,log=FALSE)
  if(r==0){y[length(y)] <- 1}
  y <- y/max(y,na.rm=TRUE)
  if(r>0){y[length(y)] <- 0}
  points(a,y,type='l',lwd=4,col=rain[r+1])
}
abline(v=MLE(r=4,n=8))
text(0.07,0.95,"r=8",col=rain[9])
text(0.20,0.95,"r=7",col=rain[8])
text(0.26,0.91,"r=6",col=rain[7])
text(0.29,0.83,"r=5",col=rain[6])
text(0.33,0.76,"r=4",col=rain[5])
text(0.35,0.64,"r=3",col=rain[4])
text(0.39,0.53,"r=2",col=rain[3])
text(0.45,0.41,"r=1",col=rain[2])
text(0.63,0.22,"r=0",col=rain[1])
}

plotter()
pdf(file="ninelikes.pdf")
plotter()
dev.off()
```

### Alternative production of rainbow figure, as a check {-}

```{r usehyper3,cache=TRUE}
n <- 7
a <- seq(from=0,to=1,by=0.01)
fhyper3 <- function(r,n){
  out <- hyper3()
  out['a'] <- 1
  out['b'] <- r
  for(i in (n-r):n){
    out[c(a=1,b=i)] %<>% dec
  }
  return(out)
}

plot(NA,xlim=c(0,1),ylim=c(0,1),type='n')
M <- cbind(a=a,b=1-a)
for(r in 0:7){
  y <- loglik(M,fhyper3(r,n),log=FALSE)
  y <- y/max(y, na.rm=TRUE)
points(M[,1],y,type="l",lwd=8)
}
```

# Formula 1: Checo's career arc

```{r getnumgetfoc}
readstring <- function(year){read.table(paste("formula1_",year,".txt",sep=""))}
getfoc <- function(year,comp="Perez"){  # get focal competitor
  M <- as.matrix(readstring(year))
  out <- suppressWarnings(as.numeric(M[comp,seq_len(ncol(M)-1)]))
  out[is.na(out)] <- nrow(M)
  return(out)
}
getnum <- function(year){nrow(readstring(year))} # number of competitors

perez <- lapply(2011:2023,getfoc)
print(perez)
y <- unlist(lapply(seq_along(perez),function(i){mean(perez[[i]])}))
x <- 2011:2023
plot(x,y)
```

```{r checofirst}
checo_like <- function(a){
  out <- a*0
  for(year in 2011:2023){ out <- out + f_vec_arn(a,getfoc(year)-1,getnum(year)-1,log=TRUE) }
  return(out)
}

a <- seq(from=0.45,to=0.62,by=0.01)
cL <- checo_like(a)
cL <- cL - max(cL)
plot(a,cL,type='b')
abline(h=c(0,-2))
```

```{r definef1log}
f1_logistic <- function(vec){
  alpha <- vec[1]
  beta  <- vec[2]
  out <- 0
  for(year in 2011:2023){
     LO <- alpha + beta*(year-2011)
     strength <- exp(LO)/(1+exp(LO))
     out <- out + f_vec_arn(strength,getfoc(year)-1,getnum(year)-1,log=TRUE)
  }
  return(out)
}
```

```{r calculatelogisticcontou,cache=TRUE}
a <- seq(from=-0.8,to=0.0,by=0.01)
b <- seq(from=0,to=0.2,by=0.01)
jj <- as.matrix(expand.grid(a,b))
L <- apply(jj,1,f1_logistic)
L <- L-max(L)
L <- matrix(L,length(a),length(b))
```

```{r showfilledcwithdot}
L <- pmax(L,-40)
showchec <- function(...){
contour(a,b,L,xlab=expression(alpha),ylab=expression(beta),levels=-c(2,5*(1:5)))
points(-0.27,0.0813,pch=16,cex=2)
}
showchec()
pdf(file="showchecolike.pdf")
showchec(a,b,L)
dev.off()
```

```{r uselogf1}
f1_logistic(c(0,0))
f1_logistic(c(0,0.0001))
f1_logistic(c(0.001,0))
```

```{r optimizelof1,cache=TRUE}
o <- optim(c(-0.2,0.1),fn=f1_logistic, control=list(fnscale = -1),hessian=TRUE)
```

```{r showjj}
o
o$par
o$hessian
eigen(o$hessian)
best <- o$par
f1_logistic(best)

jj <- best
jj[1] <- jj[1]*1.01
f1_logistic(jj) - f1_logistic(best)

jj <- best
jj[1] <- jj[1]*0.99
f1_logistic(jj) - f1_logistic(best)

jj <- best
jj[2] <- jj[2]*1.00001
f1_logistic(jj) - f1_logistic(best)

jj <- best
jj[2] <- jj[2]*0.99
f1_logistic(jj) - f1_logistic(best)
```

```{r testlogf1,cache=TRUE}
o_free <- optim(c(-0.2,0.1),fn=f1_logistic, control=list(fnscale = -1),hessian=TRUE)
o_constrained <- optim(c(-0.2,0.1),fn=function(v){
v[2] <- 0
return(f1_logistic(v))}, control=list(fnscale = -1),hessian=TRUE)
```

```{r showpvalue}
o_free
o_constrained
pchisq(2*(o_free$value - o_constrained$value),df=1,lower.tail=FALSE)
```

# Athletics

```{r loadolymp}
O <- read.table("olympic_athletics_mens_100m.txt",header=TRUE)
head(O)
```

```{r analuyzelmp,cache=TRUE}
jjf <- function(x){f_vec_arn(x,O$rank,O$n,log=TRUE)}
system.time(optolym <- optimize(jjf, c(0.45,0.5), maximum=TRUE))
```

```{r}
optolym
jjf(0.5)
jjf(optolym$maximum)
(LR <- 2*(jjf(optolym$maximum)-jjf(0.5)))
pchisq(LR,df=1,lower.tail=FALSE)
```

```{r makeolymplike}
a <- seq(from=0.3,to=0.65,len=45)
L <- f_vec_arn(a,O$rank,O$n,log=TRUE)
Lmax <- jjf(optolym$maximum)
Lmax
L <- L-Lmax
```

```{r findcredint}
(a_lower <- uniroot(function(x){jjf(x)+2-Lmax},interval=c(0.3,0.5))$root)
(a_upper <- uniroot(function(x){jjf(x)+2-Lmax},interval=c(0.5,0.6))$root)
```

```{r plotolymp}
plotolymp <- function(...){
plot(a,L,type='l',ylab="log-likelihood",lwd=2)
abline(h=c(0,-2))
ahat <- optolym$maximum
segments(x0=ahat,y0=-1.5,y1=0)
text(ahat,-0.91,expression(hat(a)==0.573),pos=2)
abline(v=0.5,lty=3)

segments(x0=a_lower,y0=-1.5,y1=-3.5)
segments(x0=a_upper,y0=-1.5,y1=-3.5)

text(x=a_lower,y=-3,"0.380",pos=4)
text(x=a_upper,y=-3,"0.563",pos=2)
}
plotolymp()
pdf(file="plotolymp.pdf")
plotolymp()
dev.off()
```

## Education {-}

```{r focaleducation}
rank <- c( 9, 7, 2, 3,2 ,8 ,5 ,4 ,9 )
class_size <- c(12, 17,23, 9,13,14,13,12,15)
course <- c("rings and modules","group theory","calculus","linear algebra",
"differential equations","topology","special relativity","fluid mechanics","Lie algebra")
category=c("pure","pure","applied","applied","applied","pure","applied","applied","pure")
data.frame(course,category,rank,class_size)
```

```{r seppurefromapplied}
a <- seq(from=0.2,by=0.01,to=0.8)
wp <- category=='pure'
wa <- category=='applied'
```

Pure first

```{r dopurefurst}
Lpure <- f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)
Lpure <- Lpure - max(Lpure,na.rm=TRUE)
plot(a,Lpure,ylab='pure strength')
optimize(function(a){f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
optimize(function(a){f_vec_arn(a,rank[wa]-1,class_size[wa]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
```

```{r definesupp2ed}
supp2_ed <- function(vec,place1,place2,runners1,runners2){
  out <- 0
  M1 <- cbind(place1,runners1)
  M2 <- cbind(place2,runners2)
  for(i in seq_len(nrow(M1))){
    out <- out + f_vec_arn(vec[1],r=M1[i,1]-1,n=M1[i,2]-1,log=TRUE)
  }
  for(i in seq_len(nrow(M2))){
    out <- out + f_vec_arn(vec[2],r=M2[i,1]-1,n=M2[i,2]-1,log=TRUE)
  }
  return(out)
}
```

```{r,label=fishbish,cache=TRUE}
x <- seq(from=0.1,to=0.99,by=0.01)
jj <- as.matrix(expand.grid(x,x))
L <- apply(jj,1,supp2_ed,rank[wp]-1,rank[wa]-1,class_size[wp]-1,class_size[wa]-1)
L <- L - max(L,na.rm=TRUE)
L <- matrix(L,length(x),length(x))
```

```{r,label=contourstrength}
plotpureandapplied <- function(...){
par(pty='s')
contour(x,x,L,asp=1,xlim=range(x),ylim=range(x),levels=-(0:5),
xlab='pure strength',ylab='applied strength')
abline(0,1)
}
plotpureandapplied()
pdf(file="plotpureandapplied.pdf")
plotpureandapplied()
dev.off()
```

```{r,label=defsupp2}
jj1 <-
optim(
   c(.4,.8),
   function(vec){
      supp2_ed(vec,rank[wp]-1,rank[wa]-1,class_size[wp]-1,class_size[wa]-1)},
   control=list(fnscale = -1)
)
jj1
```

```{r,label=findmaxlike}
jj2 <-
optimize(function(v){supp2_ed(c(v,v),rank[wp]-1,rank[wa]-1,class_size[wp]-1,class_size[wa]-1)},
c(0.1,0.9),maximum=TRUE)
jj2
```

```{r,label=shochisw}
pchisq(2*(jj1$value-jj2$objective),lower.tail=FALSE,df=1)
```