### 2024-03-28 inst/parkrun_idea.Rmd
tweaks

---
title: "Underdispersion in the context of community athletics"
output: bookdown::html_document2
author: "Robin K. S. Hankin"
bibliography: "hyper2.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.

```{r deff}
library("hyper2")

f <- function(n,alpha){
  jj <- 1/alpha^(1:n)
  return(jj/sum(jj))
}

getlike <- function(v){
  alpha <- v[1]
  beta <- v[2]
 c(
tabulate(replicate(1000,which(rrace(setNames(f(26,alpha),letters))==letters[beta])),nbins=26)
,alpha,beta)
}
```

```{r calculateLikemat,cache=FALSE}
alpha_try <- seq(from=1,to=6,len=20)
beta_try <- 1:26
calculate_from_scratch <- FALSE
if(calculate_from_scratch){
    M <- as.matrix(expand.grid(alpha=alpha_try, beta=beta_try))
    system.time(L <- t(apply(M,1,getlike)))
    colnames(L) <- c(letters,'alpha','beta')
    write.table(L,file="likelihoodmatrix.txt")
} else {
    L <- read.table("likelihoodmatrix.txt")
}
L_integer <- L  # counts
L[,1:26] <- L[,1:26]/mean(rowSums(L[,1:26]))
head(L_integer)
```

```{r makecontL}
colSums(L)
mL <- function(i){matrix(L[,i],length(alpha_try),length(beta_try))}
filled.contour(alpha_try,beta_try,mL(1))
filled.contour(alpha_try,beta_try,mL(5))
filled.contour(alpha_try,beta_try,mL(10))
filled.contour(alpha_try,beta_try,mL(15))
filled.contour(alpha_try,beta_try,mL(20))
filled.contour(alpha_try,beta_try,mL(26))
```

```{r}
#plot(table(L_integer[,1:26]))
```

## Now 10 in-silico races:
```{r doinsilico,cache=FALSE}
alphatrue <- 2.3
betatrue <- 4  
set.seed(1)
D <- replicate(10,which(rrace(setNames(f(26,alphatrue),letters))==letters[betatrue]))
sort(D)
```

```{r trylikeinsilico,cache=FALSE}
likelihood <- apply(L[,D],1,prod) + 0.00
likelihood <- matrix(likelihood,length(alpha_try),length(beta_try))
```

```{r showlike}
round(alpha_try,2)
matplot(alpha_try,likelihood,type='b')
matplot(beta_try,t(likelihood),type='b')
matplot(alpha_try,log(likelihood)-log(max(likelihood)),type='b',ylim=c(-20,0))
matplot(beta_try,log(t(likelihood))-log(max(likelihood)),type='b',ylim=c(-20,0))
filled.contour(alpha_try,beta_try,(likelihood))
table(likelihood==0)
plot(f(26,alphatrue))
abline(v=betatrue)
contour(alpha_try,beta_try,likelihood)
abline(v=alphatrue)
abline(h=betatrue)
contour(alpha_try,beta_try,log(likelihood) - log(max(likelihood)),levels=-(0:10))
abline(v=alphatrue)
abline(h=betatrue)
```

\subsection{Park run}

"Parkrun" is a distributed community initiative that organises
weekly timed 5 km runs/walks in parks worldwide~\cite{hindley2020}.  A
typical event will have 200 participants.  Table~\ref{parkruntable}
shows the author has completed a total of 15 parkruns to date, and
from the first pair of numbers we see that 238 runners attended that
particular parkrun, of whom the author placed 173.

\begin{table*}[t]
  \caption{Parkrun results}
\label{parkruntable}
\begin{tabular}{ccccccccccc}
\hline
rank   &173  & 165 & 172 & 199 & 181 & 229 & 177 & 222 & 206 & 142 \\
runners&238  & 238 & 196 & 242 & 242 & 318 & 259 & 305 & 297 & 241 \\ \\ \\
rank   & 118 & 224 & 128 & 115 & 183 &     &     &     &     &\\
runners& 179 & 338 & 203 & 245 & 254 &     &     &     &     &\\
\hline
\end{tabular}
\end{table*}

We may consider the author to have an unknown generalized Bradley
Terry strength $a$: Figure~\ref{parkrunsupport} shows a support curve
for $a$ with the evaluate, $\hat{a}\simeq 0.448$ shown.  We also see
two units of support~\cite{edwards1992} shown as a dotted line
illustrating a support interval of about 0.315 - 0.567; we may be
reasonably confident that the author's true BT strength lies in this
range.  In particular, note that a reasonable $H_0\colon
a=\frac{1}{2}$ may not be rejected, the support for $H_0$ being only
susceptible to very minor improvement [$\simeq 0.351$] by moving to
the evaluate.

```{asis}
\begin{figure}[t]
\includegraphics[width=4in]{plotparkrun}  % plotparkrun.pdf made in very_simplified_likelihood.Rmd
\caption{Author's strength\label{parkrunsupport}}
\end{figure}
```