### 2024-04-12 inst/formula1_sequential_correlation.Rmd
start work on sequential correlatoin

---
title: "formula1: sequential correlation"
output: html_document
date: "2024-03-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

Formula 1 grand prix racing, 2022 season.  We wonder whether placing
well at a race helps the team in some way in the next race.  The
"help" might be an $\alpha$ weighting term for just the winners, or a
$\alpha^n$ term for placing $n$-th in the previous race.  I would have
thought $\alpha>1$ [the help accruing from a win is positive] but it
is possible that $\alpha<1$ which would correspond to it being a
hindrance in some way.  Perhaps winning leads to overconfidence or
complacency in the next one?

```{r loadfinish}
setwd("~/rstudio/hyper2/inst")
finish <- read.table("formula1_2023.txt",header=TRUE)
finish <- finish[,seq_len(ncol(finish)-1)]
finish[1:9,1:5]
```

```{r starthyper2}
f <- function(i){ # this is column 'i' of finish
    ordertable2supp(finish[,i,drop=FALSE])
}
g <- function(i){
    suppressWarnings(out <- as.numeric(finish[,i,drop=TRUE]))
    out[is.na(out)] <- length(out)
    names(out) <- rownames(finish)
    return(out)
}
```

```{r muckaboutwithlikelihood}
	weightedlikelihood <- function(alpha){
		out <- hyper3()
		for(i in seq(from=2,to=ncol(finish))){
			H <- f(i)
			previous <- g(i-1)
			H[names(previous)] <- as.weight(alpha^previous)
			out <- out + H
		}
		return(out)
}
```

```{r doweightedlike,cache=TRUE}
H <- weightedlikelihood(0.9)
mH <- maxp(H)
```

```{r showweightedlike}
mH
dotchart(mH)
loglik(mH,H)
```

```{r doweightedlike91,cache=TRUE}
H91 <- weightedlikelihood(0.91)
mH91 <- maxp(H91)
```

```{r showweightedlike91}
mH91
dotchart(mH91)
loglik(mH91,H91)
```

```{r definelikealpha}
likealpha <- function(alpha){
	H <- weightedlikelihood(alpha)
	mH <- maxp(H)
	return(loglik(mH,H))
}
```

```{r uselikealpha,cache=TRUE}
alpha <- seq(from=0.95,to=1.02,by=0.01)
likes <- sapply(alpha,likealpha)
```

```{r plotalpha}
plot(alpha,likes-max(likes),type='b')
abline(h=c(0,-2))
abline(v=1)
```