### 2024-05-28 inst/grad_weight.Rmd
pursuant to commit 1d7c34c

---
title: "Differentiation with respect to weights for a `hyper3` object"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("disordR")
library("magrittr")
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
We will start with a simple example.  Consider two players $p_1$ and
$p_2$ who play tennis say, $a+b$ times with $p_1$ winning $a$ times
and $p_2$ winning $b$ times.  We use the following likelihood function:

$$\frac{(wp_1)^ap_2^b}{(wp_1+p_2)^{a+b}}$$

where $w$ represents some form of competitive advantage present in the
$a+b$ trials but perhaps absent in subsequent trials [this might be
home ground advantage in soccer or the first-mover advantage in
chess].  We may represent the log-likelihood in `hyper` idiom as
follows:

```{r,label=definefw}
f <- function(w,a=5,b=5){
    H <- hyper3()
    H[c(p1=w)] <- a
    H[c(p2=1)] <- b
    H[c(p2=1,p1=w)] <- -(a+b)
    H
}
```

Thus

```{r,label=showfinuse}
f(w=1.1)
```

Here I consider how we may differentiate this support function with
respect to $w$:

```{r,label=definegws}
grad_weight_single <- function(H, probs, wpv, wanted){
  jj <- H[wpv]
  alpha <- drop(powers(jj))
  powers(jj) <- -1
  jj[setNames(1,wanted)] <- +1
  pnames(jj) <- pnames(H)
  alpha*loglik(probs,jj,log=FALSE)
  }
```

Function `grad_weight_single()` takes a `hyper3` object `H`, a vector
of non-negative probabilities `p` with unit sum, a named vector `wpv`
corresponding to one of the brackets of `H`, and a character vector
`wanted` that specifies which term(s) in bracket `wpv` have a weight
that is to be differentiated with respect to.  If the weight appears
in multiple brackets we need to call `grad_weight_single()` once for
each bracket the weight appears in, and add the results.  For example:

```{r,label=testthederiv}
w <- 1.1
dw <- 1e-4
H <- f(w)
probs <- c(p1=0.35, p2=0.65)

wpv <- c(p1=w)
wanted <- "p1"
d1 <- grad_weight_single(H, probs, wpv, wanted)

wpv <- c(p2=1,p1=w)
wanted <- "p1"
d2 <- grad_weight_single(H, probs, wpv, wanted)

LHS <- (d1+d2)                                          # differential from grad_weight_single()
RHS <- (loglik(probs,f(w+dw/2)) - loglik(probs,f(w-dw/2)))/dw  # differential via numerics
print(c(lhs=LHS,rhs=RHS,diff=RHS-LHS))
```

above we see very close agreement.

# Slightly more complicated example

```{r slightlymorecomplicated}
H <- hyper3()
f <- function(w){
  H <- hyper3()
  H[c(p1=1 , p2=w , p3=w )] <- -5
  H[c(p1=1)] <- 2
  H[c(p2=2)] <- 1
  H[c(p2=w,p3=w)] <- 2
  return(H)
}

w <- 1.4
dw <- 1e-6
(HH <- f(w))
probs <- c(p1=0.1,p2=0.6,p3=0.3)
d1 <- grad_weight_single(HH, probs, c(p1=1 , p2=w , p3=w), "p2")
d2 <- grad_weight_single(HH, probs, c(p1=1 , p2=w , p3=w), "p3")
d3 <- grad_weight_single(HH, probs, c(       p2=w , p3=w), "p2")
d4 <- grad_weight_single(HH, probs, c(       p2=w , p3=w), "p3")
LHS <- d1+d2+d3+d4
RHS <- (loglik(probs,f(w=w+dw/2)) - loglik(probs,f(w=w-dw/2)))/dw
  c(RHS,LHS,LHS-RHS)
```

## An example from chess

Two chess players, 1, 2, and 3, play chess.  Our observations (chess
games, say) comprise $a+b+c$ games in which $p_1$ played White,
winning $a$ games, drawing $b$ and losing $c$ [we write $+a=b-c$]; and
$d+e+f$ games in which $p_2$ played White, with $+d=e-f$.  There is
also a single game between $p_3$ (white) who beat $p_1$ (black).
Consider the following likelihood function:

$$
\frac{(w p_1)^a\cdot(D(p_1+p_2))^b\cdot p_2^c}{(w p_1+D(p_1+p_2)+p_2)^{a+b+c}}
\frac{(w p_2)^d\cdot(D(p_1+p_2))^e\cdot p_1^f}{(w p_2+D(p_1+p_2)+p_1)^{d+e+f}}
\frac{ w p_3}{wp_3+p_1}
$$

```{r chessexample}
w <- 1.2
D <- 0.3

brackfunc <- function(w,D){
  list(
  b1 = c(p1=w),
  b2 = c(p1=D, p2=D),	
  b3 = c(p2=1),
  b4 = c(p1=w+D,p2=1+D),
  b5 = c(p2=w),
  b6 = c(p1=1),
  b7 = c(p1=1+D,p2=w+D),
  b8 = c(p3=w),
  b9 = c(p3=w,p1=1)
  )
}

`f2` <- function(w,D, a=3,b=12,c=1,d=5,e=6,f=6){
  H <- hyper3()	      
  brackets <- brackfunc(w,D)
  H[brackets[[1]]] %<>% inc(a)	
  H[brackets[[2]]] %<>% inc(b)	
  H[brackets[[3]]] %<>% inc(c)	
  H[brackets[[4]]] %<>% dec(a+b+c)
  H[brackets[[5]]] %<>% inc(d)	
  H[brackets[[2]]] %<>% inc(e)	 # sic, the draw brackets are the same
  H[brackets[[6]]] %<>% inc(f)	
  H[brackets[[7]]] %<>% dec(d+e+f)
  H[brackets[[8]]] %<>% inc
  H[brackets[[9]]] %<>% dec
  return(H)
}

  H <- f2(w=w, D=D)

```

$$
\frac{(w p_1)^a\cdot(D(p_1+p_2))^b\cdot p_2^c}{(w p_1+D(p_1+p_2)+p_2)^{a+b+c}}
\frac{(w p_2)^d\cdot(D(p_1+p_2))^e\cdot p_1^f}{(w p_2+D(p_1+p_2)+p_1)^{d+e+f}}
\frac{ w p_3}{wp_3+p_1}
$$

Differentiate with respect to `w`:

```{r diffwrtw}
probs <- c(p1=0.15 , p2=0.45, p3=0.4)
diff <- (
 grad_weight_single(H, probs, c(p1=w), "p1") +
 grad_weight_single(H, probs, c(p1=w+D,p2=1+D), "p1") +
 grad_weight_single(H, probs, c(p2=w), "p2") +
 grad_weight_single(H, probs, c(p1=1+D,p2=w+D), "p2") +
 grad_weight_single(H, probs, c(p3=w), "p3")+
 grad_weight_single(H, probs, c(p3=w,p1=1),"p3")
 )

LHS <- diff  # using grad_weight_single()
RHS <- (loglik(probs,f2(w=w+dw,D=D)) - loglik(probs,f2(w=w,D=D)))/dw  # differential via numerics
c(LHS,RHS,LHS-RHS)
```

We may even be a little clever and use a central difference:

```{r cleverusecentral}
RHS2 <- (loglik(probs,f2(w=w+dw/2,D=D)) - loglik(probs,f2(w=w-dw/2,D=D)))/dw  # differential via numerics
c(LHS,RHS2,LHS-RHS2)
```

If we wish to differentiate with respect to `D` we need a different technique (TBA).

### References {-}