### 2024-08-13 inst/table_tennis_serve.Rmd
pair3()

---
title: "The importance of serve in table tennis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
In this vignette, object `H` is created.  It is identical to object
`table_tennis` in the package; to see it type `data(table_tennis)` at
the command line, and to see the help page, type `help(table_tennis)`.
Suppose we have three players, "A","B", and "C" who play table tennis.
We seek to quantify the effect of the serve.  The scoreline is as
follows:

* A vs B, A serves, 5-1
* A vs B, B serves, 1-3
* A vs C, A serves, 4-1
* A vs C, C serves, 1-2

(note that B does not play C).  This dataset is easily analysed with
the `hyper2` package although the model implicitly assumes that the
serve strength does not depend on who is serving.  The first step is
to define a `hyper2` object:

```{r,label=defineH,message=FALSE}
library("hyper2",quietly=TRUE)
library("magrittr")
(H <- hyper2(pnames=c("S","a","b","c")))
```

so the likelihood function is uniform.  We have players `a`, `b`, `c`
(players' names are not capitalised in R idiom for clarity) and a
reified entity `S` representing the strength of the serve.  Now we
have to input the data, and we will go through the four scorelines
above, one by one.  Note that this approach implicitly assumes
independence of scores.

### A vs B, A serves, score 5-1: {-}
```{r,label=abaserve}
H[c("a","S")]     %<>% "+"(5  ) # a+S win 5 games
H["b"]            %<>% "+"(  1) # b wins 1 game
H[c("a","b","S")] %<>% "-"(5+1) # a+b+S play 1+5=6 games
H
```

The likelihood function is of the form
$\frac{\left(a+S\right)^5b}{\left(a+b+S\right)^6}$.  We can add the
other observations to $H$:

* A vs B, B serves, score 1-3:
```{r,label=abbserve}
H[c("b","S")]     %<>% "+"(3  ) # b+S win 3 games
H["a"]            %<>% "+"(  1) # a wins 1 game 
H[c("a","b","S")] %<>% "-"(3+1) # a+b+S play 1+3=4 games
```

* A vs C, A serves, score 4-1:
```{r,label=acaserve}
H[c("a","S")]     %<>% "+"(4  ) # a+S win 4 games
H["c"]            %<>% "+"(  1) # c wins 1 game 
H[c("a","c","S")] %<>% "-"(4+1) # a+c+S play 1+4=5 games
```

* A vs C, C serves, score 1-2:
```{r,label=accserve}
H["a"]            %<>% "+"(1  ) # a wins 1 game
H[c("c","S")]     %<>% "+"(  2) # c+S win 2 games
H[c("a","c","S")] %<>% "-"(1+2) # a+c+S play 1+2=3 games
```

For the entire dataset we have

```{r,savetime}
table_tennis <- H
table_tennis
```

Here $H$ is a likelihood function on $a,b,c,S$, algebraically
proportional to

\[
\frac{(S+a)^9(S+b)^3(S+c)^2a^2bc}{(S+a+b)^{10}(S+a+c)^8}
\]

# Maximum likelihood estimate for the strengths

We now find the unconstrained MLE:

```{r,label=maxptennis,cache=TRUE}
table_tennis_maxp <- maxp(table_tennis)
table_tennis_maxp
```

# Test hypothesis that serve strength S is zero

```{r, label=testservezero,cache=TRUE}
specificp.test(table_tennis,"S",0.01)
```

Thus the tail area, pval, is about 0.022 which gives us a significant
result: the serve has nonzero strength.

## Assessment of equal player strength.

We now assess the hypothesis that the three (real) players are of zero
strength:  $H_E\colon p_A=p_B=p_C$.

```{r, label=equalplayerstrength,cache=TRUE}
samep.test(table_tennis,c("a","b","c"))
```

Above, `samep.test()` complains because it is performing a
one-dimensional optimization (we may ignore this issue).  The result
of the test is that there is no evidence to suggest a difference
between players `a`, `b`, and `c`.

# Profile likelihood for serve strength

We can give a profile likelihood for $S$ by maximizing the likelihood
of $(S,p_a,p_b,p_c)$ subject to a particular value of $S$ (in addition
to the unit sum constraint).  The following function defines a profile
likelihood function for $S$:

```{r,label=svals,cache=TRUE}
S_vals <- seq(from=0.02,to=0.85,len=30)  # possible values for S
plot(S_vals,profsupp(table_tennis,"S",S_vals))
abline(h=c(0,-2))
```

So we can see that the credible range is from about 0.07-0.8 (and in
particular excludes zero).

# Assessment of B vs C

The maximum likelihood estimate for strengths allows us to assess what
might happen when B plays C (which was not observed).  For
convenience, here is the evaluate:

```{r,label=maxtab,cache=TRUE}
maxp(table_tennis)
```

If B serves, we would have $B+S:C\simeq 0.14+0.45:0.15\equiv 59:15$,
or B wins with probability of about $0.78$.  If C serves we have
$B:C+S\simeq 0.14:0.16+0.45\equiv 14:61$, so B wins with a probability
of about $0.19$.

# Profile likelihood for B vs C

The profile likelihood technique can be applied to the question of the
strengths of players B and C, even though no competition between the
players was observed:

```{r,label=proffli,dpi=72,cache=TRUE}
proflike <- function(bc,give=FALSE){
  B <- bc[1]
  C <- bc[2]
  if(B+C>=1){return(NA)}

  objective <- function(S){     #

    jj <- c(S,A=1-(S+B+C),B,C)  # B,C fixed
    loglik(indep(jj),H)         # no minus: we use maximum=T in optimize()
  }

  maxlike_constrained <-  # single DoF is S [B,C given, A is fillup]
    optimize(             # single DoF -> use optimize() not optim()
        f = objective,         
        interval = c(0,1-(B+C)),
        maximum = TRUE
    )

  if(give){
    return(maxlike_constrained)
  } else{
    return(maxlike_constrained$objective)
  }
}

small <- 1e-5
n <- 50
p <- seq(from=small,to=1-small,length=n)
jj <- expand.grid(p,p)
support <- apply(jj,1,proflike)
support <- support-max(support,na.rm=TRUE)
supportx <- support
dim(supportx) <- c(n,n)

x <- jj[,1]+jj[,2]/2
y <- jj[,2]*sqrt(3)/2
plot(x,y,cex=(support+6)/12, 
   pch=16,asp=1,xlim=c(-0.1,1.1),ylim=c(-0.2,0.9),
   axes=FALSE,xlab="",ylab="",main="profile likelihood for B,C")
polygon(x=c(0,1/2,1),y=c(0,sqrt(3)/2,0))
text(0.07,-0.04,'A+S=1, B=0, C=0')
text(0.50,+0.90,'A+S=0, B=1, C=0')
text(0.93,-0.04,'A+S=0, B=0, C=1')
```

But what one is really interested in is the relative strength of
player B and player C.  We are indifferent between $p_B=0.1,p_C=0.2$
and $p_B=0.3,p_C=0.6$.  To this end we need the profile likelihood
curve of the log contrast, $X=\log\left(p_B/p_C\right)$.

```{r,label=logcont}
logcontrast <- apply(jj,1,function(x){log(x[1]/x[2])})
plot(logcontrast,support,xlim=c(-5,5),ylim=c(-4,0))
abline(h=-2)
```

From the graph, we can see that the two-units-of-support interval is
from about $-3$ to $+3$ which would translate into a probability of B
winning against C being in the interval
$\left(\frac{1}{1+e^{3}},\frac{1}{1+e^{-3}}\right)$, or about
$(0.047,0.953)$.  However, it is not clear if $p_B$ and $p_C$ are
meaningful in isolation, as a real table tennis point has to have a
server and the value of $S$ is itself uncertain.

# Weighted likelihoods

We now try `hyper3` formalism:

```{r,label=calcproflike,cache=TRUE}
M_home <- matrix( c(NA, 5 , 4, 3, NA, 0, 2,0,NA),byrow=TRUE,3,3)
jj <- letters[1:3]
dimnames(M_home) <- list(jj,jj)

M_away <- matrix( c(NA, 1 , 1, 1, NA, 0, 1,0,NA),byrow=TRUE,3,3)
jj <- letters[1:3]
dimnames(M_away) <- list(jj,jj)
M_home
M_away

home_away3(M_home,M_away,lambda=1.8)

f <- function(lam){
    maxp(home_away3(M_home,M_away,lambda=lam),give=TRUE)$value
}
```

```{r,makelamlike,cache=TRUE}
lam <- exp(seq(from=log(0.6),to=log(20),len=9))
L <- sapply(lam,f)
```

```{r,optimlamlike,cache=TRUE}
maxf <- optimize(f,c(2,5),maximum=TRUE)
maxf
```

```{r plotlikelam}
L <- L-max(L)
plot(lam,L,type='b',pch=16,log="x")
abline(h=c(0,-2))
```

How does the situation change if we have some more observations?
Suppose `c` plays at home to `d`, and loses.  If we use
$\hat{\lambda}\simeq 3.35$:

```{r, label=newdatacd,cache=TRUE}
(Hnew <- home_away3(M_home,M_away,lambda=3.35) + pair3(c=0,d=1,lambda=3.35))
maxp(Hnew)
```

Above we see that our _estimate_ would be that `d` has a BT strength
of 1 and the other players have a strength of zero.  Is there strong
evidence that the teams' strength differ?

```{r,label=newdatatest,cache=TRUE}
equalp.test(Hnew)
```

Above we see that there is no evidence that the teams' strengths do in
fact differ.

### Package dataset

Following lines create `table_tennis.rda`, residing in the `data/`
directory of the package.

```{r,label=savetabletennisdataset}
save(table_tennis,file="table_tennis.rda")
```

### References {-}