### 2024-10-18 inst/ternaryplot_hyper2.Rmd
fixes #137

---
title: "Ternary plots for `hyper2` and `hyper3` samples"
author: "Robin Hankin"
output: html_document
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, label=hexstickerplotter, out.width='15%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

(takes about fifteen minutes to run without cache).

To cite the `hyper2` package in publications please use
@hankin2017_rmd.  Here I show how to use the excellent `Ternary`
package [@smith2017] to generate triangular plots for `hyper2` and
`hyper3` likelihood functions with size 3.

First load the `Ternary` and `hyper2` packages:

```{r loadlibraries}
suppressPackageStartupMessages(library("Ternary"))
suppressPackageStartupMessages(library("hyper2"))
```

# The `chess` dataset (`hyper2`)

Now generate a synthetic dataset.  Here we will sample from the
posterior of a uniform prior with a likelihood function specified by
the `chess` object (this takes quite a long time):

```{r generatedata, cache=TRUE}
set.seed(0)
xx <- rp(1000,chess)
```

Using the `TernaryPlot()` and `TernaryPoints()` functions is straightforward:
```{r plotdata}
jj <- pnames(chess)
TernaryPlot(atip=jj[1],btip=jj[2],ctip=jj[3])
TernaryPoints(xx,cex=0.5)
TernaryPoints(chess_maxp,pch=16,col="red")
```

Above we see black circles being samples from the posterior and the
red point is the maximum likelihood estimate for the three players'
strengths.  We can adapt this technique to higher-dimensional `hyper2`
objects.

# The `icons` dataset (`hyper2`)

Consider the `icons` likelihood function:

```{r showicons}
icons
```

We can sample from this object and plot it in different ways:

```{r samplefromicons,cache=TRUE}
n <- 10000
x <- rp(n,icons)
x[sample(n,5),]
```

```{r, iconsmatplot}
matplot(x,type="l",lty=1)
legend("topright",col=1:6,lty=1,legend=pnames(icons))
```

```{r, selectjustbigones}
wanted <- c(1,3) # NB,PB
xx <- cbind(x[,wanted],fillup=rowSums(x[,-wanted]))
TernaryPlot(atip="NB",btip="PB",ctip="other")
TernaryPoints(xx,cex=0.5)
```

# Synthetic race data (`hyper3`)

Here we use `rrace3()` to generate repeated results from a race
between three runners.

```{r singlerrace3example}
set.seed(0)
(randomrace <- noquote(rrace3(pn = c(a=6, b=3, c=8), ps = c(a=0.3,b=0.5,c=0.2))))
(S <- ordervec2supp3(randomrace))
maxp(S)
```

Above we see `randomrace` being an order vector for a race with six
`a` clones, 3 `b` clones and 2 `c` clones of PL strengths 0.3, 0.5 and
0.2 respectively.  Thus the first place was a `b`, second place was an
`a`, third place a `c`, and so on.  Object `S` is support function for
this realisation and we also see the maimum likelihood estimate for
the three strengths.  Now we repeatedly run races with these PL
strengths and calculate the maximum likelihood estimate for each
realisation:

```{r repeatedrrace3, cache = TRUE}
x <- t(replicate(50,maxp(ordervec2supp3(rrace3(pn = c(a=16, b=13, c=18), ps = c(a=0.3,b=0.5,c=0.2))))))
```

```{r showx}
head(x)
```

```{r ternaryrrace3}
TernaryPlot(atip="a", btip="b", ctip="c")
TernaryPoints(x, pch=16, cex=0.5)
TernaryPoints(c(0.3,0.5,0.2),pch=16,col="red")
```

Above, we see the ensemble of estimates of the players' Plackett-Luce
strengths as black dots, and the true value of the strengths
(viz. `c(a=0.3,b=0.5,c=0.2)`) in red.  The estimates cluster around the true value.  It might be interesting to assess whether the estimates 
estimates are Dirichlet and, if so, what parameters the distribution
possesses.