### 2025-01-30 inst/podium.Rmd
more detail

---
title: "Does being on the podium affect Formula 1 drivers' performance in the next race?"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

<font size="1"> (almost all the time is spent on chunk `podlikevec7`,
which takes about one minute per evaluation) </font>

To cite the `hyper2` package in publications, please use
@hankin2017_rmd.  This short document defines the `podium()` function
which tests the hypothesis that a Formula 1 driver's being on the
podium at one venue affects his performance on the next.  Data from
the 2007 season are used.  It uses `hyper3` objects from the `hyper2`
package.  A detailed use-case for the top six drivers (Alonso,
Hamilton, Heidfeld, Kubica, Massa, and Raikkonen^[Umlauts pose special
problems for Rmd files]), and only the first five venues (Austria,
Malaysia, Bahrain, Spain, and Monaco) is given.

***TLDR***: Statistically significant evidence for the podium effect
was found, with about 2.7 units of support for the alternative.

Function `podium()`, defined below, takes a race table and a value of
$\lambda$ and returns a `hyper3` support function.

```{r,label=definepodiumfun}
podium <- function (lambda,x,print=FALSE){
     noscore <- c("Ret", "WD", "DNS", "DSQ", "DNP", "NC", "DNQ", "EX", "Sick")
     venues <- colnames(x)
     jj <- apply(x, 2, function(y) {
         if (any(y %in% noscore)) {
             y[y %in% noscore] <- 0
         }
         return(y)
     })
     fmat <- matrix(as.numeric(jj), byrow = TRUE, ncol = nrow(x))
     colnames(fmat) <- rownames(x)
     rownames(fmat) <- venues

     o <- fmat[1, , drop = TRUE]
     o[o > 0] <- rank(o[o > 0])
     out <- as.hyper3(ordervec2supp(o))
     for (i in seq(from=2,to=nrow(fmat))){
       if(print){cat(paste(i,"/",nrow(fmat),"\n"))}
       yesterday <- fmat[i-1, , drop = TRUE]
       yesterday[yesterday > 0] <- rank(yesterday[yesterday > 0])  # 'incomplete' functionality of ordertable2supp()
       podium_yesterday <- names(which((yesterday <= 3) & (yesterday>0)))
       d <- fmat[i, , drop = TRUE] # today
       d[d > 0] <- rank(d[d > 0])  # This is the 'incomplete' functionality of ordertable2supp()
       ## Following lines lifted from ordervec2supp()
       nd <- names(d)
       while (any(d > 0)) {
         eligible <- which(d >= 0)
         winner <- nd[d == 1]  # that is, the winner of those still racing
         if(winner %in% podium_yesterday){
           jj <- lambda
         } else {
           jj <- 1
         }
         names(jj) <- winner
         out[jj] %<>% inc    # numerator

         ## Now denominator
         jj <- rep(1,length(eligible))
         jj[nd[eligible] %in% podium_yesterday] <- lambda
         names(jj) <- names(eligible)
         out[jj] %<>% dec  # denominator

         d[d == 1] <- -1
         d[d > 0] %<>% dec
       } # ordervec2supp() lookalike closes
     }  # i loop closes
     return(out)
}
```

The first race has standard Plackett-Luce likelihood:

$$a\succ b\succ c\succ d\succ e\succ f$$

has a PL likelihood of

$$
\frac{a}{a+b+c+d+e+f}\cdot
\frac{b}{  b+c+d+e+f}\cdot
\frac{c}{    c+d+e+f}\cdot
\frac{d}{      d+e+f}\cdot
\frac{e}{        e+f}
$$

However, if the result were

$$\overline{a}\succ b\succ\overline{c}\succ\overline{d}\succ e\succ f$$

(where an overbar signifies that the driver was on the podium after
the previous race), the likelihood function would be

$$
\frac{\lambda a}{\lambda a+b+\lambda c+\lambda d+e+f}\cdot
\frac{        b}{          b+\lambda c+\lambda d+e+f}\cdot
\frac{\lambda c}{            \lambda c+\lambda d+e+f}\cdot
\frac{\lambda d}{                      \lambda d+e+f}\cdot
\frac{        e}{                                e+f}\cdot
$$

and the null of no podium effect would be $H_0\colon\lambda=0$.  Next,
we will take dataset `formula1_2007` and define a subset of it which
we will call `b`:

```{r, definesubset}
a <- read.table("formula1_2007.txt",header=TRUE)
a <- a[,seq_len(ncol(a)-1)]
b <- a[1:6,1:5]
b[,1] <- 1:6
b[,2] <- c(4,5,6,1,2,3)
b[,3] <- c(5,6,4,2,3,1)

b[,5] <- c(1,2,4,3,8,6)  # tests the 'incomplete' functionality of podium()
b
```

Taking Monaco (`MON`) as an example, the preceding race was in spain
(`ESP`) at which Massa, Hamilton, and Alonso finished on the podium.
The order statistic would be

$$
\text{Raikkonen}\succ
\overline{\text{Hamilton}}\succ
\overline{\text{Massa}}\succ
\overline{\text{Alonso}}\succ
\text{Kubica}\succ
\text{Heidfeld}
$$

where again an overbar signifies that the driver was on the podium in
the previous race.  The likelihood function for Monaco would thus be

$$
\begin{split}
\operatorname{\mathcal{L}}\left(\lambda;
p_\text{Raikkonen},\ldots,p_\text{Heidfeld}
\right)
& =
\frac{p_\text{Raikkonen}}{p_\text{Raikkonen}+\lambda p_\text{Hamilton}+\lambda p_\text{Massa}+\lambda p_\text{Alonso}+p_\text{Kubica}+p_\text{Heidfeld}}\\
&{\qquad} \cdot\frac{\lambda p_\text{Hamilton}}{\lambda  p_\text{Hamilton}+\lambda p_\text{Massa}+\lambda p_\text{Alonso}+p_\text{Kubica}+p_\text{Heidfeld}}\\
&{\qquad}\cdot\frac{\lambda p_\text{Massa}}{\lambda p_\text{Massa}+\lambda p_\text{Alonso}+p_\text{Kubica}+p_\text{Heidfeld}}\\
&{\qquad}\cdot\frac{\lambda p_\text{Alonso}}{\lambda p_\text{Alonso}+p_\text{Kubika}+p_\text{Heidfeld}}\cdot
\frac{p_\text{Kubica}}{p_\text{Kubica}+p_\text{Heidfeld}}
\end{split}
$$

```{r,usepodiu,cache=TRUE}
Ha <- podium(1.4, b)
Hb <- podium(2.1, b)
ma <- maxp(Ha, give=0, n=1)
mb <- maxp(Hb, give=0, n=1)
```

```{r}
Ha
summary(Hb)
ma
mb
```

Above we see that passing $\lambda$ values of 1.4 and 2.1 result in
different likelihoods.  The difference of about 0.4 units of support
is not significant, failing to reach the two-units-of-support
criterion of Edwards.

```{r,usefuns,cache=TRUE}
fun <- function(lambda){maxp(podium(lambda,b),give=TRUE)$value}
lam <- seq(from=0.5,to=2,len=6)
like <- lapply(lam,fun)
```

```{r, plotlik}
like <- unlist(like)
plot(lam,like-max(like),type='b')
like
```

Now a real dataset, Formula 1 2007 season:

```{r,load2007}

a <- read.table("formula1_2007.txt",header=TRUE)
a <- a[1:10,seq_len(ncol(a)-1)]  # rows Raikkonen (1) through Coulthard (10)
options(width=90)
a
```

```{r,definepodlike}
podlike <- function(lambda){maxp(podium(lambda=lambda,a),n=1,give=TRUE)$value}
```

```{r,usepodlike10,cache=TRUE}
podlike(1.8)
podlike(1.9)
```

```{r,podlikevec7,cache=TRUE}
date()
lam <- seq(from=0.2,to=1.2,len=15)
like <- lapply(lam,podlike)
date()
```

```{r,plotpodlike}
like <- unlist(like)
plot(lam,like-max(like),type="b")
abline(h=c(0,-2))
abline(v=1)
```

### References {-}