### 2025-02-10 inst/home_advantage.Rmd
merge home_advantage.Rmd and home_advantage3.Rmd

---
title: "Quantifying the home ground advantage with `hyper2` and `hyper3` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r,label=libload,echo=FALSE,print=FALSE,warning=FALSE,include=FALSE}
defaultW <- getOption("warn") 
options(warn = -1) 
library("hyper2",quietly=TRUE)
library("magrittr")
options(warn = defaultW)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

<font size="1"> (takes about fifteen minutes to run without cache)
</font>

To cite the `hyper2` package in publications, please use
@hankin2017_rmd.

Objects of class \code{hyper3} are a generalization
of \code{hyper2} objects that allow the brackets to contain weighted
probabilities.  Likelihood functions are defined on non-negative
$p_1,\ldots, p_n$ subject to the unit-sum constraint $\sum p_i=1$.
Given known weights $w^i_j$ with $1\leq i\leq j$ we have

$$\mathcal{L}\left(p_1,\ldots, p_n\right)=\prod_j\left(\sum_{i=1}^n w^i_jp_i\right)^{n_j}.$$

As a motivating example, suppose two teams (players) with
Bradley-Terry strengths $p_1,p_2$ play football where we quantify the
home-ground advantage with a term $\lambda$.  If $p_1$ plays at home
$a+b$ times with $a$ wins and $b$ losses, and plays away [so $p_2$ is
at home] $c+d$ times with $c$ wins and $d$ losses, then a sensible
likelihood function might be

$$\mathcal{L}(p_1,p_2,\lambda;A,B,C,D)=
\underbrace{
\overbrace{\left(\frac{\lambda p_1}{\lambda p_1 + p_2}\right)^{A}}^{\mbox{$A$ home wins for $p_1$}}\quad
\overbrace{\left(\frac{p_2        }{\lambda p_1 + p_2}\right)^{B}}^{\mbox{$B$ away wins for $p_2$}}
}_\mbox{$p_1$ plays at home to $p_2$ $A+B$ times}\quad
\underbrace{
\overbrace{\left(\frac{\lambda p_2}{p_1 + \lambda p_2}\right)^{C}}^{\mbox{$C$ home wins for $p_2$}}\quad
\overbrace{\left(\frac{        p_1}{p_1 + \lambda p_2}\right)^{D}}^{\mbox{$D$ away wins for $p_1$}}
}_\mbox{$p_2$ plays at home to $p_1$ $C+D$ times}
$$

where we understand that $p_1+p_2=1$, $p_1,p_2,\lambda\geqslant 0$.
Elementary techniques allow us to identify a maximum and we find

$$
\hat{p_1}=\frac{\sqrt{AD}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{p_2}=\frac{\sqrt{BC}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{\lambda}=\sqrt{\frac{AC}{BD}}
$$

Note that the evaluate is unchanged if $A,B,C,D$ are all multiplied by
a positive constant.

# Aside: `hyper2` analysis of the same scoreline

Now the likelihood function would be

$$\mathcal{L}(p_1,p_2,\lambda;A,B,C,D)=
\underbrace{
\overbrace{\left(\frac{p_1+M}{p_1 + p_2+M}\right)^{A}}^{\mbox{$A$ home wins for $p_1$}}\quad
\overbrace{\left(\frac{p_2  }{p_1 + p_2+M}\right)^{B}}^{\mbox{$B$ away wins for $p_2$}}
}_\mbox{$p_1$ plays at home to $p_2$ $A+B$ times}\quad
\underbrace{
\overbrace{\left(\frac{p_2+M}{p_1 + p_2+M}\right)^{C}}^{\mbox{$C$ home wins for $p_2$}}\quad
\overbrace{\left(\frac{p_1  }{p_1 + p_2+M}\right)^{D}}^{\mbox{$D$ away wins for $p_1$}}
}_\mbox{$p_2$ plays at home to $p_1$ $C+D$ times}
$$

The maximum would be

$$
\hat{p_1}=\frac{D}{C+D}\qquad
\hat{p_2}=\frac{B}{A+B}\qquad
\hat{M}=\frac{AC-BD}{(A+B)(C+D)}
$$

although it is not clear to me what the _constrained_ optimal point
[that is, requiring $M\geq 0$] is.

## Numerical analysis

```{r label=defineABCD}
A <- 9
B <- 5
C <- 7
D <- 3

(p1hat <- sqrt(A*D)/(sqrt(A*D) + sqrt(B*C)))
(p2hat <- sqrt(B*C)/(sqrt(A*D) + sqrt(B*C)))

(lam_hat <- sqrt((A*C)/(B*D)))
```

$$
\hat{p_1}=\frac{\sqrt{21}}{\sqrt{21} + \sqrt{35}}\simeq 0.47\qquad
\hat{p_2}=\frac{\sqrt{63}}{\sqrt{21} + \sqrt{35}}\simeq 0.53\qquad
\hat{\lambda}=\sqrt{\frac{AC}{BD}}=2.05.
$$

Moving to the monster formulation we have

$$
\hat{p_1}=\frac{3}{7+3} = 0.3\qquad
\hat{p_2}=\frac{5}{5+9}\simeq  0.36\qquad
\hat{M}=\frac{63-21}{14\cdot 10} = 0.3
$$

Keeping $A=9,B=5, C=7,D=3$, and $\lambda=2.05$ [assumed for the moment
to be known, estimation techniques are discussed later], we can
represent this information in a standard format:

```{r,label=usehomeaway3a}
library("hyper2",quietly=TRUE)
M <- matrix(c(NA, A+B*1i ,C+D*1i, NA),2,2,byrow=TRUE)
teams <- c("p1","p2")
dimnames(M) <- list("@home" = teams,"@away"=teams)
dimnames(M) <- list("@home" = teams,"@away"=teams)
print(M)
```

Above, object `M` has real parts being home wins and imaginary parts
being away wins.  Appropriate package idiom to specify a likelihood
function might be to use bespoke function `home_away3()`:

```{r,label=ha3calc,cache=TRUE}
(H <- home_away3(M,lambda=2.05))
```

Keeping $A=9,B=5, C=7,D=3$, and $\lambda=2.05$ [assumed for the moment
to be known, estimation techniques are discussed later], we may
estimate $p_1$ and $p_2$ using maximum likelihood, `maxp()` in package
idiom:

```{r,label=findtheevaluateh3,cache=TRUE}
H
maxp(H)
```

Further, we can test whether the players are in fact of equal strength:

```{r,label=testequalitynow,cache=TRUE}
equalp.test(H)
```

Showing convincingly that we may reject the null that $p_1=p_2$ and
assert that the players do in fact differ in strength, at least with
this value of $\lambda$.  Observe that a naive analysis would have
$p_1$ winning $w=A+C=`r A+C`$ games and losing $l=B+D=`r B+D`$ games,
from which we would obtain $\hat{p_1}=`r A+C`/`r A+B+C+D`$; a
likelihood ratio for $H_0\colon p_1=0.5$ would be

$$\frac
{
  {{w+l}\choose {w\, l}}\left(\frac{w}{w+l}\right)^w\left(\frac{l}{w+l}\right)^l
}{
  {{w+l}\choose {w\, l}}\left(\frac{1}{2}\right)^w\left(\frac{1}{2}\right)^l
}=\frac{2^{w+l}w^wl^l}{(w+l)^{w+l}}\simeq 1.94,
$$

not significant.

Now, how to estimate $\lambda$?

```{r,label=estimatelambda3,cache=TRUE}
f <- function(lambda){maxp(home_away3(M,lambda=lambda),give=TRUE)$value}
f(2)
f(2.1)
lam <- seq(from=1,to=10,len=17)
Supp <- sapply(lam,f)
```

```{r,label=plotsup}
plot(log(lam),Supp-max(Supp),type="b")
```

Or even

```{r,label=optimlambda,cache=TRUE}
op <- optimize(f,interval=c(2,3),maximum=TRUE)
op
```

We can proceed in two ways.  Firstly we can use $\hat{\lambda}$ directly:

```{r,ept3,cache=TRUE}
equalp.test(home_away3(M,lambda=op$maximum))
```

The flaw in this analysis is that it is conditional on the estimated
value of $\lambda$ [which is not too bad, IMO].  Secondly we can
perform a fully 2D analysis, which allows for the fact that the
evaluate might have a different value of $\lambda$ from the
$\hat{\lambda}$:

```{r,label=loopysuppcalc,cache=TRUE}
probs <- seq(from=0.1,to=0.8,len=17)
jj <- as.matrix(expand.grid(lam,probs))
S <- rep(NA,nrow(jj))
for(i in seq_len(nrow(jj))){
    lambda <- jj[i,1]
    p1 <- jj[i,2]
    p <- c(p1,1-p1)
    names(p) <- c("p1","p2")
   S[i] <- loglik(p=p,H=home_away3(M,lambda=lambda),log=TRUE)
}
S <- matrix(S,length(lam),length(probs))
```

```{r}
contour(lam,probs,S,levels=seq(from=-20,to=-10,by=0.5),xlab="lambda",ylab="p1")
abline(h=0.5)
```

#  Ternary representation

```{r,label=useternary}
library(Ternary)

par(mar = rep(0.2, 4))
TernaryPlot(alab = "p1", blab = "p2", clab = "M")

f <- function(p1,p2,M) {
pmax(-20,log(M+p1)*A + log(p2)*B - log(M+p1+p2)*(A+B+C+D) + log(p2+M)*C + log(p1)*D )
}
FunctionToContour <- function(a,b,c){f(p1=a, p2=b, M=c)}

# Compute and plot colour tiles
values <- TernaryPointValues(FunctionToContour, resolution = 24L)
ColourTernary(values)

# Add contour lines
TernaryContour(FunctionToContour, resolution = 36L)
```

A slightly more complicated example:

```{r,label=genthedata}
home_games_won <- matrix(c(
    NA, 16, 12, 11,
    19, NA, 19, 16,
    17, 12, NA, 11,
    11, 12, 12, NA),
    nrow=4,ncol=4,byrow=TRUE)

away_games_won <- matrix(c(
    NA, 05, 02, 02,
     9, NA, 10, 02,
     3, 04, NA, 07,
     8, 06, 04, NA),
    nrow=4,ncol=4,byrow=TRUE)

teams <- LETTERS[1:4]
dimnames(home_games_won) <- list("@home" = teams,"@away"=teams)
dimnames(away_games_won) <- list("@home" = teams,"@away"=teams)

home_games_won
away_games_won
```

Thus `home_games_won[1,2] == 16` means A played at home against B and
won 16 times; `home_games_won[2,1] == 19` means B played at home
against A and won 19 times.  Also, `away_games_won[1,2] == 5` means A
played away against B and won 5 times, and `away_games_won[2,1] == 2`
means B played away against A and won 2 times.  Alternatively, A
played B 16+19+5+2=45 times; A played at home 16+2=18 times and won 16
times and lost 2 times; and B played at home 19+5=24 times and won 19
times and lost 5 times.  Alternatively we may use complex numbers to
represent the same dataset:

```{r,label=showhomeawayusingcomplex}
home_games_won + 1i*away_games_won
```

We will create a hyper2 object with the teams and home ground advantage term:

```{r,label=makehyper2,cache=TRUE}
H <- home_away(home_games_won,away_games_won)
H
options("use_alabama" = FALSE) # to avoid the stupid wmmin bug
specificp.gt.test(H,"home",0)
options("use_alabama" = TRUE) # to avoid the stupid wmmin bug
```

We see strong evidence to support the contention that home advantage
is real.  Further, we may test the hypothesis that all the teams have
the same strength, after accounting for the home team advantage:

```{r,label=homegroundAgresti,cache=TRUE}
samep.test(H,teams)
```

Above, we see no evidence for a difference in team strengths.  Visually:

```{r,label=showvisH,cache=TRUE}
(mt <- maxp(H))
```

A profile likelihood diagram:

```{r,label=showprofilelikelihood, cache=TRUE}
home_strength <- seq(from=0.15,to=0.45,length=13)
plot(home_strength,profsupp(H,"home",home_strength),type="b",pch=16,
     xlab="home strength",ylab="support",main="profile likelihood for home advantage")
abline(h=c(0,-2))
```

The figure give a support interval between about 0.3 and 0.5 for home
ground advantage.

# Example in Agresti

Agresti considers a dataset of seven baseball teams who play one
another repeatedly.  Each game is played at the home ground of one
team, the other playing away.  The dataset is as follows:

```{r,label=makeAgrestimatrix}
baseball_table <- matrix(c(
NA,   4+3i, 4+2i, 4+3i, 6+1i, 4+2i, 6+0i,
3+3i, NA  , 4+2i, 4+3i, 6+0i, 6+1i, 4+3i,
2+5i, 4+3i, NA  , 2+4i, 4+3i, 4+2i, 6+0i,
3+3i, 5+1i, 2+5i, NA  , 4+3i, 4+2i, 6+1i,
5+1i, 2+5i, 3+3i, 4+2i, NA  , 5+2i, 6+0i,
2+5i, 3+3i, 3+4i, 4+3i, 4+2i, NA  , 2+4i,
2+5i, 1+5i, 1+6i, 2+4i, 1+6i, 3+4i, NA),7,7)
teams <- c("Milwaukee", "Detroit", "Toronto", "New York", "Boston","Cleveland","Baltimore")
rownames(baseball_table) <- teams
colnames(baseball_table) <- teams
M <- baseball_table # saves typing
```

Above, we represent home team wins with the real part of the matrix,
the imaginary component being away wins.  Now process it:

```{r,label=makesuppfrommatrix}
baseball <- home_away(M)
baseball
```

Maximum likelihood:

```{r,label=calculatemaxpAgresti,cache=TRUE}
baseball_maxp <- maxp(baseball)
baseball_maxp
```

Visually:

```{r,label=showmaxpAgresti}
dotchart(baseball_maxp,pch=16)
pie(baseball_maxp)

```

Above, we see that the teams are all of roughly equal strength, and
the home advantage is small.  However, we may easily assess the null
that the home strength is zero:

```{r,analyseAgresti,cache=TRUE}
specificp.gt.test(baseball,"home",0)
```

```{r,Agrestiprofsupp,cache=TRUE}
home_strength <- seq(from=0.005,to=0.1,length=20)
plot(home_strength,profsupp(baseball,"home",home_strength),type="b",pch=16,
     xlab="home strength",ylab="support",main="profile likelihood for home advantage")
abline(h=c(0,-2))
```

Further, we may test the hypothesis that the baseball teams have the same strength:

```{r,label=testhomegroundadvantage,cache=TRUE}
samep.test(baseball,teams)
```

Above, we see no evidence for the teams having differential strengths.

As an interesting aside, suppose we have some additional data from
games in which (for some reason), the home advantage was not
operating.

```{r,label=newinfotobaseball}
M2 <- matrix(c(0,1,2,0),2,2)
dimnames(M2) <- list(winner=teams[1:2],loser=teams[1:2])
M2
```

Thus we have additional data from three games in which Milwaukee won
twice and lost once.  This additional information may easily be
incorporated into our likelihood function, at least on the assumption
of independence:

```{r,label=addsomenewinfotobaseball}
baseball <- baseball + pairwise(M2)
```

We may assess the difference that the new information makes:

```{r,label=baseballtransplot,cache=TRUE}
ordertransplot(baseball_maxp[1:7],maxp(baseball)[1:7],xlab="old",ylab="new",plotlims=c(0.01,0.16))
```

Above, we see very little difference in the two evaluates, except for
Detroit and Milwaukee, as expected.

### Some further thoughts on the home ground monster and Agresti's analysis.

Considering the simplest case of two teams, a and b, with results:

* $a$ at home, $b$ away, $a,b$ wins with probability $\frac{\lambda a,b}{\lambda a+b}$ (Agresti), or $\frac{a+H,b}{a+b+H}$ (me). 
* $a$ away, $b$ at home, $a,b$ wins with probability $\frac{a,\lambda b}{\lambda a+b}$ (Agresti), or $\frac{a,b+H}{a+b+H}$ (me).

Matching the probabilities for $a$ at home gives us $H=a(\lambda-1)$
and matching the probabilities for $a$ away gives $H=b(\lambda-1)$.
These cannot both be true unless $a=b$ or $\lambda=1$.  But we can, in
principle, test which one is a better model.

# hyper3 idiom

We may apply `hyper3` idiom to the dataset, using the likelihood
function of Davidson and Beaver (1977).  First a formal test:

```{r,label=optimizebaseball,cache=TRUE}
baseball_table
f <- function(l){ maxp(home_away3(baseball_table,l),give=TRUE,n=1)$likes}
o <- optimize(f,c(0.8,1.6),maximum=TRUE)
o
lambda_baseball <- o$maximum
maxL <- f(lambda_baseball)
W <- maxL - f(1)    # f(1) being the likelihood of the null
W
pchisq(2*W,df=1,lower.tail=FALSE)
```

The pvalue is significant.  Now plot a profile likelihood (support)
curve:

```{r, label=hyper3_baseball,cache=TRUE}
jj1  <- seq(from=log(lambda_baseball),to=log(1.9),len=7)
jj2  <- seq(from=log(lambda_baseball),to=log(0.8),len=7)
l <- sort(unique(exp(c(jj1,jj2))))
lam  <- sort(c(l,lambda_baseball))
L <- sapply(lam,f) - maxL
```

```{r,label=plothyper3baseball}
plot(log(lam),L,type="b")
abline(h=c(0,-2))
abline(v=0)

plot(lam,L,type="b")
abline(h=c(0,-2))
```

# Creating a home/away results table from scratch

Function `home_away_table()` takes a data frame of results and
produces a table amenable to analysis by `home_away()` or
`home_away3()`.

```{r showdata}
a <- read.table("celtic_rangers_livingston_falkirk.txt", header=TRUE)
head(a)
teams <- c("Rangers","Celtic","Falkirk","Livingston")
(aF <- home_away_table(a, give=FALSE,teams = teams))
(aT <- home_away_table(a, give=TRUE, teams = teams))
```

(from here on we ignore draws and consider only won games)

```{r doscottish,cache=TRUE}
(H <- home_away(aF))
pnames(H) <- c(teams,"home")
H
(mH <- maxp(H))
```

```{r testrangersandceltic, cache=TRUE}
samep.test(H,c("Rangers","Celtic"))

```{r  testhomezero, cache=TRUE}
specificp.gt.test(H,"home",0)
```

```{r dohome3, cache=TRUE}
home_away3(aF,lambda=1.88)
maxp(home_away3(aF,lambda=1.88),give=1)
maxp(home_away3(aF,lambda=1.11),give=1)
```

```{r calcsuppscotty,cache=TRUE}
f <- function(l){ maxp(home_away3(aF,l),give=TRUE,n=1)$likes}
lam <- seq(from=1.4,to=2.3,len=10)
date()
loglike <- sapply(lam,f)
date()
```

```{r optscotty,cache=TRUE}
(o <- optimize(f,c(1.1,4.6),maximum=TRUE))
```

```{r scottychisq,cache=TRUE}
o
scot_max_lambda <- o$maximum
maxL <- f(scot_max_lambda)

W <- maxL - f(1)    # f(1) being the likelihood of the null
W
pchisq(2*W,df=1,lower.tail=FALSE)
```

```{r scotmax}
(H <- home_away3(home_away_table(a,give=FALSE),scot_max_lambda))  # o$maximum ~= 1.353
maxp(H,give=1)
```

```{r}
plot(lam,loglike-max(loglike),type='b')
abline(h=c(0,-2))
abline(v=1)
grid()
```

```{r rememberhyper2}
(mH2 <- round(mH*1000))
```

```{r sfasdf, cache=TRUE}
o$maximum
mH3 <- maxp(home_away3(home_away_table(a,give=FALSE),scot_max_lambda),give=FALSE)
round(mH3*1000)
```

# Rangers vs Falkirk {-}

**(Rangers at home)**

* hyper2: P(Rangers;Falkirk) = $\frac{448+67; 56}{448 + 56 + 67}\simeq (0.902; 0.098)$
* hyper3: P(Rangers;Falkirk) = $\frac{1.77*448; 56}{1.77*448+56}\simeq(0.934; 0.066)$

actual scoreline: 69(R) / 2(F) or about $(0.972; 0.028)$. Expectation
would be (64;7) (H2) or (66.3;4.7) (H3)

*(Rangers away)*

* hyper2: P(Rangers;Falkirk) = $\frac{448; 56+67}{448 + 56 + 67}\simeq (0.78;0.22)$
* hyper3: P(Rangers;Falkirk) = $\frac{448; 1.77*56}{448 + 1.77*56}\simeq(0.82; 0.18)$

actual scoreline: 57(R) /17(F) - or about $(0.77; 0.23)$

# Rangers vs Livingston {-}

**(Rangers at home)**

* hyper2: P(Rangers;Livingston) = $\frac{448+67; 39}{448 + 39 + 67}\simeq (0.930; 0.070)$
* hyper3: P(Rangers;Livingston) = $\frac{1.77*448; 39}{1.77*448+39}\simeq(0.953; 0.047)$

actual scoreline:  15(R) / 0(L) or  $(1; 0)$

*(Rangers away)*

* hyper2: P(Rangers;Livingston) = $\frac{448; 39+67}{448 + 39 + 67}\simeq (0.81;0.19)$
* hyper3: P(Rangers;Livingston) = $\frac{448; 1.77*39}{448 + 1.77*39}\simeq(0.86; 0.14)$

actual scoreline: 12(R) /2(L) - or about $(0.86; 0.14)$

# Celtic vs Falkirk {-}

*(Celtic at home)*

* hyper2: P(Celtic;Falkirk) = $\frac{390+67; 56}{390 + 56 + 67}\simeq (0.89; 0.11)$
* hyper3: P(Celtic;Falkirk) = $\frac{1.77*412; 74}{1.77*412+74}\simeq(0.92; 0.08)$

actual scoreline:  61-8 or about $(0.88; 0.12)$

*(Celtic away)*

* hyper2: P(Celtic;Falkirk) = $\frac{390; 56+67}{390 + 56 + 67}\simeq (0.76;0.24)$
* hyper3: P(Celtic;Falkirk) = $\frac{412; 1.77*74}{412 + 1.77*74}\simeq(0.76; 0.24)$

actual scoreline: 43-17 or about $(0.72; 0.28)$

# Livingston vs Falkirk {-}

*(Livingston at home)*

* hyper2: P(Livingston;Falkirk) = $\frac{39+67; 56}{39 + 67 + 56}\simeq (0.65; 0.35)$
* hyper3: P(Livingston;Falkirk) = $\frac{1.77*42; 74}{1.77*42+74}\simeq(0.50; 0.50)$

actual scoreline: 3-9 or about $(0.25; 0.75)$

*(Livingston away)*

* hyper2: P(Livingston;Falkirk) = $\frac{39; 56+67}{39 + 56 + 67}\simeq (0.24;0.76)$
* hyper3: P(Livingston;Falkirk) = $\frac{42; 1.77*74}{42 + 1.77*74}\simeq(0.24; 0.76)$

actual scoreline: 6-8 or about $(0.43; 0.57)$

# Expectations

First, hyper2:

```{r}
mH
mH["Rangers"]
mH["home"]
mH3
lambda <- scot_max_lambda
h2_exp <- matrix(0,4,4)
rownames(h2_exp) <- teams
colnames(h2_exp) <- teams
h3_exp <- h2_exp
n_matches <- Re(aF) + Im(aF)
for(hometeam in teams){
    for(awayteam in teams){

        h2_exp[hometeam,awayteam] <- (
            n_matches[hometeam,awayteam] * (mH[hometeam] + mH["home"])/
            (mH[hometeam] + mH[awayteam] + mH["home"])
        )
        h2_exp[hometeam,awayteam] %<>%
            add(1i*
                n_matches[hometeam,awayteam] * (mH[awayteam])/
                (mH[hometeam] + mH[awayteam] + mH["home"])
                )

        h3_exp[hometeam,awayteam] <- (
            n_matches[hometeam,awayteam] * (lambda* mH3[hometeam])/
            (lambda* mH3[hometeam] + mH3[awayteam])
        )

        h3_exp[hometeam,awayteam] %<>%
            add(1i*
                n_matches[hometeam,awayteam] * (mH3[awayteam])/
                (lambda* mH3[hometeam] + mH3[awayteam])
        )
    }
}
aF
round(h2_exp,1)
round(h3_exp,1)

makev <- function(M){
    out <- c(Re(M),Im(M))
    out[!is.na(out)]
}

(obs <- makev(aF))
(exp2 <- makev(h2_exp))
(exp3 <- makev(h3_exp))

plot((obs-exp2)^2/exp2)
plot((obs-exp3)^2/exp3)

chisq2 <- sum((obs-exp2)^2/exp2)
chisq3 <- sum((obs-exp3)^2/exp3)

pchisq(chisq2,df=12-4,lower.tail=FALSE)
pchisq(chisq3,df=12-4,lower.tail=FALSE)

par(pty='s')
plot(obs,exp2,log="xy",asp=1,xlim=c(1,100),ylim=c(1,100))
abline(0,1)
plot(obs,exp3,log="xy",asp=1,xlim=c(1,100),ylim=c(1,100))
abline(0,1)
```

```{r defineexpectation2}
expectation2 <- function(probs, results){
    teams <- names(probs[names(probs) != "home"])
    h2_exp <- matrix(0,4,4)
    rownames(h2_exp) <- teams
    colnames(h2_exp) <- teams
    n_matches <- Re(aF) + Im(aF)
    for(hometeam in teams){
        for(awayteam in teams){

            h2_exp[hometeam,awayteam] <- (
                n_matches[hometeam,awayteam] * (mH[hometeam] + mH["home"])/
                (mH[hometeam] + mH[awayteam] + mH["home"])
            )

            h2_exp[hometeam,awayteam] %<>%
                add(1i*
                    n_matches[hometeam,awayteam] * (mH[awayteam])/
                    (mH[hometeam] + mH[awayteam] + mH["home"])
                    )
        }
    }
    return(h2_exp)
}
round(expectation2(mH,aF),1)
```

```{r defineexp3}

expectation3 <- function(probs, results, lambda){
    teams <- names(probs)
    h3_exp <- matrix(0,4,4)
    rownames(h3_exp) <- teams
    colnames(h3_exp) <- teams
    n_matches <- Re(aF) + Im(aF)
    for(hometeam in teams){
        for(awayteam in teams){

            h3_exp[hometeam,awayteam] <- (
                n_matches[hometeam,awayteam] * (lambda* mH3[hometeam])/
                (lambda* mH3[hometeam] + mH3[awayteam])
            )

            h3_exp[hometeam,awayteam] %<>%
                add(1i*
                    n_matches[hometeam,awayteam] * (mH3[awayteam])/
                    (lambda* mH3[hometeam] + mH3[awayteam])
                    )
        }
    }
    return(h3_exp)
}

round(expectation3(mH3[c("Rangers","Celtic","Falkirk","Livingston")],aF,scot_max_lambda),1)
```

# Appendix

Here is a transcript of a mathematica session which establishes the
estimates above.  First the `hyper2` likelihood:

```
rhankin@rhrpi4:~ $ wolfram
Mathematica 12.2.0 Kernel for Linux ARM (32-bit)
Copyright 1988-2021 Wolfram Research, Inc.

In[1]:= L = Log[M+p1]*A +Log[p2]*B- Log[M+p1+p2]*(A+B+C+D) + Log[p2+M]*C + Log[p1]*D

Out[1]= D Log[p1] + A Log[M + p1] + B Log[p2] + C Log[M + p2] - (A + B + C + D) Log[M + p1 + p2]

In[2]:= LL = L /. {p2 -> 1-p1-M}

Out[2]= C Log[1 - p1] + B Log[1 - M - p1] + D Log[p1] + A Log[M + p1]

In[3]:= dp = D[LL,p1]//FullSimplify

            C      D         B          A
Out[3]=  ------- + -- + ----------- + ------
         -1 + p1   p1   -1 + M + p1   M + p1

In[4]:= dM = D[LL,M]//FullSimplify

              B          A
Out[4]= ----------- + ------
         -1 + M + p1   M + p1

In[5]:= Solve[{dp==0,dM==0},{p1,M}]

                   D            A C - B D
Out[5]=  {{p1 -> -----, M -> ---------------}}
                 C + D       (A + B) (C + D)

In[6]:= p2 = 1-p1-M /. {%} //FullSimplify

             B
Out[6]=  {{-----}}
           A + B

```

Now the `hyper3` likelihood:

```
rhankin@rhrpi4:~ $ wolfram
Mathematica 12.2.0 Kernel for Linux ARM (32-bit)
Copyright 1988-2021 Wolfram Research, Inc.

In[1]:= L = A*Log[lambda*p1] + B*Log[p2] -(A+B)*Log[lambda*p1 + p2] + C*Log[lambda*p2] + D*Log[p1] -(C+D)*Log[p1 + lambda*p2]

Out[1]= D Log[p1] + A Log[lambda p1] + B Log[p2] + C Log[lambda p2] - (A + B) Log[lambda p1 + p2] - (C + D) Log[p1 + lambda p2]

In[2]:= LL = L /. {p2 -> 1-p1}

Out[2]= B Log[1 - p1] + C Log[lambda (1 - p1)] + D Log[p1] + A Log[lambda p1] - (C + D) Log[lambda (1 - p1) + p1] -

>    (A + B) Log[1 - p1 + lambda p1]

In[3]:= dp1 = D[LL,p1] //FullSimplify

         B + C    A + D   (A + B) (-1 + lambda)    (C + D) (-1 + lambda)
Out[3]= ------- + ----- - --------------------- + -----------------------
        -1 + p1    p1     1 + (-1 + lambda) p1    lambda + p1 - lambda p1

In[4]:= dlam = D[LL,lambda]//FullSimplify

          A        C           (A + B) p1           (C + D) (-1 + p1)
Out[4]= ------ + ------ - -------------------- + -----------------------
        lambda   lambda   1 + (-1 + lambda) p1   lambda + p1 - lambda p1

In[5]:= Solve[{dp1==0,dlam==0},{p1,lambda}]//FullSimplify

[snip]

In[6]:= %[[1]]

                        1                     Sqrt[A] Sqrt[C]
Out[6]= {p1 -> -------------------, lambda -> ---------------}
                   Sqrt[B] Sqrt[C]            Sqrt[B] Sqrt[D]
               1 + ---------------
                   Sqrt[A] Sqrt[D]

```

###  References

* R. R. Davidson and R. J. Beaver 1977. "On extending the
  Bradley-Terry model to incorporate within-pair order effects"
  _Biometrics_, 33:693--702

### Package dataset {-}

Following lines create `baseball.rda`, residing in the `data/`
directory of the package.

```{r,label=savesolingdataset}
save(baseball_table, baseball, baseball_maxp, 
  file="baseball.rda")
```