### 2025-02-13 inst/sushi.Rmd
better range for lambda

---
title: "Sushi monster"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.

```{r}
a <- as.matrix(read.table("sushi.txt"))
```

```{r,defcalculate_sushi_H}
make_sushi_H <- function(a, strength){
    h <- c(1, strength) # strength of support
    sushi_types <- c("ebi", "anago", "maguro", "ika", "uni", "sake", "tamago", "toro", "tekka", "kappa")
    ec <- # equivalence classes
        c(`ebi`  = 1, `anago` = 2, `maguro` = 1, `ika` = 1, `uni` = 1,
          `sake` = 1, `tamago` = 1, `toro` = 2, `tekka-maki` = 2,
          `kappa-maki` = 1)

    H <- hyper3()

    for(i in seq_len(nrow(a))){
        H <- H + cheering3(sushi_types[a[i,]], e=ec, h=h)
    }
    return(H)
}
```

```{r tryH, cache=TRUE}
wanted <- sample(seq_len(nrow(a)), 500, replace=TRUE, prob=a[,1])
H1888 <- make_sushi_H(a[wanted,-1], strength=1.888)
H1000 <- make_sushi_H(a[wanted,-1], strength=1.000)
```

```{r maxpsushiH,cache=TRUE}
mHa <- maxp(H1888,give=1)
mHb <- maxp(H1000,give=1)
```

```{r showsushiH}
mHa
mHb
```

```{r calclikesushi, cache=TRUE}
f <- function(strength){
   H <- make_sushi_H(a[wanted,-1], strength=strength)
   maxp(H,give=1)$`Log-likelihood`
}
s <- seq(from=1.4,to=7.5,len=5)
L <- sapply(s,f)
```

```{r}
plot(s,L-max(L),type='b')
```

## References {-}