### 2025-02-16 inst/cheering.Rmd
cheering() in use

---
title: "Function `cheering()` in the `hyper2` package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("partitions")
library("pdftools")
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.

This document is a distillation and clarification of
`inst/plackett_luce_monster.Rmd`.

In the following, I use the ''race'' metaphor: five runners take part
in a race and arrive in order: $a\succ b\succ c\succ d\succ 5e.
Consider the following Plackett-Luce likelihood function for this
observation:

\begin{equation}
\frac{p_a}{p_a + p_b + p_c + p_d + p_e}\cdot
\frac{p_b}{      p_b + p_c + p_d + p_e}\cdot
\frac{p_c}{            p_c + p_d + p_e}\cdot
\frac{p_d}{                  p_d + p_e}\cdot
\frac{p_e}{                        p_e}
\end{equation}

Note the use of multiplication between the terms, indicating
conditional independence.  We may use the multiplicative
generalization of Bradley-Terry strengths to introduce an element of
non-independence between the terms.  There are many ways to do this,
but one simple case would be to define equivalence classes of the
competitors with each equivalence class comprising mutually supporting
runners.  Equivalence classes of size one correspond to unsupported
runners.  The metaphor would be that a runner who has finished the
race is able to support other members of his equivalence class by
cheering his teammates, boosting their better performance.  The
`hyper2` package includes function `cheering()` which implements this
functionality.

As an example, consider figure 0.1 This shows a partial probability
tree diagram for some of the $5!=120$ possible order statistics.  The
standard Plackett-Luce likelihoods have been modified to account for
two groups of mutually supporting runners: $\left\lbrace
a,b,c\right\rbrace$ with support term $\lambda$, and $\left\lbrace
d,e\right\rbrace$ with support term $\mu$.  From START, the first
runner to cross the finishing line is associated with standard
Plackett-Luce.  Taking the top path as an example, we see that the
likelihood function for $a\succ b\succ c\succ d\succ e$ would be

```{tikz, label=tikz3, fig.cap = "Partial probability\\label{tikzabcde} tree for five competitors $a$-$e$ with mutually supporting subsets $\\left\\lbrace a,b,c\\right\\rbrace$ [with support term $\\lambda$] and $(de)$ [with support term $\\mu$], hyper3 approach", fig.ext = 'png', echo=FALSE}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2,3);
\draw (0,0) -- (2,2);
\draw (0,0) -- (2,1);
\draw (0,0) -- (2,0);
\draw (0,0) -- (2,-1);
\node at (0,2.5) (eq1) {$\underbrace{\left\lbrace a,b,c\right\rbrace}_{\lambda}\underbrace{\left\lbrace d,e\right\rbrace}_{\mu}$};

\node at (1.5,2.5) {$a$};
\node at (1.5,1.7) {$b$};
\node at (1.5,1.0) {$c$};
\node at (1.5,0.2) {$d$};
\node at (1.5,-0.5) {$e$};

\node at (3, 3)  {$\frac{a}{a+b+c+d+e}$};
\node at (3, 2)  {$\frac{b}{a+b+c+d+e}$};
\node at (3, 1)  {$\frac{c}{a+b+c+d+e}$};
\node at (3, 0)  {$\frac{d}{a+b+c+d+e}$};
\node at (3,-1)  {$\frac{e}{a+b+c+d+e}$};

\fill  (4, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill  (4, 2) circle[radius=2pt];  % terminal node
\fill  (4, 1) circle[radius=2pt];  % terminal node
\fill  (4, 0) circle[radius=2pt];  % d finishes; paths abce
\fill  (4,-1) circle[radius=2pt];  % terminal node

\draw (4,3) -- (5,5);
\draw (4,3) -- (5,4);
\draw (4,3) -- (5,3);
\draw (4,3) -- (5,2);

\node at (4.5,4.6) {$b$};
\node at (4.5,3.7) {$c$};
\node at (4.5,3.2) {$d$};
\node at (4.5,2.7) {$e$};

\newcommand{\la}[1]{\lambda{#1}}
\newcommand{\ld}[1]{\mu    {#1}}
\node at (6, 5)  {$\frac{\la{b}}{\la{b}+\la{c}+d+e}$};
\node at (6, 4)  {$\frac{\la{c}}{\la{b}+\la{c}+d+e}$};
\node at (6, 3)  {$\frac{d     }{\la{b}+\la{c}+d+e}$};
\node at (6, 2)  {$\frac{e     }{\la{b}+\la{c}+d+e}$};

\fill (7.4, 5) circle[radius=2pt];  % ab finishes; pahts cde
\draw (7.4,5) -- (8.4,6);
\draw (7.4,5) -- (8.4,5);
\draw (7.4,5) -- (8.4,4);

\node at (7.9,5.7) {$c$};
\node at (7.9,5.2) {$d$};
\node at (7.9,4.7) {$e$};
\node at (9.3, 6) {$\frac{\la{c}}{\la{c}+d+e}$};
\node at (9.3, 5) {$\frac{d  }{\la{c}+d+e}$};
\node at (9.3, 4) {$\frac{e  }{\la{c}+d+e}$};

\fill (10.2, 6) circle[radius=2pt];  % abc finishes; paths de
\draw (10.2, 6) -- (11.2,7);
\node at (10.7,6.8) {$d$};

\node at (11.8, 7) {$\frac{d}{d+e}$};

\draw (10.2, 6) -- (11.2,6.2);
\node at (11.8, 6.2) {$\frac{e}{d+e}$};
\node at (10.7,6.3) {$e$};

\fill (10.2, 5) circle[radius=2pt];
\draw (10.2, 5) -- (11.2,5.3);
\node at (11.8, 5.3) {$\frac{\la{c}}{\la{c}+\ld{e}}$};
\node at (10.7,5.3) {$c$};

\draw (10.2, 5) -- (11.2,4.7);
\node at (11.8, 4.7) {$\frac{\ld{e}}{\la{c}+\ld{e}}$};
\node at (10.7,4.7) {$e$};

\fill (10.2, 4) circle[radius=2pt];
\draw (10.2, 4) -- (11.2,3.9);
\node at (11.8, 3.9) {$\frac{\la{c}}{\la{c}+\ld{e}}$};
\node at (10.7,4.1) {$c$};

\draw (10.2, 4) -- (11.2,3.1);
\node at (11.8, 3.1) {$\frac{\ld{e}}{\la{c}+\ld{e}}$};
\node at (10.7,3.7) {$e$};

\draw (4, 0) -- (5,1);
\node at (6, 1)  {$\frac{a}{a+b+c+\ld{e}}$};

\draw (4, 0) -- (5,0);
\node at (6, 0)  {$\frac{b}{a+b+c+\ld{e}}$};

\draw (4, 0) -- (5,-1);
\node at (6, -1)  {$\frac{c}{a+b+c+\ld{e}}$};

\draw (4, 0) -- (5,-2);
\node at (6, -2)  {$\frac{\ld{e}}{a+b+c+\ld{e}}$};

\node at (4.5,0.7) {$a$};
\node at (4.5,0.2) {$b$};
\node at (4.5,-0.3) {$c$};
\node at (4.5,-0.8) {$e$};

\fill (7.4, -2) circle[radius=2pt];  % de finishes; paths abc
\draw (7.4, -2) -- (8.4,-1);
\draw (7.4, -2) -- (8.4,-2);
\draw (7.4, -2) -- (8.4,-3);

\node at (7.9,-1.2) {$a$};
\node at (7.9,-1.8) {$b$};
\node at (7.9,-2.3) {$c$};

\node at (9.3, -1) {$\frac{a}{a+b+c}$};
\node at (9.3, -2) {$\frac{b}{a+b+c}$};
\node at (9.3, -3) {$\frac{c}{a+b+c}$};

\fill (10.2, -1) circle[radius=2pt];  % dea finishes; paths bc
\fill (10.2, -2) circle[radius=2pt];  % deb finishes; paths ac
\fill (10.2, -3) circle[radius=2pt];  % dec finishes; paths ab

\draw (10.2, -1) -- (11.2,-0);
\node at (10.6,-0.4) {$b$};
\node at (10.6,-0.8) {$c$};

\node at (11.8, 0) {$\frac{\la{b}}{\la{b}+\la{c}}$};
\draw (10.2, -1) -- (11.2,-1);

\node at (11.8, -1) {$\frac{\la{c}}{\la{b}+\la{c}}$};

\draw (10.2, -2) -- (11.2,-1.6);
\node at (11.8, -1.6) {$\frac{\la{a}}{\la{a}+\la{c}}$};

\node at (10.6,-1.7) {$a$};
\node at (10.6,-2) {$c$};

\draw (10.2, -2) -- (11.2,-2.2);
\node at (11.8, -2.2) {$\frac{\la{c}}{\la{a}+\la{c}}$};

\draw (10.2, -3) -- (11.2,-2.9);
\node at (11.8, -2.9) {$\frac{\la{a}}{\la{a}+\la{b}}$};

\draw (10.2, -3) -- (11.2,-3.4);
\node at (11.8, -3.4) {$\frac{\la{b}}{\la{a}+\la{b}}$};

\node at (10.6,-2.8) {$a$};
\node at (10.6,-3.3) {$b$};

\fill (7.4, 1) circle[radius=2pt];  % da finishes; paths bce

\draw (7.4, 1) -- (8.4,2);
\draw (7.4, 1) -- (8.4,1);
\draw (7.4, 1) -- (8.4,0);

\node at (7.9,1.7) {$b$};
\node at (7.9,1.2) {$c$};
\node at (7.9,0.7) {$e$};

\node at (9.3, 2) {$\frac{\la{b}}{\la{b}+\la{c}+e}$};
\node at (9.3, 1) {$\frac{\la{c}}{\la{b}+\la{c}+e}$};
\node at (9.3, 0) {$\frac{e     }{\la{b}+\la{c}+e}$};

\fill (10.2, 1) circle[radius=2pt];  % da finishes; paths bce
\draw (10.2, 1) -- (11.2,2);
\node at (11.8, 2) {$\frac{\la{b}}{\la{b}+\ld{e}}$};
\draw (10.2, 1) -- (11.2,1);
\node at (11.8, 1) {$\frac{\ld{e}}{\la{b}+e}$};

\end{tikzpicture}
```

\begin{equation}\label{asuccb}
\frac{p_a}{p_a + p_b + p_c + p_d + p_e}\cdot
\frac{\lambda p_b}{\lambda p_b + \lambda p_c + p_d + p_e}\cdot
\frac{\lambda p_c}{\lambda p_c + p_d + p_e}\cdot
\frac{p_d}{p_d + p_e}
\end{equation}

In equation \ref{asuccb}, the first term is standard Plackett-Luce: at
this point, nooone has finished and cheering effects are absent.  The
second term reflects the fact that competitors $b$ and $c$ are
supported by competitor $a$, who has by this point finished the race
and is supporting his teammates.

By contrast, the likelihood function for observation $d\succ a\succ
c\succ b\succ e$ would be

\begin{equation}
\frac{p_d}{p_a + p_b + p_c + p_d + p_e}\cdot
\frac{p_a}{p_a + p_b + p_c + \mu p_e}\cdot
\frac{\lambda p_c}{\lambda p_b + \lambda p_c + p_e}\cdot
\frac{\lambda p_b}{\lambda p_b + \mu p_e}
\end{equation}

where this likelihood function reflects the mutual support for
equivalence class $\left\lbrace d,e\right\rbrace$.  Note that the
final term reflects the fact that competitors $b$ and $e$ have their
support active when vying for fourth place, as members of both their
teams have finished at this point.

This probability model could easily be modified to account for
specific circumstances.  The cheering effect be asymmetrical (with $a$
helping $b$ but not vice-versa, for example).  The effect could
operate only on specific ordered pairs.  Or perhaps the effect might
have a finite lifetime: if $a$ places $n^\mathrm{th}$, then the
cheering effect is active only for competitors placing
$(n+r)^\mathrm{th}$ or above.

# Package idiom

We can investigate red bus-blue bus phenomenon (as discussed, in a
slightly different context, in `inst/red_bus_blue_bus.Rmd`).  Here, we
consider a person who is given the choice of five transport methods:

* `C`, car
* `T`, train
* `RB` a red bus
* `BB` a blue bus
* `W` walking

Now, he does not really care what colour the bus is.  If we ask him to
rank his options, it is highly probable that he will put `RB` and `BB`
consecutively (because they are essentially indistinguishable).  Can
we quantify the strength of this effect?  To do this, we define a
bespoke function `RB_BB_LF()` which returns a `hyper3` log-likelihood
function corresponding to repeated observations of our commuter's
reported ranks for the five options:

```{r,defineredbuslf}
`RB_BB_LF` <- function(s){
    ec <- c(C=1,T=2,RB=3,BB=3,W=4) # equivalence classes
    h <- c(1,1,s,1)                # strength of support
    (
        cheering3(v=c("RB","BB","C" ,"T","W"),e=ec,h=h)*3 +
        cheering3(v=c("BB","RB","T" ,"C","W"),e=ec,h=h)*2 +
        cheering3(v=c("T" ,"BB","RB","C","W"),e=ec,h=h)*2 +
        cheering3(v=c("W" ,"BB","RB","T","C"),e=ec,h=h)*4 +
        cheering3(v=c("C" ,"RB","BB","W","T"),e=ec,h=h)*4 +
        cheering3(v=c("BB","C" ,"RB","T","W"),e=ec,h=h)*3
    )
}
```

Above, we see from the function body that he reported `RB,BB,C,T,W`
three times [first row], `BB,RB,T,C,W` twice [second row], and so on;
perhaps his ranking depends on the weather or how tired he is on any
given day.  Observe that in almost every case he ranks `RB` and `BB`
consecutively.  Function `RB_BB_LF()` takes argument `s` that
quantifies the perceived similarity between `RB` and `BB`.  For
example:

```{r,examplerbbblf,cache=TRUE}
(H <- RB_BB_LF(s=1.8888))
(mH <- maxp(H,n=1))
```

Now to find a profile likelihood function for `s`:

```{r,label=proflikeredbus,cache=TRUE}
o <- function(s){maxp(RB_BB_LF(s),give=TRUE,n=1)$likes} # optimand
s <- exp(seq(from=log(1.3),to=log(47),len=17)) # putative similarity measures
L <- sapply(s,o)
L <- L-max(L)
```

We can plot these:

```{r,plotproflike}
plot(s,L,type="b")
abline(h=c(0,-2))
abline(v=1)
plot(log(s),L,type="b")
abline(h=c(0,-2))
abline(v=0)
```

And formally maximize the likelihood:

```{r,label=maxlikesim,cache=TRUE}
(osup <- optimize(o,c(6,10),maximum=TRUE))
```

So a likelihood ratio test of the null that $S=1$ would be:

```{r,label=LRT_RB,cache=TRUE}
(suppdiff <- o(osup$maximum) - o(1))
```

Easily satisfying Edwards's two-units-of-support criterion; Wilks gives us an asymptotic $p$-value:

```{r,label=RBBBpval}
pchisq(suppdiff*2,df=1,lower.tail=FALSE)
```

Now use the evaluate for the likelihood function:

```{r calcmaxmax, cache=TRUE}
maxHmax <- maxp(RB_BB_LF(s = osup$maximum))
```

```{r showmaxmax}
maxHmax
```

# University ranking analysis

Here we use a dataset of university rankings, `timesData.csv`, taken
from
`https://github.com/arnaudbenard/university-ranking/blob/master/timesData.csv`.

```{r, label=readtableandcalc,cache=TRUE}
a <- read.table("timesData.csv",sep=",", header=TRUE)
wanted <- c("California Institute of Technology", "Harvard University",
"Massachusetts Institute of Technology", "Princeton University",
"Stanford University", "University of Cambridge", "University of Oxford")
names(wanted) <- c("cal","harv","mass","prin","stan","cam","ox")

a <- a[a$university_name %in% wanted,]
a <- cbind(a,"top7rank"=0)
for(y in unique(a$year)){
    a[a$year==y,"top7rank"] <- order(
     			  as.numeric(a[a$year==y,"world_rank"]) +
                                     a[a$year==y,"research"]/1e6,
decreasing=TRUE)}
a <- a[,c("top7rank","university_name","year")]
a <- reshape(a,idvar="university_name",timevar="year",direction="wide")
for(i in seq_len(nrow(a))){
   a$university_name[i] <- names(which(wanted == a$university_name[i]))
}
rownames(a) <- a$university_name
a <- a[,-1]
colnames(a) <- paste("Y",2011:2016,sep="")
a
H <- ordertable2supp(a)
```

```{r,label=testuniv,cache=TRUE}
equalp.test(H)
samep.test(H,c("ox","cam"))
```

Start to use `hyper3` idiom:

```{r,label=defineoxcam}
H3 <- function(oxcam){
  out <- hyper3()
  for(i in seq_len(ncol(a))){
    jj <- rep("",nrow(a))
    jj[a[,i]] <- rownames(a)
    out <- out + cheering3(v=jj,e=c(ox=1,cam=1,prin=2, stan=3, mass=4, harv=5, cal=6), help=c(oxcam,1,1,1,1,1))
  }
  return(out)
}
```

```{r,label=calculateoxcam,cache=TRUE}
o <- function(oxcam){maxp(H3(oxcam),give=TRUE,n=1)$likes}
oc <- exp(seq(from=log(0.5),to=log(5),len=15))
L <- sapply(oc,o)
L <- L - max(L)
```

```{r,label=plotoxcam}
plot(log(oc),L,type="b")
abline(v=0)
```

# Five nations championship

The five nations rugby championship was held from 1910 to 1999 and
file `five_nations.txt` shows the order statistic for England (E),
Scotland (S), Ireland (I), Wales (W), and France (F).

`https://en.wikipedia.org/wiki/Six_Nations_Championship`

Here is `hyper2` analysis:

```{r,load5nationsdata,cache=TRUE}
a <- as.matrix(read.table("five_nations.txt",header=FALSE))
head(a)
H <- hyper2()
for(i in seq_len(nrow(a))){
  H <- H + race(a[i,-1])
}
mH <- maxp(H)
pie(mH)
equalp.test(H)
```

Now use `hyper3` to see whether teams do better following a win:

```{r,rugby3,cache=TRUE}
rugby <- function(lambda){
  H3a <- hyper3()
  for(i in seq(from=2,to=nrow(a))){
        last_year_winner <- a[i-1,2] # e.g. "W" or "E"
        H3a <- H3a + ordervec2supp3a(a[i,-1],nonfinishers=NULL,helped=last_year_winner,lambda=lambda)
  }
  return(H3a)
}
rugby(1.888)
rugby(1.111111)
```

```{r,rugby3max, cache=TRUE}
maxp(rugby(1.8),n=1,give=TRUE)
maxp(rugby(1.9),n=1,give=TRUE)
```

## References {-}