### 2025-02-19 inst/jss_hyper3_real.Rnw
sushi example

% -*- mode: noweb; noweb-default-code-mode: R-mode; -*-
\documentclass[article]{jss}
\usepackage{booktabs}
\usepackage{units}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}

\author{Robin K. S. Hankin\\University of Stirling}
\title{Generalized Bradley-Terry models}
\Plainauthor{Robin K. S. Hankin} %% comma-separated
\Plaintitle{Generalized Bradley-Terry models}

%% an abstract and keywords
\Abstract{

  The \code{hyper2} package provides functionality to work with
  extensions of the Bradley-Terry probability model such as reified
  entities (monsters) that account for known enhancements in
  competitive ability.  Here, I present a generalization of
  \code{hyper2} in which Bradley-Terry strengths may be enhanced using
  a multiplicative factor $\lambda$.  Generally, ordinary
  Bradley-Terry is recovered if $\lambda=1$.  This approach offers
  multiple advantages over reified Bradley-Terry, and several examples
  are given.
}
\Keywords{Plackett-Luce,
  Bradley-Terry, Mann-Whitney} \Plainkeywords{Plackett-Luce,
  Bradley-Terry, Mann-Whitney}

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Robin K. S. Hankin\\
  University of Stirling\\
  E-mail: \email{hankin.robin@gmail.com}\\
  URL: \url{https://academics.aut.ac.nz/robin.hankin}
}

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand\headercell[1]{%
   \smash[b]{\begin{tabular}[t]{@{}c@{}} #1 \end{tabular}}}
\newcommand{\draw}{\ensuremath{\mathfrak{D}}}
\newcommand{\cdraw}{\ensuremath{\mathfrak{C}}}
\newcommand{\white}{\ensuremath{\mathfrak{W}}}
\newcommand{\home}{\ensuremath{\mathfrak{H}}}
\newcommand{\oo}{\vphantom{W^{W^{W^W}}}}

\begin{document}

<<setup, cache=FALSE,echo=FALSE>>=
source("loadlibrary.R",echo=FALSE)  # to avoid stupid messages
opts_chunk$set(cache=TRUE, autodep=TRUE)
options(prompt = "R> ")
@

\section{Introduction}

The Bradley-Terry model for datasets involving paired comparisons has
wide uptake in the {R} community.  The canonical problem is to
consider~$n$ players who compete against one another; the basic
inference problem is to estimate
numbers~$\mathbf{p}=\left({p_1,\ldots,p_n}\right)$, $p_i\geqslant 0$,
$\sum p_i=1$ so that
$\Prob(\mbox{$i$~beats~$j$})=\frac{p_i}{p_i+p_j}$.  Information about
the~$p_i$ may be obtained from the results of paired comparisons
between the players.

One generalization of the Bradley-Terry model is due
to~\citet{luce1959}, in which the probability of competitor~$i$
winning in a field of~$j=1,\ldots, n$ is $\frac{p_i}{p_1+\cdots
  +p_n}$.  It is then natural to quantify some ephemeral enhancement
to competitor~$i$ using a multiplicative factor $\lambda$.  Then
$\Prob(\mbox{$i$ wins})=\frac{\lambda p_i}{p_1+\cdots+\lambda
  p_i+\cdots+p_n}$.  Examples of such enhancements include the
home-ground advantage in football and rugby, playing white in chess,
or XXXX.  Note that $\lambda < 1$ would correspond to a decrease in
competitive ability.

Further, one might follow HANKIN and consider teams of players.  In
doubles tennis, for example, $\left\lbrace p_1,p_2\right\rbrace$ might
play $\left\lbrace p_3,p_4\right\rbrace$ and win with probability
$\frac{p_1+p_2}{p_1+p_2+p_3+p_4}$.  It is natural to combine this idea
with that of a multiplicative factor.  For example, suppose $p_1$ has
a special move against $p_3$, and $p_4$ has just been given a
performance-enhancing drug, we might have a likelihood functions
proportional to $\frac{\lambda p_1+ p_2}{p_1+p_2+p_3+\mu p_4}$.  Then
the effectiveness of $p_3$'s special move and the enhancement afforded
by the drug are measured by $\lambda$ and $\mu$ respectively with
values $>1$ corresponding to a positive effect.

In this paper I present a generalization of the {\tt hyper2} suite for
Bradley-Terry likelihoods to accommodate general multiplicative
factors.  The most general likelihood function would be

\begin{equation}
  \prod_{i=1}^N
  \sum_{j=1}^n
 \left(\lambda_{ij}p_i\right)^{w_i}
  \end{equation}

where the $p_i$ are traditional Bradley-Terry strengths and the
$\lambda_{ij}\geqslant 0$ are multiplicative factors that account for
ephemeral or temporary changes in competitive ability.  In practice,
almost all the $\lambda_{ij}$ are zero.  To fix ideas, consider the
data presented in Table~\ref{oldfirm}.  Ignoring draws for the moment,
we have a likelihood proportional to

\begin{equation}
  \left(\frac{\lambda p_R}{\lambda p_R + p_C}\right)^{81}\cdot
  \left(\frac{\lambda p_R}{\lambda p_R + p_F}\right)^{69}\cdots
  \left(\frac{\lambda p_L}{\lambda p_L + p_F}\right)^{81}\cdots
  \left(\frac{p_C}{p_C + \lambda p_R}\right)^{42}\cdots
    \left(\frac{p_F}{p_F + \lambda p_L}\right)^{9}
\end{equation}

In {\tt hyper3} idiom, a term such as $\lambda p_R + p_C$ is
represented as a named numeric vector.  For example, if the home
strength were $\lambda = 1.8$ then the first bracket would be {\tt
  c(Rangers=1.8, Celtic=1)} with an associated power of $-81$.  The
computational methodology is a generalization of that used
in~\cite{hankin2022} except that the multiplicative factors are real
(as opposed to non-negative integer), and also unknown.

\subsection{The package in use}
Package idiom for creating a likelihood function for this dataset
would look something like this: Considering only Rangers and Celtic:
Rangers at home:

<<defineoldfirm, cache=TRUE>>=
oldfirm <- function(lambda){
    H <- hyper3(pnames = c("Rangers", "Celtic"))
    H[c(Rangers = lambda)            ] %<>% add(81)
    H[c(Celtic = 1)                  ] %<>% add(42)
    H[c(Rangers = lambda, Celtic = 1)] %<>% subtract(81 + 42)

    H[c(Celtic = lambda)             ] %<>% add(76)
    H[c(Rangers = 1)                 ] %<>% add(46)
    H[c(Celtic = lambda, Rangers = 1)] %<>% subtract(46 + 76)
    return(H)
}
@

Above we see package idiom for creating a log-likelihood function.  We
first create an empty {\tt hyper3} object with an appropriate names
attribute.  The next three lines, using \pkg{magrittr} assignment
pipes, give the results for Rangers at home, and the next three lines
give the result for Celtic at home.  It is then easy to create a
bespoke log-likelihood function for Rangers-Celtic matches:

<<useoldfirm, cache = TRUE>>=
oldfirm(1.88)
@

We see a log-likelihood function for all $82+42+76+46=245$
Rangers-Celtic matches that were not score draws, conditional on
assumed value $\lambda=1.88$.  We may make inferences about the true
value of $\lambda$ using profile support:

<<profilelikeoldfirm, cache=TRUE>>=
f <- function(lambda){maxp(oldfirm(lambda),give=1)$`Log-likelihood`}
lam <- seq(from=0.9, to=3, len=23)
logLike <- sapply(lam,f)
@

Thus {\tt f()} returns the support at the evaluate, conditional on a
supplied value of $\lambda$.  We may use this to create a profile
likelihood:

\begin{figure}[htbp]
  \begin{center}
<<plotproflikeoldfirm, echo=FALSE>>=
plot(lam,logLike - max(logLike),type='b')
abline(v=1,col='gray')
segments(x0=1.785,y0=0.1,y1=-1,col='red')
abline(h = c(0,-2),col='gray')
@
\caption{A\label{profilesupportlambda} profile support function for the
  home-team advantage $\lambda$.  Vertical gray line shows
  $H_0\colon\lambda =1$, horizontal lines show the
  two-units-of-support limits; small red bar shows
  $\hat{\lambda}\simeq 1.785$}
\end{center}
\end{figure}

<<optimoldfirm, cache=TRUE>>=
(lambda_max <-  optimize(f,c(1.5,2.0),maximum=TRUE)$maximum)
@

<<mleoldfirm,cache=TRUE>>=
maxp(oldfirm(lambda_max))
@

Compare with the theoretical value of $\widehat{p_{\mbox{\tiny
      C}}}=\frac{\sqrt{42\cdot 76}}{\sqrt{81\cdot 46}+\sqrt{42\cdot
    76}}\simeq 0.4807$, $\widehat{p_{\mbox{\tiny R}}}
=\frac{\sqrt{81\cdot 46}}{\sqrt{81\cdot 46}+\sqrt{42\cdot 76}}\simeq
0.5193$, and $\widehat{\lambda}=\frac{\sqrt{76\cdot 81}}{\sqrt{46\cdot
    42}}\simeq 1.785$.

\section{Score draws}

About 22\% of matches in Table~\ref{oldfirm} end in a draw and {\tt
  hyper2} formalism may be used to assess this.  Consider the
likelihood function described in Table~\ref{windrawloss}; this has a
number of desirable features.  Firstly, if we consider two players of
identical strengths and nullify the home ground effect (that is,
$\lambda=1$) then the probability of a draw is $\frac{2D}{2+2D}$; the
odds of a draw is simply $D$, regardless of team strength.  Secondly,
considering only won (that is, not score-drawn) games, the probability
of a home win is $\frac{\lambda}{1+\lambda}$, that is, the odds of a
home win is simply $\lambda$, independently of the player
strength. Such arguments give direct operational significance to
$\lambda$ and $D$.

\begin{table}
\centering
\begin{tabular}{l|c|c}\\&
{$i$ plays at home} &
{$i$ plays away} \\
  \hline
  $\Prob{(\mbox{$i$ wins})}$                             &
  $\frac{\lambda p_i}{\lambda p_i + D(p_i + p_j) + p_j}$  &
  $\frac{p_i}{p_i + D(p_i + \lambda p_j) + p_j}$          \\
  $\Prob{(\mbox{draw})}$                                 &
  $\frac{D(p_i + p_j)}{\lambda p_i+D(p_i+p_j) + p_j}$     &
  $\frac{D(p_i + p_j)}{p_i + D(p_i+p_j) + \lambda p_j}$   \\
  $\Prob{(\mbox{$i$ loses})}$                            &
  $\frac{p_j}{\lambda p_i + D(p_i+p_j) + p_j}$            &
  $\frac{\lambda p_j}{p_i + D(p_i + p_j) + \lambda p_j}$  \\
\end{tabular}
\caption{Probabilities of a win, draw, loss for the two teams under a
  multiplicative Bradley-Terry model\label{windrawloss} and a lambda thing}
\end{table}

We are now in a position to consider the entirety of
Table~\ref{oldfirm}.  Bespoke function \code{home\_draw\_away3()}
converts data in the form of Table~\ref{oldfirm} to a \code{hyper3}
likelihood function following Table~\ref{windrawloss}; it takes
numerical values for the home advantage $\lambda$ and proclivity to
draw $D$:

<<eg_RCLF3_lf, cache=TRUE>>=
home_draw_away3(RCLF3_table, lambda=1.7, D=0.3)
@

Observe that the resulting likelihood function has two distinct types
of parameters: the plain Bradley-Terry strengths $\mathbf{p}=(p_R,
p_C,p_L,p_F)$, with unit sum; but also the auxiliary parameters
$\lambda,D$.  To make inferences about these parameters we calculate a
profile likelihood function: conditional on $\lambda,D$ we maximize
the likelihood over possible values of $\mathbf{p}$.

<<defineF>>=
f <- function(v){
    H <- home_draw_away3(RCLF3_table, lambda = v[1], D=v[2])
    maxp(H,give=1)$`Log-likelihood`
}
@

Thus \code{f()} gives us a likelihood function for
$\mathbf{v}=(\lambda,D)$.  We may find the evaluate using
\code{optim()}:

<<findmaxlambdaD, cache=TRUE>>=
maxv <- optim(par=c(2, 0.3), fn=f, control=list(fnscale = -1))$par
maxv
@

Then

<<showevaluate, cache=TRUE>>=
maxp(home_draw_away3(RCLF3_table, lambda = maxv[1], D=maxv[2]))
@

Such estimates have rich predictive power.  For example, if Rangers
were to play Celtic with both teams away [thus nullifying the home
  ground advantage: $\lambda=1$ in Table~\ref{windrawloss}], the
probability of win/draw/loss for Rangers would be
$\left(\frac{p_R}{(1+D)(p_R+p_C)};\frac{D}{1+D};\frac{p_R}{(1+D)(p_R+p_C)}\right)$,
estimated at $(0.39;0.27;0.34)$.

\begin{figure}[htbp]
  \begin{center}
<<makecontour, cache=TRUE>>=
n <- 20
lambda <- seq(from=1, to=3, len=n)
D <- seq(from=0.1, to=0.6, len=n)
V <- as.matrix(expand.grid(lambda, D))
LL <- apply(V, 1, f)
LL <- pmax(-20, LL - max(LL))
contour(lambda, D, matrix(LL, n, n),levels = -2*(0:9), xlab=expression(lambda),ylab=expression(D))
points(maxv[1], maxv[2], pch=16, col='red')
@
\caption{Contours of support\label{contourlambdaD} evaluated over
  a grid.  Maximum likelihood estimate for $(\lambda, D)$ shown as
  a red dot at the evaluate}
  \end{center}
\end{figure}

<<testlambdazero, cache=TRUE, include=FALSE>>=
jj <- optimize(f=function(D){f(c(lambda=1,D))}, interval=c(0.1, 0.4), maximum=TRUE)
@

Further, we may calculate a support function over the
$(\lambda,D)$-plane, shown in figure~\ref{contourlambdaD}.  For
example, we may test the hypothesis that $\lambda=1$ [corresponding to
  the right hand side of figure~\ref{contourlambdaD}] for which the
support is about \Sexpr{round(f(maxv) - jj$objective,2)}, far
exceeding Edwards's two units of support per degree of freedom
criteron.  Alternatively, we can observe that this lies in the tail
region of its asymptotic null $\chi^2_4$ distribution with a $p$-value
of about \Sexpr{signif(pchisq(f(maxv) -
  jj$objective,lower.tail=FALSE,df=2),digits=2)}.

\section{Non-independence in successive arrivals}

In the following, I use the ``race'' metaphor: five runners take part
in a race and arrive in order: $a\succ b\succ c\succ d\succ e$.
Consider the following Plackett-Luce likelihood function for this
observation:

\begin{equation}
\frac{p_a}{p_a + p_b + p_c + p_d + p_e}\cdot
\frac{p_b}{      p_b + p_c + p_d + p_e}\cdot
\frac{p_c}{            p_c + p_d + p_e}\cdot
\frac{p_d}{                  p_d + p_e}\cdot
\frac{p_e}{                        p_e}
\end{equation}

Note the use of multiplication between the terms, indicating
conditional independence.  We may use the multiplicative
generalization of Bradley-Terry strengths to introduce an element of
non-independence between the terms.  There are many ways to do this,
but one simple case would be to define equivalence classes of the
competitors with each equivalence class comprising mutually supporting
runners.  Equivalence classes of size one correspond to unsupported
runners.  The metaphor would be that a runner who has finished the
race is able to support other members of his equivalence class by
cheering his teammates, boosting their performance.  The \pkg{hyper2}
package includes function \code{cheering()} which implements this
functionality.

As an example, consider figure~\ref{tikzabcde}.  This shows a partial
probability tree diagram for some of the $5!=120$ possible order
statistics.  The standard Plackett-Luce likelihoods have been modified
to account for two groups of mutually supporting runners:
$\left\lbrace a,b,c\right\rbrace$ with support term $\lambda$, and
$\left\lbrace d,e\right\rbrace$ with support term $\mu$.  From START,
the first runner to cross the finishing line has standard
Plackett-Luce probability.  Taking the top path as an example, we see
that the likelihood function for $a\succ b\succ c\succ d\succ e$ would
be

\begin{figure}[htbp]
  \begin{center}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw[thick] (0,0) -- (2, 3) -- (4.8, 3);
\draw[thick] (0,0) -- (2, 2) -- (4.8, 2);
\draw[thick] (0,0) -- (2, 1) -- (4.8, 1);
\draw[thick] (0,0) -- (2, 0) -- (4.8, 0);
\draw[thick] (0,0) -- (2,-1) -- (4.8,-1);
\node at (0,2.5) (eq1) {$\underbrace{\left\lbrace a,b,c\right\rbrace}_{\lambda}\underbrace{\left\lbrace d,e\right\rbrace}_{\mu}$};

\node at (4.8, 2.8) {$a$};
\node at (4.8, 1.8) {$b$};
\node at (4.8, 0.8) {$c$};
\node at (4.8,-0.2) {$d$};
\node at (4.8,-1.2) {$e$};

\node at (3.4, 3.4)  {$\frac{p_a}{p_a+p_b+p_c+p_d+p_e}$};
\node at (3.4, 2.4)  {$\frac{p_b}{p_a+p_b+p_c+p_d+p_e}$};
\node at (3.4, 1.4)  {$\frac{p_c}{p_a+p_b+p_c+p_d+p_e}$};
\node at (3.4, 0.4)  {$\frac{p_d}{p_a+p_b+p_c+p_d+p_e}$};
\node at (3.4,-0.6)  {$\frac{p_e}{p_a+p_b+p_c+p_d+p_e}$};

\fill (4.8, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill (4.8, 2) circle[radius=2pt];  % terminal node
\fill (4.8, 1) circle[radius=2pt];  % terminal node
\fill (4.8, 0) circle[radius=2pt];  % d finishes; paths abce
\fill (4.8,-1) circle[radius=2pt];  % terminal node

\draw[thick] (4.8,3) -- (5.8,5) -- (8.2,5);
\draw[thick] (4.8,3) -- (5.8,4) -- (8.2,4);
\draw[thick] (4.8,3) -- (5.8,3) -- (8.2,3);
\draw[thick] (4.8,3) -- (5.8,2) -- (8.2,2);

\newcommand{\la}[1]{\lambda{#1}}
\newcommand{\ld}[1]{\mu    {#1}}
\node at (7, 5.4)  {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (7, 4.4)  {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (7, 3.4)  {$\frac{p_d     }{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (7, 2.4)  {$\frac{p_e     }{\la{p_b}+\la{p_c}+p_d+p_e}$};

\fill (8.2,5) circle[radius=2pt];  % ab finishes; paths cde
\fill (8.2,4) circle[radius=2pt];  % ab finishes; paths cde
\fill (8.2,3) circle[radius=2pt];  % ab finishes; paths cde
\fill (8.2,2) circle[radius=2pt];  % ab finishes; paths cde

\node at (8.2,4.8) {$b$};
\node at (8.2,3.8) {$c$};
\node at (8.2,2.8) {$d$};
\node at (8.2,1.8) {$e$};

\draw[thick] (8.2,5) -- (8.9,6) -- (10.6,6);
\draw[thick] (8.2,5) -- (8.9,5) -- (10.6,5);
\draw[thick] (8.2,5) -- (8.9,4) -- (10.6,4);

\node at (9.5, 6.4) {$\frac{\la{p_c}}{\la{p_c}+p_d+p_e}$};
\node at (9.5, 5.4) {$\frac{p_d  }{\la{p_c}+p_d+p_e}$};
\node at (9.5, 4.4) {$\frac{p_e  }{\la{p_c}+p_d+p_e}$};

\fill (10.6, 6) circle[radius=2pt];  % abc finishes; paths de
\node at (10.6,5.8) {$c$};
\draw[thick] (10.6, 6) -- (11.2,7) -- (12.2,7);
\fill (12.2,7) circle[radius=2pt];
\node[anchor = west] at (12.2,7) {$a\succ b\succ c\succ d\succ e$};
\node at (11.8, 7.4) {$\frac{p_d}{p_d+p_e}$};

\draw[thick] (10.6, 6) -- (11.2,6.2) -- (12.2,6.2);
\fill (12.2,6.2) circle[radius=2pt];
\node[anchor = west] at (12.2,6.2) {$a\succ b\succ c\succ e\succ d$};
\node at (11.8, 6.6) {$\frac{p_e}{p_d+p_e}$};

\fill (10.6, 5) circle[radius=2pt];
\node at (10.6,4.8) {$d$};

\draw[thick] (10.6, 5) -- (11.2,5.3) -- (12.2, 5.3);
\fill (12.2,5.3) circle[radius=2pt];
\node[anchor = west] at (12.2,5.3) {$a\succ b\succ d\succ c\succ e$};
\node at (11.8, 5.7) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_e}}$};

\draw[thick] (10.6, 5) -- (11.2,4.7) -- (12.2, 4.7);
\fill (12.2,4.7) circle[radius=2pt];
\node[anchor = west] at (12.2,4.7) {$a\succ b\succ d\succ e\succ c$};
\node at (11.8, 5.0) {$\frac{\ld{p_e}}{\la{p_c}+\ld{p_e}}$};

\fill (10.6, 4) circle[radius=2pt];
\node at (10.6,3.8) {$e$};
\draw[thick] (10.6, 4) -- (11.2,3.9) -- (12.2, 3.9);
\fill (12.2,3.9) circle[radius=2pt];
\node[anchor = west] at (12.2,3.9) {$a\succ b\succ e\succ c\succ d$};
\node at (11.8, 4.3) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_d}}$};

\draw[thick] (10.6, 4) -- (11.2,3.1) -- (12.2,3.1);
\fill (12.2,3.1) circle[radius=2pt];
\node at (11.8, 3.5) {$\frac{\ld{p_d}}{\la{p_c}+\ld{p_d}}$};
\node[anchor = west] at (12.2,3.1) {$a\succ b\succ e\succ d\succ c$};

\draw[thick] (4.8, 0) -- (5.8,1) -- (8.2,1);
\fill (12.2,1) circle[radius=2pt];
\node[anchor = west] at (12.2,1) {$d\succ a\succ c\succ e\succ b$};
\node at (7, 1.4)  {$\frac{p_a}{p_a+p_b+p_c+\ld{p_e}}$};

\draw[thick] (4.8, 0) -- (5.8,0) -- (8.2,0);
\fill (12.2,0) circle[radius=2pt];
\node[anchor = west] at (12.2,0) {$d\succ e\succ a\succ b\succ c$};
\node at (7, 0.4)  {$\frac{p_b}{p_a+p_b+p_c+\ld{p_e}}$};

\draw[thick] (4.8, 0) -- (5.8,-1) -- (8.2,-1);
\node at (7, -0.6)  {$\frac{p_c}{p_a+p_b+p_c+\ld{p_e}}$};

\draw[thick] (4.8, 0) -- (5.8,-2) -- (8.2,-2);
\node at (7, -1.6)  {$\frac{\ld{p_e}}{p_a+p_b+p_c+\ld{p_e}}$};

\fill (8.2, -2) circle[radius=2pt];  % de finishes; paths abc
\fill (8.2, -1) circle[radius=2pt];  % de finishes; paths abc
\fill (8.2,  0) circle[radius=2pt];  % de finishes; paths abc

\node at (8.2, -2.2) {$e$};
\node at (8.2, -1.2) {$c$};
\node at (8.2, -0.2) {$b$};
\node at (8.2,  0.8) {$a$};

\draw[thick] (8.2, -2) -- (8.9,-1) -- (10.6,-1);
\draw[thick] (8.2, -2) -- (8.9,-2) -- (10.6,-2);
\draw[thick] (8.2, -2) -- (8.9,-3) -- (10.6,-3);

\fill (12.2,-1) circle[radius=2pt];
\node[anchor = west] at (12.2,-1) {$d\succ e\succ a\succ c\succ b$};

\node at (9.5, -0.6) {$\frac{p_a}{p_a+p_b+p_c}$};
\node at (9.5, -1.6) {$\frac{p_b}{p_a+p_b+p_c}$};
\node at (9.5, -2.6) {$\frac{p_c}{p_a+p_b+p_c}$};

\fill (10.6, -1) circle[radius=2pt];  % dea finishes; paths bc
\fill (10.6, -2) circle[radius=2pt];  % deb finishes; paths ac
\fill (10.6, -3) circle[radius=2pt];  % dec finishes; paths ab
\node at (10.6,-1.2) {$a$};
\node at (10.6,-2.2) {$b$};
\node at (10.6,-3.2) {$c$};

\draw[thick] (10.6, -1) -- (11.2,-0) -- (12.2,0);

\node at (11.8, 0.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}}$};
\draw[thick] (10.6, -1) -- (11.2,-1) -- (12.2,-1);

\node at (11.8, -0.6) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}}$};

\draw[thick](10.6, -2) -- (11.2,-1.8) -- (12.2,-1.8);
\fill (12.2,-1.8) circle[radius=2pt];
\node[anchor = west] at (12.2,-1.8) {$d\succ e\succ b\succ a\succ c$};
\node at (11.8, -1.4) {$\frac{\la{p_a}}{\la{p_a}+\la{p_c}}$};

\draw[thick] (10.6, -2) -- (11.2,-2.6) -- (12.2,-2.6);
\fill (12.2,-2.6) circle[radius=2pt];
\node[anchor = west] at (12.2,-2.6) {$d\succ e\succ b\succ c\succ a$};
\node at (11.8, -2.2) {$\frac{\la{p_c}}{\la{p_a}+\la{p_c}}$};

%\draw[thick] (10.6, -3) -- (11.2,-2.9);
%\fill (12.2,-2.9) circle[radius=2pt];
%\node at (11.8, -2.9) {$\frac{\la{p_a}}{\la{p_a}+\la{p_b}}$};

\draw[thick] (10.6, -3) -- (11.2,-3.4) -- (12.2, -3.4);
\fill (12.2,-3.4) circle[radius=2pt];
\node[anchor = west] at (12.2,-3.4) {$d\succ e\succ c\succ b\succ a$};

\node at (11.8, -3) {$\frac{\la{p_b}}{\la{p_a}+\la{p_b}}$};

\fill (8.2, 1) circle[radius=2pt];  % da finishes; paths bce

\draw[thick] (8.2,1) -- (8.9,2) -- (10.6,2);
\draw[thick] (8.2,1) -- (8.9,1) -- (10.6,1);
\draw[thick] (8.2,1) -- (8.9,0) -- (10.6,0);

\node at (9.5, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 1.4) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 0.4) {$\frac{p_e     }{\la{p_b}+\la{p_c}+p_e}$};
\fill (10.6, 2) circle[radius=2pt];  % da finishes; paths bce
\fill (10.6, 1) circle[radius=2pt];  % da finishes; paths bce
\fill (10.6, 0) circle[radius=2pt];  % da finishes; paths bce

\node at (10.6, 1.8) {$b$};
\node at (10.6, 0.8) {$c$};
\node at (10.6,-0.2) {$e$};

\draw[thick] (10.6, 1) -- (11.2,2) -- (12.2,2);
\node at (11.8, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\ld{p_e}}$};
\fill (12.2,2) circle[radius=2pt];
\node[anchor = west] at (12.2,2) {$d\succ a\succ c\succ b\succ e$};
\draw[thick] (10.6, 1) -- (11.2,1) -- (12.2,1);
\node at (11.8, 1.4) {$\frac{\ld{p_e}}{\la{p_b}+p_e}$};

\end{tikzpicture}
\caption{Partial probability\label{tikzabcde} tree for five
  competitors $a$-$e$ with two mutually supporting subsets
  $\left\lbrace a,b,c\right\rbrace$ and $\left\lbrace d,
  e\right\rbrace$ having support $\lambda$, $\mu$ respectively}
\end{center}
\end{figure}

\begin{equation}\label{asuccb}
\frac{p_a}{p_a + p_b + p_c + p_d + p_e}\cdot
\frac{\lambda p_b}{\lambda p_b + \lambda p_c + p_d + p_e}\cdot
\frac{\lambda p_c}{\lambda p_c + p_d + p_e}\cdot
\frac{p_d}{p_d + p_e}
\end{equation}

In equation \ref{asuccb}, the first term is standard Plackett-Luce: at
this point, no competitor has finished and cheering effects are
absent.  The second term reflects the fact that competitors $b$ and
$c$ are supported by competitor $a$, who has by this point finished
the race and is supporting his teammates.

By contrast, the likelihood function for observation $d\succ a\succ
c\succ b\succ e$ would be

\begin{equation}
\frac{p_d}{p_a + p_b + p_c + p_d + p_e}\cdot
\frac{p_a}{p_a + p_b + p_c + \mu p_e}\cdot
\frac{\lambda p_c}{\lambda p_b + \lambda p_c + p_e}\cdot
\frac{\lambda p_b}{\lambda p_b + \mu p_e}
\end{equation}

where this likelihood function reflects the mutual support for
equivalence class $\left\lbrace d,e\right\rbrace$.  Note that the
final term reflects the fact that competitors $b$ and $e$ have their
support active when vying for fourth place, as members of both their
teams have finished at this point.

This probability model could easily be modified to account for
specific circumstances.  The cheering effect could be asymmetrical
(with $a$ helping $b$ but not vice-versa, for example).  The effect
might operate only on certain ordered pairs, following a directed
graph (cyclic or acyclic) instead of equivalence classes.  Or perhaps
the effect might have a finite lifetime: if $a$ places
$n^\mathrm{th}$, then the cheering effect is active only for
competitors placing $(n+r)^\mathrm{th}$ or above, for some fixed $r$.
There is no reason that $\lambda$ could not be less than one, which
would indicate that some competitors inhibit, rather than help,
others.

The probability model of Figure~\ref{tikzabcde} is applied, in a
simple one-parameter form, to two examples: Formula 1 motor racing,
and also a preference rating.

\section{Red bus, blue bus}

The red bus-blue bus problem was originally stated by
\cite{chipman1960} and discussed by \cite{davidson1970} and
\cite[p153]{alvo2014}.  was \cite{debreu1960}.  The Holiday problem is
discussed in \cite{tverskey1972}.  A traveller has a choice of going
to work by different transport methods.  Here, we can assume that the
traveller has strong but unknown views on different forms of transport
which might incorporate a random component.  If considering a car or
bus, the colour of the bus is of no importance to him.  However, when
offered the choice between a red bus or a blue bus, the colour becomes
important (having nothing else to guide his decision).  Now, he does
not really care what colour the bus is.  If we ask him to rank his
options, it is highly probable that he will put RB and BB
consecutively (because they are essentially indistinguishable).  Can
we quantify the strength of this effect?  To do this, we define a
bespoke function \code{RB\_BB\_LF()} which returns a \code{hyper3}
log-likelihood function corresponding to repeated observations of our
commuter's reported ranks for the five options:

<<defineredbuslf>>=
`RB_BB_LF` <- function(lambda){
    ec <- c(C=1, T=2, RB=3, BB=3, W=4) 
    h <- c(1, 1, lambda, 1)           
    (
        cheering3(v=c("RB", "BB", "C" , "T", "W"), e=ec, h=h)*3 + 
        cheering3(v=c("BB", "RB", "T" , "C", "W"), e=ec, h=h)*2 + 
        cheering3(v=c("T" , "BB", "RB", "C", "W"), e=ec, h=h)*5 + 
        cheering3(v=c("W" , "BB", "RB", "T", "C"), e=ec, h=h)*4 + 
        cheering3(v=c("C" , "RB", "BB", "W", "T"), e=ec, h=h)*3 + 
        cheering3(v=c("BB", "C" , "RB", "T", "W"), e=ec, h=h)*3
    )
}
@

Above, we see from the function body that he reported $RB\succ BB\succ
C\succ T\succ W$ three times [first row], and so on; perhaps his
ranking depends on the weather or how tired he is on any given day.
Observe that in almost every case he ranks RB and BB consecutively.
Function \code{RB_BB_LF()} takes argument \code{lambda} that
quantifies the perceived similarity between RB and BB in the same way
as function \code{oldfirm()} above.  It is straightforward to find a
profile support function for $\lambda$,
Figure~\ref{profilesupportlambdaRBBB}, and formally maximize the
likelihood:

\begin{figure}[htbp]
\begin{center}
<<plotproflike>>=
o <- function(lambda){maxp(RB_BB_LF(lambda), give=1)$`Log-likelihood`} 
lambda <- exp(seq(from=log(0.9), to=log(47), len=17)) 
L <- sapply(lambda, o)
plot(log(lambda), L-max(L), type="b")
abline(h=c(0, -2))
abline(v=0)
@ 
\caption{A\label{profilesupportlambdaRBBB} profile support function
  for the red bus, blue bus parameter $\lambda$.  Vertical line
  indicates null of $\lambda=1$, horizontal lines show maximum support
  and $-2$}
\end{center}
\end{figure}

<<label=maxlikesim, cache=TRUE>>=
(osup <- optimize(o, c(10, 40), maximum=TRUE))
@

So a likelihood ratio test of the null that $\lambda=1$ would be:

<<label=LRT_RB, cache=TRUE>>=
(suppdiff <- o(osup$maximum) - o(1))
@

Easily satisfying Edwards's two-units-of-support criterion;
alternatively, we could observe that this observation lies in the tail
region of its asymptotic $\chi^2_1$ distribution (Wilks's theorem),
with a $p$-value of \Sexpr{signif(pchisq(suppdiff*2, df=1,
  lower.tail=FALSE),3)}.

\subsection{Sushi}

We now apply the ideas above to a dataset obtained from the {\tt
  preflib} library.

<<readsushi>>=
a <- as.matrix(read.table("sushi.txt"))
head(a)
@

<<defcalculate_sushi_H}>>=
make_sushi_H <- function(a, strength){
    h <- c(1, strength) # strength of support
    sushi_types <- c("ebi", "anago", "maguro", "ika", "uni", "sake", "tamago", "toro", "tekka", "kappa")
    ec <- # equivalence classes
        c(`ebi`  = 1, `anago` = 2, `maguro` = 1, `ika` = 1, `uni` = 1,
          `sake` = 1, `tamago` = 1, `toro` = 2, `tekka-maki` = 2,
          `kappa-maki` = 1)

    H <- hyper3()

    for(i in seq_len(nrow(a))){
        H <- H + cheering3(sushi_types[a[i,]], e=ec, h=h)
    }
    return(H)
}
@

<<tryH,cache=TRUE>>=
wanted <- sample(seq_len(nrow(a)), 50, replace=TRUE, prob=a[,1])
H1888 <- make_sushi_H(a[wanted,-1], strength=1.888)
H1000 <- make_sushi_H(a[wanted,-1], strength=1.000)
@

<<calclikesushi, cache=TRUE>>=
f <- function(strength){
   H <- make_sushi_H(a[wanted,-1], strength=strength)
   maxp(H,give=1)$`Log-likelihood`
}
s <- seq(from=0.4,to=4,len=5)
L <- sapply(s,f)
@

<<plotsushi>>=
plot(s,L-max(L),type='b')
@

<<justf1>>=
f1 <- f(1)
@

<<diffmaxlike>>=
f1 - max(L)
@

\section{Other stuff}

functions~\citep{hankin2010,hankin2017_rnw} account for enhancements
of competitive ability through the inclusion of reified entities
(monsters) whose Bradley-Terry strength represents the extra strength.
For example, if a player is at an advantage for some reason then a
reified entity might help him and then
$\Prob(\mbox{$i$~beats~$j$})=\frac{p_i+\mathfrak{M}}{p_i+p_j+\mathfrak{M}}$,
where $\mathfrak{M}$ is a reified entity whose strength helps
competitor $i$.
Such considerations motivate the
likelihood function in Equations~\ref{likefour}, originally set out by
XXXXX and YYYY.

\begin{align}\label{likefour}
\Prob{(\mbox{$i$ wins at home})} &= \frac{\lambda p_i}{\lambda p_i+p_j}\\
\Prob{(\mbox{$i$ wins away})} &= \frac{\lambda p_i}{\lambda p_i+p_j}\\
\Prob{(\mbox{$j$ wins at home})} &= \frac{\lambda p_j}{p_i+\lambda p_j}\\
\Prob{(\mbox{$j$ wins away})} &= \frac{p_j}{\lambda p_i+\lambda p_j}
\end{align}

Here, $\lambda\geqslant 0$ is a multiplicative constant accounting for
the home-ground advantage. If $\lambda>1$ the advantage is positive
and if $\lambda<1$ it is negative and helps the away side.

Here I introduce {\tt hyper3} formalism

However, reified Bradley-Terry techniques suffer from the following
defect, here illustrated using the context of Scottish football, and
the home advantage helping the home team and hindering the away team.
Consider the case of four teams 1-4 and represent the strength of the
home ground advantage as a home monster of strength $\mathfrak{H}$
(draws being discarded for the moment).  Now suppose that
$p_1=p_2=p_A$, and $p_3=p_4=p_B$.  The unit-sum constraint gives
$2p_A+2p_aB+\mathfrak{H}=1$.  The probability that team1 (at home)
beats team 2 (away) is then
$\frac{p_A+\mathfrak{H}}{2p_A+\mathfrak{H}}$; and the probability that
team 3 (home) beats team 4 (away) is
$\frac{p_B+\mathfrak{H}}{2p_B+\mathfrak{H}}$.  The odds ratios would
be $1+\mathfrak{H}/p_A$ and be $1+\mathfrak{H}/p_B$.  Such analysis is
incompatible with the perfectly reasonable requirement that the odds
ratio of home victory between two teams of equal strength be constant;
but here we see that weak teams are predicted to have an overwhelming
home-ground advantage (finding actual datasets that illustrate
precisely this phenomenon is not straightforward: reified
Bradley-Terry methods have a suprising amount of flexibility and can
accommodate wide ranges of strengths by dint of manipulating reified
strengths and real strengths).

However, consider the data presented in table~\ref{oldfirm} in which
we see football results from

\begin{table}
\centering
\begin{tabular}{@{} *{5}{c} @{}}
\headercell{\small{Plays at home,}\\ \small{wins}} &
\multicolumn{4}{c@{}}{\small{Plays away, loses}}\\
\cmidrule(l){2-5}
&   Rangers     &    Celtic & Falkirk & Livingston\\
\midrule
   Rangers      &   NA &    81  &    69   &      15\\
   Celtic       &   76 &    NA  &    61   &      15\\
   Falkirk      &   17 &    17  &    NA   &       6\\
   Livingston   &    2 &     2  &     3   &      NA\\
\end{tabular}
\\
\rule{1mm}{6mm}
\\
\begin{tabular}{@{} *{5}{c} @{}}
\headercell{\small{Plays at home,}\\ \small{draws}} &
\multicolumn{4}{c@{}}{\small{Plays away, draws}}\\
\cmidrule(l){2-5}
&  Rangers     &    Celtic & Falkirk & Livingston\\
\midrule
   Rangers     &       NA  &   44 &     14   &       2\\
   Celtic      &       47  &   NA &     10   &       3\\
   Falkirk     &       12  &   19 &     NA   &       4\\
   Livingston  &       4   &   4  &     4    &      NA\\
   \end{tabular}
\\
\rule{1mm}{6mm}
\\
\begin{tabular}{@{} *{5}{c} @{}}
\headercell{\small{Plays at home,}\\ \small{loses}} &
\multicolumn{4}{c@{}}{\small{Plays away, wins}}\\
\cmidrule(l){2-5}
&  Rangers     &    Celtic & Falkirk & Livingston\\
\midrule
   Rangers     &   NA  &  42  &    2  &       0\\
   Celtic      &   46  &  NA  &    8  &       0\\
   Falkirk     &   57  &  43  &   NA  &       8\\
   Livingston  &   12  &  11  &    9  &      NA\\
  \end{tabular}
\caption{The \label{oldfirm} dominance of the Old Firm: 669 matches
  between Rangers, Celtic, Falkirk and Livingston arranged by home
  wins, draws, and home losses.  Data from {\tt soccerworld}}
\end{table}

\section{Conclusions and further work}

consider Table~\ref{kka}, originally considered
by~\citet{hankin2020}.  Considering only non-drawn games for the
moment, we wish to represent the first-mover advantage using a ``White
monster'', of Bradley-Terry strength \white

\begin{table}
\centering
\begin{tabular}{@{} *{4}{c} @{}}
\headercell{\small{Plays White,}\\ \small{wins}} & \multicolumn{3}{c@{}}{\small{Plays black, loses}}\\
\cmidrule(l){2-4}
& An &  Kr & Ks        \\
\midrule
  An & -  &  18 &   6  \\
  Kr &  7 &  -  &  18  \\
  Ks & 15 &  30 &  -   \\
\end{tabular}\quad
\begin{tabular}{@{} *{4}{c} @{}}
\headercell{\small{Plays W,}\\ \small{draws}} & \multicolumn{3}{c@{}}{\small{Plays black, draws}}\\
 \cmidrule(l){2-4}
& An &  Kr & Ks        \\
\midrule
  An & -  &  20 &  20  \\
  Kr & 29 &  -  &  72  \\
  Ks & 26 &  57 &  -   \\
\end{tabular}\quad
\begin{tabular}{@{} *{4}{c} @{}}
\headercell{\small{Plays W,}\\ \small{loses}} & \multicolumn{3}{c@{}}{\small{Plays black, wins}}\\
\cmidrule(l){2-4}
& An &  Kr & Ks        \\
\midrule
  An & -  &   5 &   11  \\
  Kr & 13 &  -  &    9  \\
  Ks &  2 &   7 &  -   \\
\end{tabular}
\caption{Results of 365 chess games \label{kka} between Grandmaster
  Anand, Grandmaster Karpov, and Grandmaster Kasparov.  For example,
  Anand played white against Karpov a total of $18+20+13=51$ times and
  won 18, drew 20, lost 13}
\end{table}

\begin{table}
\centering
\begin{tabular}{l|cc|cc}\\&
  \multicolumn{2}{c}{$i$ plays at home} &
  \multicolumn{2}{c}{$i$ plays away} \\
 & Reified & $\lambda$ &
   Reified & $\lambda$ \\
  \hline
  $\Prob{(\mbox{$i$ wins})}$  &
  $\frac{p_i+\home\oo}{p_i+p_j+\home}$ &
  $\frac{\lambda p_i}{\lambda p_i+p_j}$ &
  $\frac{p_i}{p_i+p_j+\home}$ &
  $\frac{p_i}{\lambda p_i+p_j}$ \\
  $\Prob{(\mbox{$i$ loses})}$ &
  $\frac{p_j\oo}{p_i+p_j+\home}$   &
  $\frac{p_j}{p_i+\lambda p_j}$  &
  $\frac{p_j+\home}{p_i+p_j+\home}$ &
  $\frac{\lambda p_j}{p_i + \lambda p_j}$
  \\
\end{tabular}
\caption{Probabilities of a win, draw, loss for the two teams under a reified
  Bradley-Terry model\label{probabilities} and a lambda thing}
\end{table}

\section{Conclusions and further work}

Plackett-Luce likelihood functions for rank datasets have been

\bibliography{hyper2}
\end{document}