### 2025-02-20 inst/sushi.Rmd
sushi

---
title: "Sushi monster"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

<font size="1"> most of the time is taken calculating the
`calclikesushi` and `maxlike`chunks.  With a sample size of 50 in
chunk `tryH`, and 11 points on the graph, it takes about half an hour
to run without cache. </font>

To cite the `hyper2` package in publications, please use @hankin2017_rmd.

```{r}
a <- as.matrix(read.table("sushi.txt", skip=22))
set.seed(0)
wanted <- sample(seq_len(nrow(a)),50,replace=TRUE,prob=a[,1])
a <- a[wanted,-1]
```

```{r}
sushi_types <- c("ebi", "anago", "maguro", "ika", "uni", "sake", "tamago", "toro", "tekka", "kappa")
a[] <- sushi_types[a]
noquote(a)
    ec <- c("ebi"  = 1, "anago" = 2, "maguro" = 1, "ika" = 1, "uni" = 1,
          "sake" = 1, "tamago" = 1, "toro" = 2, "tekka" = 2,
          "kappa" = 1)

```

```{r,defcalculate_sushi_H}
make_sushi_H <- function(a, lambda){
    H <- hyper3()

    for(i in seq_len(nrow(a))){
        H <- H + cheering3(a[i,], e=ec, h=c(1,lambda))
    }
    return(H)
}
```

```{r tryH, cache=TRUE}
H1 <- make_sushi_H(a, lambda=1)
H2 <- make_sushi_H(a, lambda=2)
```

```{r maxpsushiH,cache=TRUE}
mH1 <- maxp(H1,give=1)
mH2 <- maxp(H2,give=1)
```

```{r showsushiH}
mH1
mH2
```

```{r calclikesushi, cache=TRUE}
f <- function(lambda){
   H <- make_sushi_H(a, lambda = lambda)
   maxp(H,give=1)$`Log-likelihood`
}

s <- exp(seq(from = log(0.9), to = log(6), len = 11))
L <- sapply(s,f)
```

```{r maxlike, cache=TRUE}
(jj <- optimize(f, 2:3, maximum=TRUE))
```

```{r plotsushilike}
plot(log(s), L-jj$objective, type='b')
```

```{r calc_sushi}
sushi <- make_sushi_H(a, lambda = jj$maximum)
```

```{r calc_sushimaxp}
(sushi_maxp <- maxp(sushi))
```

### Package dataset {-}

Following lines create `sushi.rda`, residing in the `data/` directory of the package.

```{r,label=savesushidataset}
sushi_table <- a
save(sushi_table,  sushi, sushi_maxp, file="sushi.rda")
```

## References {-}