### 2025-02-23 inst/skating.Rmd

---
title: "Ladies' figure skating in the 2002 Winter Olympics"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

<font size="1"> (takes about twenty minutes to run without cache)
</font>

To cite the `hyper2` package in publications, please use
@hankin2017_rmd).  This document analyses Bradley-Terry strengths of
the competitors in the Ladies' Figure Skating at the 2002 Winter
Olympics.  In the package, dataframe `skating_table` is copied from
Lock and Lock (see skating.Rd for more details).  It also creates file
`data/skating.rda` which contains R objects `skating`,
`skating_table`, and `skating_maxp`.

```{r,showskating}
set.seed(0)
skating_table <- ordertable(read.table("skating.txt", header=TRUE))
```

Object `skating_table` is an order table.  It is structured so that
each competitor is a row, and each judge is a column.  Function
`rank_likelihood()` considers each row to be a race [or a judge], so
we need to take a transpose.  The following R idiom is plausible but
**incorrect**:

```{r,skatingwrong}
skating_incorrect <- rank_likelihood(t(skating_table)) # incorrect
```

The above is incorrect because we are applying a rank technique to an
order table.  Correct analysis follows.  One good thing to do is to
present the dataset as an order table;

```{r,coercetoranktable}
ordertable_to_ranktable(skating_table)
```

In the above, the column names should be read "came first" (`c1`),
"came second" (`c2`) and so on to "came 23rd" (`c23`).  From the first
column, we see that Hughes came first 5 times [according to judges
1,5,7,8,9] and Slutskya came first four times [judges 2,3,4,6].  From
the last column we see that all judges except `J3` awarded Luca last
place.  File `man/rrank.Rd` has a detailed discussion of the
differences between rank and order.

We can convert an order table to a support function:

```{r,convertordertosupport}
summary(suppfun(ordertable(skating_table)))
```

Further, we might ask how many judges ranked each competitor first,
second and so on:

```{r,alljudgeorder}
oo <- sapply(seq_len(23),function(i){rowSums(skating_table==i)})
colnames(oo) <- rep(" ",23)  # nicer print
oo
```

In the above, each row is a competitor and each column corresponds to
a ranking.  Thus the first column corresponds to "first place" and
shows that five judges ranked Hughes first and four judges ranked
Slutskaya first.  The second column corresponds to "second place" and
shows that one judge gave Hughes second place, two judges placed
Slutskaya second, five gave Kawn second, and one gave Cohen second.
The first row corresponds to Hughes and we see that, of $5+1+1+2=9$
judges, 5 placed Hughes first, one placed her second, one placed her
third, and two placed her fourth.

The formal ordering used in competition is, according to Lock and
Lock, given by the median ordinal:

```{r,medianordinal}
apply(skating_table,1,median)
```

but with ties broken using a complicated hierarchy, the first of which
is the "size of the majority":

```{r,sizeofthemajority}
rowSums(sweep(skating_table,1,seq_len(23))>=0)
```

which would suggest that Slutskaya beats Kwan (5-4), Butyrskaya draws
with Robinson (7-7), and so on.  The rows of `skating_table` are in
the order given by this system.

# Data visualization

Now some data visualization.  First the MLE for the strengths:

```{r,maxshow, cache = TRUE}
skating <- ordertable2supp(skating_table)
options("use_alabama" = TRUE)
skating_maxp <- maxp(skating,startp=
c(5.0493913251e-06, 0.0053037393083, 0.090240695999, 
0.00036893464900, 1.1397045995e-05, 8.3424027249e-05, 
9.8129903007e-06, 0.00026960143514, 0.29831585124, 
0.001376333090, 2.9434081237e-06, 0.26550514587, 
0.00051946513092, 1.7e-06, 0.00080745068338, 
0.00029358481485, 0.00048852120227, 0.0053075302458, 
0.0025877438737, 0.30889220253, 1.2869403350e-05, 
0.0184563624305147))    # values from previous repeated runs of maxp()
m <- skating_maxp # for ease of  typing
m
```

```{r equalpskatingtest,cache=TRUE}
equalp.test(skating)
```

```{r,dotchartm,dpi=72}
dotchart(m)
dotchart(log(m))
```

## Evidence for medal positions

Looking at the dotcharts, it seems that the medallists---Hughes (gold),
Slutskya (silver), Kwan (bronze)---were considerably higher in
strength than the rest of the field.  Here I will test the hypothesis
that the medallists were in fact the strongest three competitors.
Technically you need to optimize over the union of the possibilities
that one of the three medallists did not come in the top three; but
this is hard.  We will do something much easier but numerically
equivalent: optimize over the union of outcomes where either Cohen or
Suguri (who placed fourth and fifth respectively) had higher strength
than any of the medallists.

```{r}
f <- function(smaller,bigger){
  jj <- rep(0,22)
  names(jj) <- pnames(skating)[-23]  #  no fillup
  jj[smaller] <- -1
  jj[bigger ] <- +1
  return(jj)
}

problem_constraints <-  NULL
problem_constraints %<>% rbind(f("hughes"   ,"cohen" ))
problem_constraints %<>% rbind(f("slutskaya","cohen" ))
problem_constraints %<>% rbind(f("kwan"     ,"cohen" ))
problem_constraints %<>% rbind(f("hughes"   ,"suguri"))
problem_constraints %<>% rbind(f("slutskaya","suguri"))
problem_constraints %<>% rbind(f("kwan"     ,"suguri"))

rownames(problem_constraints) <-
c("hug < coh", "slu < coh", "kwa < coh", "hug < sug", "slu < sug", "kwa < sug" )
colnames(problem_constraints) %<>% substr(1,3)
problem_constraints

small <- 0.01
big <- 0.1
start_vector <- rep(small,22)
names(start_vector) <- pnames(skating)[-23]  #  no fillup
start_vector["cohen"] <- big
start_vector["suguri"] <- big
```

```{r,longthing,cache=TRUE}

out <- rep(0,nrow(problem_constraints))
fullout <- list()

mle_skaters <- NULL

for(i in seq_len(nrow(problem_constraints))){
  jj <- maxp(skating, startp=start_vector, give=TRUE,fcm=problem_constraints[i,], fcv=0,n=10,SMALL=1e-5)
   fullout[[i]] <- jj
   out[i] <- jj$value
   mle_skaters <- cbind(mle_skaters,jj$par)

}
out
out-max(out)
colnames(mle_skaters) <- rownames(problem_constraints)
mle_skaters
```

Now compare these values with the unconstrained maximum likelihoods:

```{r,unconstrainedlikes, cache=TRUE}
options(use_alabama = TRUE)
mgv <- maxp(skating, give=TRUE, n=1)$value
mgv
mgv  - out
mgv  - max(out)
```

that is, a little over 7 units of support, clearly exceeding the two
units suggested by Edwards.  We have strong evidence to suggest that
the medallists were indeed the strongest three competitors.  Observe
that the maximum likelihood among the six alternative hypotheses is
that of number 3, in which the maximization was constrained to obey
`Kwan < Cohen`.  If we _have_ to change the medallists, then
exchanging the order of Kwan and Cohen incurs the lowest likelihood
penalty.  If, instead, we ask whether there is evidence that Suguri
[who was much weaker than Cohen] should not have been a medallist, we
find

```{r subtractmaxfrommgv}
mgv  - max(out[4:6])
```

Much stronger.

```{r plotrestrictedlikes}
plot(out,ylab="log likelihood",axes=FALSE)
axis(2)
axis(side=1,at=1:6,srt=45,labels=c(
"Hug<Coh", "Slu<Coh", "Kwa<Coh",
"Hug<Sug", "Slu<Sug", "Kwa<Sug"
))
```

In the plot above, the vertical axis shows the support.  The six
points on the x-axis correspond to the six rows of
`problem_constraints`; names have been abbreviated to the first three
letters.  Thus the first three points are maximum likelihoods for
`Hughes < Cohen`, `Slutskya < Cohen`, and `Kwan < Cohen` respectively.

We may test the hypothesis that Slutskaya == Kwan == Hughes:

```{r,testtop3,cache=TRUE}
samep.test(skating,c("slutskaya","kwan","hughes"))
```

## Ranking methodologies.

maximum likelihood estimated strengths furnish an ordering for the
competitors.  We can compare ranking by strengths with the official
point-tallying method:

```{r rLrO,fig.cap="Official ranks vs likelihood rank, Ladies free skating, 2002 Winter Olympics"}
 rL <- sort(skating_maxp,decreasing=TRUE)
 rL[] <- seq_along(rL)
 rO <- seq_len(nrow(skating_table))
 names(rO) <- rownames(skating_table)
 ordertransplot(rO,rL,xlab="offical rank",ylab="likelihood rank",main="Ladies free skating, 2002 Winter Olympics")
```

In figure \@ref(fig:rLrO), there is close agreement between the two
methods in the large, but differences in detail.  For example, the
official ordering is hughes first, then slutskaya second, then kwan
third; likelihood winner is slutskya, followed by hughes and then
kwan.  We can gain some understanding of this result by looking at the
raw judging results for the three medallists:

```{r,threemedallists}
head(skating_table,2)
```

and then calculating a table of the results:

```{r, tableofresults}
hughes    <- unlist(skating_table[1,])
slutskaya <- unlist(skating_table[2,])
table(hughes)
table(slutskaya)
```

Looking at the above, we see that Hughes has five judges who gave her
"1"s, while Slutskaya has only 4, which is why she was considered to
be the top.  It seems that Hughes's two "4"s (compared with
Slutskaya's one) have cost her more likelihood-based strength than her
extra "1" gave her; note also that the two judges who placed Hughes
fourth (viz `J2` and `J4`) placed Slutskaya first.  Also:

```{r,compareslutskayahughes}
table(hughes < slutskaya)
```

(note that equality is disallowed) showing that five judges preferred Hughes
to Slutskaya and four preferred Slutsakaya to Hughes.

# Just the top seven competitors

```{r justtop7}
skating_table
o <- ordertable_to_ranktable(skating_table)
o
```

```{r setclasstomatrix}
class(o) <- "matrix"
o
```

Select just the top seven:
```{r selectthetop7}
o <- o[,1:7]
o
```

and now replace ranks with internal ranks (that is, rank among the top
seven):

```{r replaceranks}
o <- t(apply(o,1,rank))
colnames(o) <- rownames(skating_table)[1:7]
o
```

Now coece to a `ranktable`:

```{r coercetoranktble}
class(o) <- "ranktable"
o
```

Thence to an order table:

```{r thenceto}
o <- ranktable_to_ordertable(o)
o
```

Now create a likelihood function:

```{r makelikef}
o <- ordertable2supp(o)
o
```
and have some fun with it:

```{r plotapieofo,cache=TRUE}
(o <- maxp(o))
pie(o)
```

### Package dataset {-}

Following lines create `skating.rda`, residing in the `data/`
directory of the package.

```{r,label=saveskatingdataset}
save(skating_table, skating, skating_maxp, file="skating.rda")
```

### References {-}