### 2025-02-27 inst/salad.Rmd

---
title: "salad"
output: bookdown::html_document2
author: "R. K. S. Hankin"
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
library("prefmod")
library("PlackettLuce")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document shows how to apply `hyper2` idiom to the `salad`
dataset of the `prefmod` package [@hatzinger2012].  From `salad.Rd`:

```
The dataset contains the rankings of four salad dressings
concerning tartness by 32 judges, with values ranging
from 1 (most tart) to 4 (least tart).
```

```{r showsalad}
head(salad)
nrow(salad)
```

From row 3, for example, we see that salad `A` was ranked as second
most tart, `B` the most tart, `C` the third most tart, and `D` the
least tart of the four.  We may process this using two methods,
explicit and slick.  First, explicit:

```{r useexplicitmethod,cache=TRUE}
H1 <- hyper2()
for(i in seq_len(nrow(salad))){
   H1 <-  H1 + ordervec2supp(as.matrix(salad)[i,])
}
H1
```

And second, a slick method.  We take the transpose of `salad` (things
are clearer if we name the rows):

```{r}
rownames(salad) <-
   paste("j",formatC(seq_len(nrow(salad)),width=2,format="d",flag="0"),sep="")
ts <- ordertable(t(salad))
```

Above, we see that object `ts` is an order table.  It is just like the
formula 1 result table `F1_table_2016`, except that venues are judges
and drivers are the salads [which are "competing" for the "who has the
most tart dressing" prize].  Converting this to a support function is
accomplished with `ordertable2supp()`:

```{r useslickmethod,cache=TRUE}
(H2 <- suppfun(ts))
```

Just to check:

```{r checkequalityofsalads}
H1 == H2
```

```{r equalsaladtart, cache=TRUE}
equalp.test(H1)
```

The null estimate agrees to six places of decimals with that
presented by @turner2020_nomarkup:

```{r usepladmm, cache=TRUE}
standardPL_PlackettLuce <- PlackettLuce(salad, npseudo = 0)
(p1 <- itempar(standardPL_PlackettLuce))
(p2 <- maxp(H1))
p1-p2
```

We may use `hyper2` to test some hypotheses.  Firstly, $H_0\colon
p_B=\frac{1}{4}$, which we test with `specificp.test()`:

```{r specificptestB,cache=TRUE}
specificp.test(H2,"B",1/4)
```

Above we see a $p$-value of about $10^{-7}$, clearly significant.
However, we may also test the hypothesis that $p_B$ is less than the
second strongest, $p_C$:

```{r sameptestBC,cache=TRUE}
samep.test(H2,c("B","C"))
```

We see a $p$-value of about $7\times 10^{-4}$, much higher than the
value from `specificp.test()` of $H_0\colon p_B=\frac{1}{4}$.  This is
because `samep.test()` allows one to change the value of $p_C$ and we
would expect a less significant result.

# Acid analysis

Turner goes on to assess whether a model of tartness using acetic and
gluconic acids is appropriate.  This is possible using the `hyper2`
formalism as follows.

```{r}
features <- data.frame(salad = LETTERS[1:4],
                       acetic = c(0.5, 0.5, 1, 0),
                       gluconic = c(0, 10, 0, 10))
features
(M <- as.matrix(features[,-1]))
makestrengths <- function(v){
    out <- exp(M %*% v)
    out <- c(out/sum(out))
    out <- pmax(out,0)
    names(out) <- LETTERS[1:4]
    return(out)
}
makestrengths(c(2,1))
f <- function(v){loglik(makestrengths(v),H2)}
(maxo <- optim(par=c(3, 0.3), fn=f,control=list(fnscale = -1)))
makestrengths(maxo$par)
loglik(makestrengths(maxo$pa),H1)
maxp(H1,give=1)
```

# References