### 2025-03-06 inst/jss_hyper3_real.Rnw
more reworking

% -*- mode: noweb; noweb-default-code-mode: R-mode; -*-
\documentclass[article]{jss}
\usepackage{booktabs}
\usepackage{units}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{wrapfig}

%  Process with:
%  
%  Rscript -e 'library("knitr") ; knit("jss_hyper3_real.Rnw")' && pdflatex jss_hyper3_real && open jss_hyper3_real.pdf
%
%  (also the Makefile has an entry for this)

\author{Robin K. S. Hankin\\University of Stirling}
\title{Generalized Bradley-Terry models}
\Plainauthor{Robin K. S. Hankin} %% comma-separated
\Plaintitle{Generalized Bradley-Terry models}

%% an abstract and keywords
\Abstract{

  The \code{hyper2} package provides functionality to work with
  extensions of the Bradley-Terry probability model such as reified
  entities (monsters) that account for known enhancements in
  competitive ability.  Here, I present a generalization of
  \code{hyper2} in which Bradley-Terry strengths may be enhanced using
  a multiplicative factor $\lambda$.  Generally, ordinary
  Bradley-Terry is recovered if $\lambda=1$.  This approach offers
  multiple advantages over reified Bradley-Terry, and several examples
  are given.
}
\Keywords{Plackett-Luce,
  Bradley-Terry, Mann-Whitney} \Plainkeywords{Plackett-Luce,
  Bradley-Terry, Mann-Whitney}

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Robin K. S. Hankin\\
  University of Stirling\\
  E-mail: \email{hankin.robin@gmail.com}\\
  URL: \url{https://academics.aut.ac.nz/robin.hankin}
}

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand\headercell[1]{%
   \smash[b]{\begin{tabular}[t]{@{}c@{}} #1 \end{tabular}}}
\newcommand{\draw}{\ensuremath{\mathfrak{D}}}
\newcommand{\cdraw}{\ensuremath{\mathfrak{C}}}
\newcommand{\white}{\ensuremath{\mathfrak{W}}}
\newcommand{\home}{\ensuremath{\mathfrak{H}}}
\newcommand{\oo}{\vphantom{W^{W^{W^W}}}}

\begin{document}

<<setup, cache=FALSE,echo=FALSE>>=
source("loadlibrary.R",echo=FALSE)  # to avoid stupid messages
opts_chunk$set(cache=TRUE, autodep=TRUE)
options(prompt = "R> ")
@

\section{Introduction}

\begin{wrapfigure}{r}{0.1\textwidth}
    \centering
    \includegraphics[width=0.1\textwidth]{hyper2.png}\\
\end{wrapfigure}

The Bradley-Terry model for datasets involving paired comparisons has
wide uptake in the {R} community.  The canonical problem is to
consider~$n$ players who compete against one another; the basic
inference problem is to estimate
numbers~$\mathbf{p}=\left({p_1,\ldots,p_n}\right)$, $p_i\geqslant 0$,
$\sum p_i=1$ so that
$\Prob(\mbox{$i$~beats~$j$})=\frac{p_i}{p_i+p_j}$.  Information about
the~$p_i$ may be obtained from the results of paired comparisons
between the players.

One generalization of the Bradley-Terry model is due
to~\citet{luce1959}, in which the probability of competitor~$i$
winning in a field of~$j=1,\ldots, n$ is $\frac{p_i}{p_1+\cdots
  +p_n}$.  It is then natural to quantify some ephemeral enhancement
to competitor~$i$ using a multiplicative factor $\lambda$.  Then
$\Prob(\mbox{$i$ wins})=\frac{\lambda p_i}{p_1+\cdots+\lambda
  p_i+\cdots+p_n}$.  Examples of such enhancements include the
home-ground advantage in football and rugby, or playing white in
chess~\citep{hankin2020}.  Note that $\lambda < 1$ would correspond to
a decrease in competitive ability.

Further, one might follow~\cite{hankin2017_rnw} and consider teams of
players.  In doubles tennis, for example, $\left\lbrace
p_1,p_2\right\rbrace$ might play $\left\lbrace p_3,p_4\right\rbrace$
and win with probability
$\frac{p_1+p_2}{p_1+p_2+p_3+p_4}$~\citep{hankin2010}.  It is natural to
combine this idea with that of a multiplicative factor.  For example,
suppose $p_1$ has a special move against $p_3$, and $p_4$ has just
been given a performance-enhancing drug, we might have a likelihood
functions proportional to $\frac{\lambda p_1+ p_2}{\lambda
  p_1+p_2+p_3+\mu p_4}$.  Then the effectiveness of $p_3$'s special
move and the enhancement afforded by the drug are measured by
$\lambda$ and $\mu$ respectively with values $>1$ corresponding to a
positive effect.

In this paper I present a generalization of the {\tt hyper2} suite for
Bradley-Terry likelihoods to accommodate general multiplicative
factors.  The most general likelihood function would be

\begin{equation}
  \prod_{i=1}^N
  \sum_{j=1}^n
 \left(\lambda_{ij}p_i\right)^{w_i}
  \end{equation}

where the $p_i$ are traditional Bradley-Terry strengths and the
$\lambda_{ij}\geqslant 0$ are multiplicative factors that account for
ephemeral or temporary changes in competitive ability.  In practice,
almost all the $\lambda_{ij}$ are zero.  To fix ideas, consider the
data presented in Table~\ref{oldfirm}.  Ignoring draws for the moment,
we have a likelihood proportional to

\begin{table}
\centering
\begin{tabular}{@{} *{5}{c} @{}}
\headercell{\small{Plays at home,}\\ \small{wins}} &
\multicolumn{4}{c@{}}{\small{Plays away, loses}}\\
\cmidrule(l){2-5}
&   Rangers     &    Celtic & Falkirk & Livingston\\
\midrule
   Rangers      &   NA &    81  &    69   &      15\\
   Celtic       &   76 &    NA  &    61   &      15\\
   Falkirk      &   17 &    17  &    NA   &       6\\
   Livingston   &    2 &     2  &     3   &      NA\\
\end{tabular}
\\
\rule{1mm}{6mm}
\\
\begin{tabular}{@{} *{5}{c} @{}}
\headercell{\small{Plays at home,}\\ \small{draws}} &
\multicolumn{4}{c@{}}{\small{Plays away, draws}}\\
\cmidrule(l){2-5}
&  Rangers     &    Celtic & Falkirk & Livingston\\
\midrule
   Rangers     &       NA  &   44 &     14   &       2\\
   Celtic      &       47  &   NA &     10   &       3\\
   Falkirk     &       12  &   19 &     NA   &       4\\
   Livingston  &       4   &   4  &     4    &      NA\\
   \end{tabular}
\\
\rule{1mm}{6mm}
\\
\begin{tabular}{@{} *{5}{c} @{}}
\headercell{\small{Plays at home,}\\ \small{loses}} &
\multicolumn{4}{c@{}}{\small{Plays away, wins}}\\
\cmidrule(l){2-5}
&  Rangers     &    Celtic & Falkirk & Livingston\\
\midrule
   Rangers     &   NA  &  42  &    2  &       0\\
   Celtic      &   46  &  NA  &    8  &       0\\
   Falkirk     &   57  &  43  &   NA  &       8\\
   Livingston  &   12  &  11  &    9  &      NA\\
  \end{tabular}
\caption{The \label{oldfirm} dominance of the Old Firm: 669 matches
  between Rangers, Celtic, Falkirk and Livingston arranged by home
  wins, draws, and home losses.  Data from {\tt soccerworld}}
\end{table}

\begin{equation}
  \left(\frac{\lambda p_R}{\lambda p_R + p_C}\right)^{81}\cdot
  \left(\frac{\lambda p_R}{\lambda p_R + p_F}\right)^{69}\cdots
  \left(\frac{\lambda p_L}{\lambda p_L + p_F}\right)^{81}\cdots
  \left(\frac{p_C}{p_C + \lambda p_R}\right)^{42}\cdots
    \left(\frac{p_F}{p_F + \lambda p_L}\right)^{9}
\end{equation}

In {\tt hyper3} idiom, a term such as $\lambda p_R + p_C$ is
represented as a named numeric vector.  For example, if the home
strength were $\lambda = 1.8$ then the first bracket would be {\tt
  c(Rangers=1.8, Celtic=1)} with an associated power of $-81$.  The
computational methodology is a generalization of that used
in~\cite{hankin2024_hyper3} in which the multiplicative factors are
known non-negative integers.

\subsection{The package in use}
Package idiom for creating a likelihood function for this dataset
would look something like this: Considering only Rangers and Celtic:
Rangers at home:

<<defineoldfirm, cache=TRUE>>=
oldfirm <- function(lambda){
    H <- hyper3(pnames = c("Rangers", "Celtic"))
    H[c(Rangers = lambda)            ] %<>% add(81)
    H[c(Celtic = 1)                  ] %<>% add(42)
    H[c(Rangers = lambda, Celtic = 1)] %<>% subtract(81 + 42)

    H[c(Celtic = lambda)             ] %<>% add(76)
    H[c(Rangers = 1)                 ] %<>% add(46)
    H[c(Celtic = lambda, Rangers = 1)] %<>% subtract(46 + 76)
    return(H)
}
@

Above we see package idiom for creating a log-likelihood function.  We
first create an empty {\tt hyper3} object with an appropriate names
attribute.  The next three lines, using \pkg{magrittr} assignment
pipes, give the results for Rangers at home, and the next three lines
give the result for Celtic at home.  It is then easy to create a
bespoke log-likelihood function for Rangers-Celtic matches:

<<useoldfirm, cache = TRUE>>=
oldfirm(1.88)
@

We see a log-likelihood function for all $82+42+76+46=245$
Rangers-Celtic matches that were not score draws, conditional on
assumed value $\lambda=1.88$.  We may make inferences about the true
value of $\lambda$ using profile support:

<<profilelikeoldfirm, cache=TRUE>>=
f <- function(lambda){maxp(oldfirm(lambda),give=1)$`Log-likelihood`}
lam <- seq(from=0.9, to=3, len=23)
logLike <- sapply(lam,f)
@

Thus {\tt f()} returns the support at the evaluate, conditional on a
supplied value of $\lambda$.  We may use this to create a profile
likelihood:

\begin{figure}[htbp]
  \begin{center}
<<plotproflikeoldfirm, echo=FALSE>>=
plot(lam,logLike - max(logLike),type='b')
abline(v=1,col='gray')
segments(x0=1.785,y0=0.1,y1=-1,col='red')
abline(h = c(0,-2),col='gray')
@
\caption{A\label{profilesupportlambda} profile support function for the
  home-team advantage $\lambda$.  Vertical gray line shows
  $H_0\colon\lambda =1$, horizontal lines show the
  two-units-of-support limits; small red bar shows
  $\hat{\lambda}\simeq 1.785$}
\end{center}
\end{figure}

<<optimoldfirm, cache=TRUE>>=
(lambda_max <-  optimize(f,c(1.5,2.0),maximum=TRUE)$maximum)
@

<<mleoldfirm,cache=TRUE>>=
maxp(oldfirm(lambda_max))
@

Compare with the theoretical value of $\widehat{p_{\mbox{\tiny
      C}}}=\frac{\sqrt{42\cdot 76}}{\sqrt{81\cdot 46}+\sqrt{42\cdot
    76}}\simeq 0.4807$, $\widehat{p_{\mbox{\tiny R}}}
=\frac{\sqrt{81\cdot 46}}{\sqrt{81\cdot 46}+\sqrt{42\cdot 76}}\simeq
0.5193$, and $\widehat{\lambda}=\frac{\sqrt{76\cdot 81}}{\sqrt{46\cdot
    42}}\simeq 1.785$.

\section{Score draws}

About 22\% of matches in Table~\ref{oldfirm} end in a draw and {\tt
  hyper2} formalism may be used to assess this.  Consider the
likelihood function described in Table~\ref{windrawloss}; this has a
number of desirable features.  Firstly, if we consider two players of
identical strengths and nullify the home ground effect (that is,
$\lambda=1$) then the probability of a draw is $\frac{2D}{2+2D}$; the
odds of a draw is simply $D$, regardless of team strength.  Secondly,
considering only won (that is, not score-drawn) games, the probability
of a home win is $\frac{\lambda}{1+\lambda}$, that is, the odds of a
home win is simply $\lambda$, independently of the player
strength. Such arguments give direct operational significance to
$\lambda$ and $D$.

\begin{table}
\centering
\begin{tabular}{l|c|c}\\&
{$i$ plays at home} &
{$i$ plays away} \\
  \hline
  $\Prob{(\mbox{$i$ wins})}$                             &
  $\frac{\lambda p_i}{\lambda p_i + D(p_i + p_j) + p_j}$  &
  $\frac{p_i}{p_i + D(p_i + \lambda p_j) + p_j}$          \\
  $\Prob{(\mbox{draw})}$                                 &
  $\frac{D(p_i + p_j)}{\lambda p_i+D(p_i+p_j) + p_j}$     &
  $\frac{D(p_i + p_j)}{p_i + D(p_i+p_j) + \lambda p_j}$   \\
  $\Prob{(\mbox{$i$ loses})}$                            &
  $\frac{p_j}{\lambda p_i + D(p_i+p_j) + p_j}$            &
  $\frac{\lambda p_j}{p_i + D(p_i + p_j) + \lambda p_j}$  \\
\end{tabular}
\caption{Probabilities of a win, draw, loss for the two teams under a
  multiplicative Bradley-Terry model\label{windrawloss} and a lambda thing}
\end{table}

We are now in a position to consider the entirety of
Table~\ref{oldfirm}.  Bespoke function \code{home\_draw\_away3()}
converts data in the form of Table~\ref{oldfirm} to a \code{hyper3}
likelihood function following Table~\ref{windrawloss}; it takes
numerical values for the home advantage $\lambda$ and proclivity to
draw $D$:

<<eg_RCLF3_lf, cache=TRUE>>=
home_draw_away3(RCLF3_table, lambda=1.7, D=0.3)
@

Observe that the resulting likelihood function has two distinct types
of parameters: the plain Bradley-Terry strengths $\mathbf{p}=(p_R,
p_C,p_L,p_F)$, with unit sum; but also the auxiliary parameters
$\lambda,D$.  To make inferences about these parameters we calculate a
profile likelihood function: conditional on $\lambda,D$ we maximize
the likelihood over possible values of $\mathbf{p}$.

<<defineF>>=
f <- function(v){
    H <- home_draw_away3(RCLF3_table, lambda = v[1], D=v[2])
    maxp(H,give=1)$`Log-likelihood`
}
@

Thus \code{f()} gives us a likelihood function for
$\mathbf{v}=(\lambda,D)$.  We may find the evaluate using
\code{optim()}:

<<findmaxlambdaD, cache=TRUE>>=
maxv <- optim(par=c(2, 0.3), fn=f, control=list(fnscale = -1))$par
maxv
@

Then

<<showevaluate, cache=TRUE>>=
maxp(home_draw_away3(RCLF3_table, lambda = maxv[1], D=maxv[2]))
@

Such estimates have rich predictive power.  For example, if Rangers
were to play Celtic with both teams away [thus nullifying the home
  ground advantage: $\lambda=1$ in Table~\ref{windrawloss}], the
probability of win/draw/loss for Rangers would be
$\left(\frac{p_R}{(1+D)(p_R+p_C)};\frac{D}{1+D};\frac{p_R}{(1+D)(p_R+p_C)}\right)$,
estimated at $(0.39;0.27;0.34)$.

\begin{figure}[htbp]
  \begin{center}
<<makecontour, cache=TRUE>>=
n <- 20
lambda <- seq(from=1, to=3, len=n)
D <- seq(from=0.1, to=0.6, len=n)
V <- as.matrix(expand.grid(lambda, D))
LL <- apply(V, 1, f)
LL <- pmax(-20, LL - max(LL))
contour(lambda, D, matrix(LL, n, n),levels = -2*(0:9), xlab=expression(lambda),ylab=expression(D))
points(maxv[1], maxv[2], pch=16, col='red')
@
\caption{Contours of support\label{contourlambdaD} evaluated over
  a grid.  Maximum likelihood estimate for $(\lambda, D)$ shown as
  a red dot at the evaluate}
  \end{center}
\end{figure}

<<testlambdazero, cache=TRUE, include=FALSE>>=
jj <- optimize(f=function(D){f(c(lambda=1,D))}, interval=c(0.1, 0.4), maximum=TRUE)
@

Further, we may calculate a support function over the
$(\lambda,D)$-plane, shown in figure~\ref{contourlambdaD}.  For
example, we may test the hypothesis that $\lambda=1$ [corresponding to
  the right hand side of figure~\ref{contourlambdaD}] for which the
support is about \Sexpr{round(f(maxv) - jj$objective,2)}, far
exceeding the two units of support per degree of freedom
criteron~\citep{edwards1992}.  Alternatively, we can observe that this
lies in the tail region of its asymptotic null $\chi^2_4$ distribution
with a $p$-value of about \Sexpr{signif(pchisq(f(maxv) -
  jj$objective,lower.tail=FALSE,df=2),digits=2)}.

\section{Non-independence in successive arrivals}

In the following, I use the ``race'' metaphor: five runners take part
in a race and arrive in order: $a\succ b\succ c\succ d\succ e$.
Consider the following Plackett-Luce likelihood function for this
observation:

\begin{equation}
\frac{p_a}{p_a + p_b + p_c + p_d + p_e}\cdot
\frac{p_b}{      p_b + p_c + p_d + p_e}\cdot
\frac{p_c}{            p_c + p_d + p_e}\cdot
\frac{p_d}{                  p_d + p_e}\cdot
\frac{p_e}{                        p_e}
\end{equation}

Note the use of multiplication between the terms, indicating
conditional independence.  We may use the multiplicative
generalization of Bradley-Terry strengths to introduce an element of
non-independence between the terms.  There are many ways to do this,
but one simple case would be to define equivalence classes of the
competitors with each equivalence class comprising mutually supporting
runners.  Equivalence classes of size one correspond to unsupported
runners.  The metaphor would be that a runner who has finished the
race is able to support other members of his equivalence class by
cheering his teammates, boosting their performance.  The \pkg{hyper2}
package includes function \code{cheering()} which implements this
functionality.

As an example, consider figure~\ref{tikzabcde}.  This shows a partial
probability tree diagram for some of the $5!=120$ possible order
statistics.  The standard Plackett-Luce likelihoods have been modified
to account for two groups of mutually supporting runners:
$\left\lbrace a,b,c\right\rbrace$ with support term $\lambda$, and
$\left\lbrace d,e\right\rbrace$ with support term $\mu$.  From START,
the first runner to cross the finishing line has standard
Plackett-Luce probability.  Taking the top path as an example, we see
that the likelihood function for $a\succ b\succ c\succ d\succ e$ would
be

\begin{figure}[htbp]
  \begin{center}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw[thick] (0,0) -- (2, 3) -- (4.8, 3);
\draw[thick] (0,0) -- (2, 2) -- (4.8, 2);
\draw[thick] (0,0) -- (2, 1) -- (4.8, 1);
\draw[thick] (0,0) -- (2, 0) -- (4.8, 0);
\draw[thick] (0,0) -- (2,-1) -- (4.8,-1);
\node at (0,2.5) (eq1) {$\underbrace{\left\lbrace a,b,c\right\rbrace}_{\lambda}\underbrace{\left\lbrace d,e\right\rbrace}_{\mu}$};

\node at (4.8, 2.8) {$a$};
\node at (4.8, 1.8) {$b$};
\node at (4.8, 0.8) {$c$};
\node at (4.8,-0.2) {$d$};
\node at (4.8,-1.2) {$e$};

\node at (3.4, 3.4)  {$\frac{p_a}{p_a+p_b+p_c+p_d+p_e}$};
\node at (3.4, 2.4)  {$\frac{p_b}{p_a+p_b+p_c+p_d+p_e}$};
\node at (3.4, 1.4)  {$\frac{p_c}{p_a+p_b+p_c+p_d+p_e}$};
\node at (3.4, 0.4)  {$\frac{p_d}{p_a+p_b+p_c+p_d+p_e}$};
\node at (3.4,-0.6)  {$\frac{p_e}{p_a+p_b+p_c+p_d+p_e}$};

\fill (4.8, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill (4.8, 2) circle[radius=2pt];  % terminal node
\fill (4.8, 1) circle[radius=2pt];  % terminal node
\fill (4.8, 0) circle[radius=2pt];  % d finishes; paths abce
\fill (4.8,-1) circle[radius=2pt];  % terminal node

\draw[thick] (4.8,3) -- (5.8,5) -- (8.2,5);
\draw[thick] (4.8,3) -- (5.8,4) -- (8.2,4);
\draw[thick] (4.8,3) -- (5.8,3) -- (8.2,3);
\draw[thick] (4.8,3) -- (5.8,2) -- (8.2,2);

\newcommand{\la}[1]{\lambda{#1}}
\newcommand{\ld}[1]{\mu    {#1}}
\node at (7, 5.4)  {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (7, 4.4)  {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (7, 3.4)  {$\frac{p_d     }{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (7, 2.4)  {$\frac{p_e     }{\la{p_b}+\la{p_c}+p_d+p_e}$};

\fill (8.2,5) circle[radius=2pt];  % ab finishes; paths cde
\fill (8.2,4) circle[radius=2pt];  % ab finishes; paths cde
\fill (8.2,3) circle[radius=2pt];  % ab finishes; paths cde
\fill (8.2,2) circle[radius=2pt];  % ab finishes; paths cde

\node at (8.2,4.8) {$b$};
\node at (8.2,3.8) {$c$};
\node at (8.2,2.8) {$d$};
\node at (8.2,1.8) {$e$};

\draw[thick] (8.2,5) -- (8.9,6) -- (10.6,6);
\draw[thick] (8.2,5) -- (8.9,5) -- (10.6,5);
\draw[thick] (8.2,5) -- (8.9,4) -- (10.6,4);

\node at (9.5, 6.4) {$\frac{\la{p_c}}{\la{p_c}+p_d+p_e}$};
\node at (9.5, 5.4) {$\frac{p_d  }{\la{p_c}+p_d+p_e}$};
\node at (9.5, 4.4) {$\frac{p_e  }{\la{p_c}+p_d+p_e}$};

\fill (10.6, 6) circle[radius=2pt];  % abc finishes; paths de
\node at (10.6,5.8) {$c$};
\draw[thick] (10.6, 6) -- (11.2,7) -- (12.2,7);
\fill (12.2,7) circle[radius=2pt];
\node[anchor = west] at (12.2,7) {$a\succ b\succ c\succ d\succ e$};
\node at (11.8, 7.4) {$\frac{p_d}{p_d+p_e}$};

\draw[thick] (10.6, 6) -- (11.2,6.2) -- (12.2,6.2);
\fill (12.2,6.2) circle[radius=2pt];
\node[anchor = west] at (12.2,6.2) {$a\succ b\succ c\succ e\succ d$};
\node at (11.8, 6.6) {$\frac{p_e}{p_d+p_e}$};

\fill (10.6, 5) circle[radius=2pt];
\node at (10.6,4.8) {$d$};

\draw[thick] (10.6, 5) -- (11.2,5.3) -- (12.2, 5.3);
\fill (12.2,5.3) circle[radius=2pt];
\node[anchor = west] at (12.2,5.3) {$a\succ b\succ d\succ c\succ e$};
\node at (11.8, 5.7) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_e}}$};

\draw[thick] (10.6, 5) -- (11.2,4.7) -- (12.2, 4.7);
\fill (12.2,4.7) circle[radius=2pt];
\node[anchor = west] at (12.2,4.7) {$a\succ b\succ d\succ e\succ c$};
\node at (11.8, 5.0) {$\frac{\ld{p_e}}{\la{p_c}+\ld{p_e}}$};

\fill (10.6, 4) circle[radius=2pt];
\node at (10.6,3.8) {$e$};
\draw[thick] (10.6, 4) -- (11.2,3.9) -- (12.2, 3.9);
\fill (12.2,3.9) circle[radius=2pt];
\node[anchor = west] at (12.2,3.9) {$a\succ b\succ e\succ c\succ d$};
\node at (11.8, 4.3) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_d}}$};

\draw[thick] (10.6, 4) -- (11.2,3.1) -- (12.2,3.1);
\fill (12.2,3.1) circle[radius=2pt];
\node at (11.8, 3.5) {$\frac{\ld{p_d}}{\la{p_c}+\ld{p_d}}$};
\node[anchor = west] at (12.2,3.1) {$a\succ b\succ e\succ d\succ c$};

\draw[thick] (4.8, 0) -- (5.8,1) -- (8.2,1);
\fill (12.2,1) circle[radius=2pt];
\node[anchor = west] at (12.2,1) {$d\succ a\succ c\succ e\succ b$};
\node at (7, 1.4)  {$\frac{p_a}{p_a+p_b+p_c+\ld{p_e}}$};

\draw[thick] (4.8, 0) -- (5.8,0) -- (8.2,0);
\fill (12.2,0) circle[radius=2pt];
\node[anchor = west] at (12.2,0) {$d\succ e\succ a\succ b\succ c$};
\node at (7, 0.4)  {$\frac{p_b}{p_a+p_b+p_c+\ld{p_e}}$};

\draw[thick] (4.8, 0) -- (5.8,-1) -- (8.2,-1);
\node at (7, -0.6)  {$\frac{p_c}{p_a+p_b+p_c+\ld{p_e}}$};

\draw[thick] (4.8, 0) -- (5.8,-2) -- (8.2,-2);
\node at (7, -1.6)  {$\frac{\ld{p_e}}{p_a+p_b+p_c+\ld{p_e}}$};

\fill (8.2, -2) circle[radius=2pt];  % de finishes; paths abc
\fill (8.2, -1) circle[radius=2pt];  % de finishes; paths abc
\fill (8.2,  0) circle[radius=2pt];  % de finishes; paths abc

\node at (8.2, -2.2) {$e$};
\node at (8.2, -1.2) {$c$};
\node at (8.2, -0.2) {$b$};
\node at (8.2,  0.8) {$a$};

\draw[thick] (8.2, -2) -- (8.9,-1) -- (10.6,-1);
\draw[thick] (8.2, -2) -- (8.9,-2) -- (10.6,-2);
\draw[thick] (8.2, -2) -- (8.9,-3) -- (10.6,-3);

\fill (12.2,-1) circle[radius=2pt];
\node[anchor = west] at (12.2,-1) {$d\succ e\succ a\succ c\succ b$};

\node at (9.5, -0.6) {$\frac{p_a}{p_a+p_b+p_c}$};
\node at (9.5, -1.6) {$\frac{p_b}{p_a+p_b+p_c}$};
\node at (9.5, -2.6) {$\frac{p_c}{p_a+p_b+p_c}$};

\fill (10.6, -1) circle[radius=2pt];  % dea finishes; paths bc
\fill (10.6, -2) circle[radius=2pt];  % deb finishes; paths ac
\fill (10.6, -3) circle[radius=2pt];  % dec finishes; paths ab
\node at (10.6,-1.2) {$a$};
\node at (10.6,-2.2) {$b$};
\node at (10.6,-3.2) {$c$};

\draw[thick] (10.6, -1) -- (11.2,-0) -- (12.2,0);

\node at (11.8, 0.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}}$};
\draw[thick] (10.6, -1) -- (11.2,-1) -- (12.2,-1);

\node at (11.8, -0.6) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}}$};

\draw[thick](10.6, -2) -- (11.2,-1.8) -- (12.2,-1.8);
\fill (12.2,-1.8) circle[radius=2pt];
\node[anchor = west] at (12.2,-1.8) {$d\succ e\succ b\succ a\succ c$};
\node at (11.8, -1.4) {$\frac{\la{p_a}}{\la{p_a}+\la{p_c}}$};

\draw[thick] (10.6, -2) -- (11.2,-2.6) -- (12.2,-2.6);
\fill (12.2,-2.6) circle[radius=2pt];
\node[anchor = west] at (12.2,-2.6) {$d\succ e\succ b\succ c\succ a$};
\node at (11.8, -2.2) {$\frac{\la{p_c}}{\la{p_a}+\la{p_c}}$};

%\draw[thick] (10.6, -3) -- (11.2,-2.9);
%\fill (12.2,-2.9) circle[radius=2pt];
%\node at (11.8, -2.9) {$\frac{\la{p_a}}{\la{p_a}+\la{p_b}}$};

\draw[thick] (10.6, -3) -- (11.2,-3.4) -- (12.2, -3.4);
\fill (12.2,-3.4) circle[radius=2pt];
\node[anchor = west] at (12.2,-3.4) {$d\succ e\succ c\succ b\succ a$};

\node at (11.8, -3) {$\frac{\la{p_b}}{\la{p_a}+\la{p_b}}$};

\fill (8.2, 1) circle[radius=2pt];  % da finishes; paths bce

\draw[thick] (8.2,1) -- (8.9,2) -- (10.6,2);
\draw[thick] (8.2,1) -- (8.9,1) -- (10.6,1);
\draw[thick] (8.2,1) -- (8.9,0) -- (10.6,0);

\node at (9.5, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 1.4) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 0.4) {$\frac{p_e     }{\la{p_b}+\la{p_c}+p_e}$};
\fill (10.6, 2) circle[radius=2pt];  % da finishes; paths bce
\fill (10.6, 1) circle[radius=2pt];  % da finishes; paths bce
\fill (10.6, 0) circle[radius=2pt];  % da finishes; paths bce

\node at (10.6, 1.8) {$b$};
\node at (10.6, 0.8) {$c$};
\node at (10.6,-0.2) {$e$};

\draw[thick] (10.6, 1) -- (11.2,2) -- (12.2,2);
\node at (11.8, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\ld{p_e}}$};
\fill (12.2,2) circle[radius=2pt];
\node[anchor = west] at (12.2,2) {$d\succ a\succ c\succ b\succ e$};
\draw[thick] (10.6, 1) -- (11.2,1) -- (12.2,1);
\node at (11.8, 1.4) {$\frac{\ld{p_e}}{\la{p_b}+p_e}$};

\end{tikzpicture}
\caption{Partial probability\label{tikzabcde} tree for five
  competitors $a$-$e$ with two mutually supporting subsets
  $\left\lbrace a,b,c\right\rbrace$ and $\left\lbrace d,
  e\right\rbrace$ having support $\lambda$, $\mu$ respectively}
\end{center}
\end{figure}

\begin{equation}\label{asuccb}
\frac{p_a}{p_a + p_b + p_c + p_d + p_e}\cdot
\frac{\lambda p_b}{\lambda p_b + \lambda p_c + p_d + p_e}\cdot
\frac{\lambda p_c}{\lambda p_c + p_d + p_e}\cdot
\frac{p_d}{p_d + p_e}
\end{equation}

In equation \ref{asuccb}, the first term is standard Plackett-Luce: at
this point, no competitor has finished and cheering effects are
absent.  The second term reflects the fact that competitors $b$ and
$c$ are supported by competitor $a$, who has by this point finished
the race and is supporting his teammates.

By contrast, the likelihood function for observation $d\succ a\succ
c\succ b\succ e$ would be

\begin{equation}
\frac{p_d}{p_a + p_b + p_c + p_d + p_e}\cdot
\frac{p_a}{p_a + p_b + p_c + \mu p_e}\cdot
\frac{\lambda p_c}{\lambda p_b + \lambda p_c + p_e}\cdot
\frac{\lambda p_b}{\lambda p_b + \mu p_e}
\end{equation}

where this likelihood function reflects the mutual support for
equivalence class $\left\lbrace d,e\right\rbrace$.  Note that the
final term reflects the fact that competitors $b$ and $e$ have their
support active when vying for fourth place, as members of both their
teams have finished at this point.

This probability model could easily be modified to account for
specific circumstances.  The cheering effect could be asymmetrical
(with $a$ helping $b$ but not vice-versa, for example).  The effect
might operate only on certain ordered pairs, following a directed
graph (cyclic or acyclic) instead of equivalence classes.  Or perhaps
the effect might have a finite lifetime: if $a$ places
$n^\mathrm{th}$, then the cheering effect is active only for
competitors placing $(n+r)^\mathrm{th}$ or above, for some fixed $r$.
There is no reason that $\lambda$ could not be less than one, which
would indicate that some competitors inhibit, rather than help,
others.

The probability model of Figure~\ref{tikzabcde} is applied, in a
simple one-parameter form, to two datasets: one a set of preferences
of different types of sushi, and one in the field of e-sports.

\section{Red bus, blue bus}

The red bus-blue bus problem was originally stated by
\cite{chipman1960} and discussed by \cite{davidson1970} and
\cite[p153]{alvo2014}.  
by \cite{tverskey1972}.
The canonical problem is as follows. A traveller has a choice of going
to work by different transport methods.  Here, we can assume that the
traveller has strong but unknown views on different forms of transport
which might incorporate a random component.  If considering whether to
use a car or a bus, the colour of the bus is of no importance to him.
Such preferences are problematic for Bradley-Terry: If our subject is
indifferent between cars and buses, say, $p_{C} = p_{RB}$ and also
indifferent to colour of bus, so $p_{RB}=p_{BB}$, the following
paradox occurs.  Bradley-Terry strengths must be equal so $p_{C} =
p_{RB} = p_{BB}=\frac{1}{3}$.  Thus $\Prob(C| \left\lbrace
C,RB\right\rbrace)=\frac{1}{2}$.  Now we introduce a blue bus as an
option.  Any realistic choice model will have $\Prob(C | \left\lbrace
C,RB,BB\right\rbrace)=\frac{1}{2}$ but Bradley-Terry gives us
$\Prob(C| \left\lbrace C,RB\right\rbrace)=\frac{1}{3}$.  Various
solutions to this problem have been proposed.  \cite{tverskey1972},
for example, presents an approach involving overlapping aspects;

If we ask him to
rank his options, it is highly probable that he will put RB and BB
consecutively (because they are essentially indistinguishable).  Can
we quantify the strength of this effect?  To do this, we define a
bespoke function \code{RB\_BB\_LF()} which returns a \code{hyper3}
log-likelihood function corresponding to repeated observations of our
commuter's reported ranks for the five options:

<<defineredbuslf>>=
`RB_BB_LF` <- function(lambda){
    ec <- c(C=1, T=2, RB=3, BB=3, W=4) 
    h <- c(1, 1, lambda, 1)           
    (
        cheering3(v=c("RB", "BB", "C" , "T", "W"), e=ec, h=h)*3 + 
        cheering3(v=c("BB", "RB", "T" , "C", "W"), e=ec, h=h)*2 + 
        cheering3(v=c("T" , "BB", "RB", "C", "W"), e=ec, h=h)*5 + 
        cheering3(v=c("W" , "BB", "RB", "T", "C"), e=ec, h=h)*4 + 
        cheering3(v=c("C" , "RB", "BB", "W", "T"), e=ec, h=h)*3 + 
        cheering3(v=c("BB", "C" , "RB", "T", "W"), e=ec, h=h)*3
    )
}
@

Above, we see from the function body that he reported $RB\succ BB\succ
C\succ T\succ W$ three times [first row], and so on; perhaps his
ranking depends on the weather or how tired he is on any given day.
Observe that in almost every case he ranks RB and BB consecutively.
Function \code{RB_BB_LF()} takes argument \code{lambda} that
quantifies the perceived similarity between RB and BB in the same way
as function \code{oldfirm()} above.  It is straightforward to find a
profile support function for $\lambda$,
Figure~\ref{profilesupportlambdaRBBB}, and formally maximize the
likelihood:

\begin{figure}[htbp]
\begin{center}
<<plotproflike>>=
o <- function(lambda){maxp(RB_BB_LF(lambda), give=1)$`Log-likelihood`} 
lambda <- exp(seq(from=log(0.9), to=log(47), len=17)) 
L <- sapply(lambda, o)
plot(log(lambda), L-max(L), type="b")
abline(h=c(0, -2))
abline(v=0)
@ 
\caption{A\label{profilesupportlambdaRBBB} profile support function
  for the red bus, blue bus parameter $\lambda$.  Vertical line
  indicates null of $\lambda=1$, horizontal lines show maximum support
  and $-2$}
\end{center}
\end{figure}

<<label=maxlikesim, cache=TRUE>>=
(osup <- optimize(o, c(10, 40), maximum=TRUE))
@

So a likelihood ratio test of the null that $\lambda=1$ would be:

<<label=LRT_RB, cache=TRUE>>=
(suppdiff <- o(osup$maximum) - o(1))
@

Easily satisfying \citeauthor{edwards1992}'s two-units-of-support
criterion; alternatively, we could observe that this observation lies
in the tail region of its asymptotic $\chi^2_1$ distribution (Wilks's
theorem), with a $p$-value of \Sexpr{signif(pchisq(suppdiff*2, df=1,
  lower.tail=FALSE),3)}.

\subsection{Sushi}

We now apply the ideas above to a dataset obtained from the {\tt
  preflib} library~\citep{MaWa13a} in which ten different types of
sushi were placed in order of preference by 5000 judges.  Three of the
sushis (maguro, toro, tekka-maki) were made with tuna, and it is
reasonable to view these three types as being exchangeable in much the
same way as the red and blue buses.  To proceed, we need to define
equivalence classes of sushi---one for tuna-based sushis and one for
everything else.  Of course, one can also imagine the opposite
phenomenon whereby the judge decides he has had enough tuna for the
moment and this disincentivises further tuna choices.  This mechanism
would be modelled by having $\lambda<1$.

Implementation in \code{hyper3} formalism is straightforward.  The
dataset used is \code{sushi\_table}, part of the \code{hyper2} package:

<<showsushitable>>=
head(noquote(sushi_table))
@

Thus, from the top line, we see that the first judge placed tamago
first, followed by anago and ika.  The equivalence classes are given
by

<<show_sushi_eq_classes>>=
sushi_eq_classes
@

so class 2 is the tuna group and class 1 everything else.  We can now
use these datapoint to define a bespoke likelihood function, using
$\lambda$ as the argument:

<<defcalculate_sushi_H>>=
make_sushi_H <- function(lambda){
    H <- hyper3()
    for(i in seq_len(nrow(sushi_table))){
        H <- H + cheering3(sushi_table[i,], e=sushi_eq_classes, h=c(1, lambda))
    }
    return(H)
}
@

Then, as before, we may define a profile likelihood function of
$\lambda$, Figure~\ref{profilesupportsushi}.

\begin{figure}[htbp]
  \begin{center}
<<calclikesushi, cache=TRUE>>=
f <- function(lambda){
   H <- make_sushi_H(lambda)
   maxp(H,give=1)$`Log-likelihood`
}
s <- exp(seq(from = log(0.9), to = log(10), len = 12))
L <- sapply(s,f)
plot(log(s),L-max(L),type='b',
     xlab=expression(log(lambda)), ylab='Log-likelihood')
@ 
\caption{A\label{profilesupportsushi} profile support function for the
  exchangeability term $\lambda$.  Vertical gray line shows
  $H_0\colon\lambda =1$, horizontal lines show the
  two-units-of-support limits}
\end{center}
\end{figure}

\subsection{Counterstrike}

\newcommand{\skad}{{\sc Skadoodle}}
\newcommand{\rush}{{\sc RUSH}}
\newcommand{\tarik}{{\sc tarik}}
\newcommand{\stew}{{\sc Stewie2K}}

{\em Counter Strike: Global Offensive (CS:GO)} is a multiplayer
first-person shooter game in which two teams of five compete in an
immersive virtual reality combat environment.  Gameplay is typically
complex, including features such as cooperative teamwork and real-time
strategic fluidity.

Here I analyse a dataset taken from recent e-sports tournament {\em
  ELEAGUE Major: Boston 2018}.  This Valve-sponsored Major featured 24
elite teams competing for a million dollar prize pool; peak concurrent
viewership exceeded $1.3\times 10^6$, the highest in CS:GO history at
the time.  I study the performance of tournament champions {\em
  Cloud9} who, as underdogs, narrowly defeated aggressive European
rivals {\em FaZe Clan} in the final.

Typically when a player is killed, the team suffers a number of
strategic disadvantages; player deaths are thus a quantifiable proxy
for both individual ability and also both teams' tactical execution.
Here I analyse the order statistic for number of deaths of each of the
five {\em Cloud9} players.

<<echo=FALSE>>=
a <- read.table("cloud9_boston_2018.txt", header=TRUE)
a[,-1]
@

From the first line we see that {\em Cloud9} played {\em FaZe Clan} on
the {\em inferno} map, won the match, and \skad\ had the least number
of deaths, followed by \rush, then \tarik, and so on.  Consider two
{\em Cloud9} players: \skad\ and \stew.  These teammates, although
comparable in skill [we fail to reject $H_0\colon p_{\mbox{\tiny\sc
      Skadoodle}}\neq p_{\mbox{\tiny\sc Stewie2k}}$, Plackett-Luce,
  $p=0.28$] have playing styles that stand in dramatic contrast.
Commentators characterise \skad\ as emphasising reactive map control
while \stew\ is more of a disruptive presence engaging in
high-variance play.

Given that these two players offer equally attractive and accessible
targets to the opposing team, it is possible that some tactical
situations might result in a systematic preference for targetting
\skad, and some a preference for killing \stew.  If this is so, the
deaths will be overdispersed and one would have a red bus-blue bus
situation, but with $\lambda<1$.

<<defredbusfunc, echo=FALSE>>=
mk <- function(lam){
  H3 <- hyper3()
  for(i in seq_len(nrow(a))){

    d <- strsplit(a$deathorder[i],", ")[[1]]
    H3 <- H3 + cheering3(d, e=ec, help=c(1,lam))
  }
  return(H3)
}
f <- function(l){maxp(mk(l), give=1)$`Log-likelihood`}
@

<<findopt, cache=TRUE, echo=FALSE>>=
optlam <- optimize(f,c(0.09, 0.14), maximum = TRUE)
evaluate <- optlam$maximum
support <- optlam$objective - f(1)
pvalue <- pchisq(support,df=1,lower.tail=FALSE)
@

<<calcklikelam, cache=TRUE, echo=FALSE>>=
lam <- exp(seq(from = -4,to = log(1.1), len = 23))
L <- sapply(lam, f)
@

<<findinterval, cache=TRUE, echo=FALSE>>=
g <- function(l){f(l) - optlam$objective  + 2}
jj1 <- uniroot(g, interval = exp(c(-4,-3)))
jj2 <- uniroot(g, interval = exp(c(-1,-0)))
lowerlimit <- jj1$root
upperlimit <- jj2$root
@

\begin{figure}[htbp]
  \begin{center}
<<plotlikelam, echo=FALSE>>=
plot(log(lam), L - max(L), type='b')
segments(x0=log(optlam$maximum),y0 = -0.5, y1 = 0.1,col='red')
abline(h=c(0,-2), col = 'gray')
segments(x0 = log(jj1$root), y0 = -2.2, y1 = -1.8, col = 'blue')
segments(x0 = log(jj2$root), y0 = -2.2, y1 = -1.8, col = 'blue')
abline(v = 0, lty = 2)
@ 
\caption{A\label{profsupcounterstrike} profile support function for
  the red bus-blue bus parameter $\log\lambda$ applied to mutually
  supporting equivalence class
  $\left\lbrace\mbox{\skad},\mbox{\stew}\right\rbrace$.  Vertical gray
  line shows $H_0\colon\lambda =1$, for which the support is about
  $\Sexpr{round(support,2)}$ relative to the evaluate
  $\hat{\lambda}\simeq\Sexpr{round(evaluate,3)}$ (small red bar),
  corresponding to a $p$-value of about $\Sexpr{round(pvalue,3)}$.
  Horizontal lines show the two-units-of-support credible interval of
  $(\Sexpr{round(lowerlimit,3)},\Sexpr{round(upperlimit,3)})$, marked
  in blue}
\end{center}
\end{figure}

To assess this, we follow exactly the same protocol as for the sushi
preference dataset but using the deathorder statistic from the dataset
above.  We specify two equivalence classes: one comprising \skad\ and
\stew, and one everyone else.  Figure~\ref{profsupcounterstrike} gives
a support function for $\lambda$ showing, among other things, that we
may reject the hypothesis that $\lambda=1$ in favour of $\lambda<1$ by
a likelihood ratio of about $\Sexpr{round(support,2)}$.  We observe
that this exceeds the two units of support criterion of
\cite{edwards1992} [asymptotic $p$-value
  $\simeq\Sexpr{round(pvalue,3)}$] so we may infer that there is
indeed some systematic mechanism---perhaps a positive feedback
loop---which tends to magnify differences in death statistic.

\section{Conclusions and further work}

Plackett-Luce likelihood functions for rank datasets have been

\bibliography{hyper2}
\end{document}