### 2025-04-02 inst/exponential_BT.Rmd
kickoff for exponential BT

---
title: "exponential_BT"
author: "robin hankin"
date: "2025-04-02"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(hyper2)
library(stringr)
set.seed(0)
```

Basic idea is to have $n$ competitors with BT strengths $\beta, \beta
x,\ldots \beta x^{n-1}$ with $x\in[0,1]$.  The unit sum constraint
gives $\beta=\frac{1-x}{1-x^n}$ but this usually cancels out.  One of
the competitors is the _focal_ competitor with BT strength $\beta
x^{r-1}$ for some $r$ with $1\leqslant r < n$.  Thus $r=1$ means that
the focal competitor is the strongest among the $n$, $r=2$ means he is
the second, and so on.

```{r}
BT <- function(n,x){
    out <- x^(0:(n-1))
    if(x != 1){
        out <- out*(1-x)/(1-x^n)
    } else {
        out <- out/n
    }
    names(out) <- paste0("p",str_pad(1:n,ceiling(log10(n))))
    out
}
BT(4,0.5)
sum(BT(4,0.5))
```

We will consider $n=9$, $x=0.5$, $r=4$ and make repeated observations:

```{r}
(jj <- BT(9,0.5))
o <- tabulate(replicate(1e4,which(rrace(jj) == "p4")))
o
plot(o,type='h')
```

```{r, cache=TRUE}
getprob <- function(n, r, x, N=1e4){
    jj <- BT(n, x)
    wanted <- paste0("p", str_pad(r ,ceiling(log10(n))))
    c(r=r, x=x, setNames(tabulate(replicate(N, which(rrace(jj) == wanted)),nbins=n), paste0("p", 1:n)))
}
rbind(
getprob(9,1,0.5,N=100),
getprob(9,2,0.5,N=100),
getprob(9,3,0.5,N=100),
getprob(9,4,0.5,N=100),
getprob(9,5,0.5,N=100),
getprob(9,6,0.5,N=100),
getprob(9,7,0.5,N=100),
getprob(9,8,0.5,N=100),
getprob(9,9,0.5,N=100)
)

rbind(
getprob(9,1,0.2,N=100),
getprob(9,2,0.2,N=100),
getprob(9,3,0.2,N=100),
getprob(9,4,0.2,N=100),
getprob(9,5,0.2,N=100),
getprob(9,6,0.2,N=100),
getprob(9,7,0.2,N=100),
getprob(9,8,0.2,N=100),
getprob(9,9,0.2,N=100)
)
rbind(
getprob(9,1,0.9,N=100),
getprob(9,2,0.9,N=100),
getprob(9,3,0.9,N=100),
getprob(9,4,0.9,N=100),
getprob(9,5,0.9,N=100),
getprob(9,6,0.9,N=100),
getprob(9,7,0.9,N=100),
getprob(9,8,0.9,N=100),
getprob(9,9,0.9,N=100)
)
```

```{r makeII,cache=TRUE}
r_try <- 1:9
x_try <- seq(from = 0.1, to = 1, by = 0.1)
jj <- as.matrix(expand.grid(r_try, x_try))
II <- t(apply(jj,1,function(v){getprob(9,v[1],v[2],N=1e4)}))
head(II)
```

```{r makesupp, cache=TRUE}
r_true <- 3
x_true <- 0.85
o <- getprob(9, r_true, x_true, 17)  # observations
o[1:2]
o[-(1:2)]

support_func <- function(o,v){
    v <- v[-(1:2)]
    v <- v/sum(v)
    dmultinom(o[-(1:2)], prob=v, log=TRUE)
}
support <- rep(0,nrow(II))
for(i in seq_len(nrow(II))){
    support[i] <- support_func(o,II[i,])
}
```

```{r}
support <- support-max(support)
support <- pmax(support,-10)
support  <- matrix(support,length(r_try),length(x_try))
contour(r_try, x_try, support, xlab="r", ylab="x", levels = -(1:5)*2)
abline(v = r_true)
abline(h = x_true)
abline(v = 1:9, col = 'gray', lty = 3, lwd = 0.3)
filled.contour(r_try, x_try, support, xlab = "r", ylab = "x")
```

```{R showjj}
jj <- cbind(II[,1:2],c(support))
jj[jj[,2] == 1,]
```

#  Analysis from a single observation

We translate $\alpha\in [0,1]$ to a number from 1 to $n$ inclusive
with $\lceil n\alpha\rceil$

Write a likelihood function for 9,12 [that is, came 9th in a class of
12] There are 8 better and 3 worse than the focal student.

We need a likelihood function for the observation (9) of $\alpha,x$.
Let's try again.

```{r ninetwelve, cache=TRUE}
n <- 12
a_try <- seq(from=0.01,to=0.99,len=10)
p_try <- ceiling(n * a_try)
x_try <- seq(from = 0.1, to = 1, by = 0.1)
jj <- as.matrix(expand.grid(p_try, x_try))
II <- t(apply(jj,1,function(v){getprob(n, v[1], v[2], N=1e4)}))
head(II)
tail(II)
S912 <- matrix(II[,11],length(a_try),length(x_try))
```

```{r show912}
S912 <- S912/max(S912)
contour(p_try,a_try,log(S912))
```

# Variable class sizes

Let's try this on the apocryphal education dataset

```{r}
M <- matrix(c(
    9, 7, 2, 3, 2, 8, 3, 4, 9,
    12,17,23,09,13,14,13,12,15
),nrow=2,byrow=TRUE)
M
```