### 2025-04-08 inst/imo.Rmd
consistentification

---
title: "imo"
author: "robin hankin"
date: "2025-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(hyper2)
library(stringr)
```

## R Markdown

```{r}
oecd <- as.character(read.csv("oecd.txt",header=FALSE)$V1)
oecd

oecd <- oecd[!(oecd %in% c("Chile", "Democratic People's Republic of Korea", "Turkiye"))]

years <- 2001:2024

jj <- read.table("imo.txt", header=TRUE)
ranks <- list()
for(i in seq_along(years)){
    ranks[[i]] <- subset(jj, jj$year == years[[i]])$country
}
names(ranks) <- years
```

```{r}
for(i in seq_along(ranks)){
    jj <- ranks[[i]]
    ranks[[i]] <- jj[jj %in% oecd]
}
```

```{r}
uk <- seq_along(ranks)*0
for(i in seq_along(ranks)){
    uk[i] <- which(ranks[[i]] == "Australia")
}
uk
```

```{r checo}
BT <- function(n,x){
    out <- x^(0:(n-1))
    if(x != 1){
        out <- out*(1-x)/(1-x^n)
    } else {
        out <- out/n
    }
    names(out) <- paste0("p",str_pad(1:n,ceiling(log10(n))))
    out
}

getprob <- function(n, r, x, N=1e2){
    jj <- BT(n, x)
    wanted <- paste0("p", str_pad(r ,ceiling(log10(n))))
    c(r=r, x=x, setNames(tabulate(replicate(N, which(rrace(jj) == wanted)),nbins=n), paste0("p", 1:n)))
}
n <- length(oecd)
a_try <- seq(from = 0.01, to = 0.99, len = 20)
x_try <- seq(from = 0.2, to = 1, by = 0.05)
multiple_like_fun <- function(n,           # n competitors
                            a_try, x_try,  # values to try
                            N=1e2){        # N is number of trials
    p_try <- ceiling(n * a_try)
    jj <- as.matrix(expand.grid(p_try, x_try))
    t(apply(jj,1,function(v){getprob(n, v[1], v[2], N)}))
}
```

```{r trymultxy, cache=TRUE}
print(system.time(II <- multiple_like_fun(length(oecd), a_try, x_try, 1e4)))
head(II)
nrow(II)
length(a_try)
length(x_try)
```

```{r calcmultxy, cache=TRUE}
like <- list()
dim(II)
o <- uk # o for observation
o
for(i in seq_along(o)){
    like[[i]] <-  matrix(II[, o[i] + 2], length(a_try), length(x_try))
}
```

```{r showlike}
alllike <- Reduce("*",like)
S <- log(alllike)
S <- S-max(S)
S <- pmax(S, -9.999)
jj <- S
jj[jj < -9.99] <- NA
round(jj,2)
contour(a_try, x_try, S,
        xlab = "BT strength (scaled)", ylab = "x",
        levels = -(0:6)*2, lwd=c(1,4,1,1,1,1,1))
dim(S)
contour(a_try[1:8], x_try[6:17], S[1:8, 6:17], levels = -(0:6)*2, lwd=c(1,4,1,1,1,1,1))
filled.contour(S)
```

```{r}
like[[1]]
```