### 2025-12-03 inst/tennis.Rmd
more work on hyper3

---
title: "Tennis doubles and ghosts"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document discusses a small dataset first presented by
@hankin2010.  There, results from tennis games among four players was
considered:

```
       {p1,p2} vs {p3,p4}   9-2  
       {p1,p3} vs {p2,p4}   4-4  
       {p1,p4} vs {p2,p3}   6-7  
          {p1} vs {p3}     10-14 
          {p2} vs {p3}     12-14 
          {p1} vs {p4}     10-14 
          {p2} vs {p4}     11-10 
          {p3} vs {p4}     13-13

```

It is suspected that $p_1$ and $p_2$ have some form of team cohesion
and play better when paired than when either solo or with other
players.  As the scores show, each player and, apart from the first
line, each doubles partnership is of approximately the same strength.

We will create two likelihood functions: a regular one that assumes
the stength of a doubles partnership is the sum of the strength of the
players; and one that accounts for team cohesion.

```{r defineHhyper2}
H <- hyper2()

## 1&3 vs 2&4, scoreline 4-4:
H[c("p1","p3")] %<>% inc(4)
H[c("p2","p4")] %<>% inc(4)
H[c("p1","p2","p3","p4")] %<>% dec(8)

## 1&4 vs 2&3, scoreline 6-7:
H[c("p1","p4")] %<>% inc(6)
H[c("p2","p3")] %<>% inc(7)
H[c("p1","p2","p3","p4")] %<>% dec(13)

## 1 vs 3, scoreline 10-14:
H["p1"] %<>% inc(10)
H["p3"] %<>% inc(14)
H[c("p1","p3")] %<>% dec(24)

## 2 vs 3, scoreline 12-14:
H["p2"] %<>% inc(12)
H["p3"] %<>% inc(14)
H[c("p2","p3")] %<>% dec(26)

## 1 vs 4, scoreline 10-14:
H["p1"] %<>% inc(10)
H["p4"] %<>% inc(14)
H[c("p1","p4")] %<>% dec(24)

## 2 vs 4, scoreline 11-10:
H["p2"] %<>% inc(11)
H["p4"] %<>% inc(10)
H[c("p2","p4")] %<>% dec(21)

## 3 vs 4, scoreline 13-13:
H["p3"] %<>% inc(13)
H["p4"] %<>% inc(13)
H[c("p3","p4")] %<>% dec(26)
```

Thus `H` is a hyper2 object corresponding to results except the first
line.  We include this in two ways.  Firstly, treating the first
scoreline like the others:

```{r addtohyper2H}
tennis <- H

## 1&2 vs 3&4, scoreline 9-2:
tennis[c("p1","p2")] %<>% inc(9)
tennis[c("p3","p4")] %<>% inc(2)
tennis[c("p1","p2","p3","p4")] %<>% dec(11)

tennis
```

And secondly, accounting for any possible team cohesion by way of a
ghost player `G`:

```{r doghoststuff}
tennis_ghost <- H

## 1&2 vs 3&4 (NB: includes ghost!), scoreline 9-2 (again):
tennis_ghost[c("p1","p2","G")] %<>% inc(9)
tennis_ghost[c("p3","p4")] %<>% inc(2)
tennis_ghost[c("p1","p2","p3","p4","G")] %<>% dec(11)

tennis_ghost
```

standard evaluates:

```{r maximizeghostlike,cache=TRUE}
(tennis_maxp <- maxp(tennis))
(tennis_ghost_maxp <- maxp(tennis_ghost))
```

First test the null of equal strength:

```{r doequalptest,cache=TRUE}
equalp.test(tennis)
```

not significant.  Is the ghost real?

```{r specificptesthere,cache=TRUE}
specificp.gt.test(tennis_ghost,"G",0)  
```

so we can infer that the ghost is indeed real as we reject $H_0\colon
p_G=0$.

We can use the Hessian matrix:

```{r hessianmatrix}
(M <- hessian(tennis,indep(tennis_maxp),border=TRUE))
```

See how the fillup value `p4` takes up the unit sum constraint [it is
defined as $p_4=1-p_1-p_2-p_3$].  Matrix `M` should be
negative-semidefinite:

```{r eigenwork}
eigen(M,TRUE,TRUE)$values # one value is zero to numerical precision; the rest should be <0
```

See how one eigenvalue is zero (to numerical precision) and the rest
are strictly negative, corresponding to the support function having a
well-defined maximum.

### Analysis with `hyper3` {-}

The likelihood for the 9-2 scoreline between $\left\lbrace
p_1,p_2\right\rbrace$ and $\left\lbrace p_3,p_4\right\rbrace$ would
have a likelihood function of

$$
\frac{
\left(\lambda p_1 +\lambda p_2\right)^{9}
\left(p_2 + p_3\right)^{2}
}{
\left(\lambda p_1 +\lambda p_2+p_2 + p_3\right)^{11}
}
$$

```{r defineadd12vs34}

## 1&2 vs 3&4;  scoreline 9-2 (again)
add12vs34 <- function(lambda){
    H3 <- hyper3()
    H3[c(p1=lambda,p2=lambda)] %<>% inc(9)
    H3[c(p3=1,p2=1)] %<>% inc(2)
    H3[c(p1=lambda,p2=lambda,p3=1,p2=1)] %<>% dec(11)
    return(H3 + as.hyper3(H))
}
```

```{r useadd12vs34, cache=TRUE}
add12vs34(1.88)
maxp(add12vs34(2.0), give=TRUE)$value
maxp(add12vs34(2.1), give=TRUE)$value
```

```{r defineprofilelikelihood12vs34}
L <- function(lambda){maxp(add12vs34(lambda), give=TRUE)$value}
```

```{r useprofilelikelihood12vs34, cache=TRUE}
lam <- exp(seq(from=log(0.9), to=log(60), len=20))
support <- sapply(lam, L)
```

```{r optimize12v34, cache=TRUE}
o <- optimize(L, interval=c(exp(1.5),exp(2.5)), maximum=TRUE)
o
lambda_max <- o$maximum
```

```{r plotproflike}
support <- support - max(support)
plot(log(lam), support,
     pch=16, type="b",
     xlab=expression(log(lambda)), ylab="log-Likelihood"
     )
abline(h=c(0,-2))
abline(v=c(0,log(lambda_max)), lty=2)
```

```{r maxtennis3, cache=TRUE}
tennis_hyper3_maxp <- maxp(add12vs34(lambda_max))
```

Compare the `hyper2` evaluates:

```{r, showallmax}
tennis_maxp
tennis_ghost_maxp
tennis_hyper3_maxp
```

Visuals:

```{r plotallmax}
x1 <- tennis_maxp
x2 <- tennis_ghost_maxp[-1]
x2 <- x2/sum(x2)
x3 <- tennis_hyper3_maxp
x1
x2
x3
```

Above, observe how x2 and x3 are identical: evidently, in this case,
`hyper2` formalism is formally identical to `hyper3` formalism.
Visually:

```{r showall}
par(pty="s")
l <- c(0.15,0.3)
plot(x1, x2, asp=1, xlim=l, ylim=l, pch=16)
abline(0,1)
```

### Package dataset {-}

Following lines create `tennis.rda`, residing in the `data/` directory of the package.

```{r,label=saveghosttennisdataset}
save(tennis, tennis_maxp, tennis_ghost, tennis_ghost_maxp, file="tennis.rda")
```

### References {-}