### 2025-12-06 inst/pwa3.Rmd
pwa3() example

---
title: "Person with advantage functionality for `hyper3` objects: `pwa3()`"
output: bookdown::html_document2
author: "Robin K. S. Hankin"
link-citations: true
bibliography: hyper2.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

```{r}
pwa3
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.

```{r definelamf}
lam <- function(lambda){
  race_nohelp <-
    race(letters[c(1,2,4,3)]) + 
    race(letters[c(3,1,2,4)]) +
    race(letters[c(2,4,1,3)])

  race_help <- 
    race(letters[c(2,3,1,4)]) |> pwa3("b", lambda) + 
    race(letters[c(1,3,4,2)]) |> pwa3("b", lambda) +
    race(letters[c(3,2,1,4)]) |> pwa3("b", lambda)

  maxp(as.hyper3(race_nohelp) + race_help, give=TRUE)$value
}
```

```{r calclamf, cache=TRUE}
l <- exp(seq(from=log(0.02), to=log(8), len=17))
like <- sapply(l,lam)
```

```{r optimizelamf, cache=TRUE}
optimal_sol <- optimize(lam, c(0.01, 3), maximum=TRUE)
```

```{r showoptimlam}
optimal_sol
```

```{r lamfminus2, cache=TRUE}
f <- function(l){lam(l) - optimal_sol$objective + 2}
lower <- uniroot(f, interval = c(0.01, 1))
upper <- uniroot(f, interval = c(1, 10))
```

```{r showroots}
lower
upper
```

```{r plotlamf}
plot(log(l), like - optimal_sol$objective, type='b')
abline(h=c(0, -2))
abline(v = log(optimal_sol$maximum))
segments(x0 = log(c(lower$root, upper$root)), y0 = -2.25, y1 = -1.75)
```