### 2025-12-06 inst/pwa3.Rmd
better example

---
title: "Person with advantage functionality for `hyper3` objects: `pwa3()`"
output: bookdown::html_document2
author: "Robin K. S. Hankin"
link-citations: true
bibliography: hyper2.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

```{r}
pwa3
```

To cite the `hyper2` package in publications, please use
@hankin2017_rmd.  Here we analyse some simple Plackett-Luce
observations, using `hyper3` likelihood functions.

Our observations are a total of six order statistics for four
competitors `a,b,c,d`.  These competitors race three times under
Plackett-Luce probabilities.  They then race three times, but with
competitor `b` having some advantage (steroids?), parametrized by his
Plackett-Luce strength being $\lambda b$.

Our object of inference is $\lambda$; the Plackett-Luce strengths
`a,b,c,d` are nuisance parameters.

```{r definelamf}
lam <- function(lambda){

  race_nohelp <-
    race(c("b","a","d","c")) + 
    race(c("c","b","a","d")) + 
    race(c("b","d","a","c"))

  race_help <- 
    race(c("a","c","d","b")) |> pwa3("b", lambda) + 
    race(c("c","d","b","a")) |> pwa3("b", lambda) + 
    race(c("d","c","a","b")) |> pwa3("b", lambda)

  maxp(as.hyper3(race_nohelp) + race_help, give=TRUE)$value
}
```

Above we see that in the unaided case (the first three races), `b`
does quite well, coming first twice and second once.  In the aided
case, `b` does much worse, coming either third or fourth.  Evidently,
the "help" actually hinders `b` and we suspect that $\lambda<1$.

```{r calclamf, cache=TRUE}
l <- exp(seq(from=log(0.0002), to=log(2), len=7))
like <- sapply(l,lam)
```

```{r optimizelamf, cache=TRUE}
optimal_sol <- optimize(lam, c(0.01, 3), maximum=TRUE)
```

```{r showoptimlam}
optimal_sol
```

```{r lamfminus2, cache=TRUE}
f <- function(l){lam(l) - optimal_sol$objective + 2}
lower <- uniroot(f, interval = c(exp(-8), exp(-7)))
upper <- uniroot(f, interval = c(exp(-2), exp( 0)))
```

```{r showroots}
lower
upper
```

```{r plotlamf, echo=TRUE,fig.cap="Log-likelihood function for lambda"}
plot(log(l), like - optimal_sol$objective, type='b')
abline(h = c(0, -2))
abline(v = log(optimal_sol$maximum))
abline(v = 0)
segments(x0 = log(c(lower$root, upper$root)), y0 = -2.25, y1 = -1.75, col='gray')
```

Figure \@ref(fig:plotlamf) shows