### 2025-12-09 inst/grid.Rmd
grid placing work

---
title: "The starting grid in Formula 1: 2025"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy from vignettes/

---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.

```{r}
x <- readLines("grid_2025.txt")
```

```{r}
drivers <- function(i){strsplit(x[i], " ")[[1]][-1]}
venue   <- function(i){strsplit(x[i], " ")[[1]][ 1]}
drivers(11)
allvenues <- sapply(1:24,venue)
alldrivers <- NULL
for(i in 1:24){alldrivers <- c(alldrivers,drivers(i))}
alldrivers <- sort(unique(alldrivers))
M <- matrix(0,nrow=24, ncol=length(alldrivers))
colnames(M) <- alldrivers
rownames(M) <- allvenues
for(i in 1:24){M[i,] <- alldrivers %in%  drivers(i)    }
M
rowSums(M)
colSums(M)   
```

Above we see that Colapinto and Doohan were the only drivers that did
not compete at every venue.  We will have to remove these two drivers
from the data:

```{r}
grid_order <- list()
unwanted <- c("Colapinto", "Doohan")
for(i in 1:24){
    jj <- drivers(i)
    grid_order[[i]] <- jj[!(jj %in% unwanted)]
}
grid_order
```

We are going to use the 2025 finishing dataset:

```{r, read2025, cache=TRUE}
a <- read.table("~/rstudio/hyper2/inst/formula1_2025.txt")
a <- a[1:19,-ncol(a)]   # remove Colapinto and Doohan; and the points
```

We need to iterate through each venue, and use `pwa3()` to modify the
likelihood function for each one.  Then add the log-likelihoods for
each venue together.

```{r f1griddef}

geometric <- function(lambda, x){setNames(lambda^(seq_along(x)-1),x)}

f1grid <- function(lambda){

    H <- list()
    for(i in 1:24){
        vec <- as.numeric(a[,1])
        vec[is.na(vec)] <- 0
        jj <- setNames(vec, rownames(a))
        H[[i]] <- ordervec2supp(jj) |> pwa3(geometric(lambda,grid_order[[i]]))
    }
    out <- hyper3()
    for(i in 1:24){
        out <- out + H[[i]]
    }
    return(out)
}
```

```{r calcpoll, cache=TRUE}
jj <- f1grid(0.5)
```

```{r maxcalc, cache=TRUE}
mH <- maxp(jj)
```

```{r}
mH
```