### 2025-12-12 inst/grid.Rmd
more output

---
title: "The starting grid in Formula 1: 2025"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy from vignettes/

---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017_rmd.

```{r}
x <- readLines("grid_2025.txt")
```

```{r}
drivers <- function(i){strsplit(x[i], " ")[[1]][-1]}
venue   <- function(i){strsplit(x[i], " ")[[1]][ 1]}
drivers(11)
allvenues <- sapply(1:24,venue)
alldrivers <- NULL
for(i in 1:24){alldrivers <- c(alldrivers,drivers(i))}
alldrivers <- sort(unique(alldrivers))
M <- matrix(0,nrow=24, ncol=length(alldrivers))
colnames(M) <- alldrivers
rownames(M) <- allvenues
for(i in 1:24){M[i,] <- alldrivers %in%  drivers(i)    }
M
rowSums(M)
colSums(M)   
```

Above we see that Colapinto and Doohan were the only drivers that did
not compete at every venue.  We will have to remove these two drivers
from the data:

```{r}
grid_order <- list()
unwanted <- c("Colapinto", "Doohan")
for(i in 1:24){
    jj <- drivers(i)
    grid_order[[i]] <- jj[!(jj %in% unwanted)]
}
grid_order
```

We are going to use the 2025 finishing dataset:

```{r, read2025, cache=TRUE}
a <- read.table("~/rstudio/hyper2/inst/formula1_2025.txt")
a <- a[1:19,-ncol(a)]   # remove Colapinto and Doohan; and the points
```

We need to iterate through each venue, and use `pwa3()` to modify the
likelihood function for each one.  Then add the log-likelihoods for
each venue together.

```{r f1griddef}

geometric <- function(lambda, x){setNames(lambda^(seq_along(x)-1),x)}

f1grid <- function(lambda){

    H <- list()
    for(i in 1:24){
        suppressWarnings(vec <- as.numeric(a[,1]))
        vec[is.na(vec)] <- 0
        jj <- setNames(vec, rownames(a))
        H[[i]] <- ordervec2supp(jj) |> pwa3(geometric(lambda,grid_order[[i]]))
    }
    out <- hyper3()
    for(i in 1:24){
        out <- out + H[[i]]
    }
    return(out)
}
```

```{r calcpolltime, cache=TRUE}
supptime  <- system.time(jj <- f1grid(0.5))
```

```{r maxcalctime, cache=TRUE}
maxtime <- system.time(mH <- maxp(jj))
```

```{r showresultstime}
supptime
maxtime
mH
pie(mH)
```

```{r domore, cache=TRUE}
o <- function(lambda){maxp(f1grid(lambda), give=TRUE)$value}
timeo1 <- system.time(l1 <- o(1.01))
timeo2 <- system.time(l2 <- o(1.02))
```

```{r}
timeo1
timeo2
l1
l2
```

```{r domanyLikes, cache=TRUE}
lam <- seq(from=0.9, to=1, len=17)
L <- sapply(lam, o)
```

```{r plotF1lamf, echo=TRUE, fig.cap="Log-likelihood function for lambda, the pole position measure"}
plot(lam, L - max(L), pch=16, type="b")
```

Figure \@ref(fig:plotF1lamf) shows a log-likelihood function for
$\lambda$.  It is very irregular, indicating that the optimization
routine is having difficulty finding a global maximum for all values
of $\lambda$.